### 2024-01-16 theart.qmd
quarto document initiated

---
title: "Level-1 PDD criteria something"
format: pdf
---

### 2024-01-16 theart.qmd
quarto doc template

@@
 ---
-title: "Level-1 PDD criteria something"
-format: pdf
+title: "Preoperative Cognitive Profile Predictive of Cognitive Decline after Subthalamic Deep Brain Stimulation in Parkinson’s Disease"
+shorttitle: "Cognition in PD after STN DBS"
+author:
+  - name: Josef Mana
+    corresponding: false
+    orcid: "0000-0002-7817-3978"
+    email: "josef.mana@protonmail.com"
+    role:
+      - Conceptualization
+      - Data curation
+      - Investigation
+      - Formal analysis
+      - Methodology
+      - Software
+      - Visualization
+      - Writing - original draft
+      - Writing - review & editing
+    affiliations:
+      - id: "1FM"
+        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
+        department: "Department of Neurology and Centre of Clinical Neuroscience"
+  - name: Ondrej Bezdicek
+    corresponding: true
+    orcid: "0000-0002-5108-0181"
+    email: "ondrej.bezdicek@gmail.com"
+    role:
+      - Conceptualization
+      - Data curation
+      - Investigation
+      - Methodology
+      - Supervision
+      - Writing - original draft
+      - Writing - review & editing
+    affiliations:
+      - id: "1FM"
+        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
+        department: "Department of Neurology and Centre of Clinical Neuroscience"
+  - name: Filip Ruzicka
+    corresponding: false
+    email: "filr@seznam.cz"
+    role:
+      - Investigation
+      - Writing - review & editing
+    affiliations:
+      - id: "1FM"
+        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
+        department: "Department of Neurology and Centre of Clinical Neuroscience"
+  - name: Anna Fecikova
+    corresponding: false
+    email: "Anna.Fecikova@vfn.cz"
+    role:
+      - Investigation
+    affiliations:
+      - id: "1FM"
+        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
+        department: "Department of Neurology and Centre of Clinical Neuroscience"
+  - name: Olga Klempirova
+    corresponding: false
+    email: "novakovaol@seznam.cz"
+    role:
+      - Investigation
+    affiliations:
+      - id: "1FM"
+        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
+        department: "Department of Neurology and Centre of Clinical Neuroscience"
+  - name: Tomas Nikolai
+    corresponding: false
+    email: "nikolai@centrum.cz"
+    role:
+      - Investigation
+      - Writing - review & editing
+    affiliations:
+      - id: "1FM"
+        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
+        department: "Department of Neurology and Centre of Clinical Neuroscience"
+  - name: Tereza Uhrova
+    corresponding: false
+    email: "tereza.uhrova@vfn.cz"
+    role:
+      - Investigation
+    affiliations:
+      - id: "1FM"
+        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
+        department: "Department of Neurology and Centre of Clinical Neuroscience"
+  - name: Evzen Ruzicka
+    corresponding: false
+    email: "eruzi@lf1.cuni.cz"
+    role:
+      - Conceptualization
+      - Funding acquisition
+      - Investigation
+      - Writing - review & editing
+    affiliations:
+      - id: "1FM"
+        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
+        department: "Department of Neurology and Centre of Clinical Neuroscience"
+  - name: Dusan Urgosik
+    corresponding: false
+    email: "urgosik@gmail.com"
+    role:
+      - Investigation
+    affiliations:
+      - id: "NHH"
+        name: "Na Homolce Hospital, Prague, Czech Republic"
+        department: "Department of stereotactic and radiation neurosurgery"
+  - name: Robert Jech
+    corresponding: false
+    orcid: "0000-0002-9732-8947"
+    email: "jech@cesnet.cz"
+    role:
+      - Conceptualization
+      - Data curation
+      - Funding acquisition
+      - Investigation
+      - Resources
+      - Supervision
+      - Writing - review & editing
+    affiliations:
+      - id: "1FM"
+        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
+        department: "Department of Neurology and Centre of Clinical Neuroscience"
+abstract: ""
+keywords:
+  - Parkinson’s disease
+  - dementia
+format:
+ apaquarto-pdf:
+   documentmode: man
+   a4paper: true
+   pdf-engine: lualatex
+   floatsintext: false
+   keep-tex: false
 ---
+
+{{< include _extensions/wjschne/apaquarto/_apa_title.qmd >}}
+
+# Introduction
+
+# Methods
+
+# Results
+
+# Discussion
+
+# References

### 2024-01-16 theart.qmd
theatre initiated

--- 2024-01-16 theart.qmd
@@
 ---
-title: "Level-1 PDD criteria something"
-format: pdf
+title: "Updated Criteria for the Diagnostic Procedure for Parkinson's Disease Dementia on Level I"
+shorttitle: "Parkinson's Disease Dementia Level I Criteria"
+author:
+  - name: Martina Kvapilova
+    corresponding: false
+    orcid: 
+    email: 
+    role:
+      - Conceptualization
+      - Data curation
+      - Writing - original draft
+    affiliations:
+      - id: "1FM"
+        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
+        department: "Department of Neurology and Centre of Clinical Neuroscience"
+  - name: Josef Mana
+    corresponding: true
+    orcid: "0000-0002-7817-3978" 
+    email: "josef.mana@protonmail.com"
+    role:
+      - Conceptualization
+      - Data curation
+      - Investigation
+      - Formal analysis
+      - Software
+      - Methodology
+      - Project administration
+      - Validation
+      - Writing - original draft
+    affiliations:
+      - id: "1FM"
+        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
+        department: "Department of Neurology and Centre of Clinical Neuroscience"
+  - name: Martin Cihak
+    corresponding: false
+    email: "martin.cih@gmail.com"
+    role:
+      - Conceptualization
+      - Investigation
+      - Methodology
+      - Writing - original draft
+    affiliations:
+      - id: "1FM"
+        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
+        department: "Department of Neurology and Centre of Clinical Neuroscience"
+  - name: Ondrej Bezdicek
+    corresponding: false
+    email: "ondrej.bezdicek@gmail.com"
+    role:
+      - Investigation
+      - Data curation
+      - Funding acquisition
+      - Conceptualization
+      - Project administration
+      - Supervision
+      - Writing - original draft
+    affiliations:
+      - id: "1FM"
+        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
+        department: "Department of Neurology and Centre of Clinical Neuroscience"
+  - name: Tereza Uhrova
+    corresponding: false
+    email: "tereza.uhrova@vfn.cz"
+    role:
+      - Investigation
+    affiliations:
+      - id: "1FM"
+        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
+        department: "Department of Neurology and Centre of Clinical Neuroscience"
+  - name: Robert Jech
+    corresponding: false
+    orcid: "0000-0002-9732-8947"
+    email: "jech@cesnet.cz"
+    role:
+      - Funding acquisition
+      - Resources
+      - Writing - review & editing
+    affiliations:
+      - id: "1FM"
+        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
+        department: "Department of Neurology and Centre of Clinical Neuroscience"
+abstract: ""
+keywords:
+  - Parkinson’s disease
+  - dementia
+  - level I criteria
+format:
+ apaquarto-pdf:
+   documentmode: man
+   a4paper: true
+   pdf-engine: lualatex
+   floatsintext: false
+   keep-tex: false
+bibliography: references.bib
 ---
+
+{{< include _extensions/wjschne/apaquarto/_apa_title.qmd >}}
+
+# Introduction
+
+# Methods
+
+Participants
+
+The data of patients with idiopathic PD diagnosed by a movement disorder specialist fulfilling the Movement Disorder Society (MDS) Clinical Diagnostic Criteria for Parkinson's disease (PD) ([@postuma2015]) were retrospectively gathered from clinical records acquired between January 2014 and December 2023.
+All the patients underwent a neuropsychological assessment by a trained clinical psychologist (OB) during a routine examination of cognitive functions as a part of the evaluation process for the indication of Deep Brain Stimulation (DBS) at General University Hospital in Prague.
+The Ethics Committee of the General University Hospital in Prague had approved the study protocol.
+All patients provided written informed consent prior to the examination.
+
+Neuropsychological Assessment
+
+Participants were assessed with both MMSE (Stepankova et al., 2014; Folstein et al., 1975) and MoCA (Kopecek et al., 2017; Nassredine et al., 2005) to measure overall cognitive performance.
+Moreover, a comprehensive neuropsychological assessment was performed in accordance with MDS Task Force Level 2 criteria for MCI in PD (Litvan et al., 2012).
+We described our battery including a regression based calculator for normative scores in another study (Bezdicek et al., 2017).
+Besides other neuropsychological tests, the comprehensive assessment employed Clock Drawing Test (CDT) and Letter Fluency.
+To establish cognitive performance in individual cognitive domains according to Level 1 criteria for PDD, we used test scores proposed by Dubois et al., (2007) and their analogues in MoCA, see Table A for details.
+
+# Results
+
+# Discussion
+
+# References

### 2024-01-16 theart.qmd
methods section added to theart

--- 2024-01-16 theart.qmd
@@
 ---
-title: "Level-1 PDD criteria something"
-format: pdf
+title: "Updated Criteria for the Diagnostic Procedure for Parkinson's Disease Dementia on Level I"
+shorttitle: "Parkinson's Disease Dementia Level I Criteria"
+author:
+  - name: Martina Kvapilova
+    corresponding: false
+    role:
+      - Conceptualization
+      - Data curation
+      - Writing - original draft
+    affiliations:
+      - id: "1FM"
+        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
+        department: "Department of Neurology and Centre of Clinical Neuroscience"
+  - name: Josef Mana
+    corresponding: true
+    orcid: "0000-0002-7817-3978" 
+    email: "josef.mana@protonmail.com"
+    role:
+      - Conceptualization
+      - Data curation
+      - Investigation
+      - Formal analysis
+      - Software
+      - Methodology
+      - Project administration
+      - Validation
+      - Writing - original draft
+    affiliations:
+      - id: "1FM"
+        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
+        department: "Department of Neurology and Centre of Clinical Neuroscience"
+  - name: Martin Cihak
+    corresponding: false
+    email: "martin.cih@gmail.com"
+    role:
+      - Conceptualization
+      - Investigation
+      - Methodology
+      - Writing - original draft
+    affiliations:
+      - id: "1FM"
+        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
+        department: "Department of Neurology and Centre of Clinical Neuroscience"
+  - name: Ondrej Bezdicek
+    corresponding: false
+    email: "ondrej.bezdicek@gmail.com"
+    role:
+      - Investigation
+      - Data curation
+      - Funding acquisition
+      - Conceptualization
+      - Project administration
+      - Supervision
+      - Writing - original draft
+    affiliations:
+      - id: "1FM"
+        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
+        department: "Department of Neurology and Centre of Clinical Neuroscience"
+  - name: Tereza Uhrova
+    corresponding: false
+    email: "tereza.uhrova@vfn.cz"
+    role:
+      - Investigation
+    affiliations:
+      - id: "1FM"
+        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
+        department: "Department of Neurology and Centre of Clinical Neuroscience"
+  - name: Robert Jech
+    corresponding: false
+    orcid: "0000-0002-9732-8947"
+    email: "jech@cesnet.cz"
+    role:
+      - Funding acquisition
+      - Resources
+      - Writing - review & editing
+    affiliations:
+      - id: "1FM"
+        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
+        department: "Department of Neurology and Centre of Clinical Neuroscience"
+abstract: ""
+keywords:
+  - Parkinson’s disease
+  - dementia
+  - level I criteria
+format:
+ apaquarto-pdf:
+   documentmode: man
+   a4paper: true
+   pdf-engine: lualatex
+   floatsintext: false
+   keep-tex: false
+bibliography: references.bib
 ---
+
+{{< include _extensions/wjschne/apaquarto/_apa_title.qmd >}}
+
+# Introduction
+
+# Methods
+
+## Participants
+
+The data of patients with idiopathic PD diagnosed by a movement disorder specialist fulfilling the Movement Disorder Society (MDS) Clinical Diagnostic Criteria for Parkinson's disease (PD) [@postuma2015] were retrospectively gathered from clinical records acquired between January 2014 and December 2023.
+All the patients underwent a neuropsychological assessment by a trained clinical psychologist (OB) during a routine examination of cognitive functions as a part of the evaluation process for the indication of Deep Brain Stimulation (DBS) at General University Hospital in Prague.
+The Ethics Committee of the General University Hospital in Prague had approved the study protocol.
+All patients provided written informed consent prior to the examination.
+
+## Neuropsychological Assessment
+
+Participants were assessed with both MMSE (Stepankova et al., 2014; [@folstein1975] and MoCA [@kopecek2016]; [@nasreddine2005] to measure overall cognitive performance.
+Moreover, a comprehensive neuropsychological assessment was performed in accordance with MDS Task Force Level 2 criteria for MCI in PD [@litvan2012].
+We described our battery including a regression based calculator for normative scores in another study [@bezdicek2017].
+Besides other neuropsychological tests, the comprehensive assessment employed Clock Drawing Test (CDT) and Letter Fluency.
+To establish cognitive performance in individual cognitive domains according to Level 1 criteria for PDD, we used test scores proposed by Dubois et al., [@dubois2007] and their analogues in MoCA, see Table A for details.
+
+Insert Table
+
+To measure independence in activities of daily living, we administered Functional Activities Questionnaire (FAQ; Bezdicek et al., 201; [@pfeffer1982].
+To assess neuropsychiatric functioning, we used Beck Depression Inventory II (BDI-II; [@ciharova2020]; [@beck1996] and State-Trait Anxiety Inventory (STAI; Spielberg et al., 1970, Mullner et al., 1980).
+Psychotic symptoms were assessed in an interview by a trained psychiatrist.
+
+# Results
+
+# Discussion
+
+# References

### 2024-01-18 theart.qmd
bib

@@
 
-Participants were assessed with both MMSE (Stepankova et al., 2014; [@folstein1975] and MoCA [@kopecek2016]; [@nasreddine2005] to measure overall cognitive performance.
+Participants were assessed with both MMSE [@stepankova2014; @folstein1975] and MoCA [@kopecek2016]; [@nasreddine2005] to measure overall cognitive performance.
 Moreover, a comprehensive neuropsychological assessment was performed in accordance with MDS Task Force Level 2 criteria for MCI in PD [@litvan2012].
@@
 Besides other neuropsychological tests, the comprehensive assessment employed Clock Drawing Test (CDT) and Letter Fluency.
-To establish cognitive performance in individual cognitive domains according to Level 1 criteria for PDD, we used test scores proposed by Dubois et al., [@dubois2007] and their analogues in MoCA, see Table A for details.
+To establish cognitive performance in individual cognitive domains according to Level 1 criteria for PDD, we used test scores proposed by [@dubois2007] and their analogues in MoCA, see Table A for details.
 
@@
 
-To measure independence in activities of daily living, we administered Functional Activities Questionnaire (FAQ; Bezdicek et al., 201; [@pfeffer1982].
-To assess neuropsychiatric functioning, we used Beck Depression Inventory II (BDI-II; [@ciharova2020]; [@beck1996] and State-Trait Anxiety Inventory (STAI; Spielberg et al., 1970, Mullner et al., 1980).
+To measure independence in activities of daily living, we administered Functional Activities Questionnaire (FAQ; [@bezdicek2011; @pfeffer1982].
+To assess neuropsychiatric functioning, we used Beck Depression Inventory II (BDI-II; [@ciharova2020]; [@beck1996] and State-Trait Anxiety Inventory (STAI; Spielberg et al., 1970, @mullner1980\].
 Psychotic symptoms were assessed in an interview by a trained psychiatrist.

### 2024-01-18 theart.qmd
citations for methods finished

@@
   - name: Josef Mana
-    corresponding: true
+    corresponding: false
     orcid: "0000-0002-7817-3978" 
@@
     corresponding: false
+    orcid: "0000-0001-7765-4799"
     email: "martin.cih@gmail.com"
@@
     corresponding: false
+    orcid: "0000-0002-5108-0181"
     email: "ondrej.bezdicek@gmail.com"
@@
 
-Participants were assessed with both MMSE [@stepankova2014; @folstein1975] and MoCA [@kopecek2016]; [@nasreddine2005] to measure overall cognitive performance.
+Participants were assessed with both MMSE [@stepankova2014; @folstein1975] and MoCA [@kopecek2016; @nasreddine2005] to measure overall cognitive performance.
 Moreover, a comprehensive neuropsychological assessment was performed in accordance with MDS Task Force Level 2 criteria for MCI in PD [@litvan2012].
@@
 
-To measure independence in activities of daily living, we administered Functional Activities Questionnaire (FAQ; [@bezdicek2011; @pfeffer1982].
-To assess neuropsychiatric functioning, we used Beck Depression Inventory II (BDI-II; [@ciharova2020]; [@beck1996] and State-Trait Anxiety Inventory (STAI; Spielberg et al., 1970, @mullner1980\].
+To measure independence in activities of daily living, we administered Functional Activities Questionnaire (FAQ) [@bezdicek2011; @pfeffer1982].
+To assess neuropsychiatric functioning, we used Beck Depression Inventory II (BDI) [@ciharova2020; @beck1996] and State-Trait Anxiety Inventory (STAI) [@spielberg1983; @mullner1980].
 Psychotic symptoms were assessed in an interview by a trained psychiatrist.

### 2024-04-10 theart.qmd
Insert Table_comparison old and new crit

@@
 
-Insert Table
+```{r}
+#| label: tbl-martin
+#| tbl-cap: Comparison of criteia.
+#| apa-note: The note below the table.
+
+tibble::tribble(
+  ~"Criteria", ~"Former", ~"New",
+  "", "MMSE < 26", "MoCA < 26", 
+  "Attention", "Serial 7's/months reversed", "Serial 7's",
+  "Executive functions", "Lexical Fluency (S)/ Clock Drawing Test", "Lexical Fluency (K)/ Clock Drawing Test",
+  "Visuo-spatial functions", "Drawing of Pentagons", "Drawing of Cube",
+  "Memory", "3-Word Recall", "5-Word Recall",
+  "Language", "-", "Naming of Animals",
+  ) |>
+  flextable::flextable() |>
+  flextable::theme_apa()
+```
 

### 2024-06-12 theart.qmd
Uploaded article, introduction added

@@
 ---
-title: "Updated Criteria for the Diagnostic Procedure for Parkinson's Disease Dementia on Level I"
+lang: en
+title: "Updated Criteria for the Diagnostic Procedure for Parkinson’s Disease Dementia on Level I and their Validity in Deep Brain Stimulation Cohort"
 shorttitle: "Parkinson's Disease Dementia Level I Criteria"
@@
     corresponding: false
+    orcid: "0009-0007-4665-3946"
     role:
@@
         department: "Department of Neurology and Centre of Clinical Neuroscience"
-  - name: Ondrej Bezdicek
-    corresponding: false
-    orcid: "0000-0002-5108-0181"
-    email: "ondrej.bezdicek@gmail.com"
-    role:
-      - Investigation
-      - Data curation
-      - Funding acquisition
-      - Conceptualization
-      - Project administration
-      - Supervision
-      - Writing - original draft
-    affiliations:
-      - id: "1FM"
-        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
-        department: "Department of Neurology and Centre of Clinical Neuroscience"
   - name: Tereza Uhrova
@@
         department: "Department of Neurology and Centre of Clinical Neuroscience"
+  - name: Ondrej Bezdicek
+    corresponding: true
+    orcid: "0000-0002-5108-0181"
+    email: "ondrej.bezdicek@gmail.com"
+    role:
+      - Investigation
+      - Data curation
+      - Funding acquisition
+      - Conceptualization
+      - Project administration
+      - Supervision
+      - Writing - original draft
+    affiliations:
+      - id: "1FM"
+        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
+        department: "Department of Neurology and Centre of Clinical Neuroscience"
 abstract: ""
@@
   - Parkinson’s disease
-  - dementia
+  - Parkinson's disease dementia
   - level I criteria
 format:
- apaquarto-pdf:
+ apaquarto-docx:
    documentmode: man
@@
 # Introduction
+
+Parkinson’s disease (PD) is a neurodegenerative disorder characterised by the slow, progressive onset of rigidity, bradykinesia, and resting tremors that ultimately advance to motor and non-motor impairment [@postuma2015].
+
+The diagnostic criteria for Parkinson’s disease dementia (PDD) were developed in 2006 and their procedure followed shortly after [@emre2007; @dubois2007].
+At that period the PDD criteria were oriented at criteria originating in Alzheimer’s disease (AD) since there was a lack of PD biomarkers to promote a biological classification of PD [@yamashita2023; @höglinger2024].
+
+The newer Parkinson’s disease mild cognitive criteria (PD-MCI) led to an advance in different aspects of PD research [@litvan2012].
+These included progress in clinical, genetic, therapeutic modalities, clinical trials or conversion rates to PDD [@hoogland2017; @hoogland2019; @aarsland2021].
+
+At present, there is an ongoing effort to revisit the PDD criteria to secure the comparability of studies and clinical trials across different sites [@kulisevsky2024].
+In our previous research, we aimed at establishing rigorous psychometric procedure to differentiate PD-MCI from PDD [@bezdicek2017; @bezdicek2016].
+Hence the diagnosis of PDD is an exclusion criterion in deep brain stimulation (DBS), we decided to test the revisited PDD criteria in our PD cohort selected for undergoing DBS [@deuschl2006].
+
+The current study aims at showing the agreement and differences between establishing the diagnosis using only Level 1 criteria as suggested by previous diagnostic procedure (the Movement Disorders Society (MDS)-Task Force screening checklist for detecting PD) [@emre2007; @dubois2007] in comparison to newly revised [@kulisevsky2024] one to validate them.
+The second objective of our study was to determine the agreement of revised PDD criteria with full neuropsychological testing at Level 2.
 
@@
 
-To measure independence in activities of daily living, we administered Functional Activities Questionnaire (FAQ) [@bezdicek2011; @pfeffer1982].
-To assess neuropsychiatric functioning, we used Beck Depression Inventory II (BDI) [@ciharova2020; @beck1996] and State-Trait Anxiety Inventory (STAI) [@spielberg1983; @mullner1980].
+To measure independence in activities of daily living, we administered the Functional Activities Questionnaire (FAQ) [@bezdicek2011; @pfeffer1982].
+To assess neuropsychiatric functioning, we used Beck Depression Inventory-II (BDI) [@ciharova2020; @beck1996] and State-Trait Anxiety Inventory (STAI) [@spielberg1983; @mullner1980].
 Psychotic symptoms were assessed in an interview by a trained psychiatrist.

### 2024-11-24 texts/theart.qmd
Re-working the pipeline to targets

--- 2024-06-12 theart.qmd
@@
 format:
- apaquarto-docx:
+ apaquarto-pdf:
    documentmode: man

### 2024-11-24 R/data_wranggling.R

@@ -1,0 +1,265 @@
+# This is a script containing functions for data import and wranggling before the main analysis.
+
+
+# LIST PATH TO A FILE----
+list_path <- function(folder, file) here(folder, file)
+
+
+# READ DATA FROM CSV ----
+read_data <- function(path, separator) read.csv(path, sep = separator)
+
+
+# IMPORT ITEM-LEVEL DATA ----
+import_outcome_data <- function(data, names, return = "data") {
+  
+  # rename columns and mutate variables such that they are well aligned with REDCap data
+  d0 <- data %>%
+    
+    # rename variables:
+    #   first (mostly) item-data-specific variables
+    #   then item-data/metadata shared variables
+    #   finally convert to lowercase
+    rename("id" = "IPN", "born" = "born_NA_RC", "sex" = "gender_NA_RC", "hand" = "hand_NA_RC") %>%
+    rename("mmse" = "MMSE_tot", "faq" = "FAQ_seb", "bdi" = "BDI.II", "staix1" = "STAI_1", "staix2" = "STAI_2") %>%
+    `colnames<-`( tolower( colnames(.) ) ) %>%
+    
+    # re-code some
+    mutate(
+      sex = case_when( sex == "F" ~ "female", sex == "M" ~ "male" ),
+      hand = case_when( hand == "R" ~ "right", hand == "L" ~ "left" ),
+      across( c("firstname","surname"), ~ make_clean_names( .x, allow_dupes = T ) ),
+      assdate = as.Date(assdate)
+    ) %>%
+    
+    # add inclusion indicator column
+    mutate(incl = 1, .after = id)
+  
+  # !duplicated cases, rows selected manually:
+  # IPN138: keep the first assessment because it is the 'screening' in REDCap
+  # IPN347: keep the first assessment because the second one was just one year later & the first one is REDCap's "screening"
+  # IPN661: keep the second assessment which was three years after the first one & is REDCap's "screening"
+  d0[with(d0, id == "IPN138" & assdate == "2018-02-07"), "incl"] <- 0
+  d0[with(d0, id == "IPN347" & assdate == "2021-01-25"), "incl"] <- 0
+  d0[with(d0, id == "IPN661" & assdate == "2019-10-23"), "incl"] <- 0
+  
+  # IPN225's name from item-data is consistent with IPN335 in REDCap (i.e., metadata)
[…]

### 2024-11-24 _targets.R

@@ -1,0 +1,43 @@
+# This script runs targets pipeline of the R section of the project.
+
+# Load packages required to define the pipeline:
+library(targets)
+
+# Set target options:
+tar_option_set(
+  
+  packages = c(
+    
+    "here", # for path listing
+    "tidyverse", # for data wrangling
+    "janitor" # for cleaning names of diacritics
+    #"purr"
+    
+  )
+)
+
+# Load all in-house functions:
+tar_source()
+
+# List the targets:
+list(
+  
+  # FILES PREP ----
+  tar_target( outcome_path, list_path("_raw", "PDD_cr1t2.0.csv"), format = "file" ), # item-specific data
+  tar_target( meta_path, list_path("_raw","ITEMPO-ManaExportNeuropsych_DATA_2024-01-23_1525.csv"), format = "file" ), # meta-data
+  tar_target( id_path, list_path("_raw", "ITEMPO_DATA_2024-01-17_1153.csv"), format = "file" ), # patients' IDs
+  tar_target( scoring_path, list_path("helpers","test_scoring.csv"), format = "file" ), # tests' scoring
+  
+  # DATA IMPORT ----
+  tar_target( outcome_file, read_data(outcome_path, ";") ), # item-specific data
+  tar_target( meta_file, read_data(meta_path, ",") ), # meta-data
+  tar_target( IDs, read_data(id_path, ",") ), # patients' IDs
+  tar_target( scoring, read_data(scoring_path, ";") ), # tests' scoring
+  
+  tar_target( names, import_outcome_data(outcome_file, IDs, "names") ), # discrepancies between outcome- and meta-data
+  tar_target( outcomes, import_outcome_data(outcome_file, IDs, "data") ), # outcome item-level data
+  tar_target( metadata, import_metadata(meta_file, outcomes, scoring) ), # extract meta-data
+  tar_target( discrepancies, compatibility_check(outcomes, metadata) ), # discrepancies, ought to be empty
+  tar_target( data, merge_data(outcomes, metadata) ) # merge outcome item-level data and meta-data
+  
+)

### 2024-11-24 _targets_packages.R

@@ -1,0 +1,13 @@
+# Generated by targets::tar_renv(): do not edit by hand
+library(bslib)
+library(crew)
+library(gt)
+library(here)
+library(janitor)
+library(markdown)
+library(rstudioapi)
+library(shiny)
+library(shinyWidgets)
+library(shinybusy)
+library(tidyverse)
+library(visNetwork)

### 2024-11-24 scripts/00_import.R

@@ -1,300 +1,0 @@
-# This is a script used to prepare the data for the level-I PDD criteria study.
-
-# clear the environment
-rm( list = ls() )
-gc()
-
-# list packages to be used
-pkgs <- c("here","tidyverse","janitor","purrr")
-
-# load or install each of the packages as needed
-for ( i in pkgs ) {
-  if ( i %in% rownames( installed.packages() ) == F ) install.packages(i) # install if it ain't installed yet
-  if ( i %in% names( sessionInfo()$otherPkgs ) == F ) library( i , character.only = T ) # load if it ain't loaded yet
-}
-
-# prepare a data folder
-if ( !dir.exists("_data") ) dir.create("_data")
-
-
-# IN-HOUSE FUNCTIONS ---
-
-# printing rounded numbers
-rprint <- function( x , dec = 2 ) sprintf( paste0("%.",dec,"f"), round( x , dec ) )
-
-
-# DATA READ ----
-
-# data
-d0 <- read.csv( here("_raw","PDD_cr1t2.0.csv"), sep = ";" ) # the PDD criteria specific data
-d1 <- read.csv( here("_raw","ITEMPO-ManaExportNeuropsych_DATA_2024-01-23_1525.csv"), sep = "," ) # REDCap data
-
-# helpers
-nm <- read.csv( here("_raw","ITEMPO_DATA_2024-01-17_1153.csv"), sep = "," ) # patient's identificators
-sc <- read.csv( here("helpers","test_scoring.csv"), sep = ";" ) # scoring for all tests included
-
-
-# DATA PRE-PROCESSING ----
-
-# ---- PDDcrit specific data ----
-
-# rename columns and mutate variables such that they are well aligned with REDCap data
-d0 <-
-  
-  d0 %>%
-  
-  # renaming, first (mostly) d0-specific variables, then d0/d1 shared variables, finally converting to lowercase
-  rename( "id" = "IPN", "born" = "born_NA_RC", "sex" = "gender_NA_RC", "hand" = "hand_NA_RC" ) %>%
-  rename( "mmse" = "MMSE_tot", "faq" = "FAQ_seb", "bdi" = "BDI.II", "staix1" = "STAI_1", "staix2" = "STAI_2") %>%
-  `colnames<-`( tolower( colnames(.) ) ) %>%
-  
-  # re-code some
-  mutate(
-    sex = case_when( sex == "F" ~ "female", sex == "M" ~ "male" ),
[…]

### 2025-01-05 texts/theart.qmd
Updating Martaska

@@
 format:
- apaquarto-pdf:
+ apaquarto-doc:
    documentmode: man

### 2025-01-05 texts/theart.qmd
Text update

@@
 
-Parkinson’s disease (PD) is a neurodegenerative disorder characterised by the slow, progressive onset of rigidity, bradykinesia, and resting tremors that ultimately advance to motor and non-motor impairment [@postuma2015].
+Parkinson’s disease (PD) is a neurodegenerative disorder characterized by a gradual and progressive onset of motor symptoms, including rigidity, bradykinesia, and resting tremors, which eventually extend to both motor and non-motor impairments [@postuma2015].
+Beyond these hallmark features, cognitive decline is a critical aspect of the disease trajectory, culminating in Parkinson’s disease dementia (PDD) in a substantial subset of patients.
 
-The diagnostic criteria for Parkinson’s disease dementia (PDD) were developed in 2006 and their procedure followed shortly after [@emre2007; @dubois2007].
-At that period the PDD criteria were oriented at criteria originating in Alzheimer’s disease (AD) since there was a lack of PD biomarkers to promote a biological classification of PD [@yamashita2023; @höglinger2024].
+The diagnostic criteria for PDD, first formalized in 2006, were heavily influenced by frameworks established for Alzheimer’s disease (AD) due to the absence of PD-specific biomarkers that could support a biologically informed classification system [@emre2007; @dubois2007].
+While effective at the time, these criteria lacked the granularity needed to address the unique pathophysiological and cognitive features of PD [@yamashita2023].
 
-The newer Parkinson’s disease mild cognitive criteria (PD-MCI) led to an advance in different aspects of PD research [@litvan2012].
-These included progress in clinical, genetic, therapeutic modalities, clinical trials or conversion rates to PDD [@hoogland2017; @hoogland2019; @aarsland2021].
+Subsequent advancements in the field have introduced the concept of Parkinson’s disease mild cognitive impairment (PD-MCI), providing a nuanced understanding of cognitive deficits in PD.
+The PD-MCI criteria, published in 2012, have propelled progress in diverse areas, including clinical characterization, genetic correlates, therapeutic strategies, clinical trial design, and the evaluation of progression rates to PDD [@litvan2012; @hoogland2017; @hoogland2019; @aarsland2021].
 
-At present, there is an ongoing effort to revisit the PDD criteria to secure the comparability of studies and clinical trials across different sites [@kulisevsky2024].
-In our previous research, we aimed at establishing rigorous psychometric procedure to differentiate PD-MCI from PDD [@bezdicek2017; @bezdicek2016].
-Hence the diagnosis of PDD is an exclusion criterion in deep brain stimulation (DBS), we decided to test the revisited PDD criteria in our PD cohort selected for undergoing DBS [@deuschl2006].
+Currently, efforts are underway to refine the PDD diagnostic framework to enhance its consistency and utility in research and clinical contexts across multiple centers [@kulisevsky2024].
+Our previous investigations have emphasized the importance of rigorous psychometric methodologies in differentiating PD-MCI from PDD, especially given that PDD remains a contraindication for deep brain stimulation (DBS) [@bezdicek2016; @deuschl2006].
 
-The current study aims at showing the agreement and differences between establishing the diagnosis using only Level 1 criteria as suggested by previous diagnostic procedure (the Movement Disorders Society (MDS)-Task Force screening checklist for detecting PD) [@emre2007; @dubois2007] in comparison to newly revised [@kulisevsky2024] one to validate them.
-The second objective of our study was to determine the agreement of revised PDD criteria with full neuropsychological testing at Level 2.
+The present study evaluates the diagnostic concordance between the original Level I PDD criteria, as outlined by the Movement Disorder Society (MDS) Task Force [@emre2007; @dubois2007] and the recently revised criteria [@kulisevsky2024] within a PD cohort selected for DBS.
+Specifically, we aimed to:
+
+Compare diagnostic outcomes using the original and revised Level I criteria and assess the agreement of revised Level I criteria with comprehensive neuropsychological assessment at Level II.Compare diagnostic outcomes using the original and revised Level I criteria.
+
+By addressing these objectives, the study seeks to validate the revised PDD criteria and explore their applicability in the context of DBS eligibility, thereby contributing to the refinement of cognitive assessment protocols in PD.
 
@@
 
-The data of patients with idiopathic PD diagnosed by a movement disorder specialist fulfilling the Movement Disorder Society (MDS) Clinical Diagnostic Criteria for Parkinson's disease (PD) [@postuma2015] were retrospectively gathered from clinical records acquired between January 2014 and December 2023.
-All the patients underwent a neuropsychological assessment by a trained clinical psychologist (OB) during a routine examination of cognitive functions as a part of the evaluation process for the indication of Deep Brain Stimulation (DBS) at General University Hospital in Prague.
-The Ethics Committee of the General University Hospital in Prague had approved the study protocol.
-All patients provided written informed consent prior to the examination.
+The study included retrospective data from patients diagnosed with idiopathic PD by a movement disorder specialist according to the MDS Clinical Diagnostic Criteria for PD [@postuma2015] Clinical records from January 2014 to December 2023 were analyzed.
+All participants underwent neuropsychological evaluations conducted by a trained clinical psychologist as part of routine cognitive assessments for determining eligibility for Deep Brain Stimulation (DBS) at the General University Hospital in Prague.
+
+The study protocol was approved by the Ethics Committee of the General University Hospital in Prague.
+Informed consent was obtained from all patients prior to their neuropsychological assessments, in compliance with ethical research standards.
 
@@
 
-Participants were assessed with both MMSE [@stepankova2014; @folstein1975] and MoCA [@kopecek2016; @nasreddine2005] to measure overall cognitive performance.
-Moreover, a comprehensive neuropsychological assessment was performed in accordance with MDS Task Force Level 2 criteria for MCI in PD [@litvan2012].
-We described our battery including a regression based calculator for normative scores in another study [@bezdicek2017].
-Besides other neuropsychological tests, the comprehensive assessment employed Clock Drawing Test (CDT) and Letter Fluency.
-To establish cognitive performance in individual cognitive domains according to Level 1 criteria for PDD, we used test scores proposed by [@dubois2007] and their analogues in MoCA, see Table A for details.
+Cognitive performance was evaluated using both the Mini-Mental State Examination (MMSE) [@stepankova2014; @folstein1975] and the Montreal Cognitive Assessment (MoCA) [@kopecek2016; @nasreddine2005] providing measures of global cognitive functioning.
+A comprehensive neuropsychological battery was administered in accordance with Level 2 criteria for Parkinson’s disease mild cognitive impairment (PD-MCI) proposed by the MDS Task Force [@litvan2012].
+The details of this battery, including the regression-based calculator for normative scoring, have been outlined in a previous publication [@bezdicek2017].
 
+In addition to other neuropsychological measures, the assessment incorporated the Clock Drawing Test (CDT) and Letter Fluency tasks to evaluate specific cognitive domains.
+For determining cognitive performance aligned with Level 1 criteria for Parkinson’s disease dementia (PDD), test scores defined by Dubois et al.
+[@dubois2007] nd their MoCA equivalents were utilized.
+Details of the scoring methods and their alignment with diagnostic criteria are summarized in Table A.
 ```{r}

### 2025-03-02 texts/theart.qmd
Introduction adjustement and operationalisation set up

@@
       - Validation
-      - Writing - original draft
-    affiliations:
-      - id: "1FM"
-        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
-        department: "Department of Neurology and Centre of Clinical Neuroscience"
-  - name: Martin Cihak
-    corresponding: false
-    orcid: "0000-0001-7765-4799"
-    email: "martin.cih@gmail.com"
-    role:
-      - Conceptualization
-      - Investigation
-      - Methodology
       - Writing - original draft
@@
 format:
- apaquarto-doc:
+ apaquarto-docx:
    documentmode: man
@@
 
-{{< include _extensions/wjschne/apaquarto/_apa_title.qmd >}}
-
 # Introduction
@@
 
-The diagnostic criteria for PDD, first formalized in 2006, were heavily influenced by frameworks established for Alzheimer’s disease (AD) due to the absence of PD-specific biomarkers that could support a biologically informed classification system [@emre2007; @dubois2007].
-While effective at the time, these criteria lacked the granularity needed to address the unique pathophysiological and cognitive features of PD [@yamashita2023].
+The diagnostic criteria for PDD, first formalized in 2006, were heavily influenced by frameworks established for Alzheimer’s disease (AD) owing to the absence of PD-specific biomarkers that could facilitate a biologically grounded diagnostic system [@emre2007; @dubois2007].
+Although these criteria provided a valuable initial foundation, they lacked the specificity required to capture the distinct pathophysiological and cognitive features of PD-related dementia [@yamashita2023].
 
-Subsequent advancements in the field have introduced the concept of Parkinson’s disease mild cognitive impairment (PD-MCI), providing a nuanced understanding of cognitive deficits in PD.
-The PD-MCI criteria, published in 2012, have propelled progress in diverse areas, including clinical characterization, genetic correlates, therapeutic strategies, clinical trial design, and the evaluation of progression rates to PDD [@litvan2012; @hoogland2017; @hoogland2019; @aarsland2021].
+An important feature of the original criteria was the provision of an algorithm that allowed for flexibility in test selection [@dubois2007].
+Specifically, clinicians could choose between months reversed or seven backwards for attention assessment, lexical fluency or clock drawing for executive function evaluation, MMSE pentagons for visuospatial ability, and three-word recall for memory assessment.
+The availability of multiple operationalization options for PDD enhances the ability to examine the psychometric properties of the construct.
+Agreement across different criteria allows for the parallel computation of inter-rater reliability , which, in turn, facilitates the calculation of construct validity [@conway1995] further strengthening the diagnostic framework for PDD.
 
-Currently, efforts are underway to refine the PDD diagnostic framework to enhance its consistency and utility in research and clinical contexts across multiple centers [@kulisevsky2024].
-Our previous investigations have emphasized the importance of rigorous psychometric methodologies in differentiating PD-MCI from PDD, especially given that PDD remains a contraindication for deep brain stimulation (DBS) [@bezdicek2016; @deuschl2006].
+Subsequent advancements have introduced the concept of Parkinson’s disease mild cognitive impairment (PD-MCI), refining the understanding of cognitive dysfunction in PD.
+The PD-MCI criteria, published in 2012, have propelled progress in diverse areas, enabling improvements in clinical characterization, identification of genetic correlates, therapeutic interventions, clinical trial design, and the assessment of progression risk to PDD [@litvan2012; @hoogland2017; @hoogland2019; @aarsland2021].
 
-The present study evaluates the diagnostic concordance between the original Level I PDD criteria, as outlined by the Movement Disorder Society (MDS) Task Force [@emre2007; @dubois2007] and the recently revised criteria [@kulisevsky2024] within a PD cohort selected for DBS.
+Currently, efforts are focused on refining the PDD diagnostic framework to improve its consistency and applicability in both research and clinical contexts across multiple centers [@kulisevsky2024].
+Our prior investigations have underscored the importance of employing rigorous psychometric methodologies to differentiate PD-MCI from PDD, particularly given that a diagnosis of PDD remains a contraindication for deep brain stimulation (DBS) [@bezdicek2016; @deuschl2006].
+
+The present study evaluates the diagnostic concordance between the original Level I PDD criteria, as established by the Movement Disorder Society (MDS) Task Force [@emre2007; @dubois2007] and the recently revised criteria [@kulisevsky2024] within a PD cohort selected for DBS.
 Specifically, we aimed to:
 
-Compare diagnostic outcomes using the original and revised Level I criteria and assess the agreement of revised Level I criteria with comprehensive neuropsychological assessment at Level II.Compare diagnostic outcomes using the original and revised Level I criteria.
+Compare diagnostic outcomes utilizing the original and revised Level I criteria and assess the agreement of revised Level I criteria with comprehensive neuropsychological assessment at Level II.Compare diagnostic outcomes using the original and revised Level I criteria.
 
-By addressing these objectives, the study seeks to validate the revised PDD criteria and explore their applicability in the context of DBS eligibility, thereby contributing to the refinement of cognitive assessment protocols in PD.
+By addressing these objectives, this study seeks to validate the revised PDD criteria and evaluate their relevance in the context of DBS eligibility, thereby contributing to the refinement of cognitive assessment protocols in PD.
 
@@
 Details of the scoring methods and their alignment with diagnostic criteria are summarized in Table A.
+
 ```{r}
@@
 
+## Operationalization of dementia
+
+In this study, we applied three different types of criteria for diagnosing PDD on level I.
+The first was based on the original framework [@dubois2007], which included assessments of attention, executive function, visuospatial abilities, and memory and was based Mini-Mental State Examination [@folstein1975; @stepankova2014].
+The second was based on the recent call for change of diagnostic guidelines [@kulisevsky2024], refining cognitive domain assessments to improve sensitivity to PD-related dementia, with specific items from the Montreal Cognitive Assessment (MoCA) utilized for this purpose.
+The third approach applied the Czech version of the shortened Montreal Cognitive Assessment (sMoCA) [@bezdicek2020], testing whether equivalent cognitive impairments can be reliably identified using a more time-efficient testing protocol.
+
+Below, readers can find a structured table summarizing all three diagnostic criteria, facilitating a clearer comparison of their components and applicability.
+
+Insert Table 2
+
 # Results

### 2025-03-03 texts/theart.qmd
Rephrasing of the operationalization

@@
 
-The study included retrospective data from patients diagnosed with idiopathic PD by a movement disorder specialist according to the MDS Clinical Diagnostic Criteria for PD [@postuma2015] Clinical records from January 2014 to December 2023 were analyzed.
-All participants underwent neuropsychological evaluations conducted by a trained clinical psychologist as part of routine cognitive assessments for determining eligibility for Deep Brain Stimulation (DBS) at the General University Hospital in Prague.
+This study retrospectively analyzed clinical data from patients diagnosed with idiopathic PD by a movement disorder specialist according to the MDS Clinical Diagnostic Criteria for PD [@postuma2015] Clinical records spanning January 2014 to December 2023 were examined.
+All participants underwent neuropsychological evaluations conducted by a trained clinical psychologist as part of standard preoperative cognitive assessments for DBS eligibility at the General University Hospital in Prague.
 
-The study protocol was approved by the Ethics Committee of the General University Hospital in Prague.
-Informed consent was obtained from all patients prior to their neuropsychological assessments, in compliance with ethical research standards.
+Ethical approval for the study protocol was obtained from the Ethics Committee of the General University Hospital in Prague.
+Informed consent was secured from all patients prior to their neuropsychological assessments, in adherence to ethical research guidelines.
 
@@
 Cognitive performance was evaluated using both the Mini-Mental State Examination (MMSE) [@stepankova2014; @folstein1975] and the Montreal Cognitive Assessment (MoCA) [@kopecek2016; @nasreddine2005] providing measures of global cognitive functioning.
-A comprehensive neuropsychological battery was administered in accordance with Level 2 criteria for Parkinson’s disease mild cognitive impairment (PD-MCI) proposed by the MDS Task Force [@litvan2012].
-The details of this battery, including the regression-based calculator for normative scoring, have been outlined in a previous publication [@bezdicek2017].
+A comprehensive neuropsychological battery was administered in accordance with the Level II criteria for PD-MCI, as delineated by the MDS Task Force  [@litvan2012].
+The methodological details, including a regression-based normative scoring approach, have been described in previous study [@bezdicek2017].
 
-In addition to other neuropsychological measures, the assessment incorporated the Clock Drawing Test (CDT) and Letter Fluency tasks to evaluate specific cognitive domains.
-For determining cognitive performance aligned with Level 1 criteria for Parkinson’s disease dementia (PDD), test scores defined by Dubois et al.
-[@dubois2007] nd their MoCA equivalents were utilized.
-Details of the scoring methods and their alignment with diagnostic criteria are summarized in Table A.
+In addition to core cognitive assessments, specific tasks such as the Clock Drawing Test (CDT) and Letter Fluency tasks were incorporated to evaluate domain-specific impairments.
+The diagnostic classification of PDD using Level I criteria was based on established scoring thresholds from Dubois et al.
+[@dubois2007] and their corresponding MoCA equivalents.
+A summary of the scoring methodologies and their alignment with diagnostic criteria is provided in Table A.
 
@@
 
-To measure independence in activities of daily living, we administered the Functional Activities Questionnaire (FAQ) [@bezdicek2011; @pfeffer1982].
-To assess neuropsychiatric functioning, we used Beck Depression Inventory-II (BDI) [@ciharova2020; @beck1996] and State-Trait Anxiety Inventory (STAI) [@spielberg1983; @mullner1980].
-Psychotic symptoms were assessed in an interview by a trained psychiatrist.
+To assess functional impairment, the Functional Activities Questionnaire (FAQ) [@bezdicek2011; @pfeffer1982].
+was administered.
+Additionally, neuropsychiatric status was evaluated using the Beck Depression Inventory-II (BDI) [@ciharova2020; @beck1996] and State-Trait Anxiety Inventory (STAI) [@spielberg1983; @mullner1980].
+Psychotic symptoms were assessed through structured psychiatric interviews conducted by a trained psychiatrist.
 
@@
 In this study, we applied three different types of criteria for diagnosing PDD on level I.
-The first was based on the original framework [@dubois2007], which included assessments of attention, executive function, visuospatial abilities, and memory and was based Mini-Mental State Examination [@folstein1975; @stepankova2014].
-The second was based on the recent call for change of diagnostic guidelines [@kulisevsky2024], refining cognitive domain assessments to improve sensitivity to PD-related dementia, with specific items from the Montreal Cognitive Assessment (MoCA) utilized for this purpose.
+The first was based on the original framework [@dubois2007], which was based on MMSE and included assessments of attention, executive function, visuospatial abilities, and memory.
+The second was based on the recent call for change of dementia diagnostic guidelines [@kulisevsky2024], refining cognitive domain assessments to improve sensitivity to PD-related dementia, with specific items from the MoCA utilized for this purpose.
 The third approach applied the Czech version of the shortened Montreal Cognitive Assessment (sMoCA) [@bezdicek2020], testing whether equivalent cognitive impairments can be reliably identified using a more time-efficient testing protocol.
 
-Below, readers can find a structured table summarizing all three diagnostic criteria, facilitating a clearer comparison of their components and applicability.
+To facilitate a clear comparison of these diagnostic frameworks, a structured table summarizing the components and applicability of each criterion is provided below.
 

### 2025-04-11 texts/theart.qmd
Úprava manuskriptu - sekce operacionalizace a neuropsychologické baterie

@@
 author:
-  - name: Martina Kvapilova
+  - name: Martina Mana
     corresponding: false
@@
 
-Cognitive performance was evaluated using both the Mini-Mental State Examination (MMSE) [@stepankova2014; @folstein1975] and the Montreal Cognitive Assessment (MoCA) [@kopecek2016; @nasreddine2005] providing measures of global cognitive functioning.
-A comprehensive neuropsychological battery was administered in accordance with the Level II criteria for PD-MCI, as delineated by the MDS Task Force  [@litvan2012].
-The methodological details, including a regression-based normative scoring approach, have been described in previous study [@bezdicek2017].
+Cognitive performance was evaluated at Level I using the Mini-Mental State Examination (MMSE) [@stepankova2014; @folstein1975] and the Montreal Cognitive Assessment (MoCA) [@kopecek2016; @nasreddine2005], both of which provide measures of global cognitive functioning.
 
-In addition to core cognitive assessments, specific tasks such as the Clock Drawing Test (CDT) and Letter Fluency tasks were incorporated to evaluate domain-specific impairments.
-The diagnostic classification of PDD using Level I criteria was based on established scoring thresholds from Dubois et al.
-[@dubois2007] and their corresponding MoCA equivalents.
-A summary of the scoring methodologies and their alignment with diagnostic criteria is provided in Table A.
+A comprehensive neuropsychological battery was administered in accordance with the Level II criteria for Parkinson's Disease Mild Cognitive Impairment (PD-MCI), as outlined by the MDS Task Force [@litvan2012].
+The methodology, including the use of a regression-based normative scoring approach, has been detailed in a prior study [@bezdicek2017].
+The neuropsychological assessment covered five cognitive domains, each evaluated through specific tests as follows: **attention and working memory**: assessed using Trail Making Test Part A (TMT-A) [@reitan2004], Prague Stroop Test – Dots (PST-D) [@bezdicek2021], Letter Number Sequencing (LNS) [@wechsler1997], WAIS Digit Span Backward (WAIS DSB) [@wechsler1997], and WAIS Corsi Block Backward (WAIS CB) [@wechsler1997]; **executive Function**: evaluated via the Tower of London (ToL) [@specific1982], Categorical Verbal Fluency (CF) [@benton1989], Trail Making Test Part B (TMT-B) [@reitan2004], subtests from the Prague Stroop Test – Words (PST-W) and Colors (PST-C) [@bezdicek2021]; **language**: measured with the WAIS Similarities subtest [@wechsler1997] and the Boston Naming Test (BNT-60) [@kaplan1983]; **memory**: examined using the Rey Auditory Verbal Learning Test (RAVLT) [@rey1964] for immediate and delayed recall (including List B and recognition trials), and the Brief Visual Memory Test–Revised (BVMTR) [@benedict1997], including delayed recall and forced choice recognition; **Visuospatial Function**: assessed through the Judgment of Line Orientation Test (JoLO) [@benton1983], Clock Drawing Test (CLOX) [@royall1998], and the Grooved Pegboard Test (GPT) [@klove1963].
 
-```{r}
-#| label: tbl-martin
-#| tbl-cap: Comparison of criteia.
-#| apa-note: The note below the table.
+In addition to the core cognitive assessments, tasks such as the Clock Drawing Test (CDT) and Letter Fluency tasks were included to capture domain-specific impairments.
+The classification of Parkinson’s Disease Dementia (PDD) based on Level I criteria was determined using established scoring thresholds from Dubois et al.
+[@dubois2007] with corresponding MoCA equivalents.
 
-tibble::tribble(
-  ~"Criteria", ~"Former", ~"New",
-  "", "MMSE < 26", "MoCA < 26", 
-  "Attention", "Serial 7's/months reversed", "Serial 7's",
-  "Executive functions", "Lexical Fluency (S)/ Clock Drawing Test", "Lexical Fluency (K)/ Clock Drawing Test",
-  "Visuo-spatial functions", "Drawing of Pentagons", "Drawing of Cube",
-  "Memory", "3-Word Recall", "5-Word Recall",
-  "Language", "-", "Naming of Animals",
-  ) |>
-  flextable::flextable() |>
-  flextable::theme_apa()
-```
-
-To assess functional impairment, the Functional Activities Questionnaire (FAQ) [@bezdicek2011; @pfeffer1982].
-was administered.
+To assess functional impairment, the Functional Activities Questionnaire (FAQ)[@bezdicek2011; @pfeffer1982] was administered.
 Additionally, neuropsychiatric status was evaluated using the Beck Depression Inventory-II (BDI) [@ciharova2020; @beck1996] and State-Trait Anxiety Inventory (STAI) [@spielberg1983; @mullner1980].
@@
 
-## Operationalization of dementia
+Insert Table
 
-In this study, we applied three different types of criteria for diagnosing PDD on level I.
-The first was based on the original framework [@dubois2007], which was based on MMSE and included assessments of attention, executive function, visuospatial abilities, and memory.
-The second was based on the recent call for change of dementia diagnostic guidelines [@kulisevsky2024], refining cognitive domain assessments to improve sensitivity to PD-related dementia, with specific items from the MoCA utilized for this purpose.
-The third approach applied the Czech version of the shortened Montreal Cognitive Assessment (sMoCA) [@bezdicek2020], testing whether equivalent cognitive impairments can be reliably identified using a more time-efficient testing protocol.
+## Operationalization of Parkinson's Disease Dementia
 
-To facilitate a clear comparison of these diagnostic frameworks, a structured table summarizing the components and applicability of each criterion is provided below.
+In this study, we applied three distinct sets of diagnostic criteria for PDD at Level I.
+The first set was based on the original framework [@dubois2007], which utilized the Mini-Mental State Examination (MMSE) as a global cognitive screening tool, supplemented by assessments of attention, executive function, visuospatial abilities, and memory.
+The second set of criteria was drawn from the recent call for change of dementia diagnostic guidelines [@kulisevsky2024], which advocate for more sensitive cognitive domain assessments in the context of Parkinson's disease (PD).
+This updated approach incorporated specific items from the Montreal Cognitive Assessment (MoCA) to better detect PD-related dementia.
+The third approach applied the Czech version of the shortened Montreal Cognitive Assessment (sMoCA) [@bezdicek2020], a time-efficient modification designed to ascertain whether equivalent cognitive impairments could be reliably identified using a reduced testing protocol.
+Lastly, the fourth approach followed the Level II battery protocol [@bezdicek2017], which is commonly used in the evaluation of PD-MCI.
 
-Insert Table 2
+For each of these diagnostic approaches, we applied two operationalization strategies based on deficits in Instrumental Activities of Daily Living (IADL).
+First, we utilized FAQ item 9, which corresponds to the pill questionnaire from the original criteria [@dubois2007].
+Second, we applied the entire Functional Activities Questionnaire (FAQ), employing a cut-off score of 7 points or higher as suggested in the call for change [@kulisevsky2024].
+These methodologies resulted in a total of 68 operationalisations, which were distributed across different diagnostic criteria: 4 for MMSE, 60 for MoCA, 2 for sMoCA, and 2 for the Level II battery.
+
+To facilitate a clear and structured comparison of these diagnostic frameworks, a detailed table summarizing the components, scoring thresholds, and applicability of each diagnostic criterion is provided below.
+
+Inser Table
 
 # Results
+
+We described continuous variables using mean, standard deviation, median, minimum, and maximum values, while categorical variables were summarized by the number of patients in each category.
+As presented in Table 3, we reported the prevalence of PDD according to each set of diagnostic criteria.
+Specifically, the strictest criteria were found in the sMoCA, which was characterized by a lower threshold for cognitive impairment.
+
+For each of the 68 operationalisations, PDD diagnoses were generated for each patient, and the prevalence of PDD in the entire sample was estimated as a percentage based on the given criteria.
+This approach allowed for a detailed comparison of the different diagnostic frameworks.
+Table 3 displays the estimated prevalence rates of probable PDD across different sets of diagnostic criteria/To enhance clarity, the results presented in Table 3 offer a detailed summary of these operationalisations, showing how variations in diagnostic criteria lead to different estimates of PDD prevalence:
+
+Insert table 3
+
+These findings underscore the variability in prevalence estimates depending on the diagnostic framework employed.
+The stricter sMoCA criteria identified the largest cohort of individuals with PDD, while the MMSE-based criteria, identified the least patients.
+This suggests that different operationalisations of cognitive deficits and functional impairments may yield differing estimates of PDD prevalence within the same sample.
+
+To enhance clarity, the results presented in Table 3 offer a detailed summary of these operationalisations, showing how variations in diagnostic criteria lead to different estimates of PDD prevalence.
 

### 2025-04-11 texts/theart.qmd
Adding citations

@@
 The methodology, including the use of a regression-based normative scoring approach, has been detailed in a prior study [@bezdicek2017].
-The neuropsychological assessment covered five cognitive domains, each evaluated through specific tests as follows: **attention and working memory**: assessed using Trail Making Test Part A (TMT-A) [@reitan2004], Prague Stroop Test – Dots (PST-D) [@bezdicek2021], Letter Number Sequencing (LNS) [@wechsler1997], WAIS Digit Span Backward (WAIS DSB) [@wechsler1997], and WAIS Corsi Block Backward (WAIS CB) [@wechsler1997]; **executive Function**: evaluated via the Tower of London (ToL) [@specific1982], Categorical Verbal Fluency (CF) [@benton1989], Trail Making Test Part B (TMT-B) [@reitan2004], subtests from the Prague Stroop Test – Words (PST-W) and Colors (PST-C) [@bezdicek2021]; **language**: measured with the WAIS Similarities subtest [@wechsler1997] and the Boston Naming Test (BNT-60) [@kaplan1983]; **memory**: examined using the Rey Auditory Verbal Learning Test (RAVLT) [@rey1964] for immediate and delayed recall (including List B and recognition trials), and the Brief Visual Memory Test–Revised (BVMTR) [@benedict1997], including delayed recall and forced choice recognition; **Visuospatial Function**: assessed through the Judgment of Line Orientation Test (JoLO) [@benton1983], Clock Drawing Test (CLOX) [@royall1998], and the Grooved Pegboard Test (GPT) [@klove1963].
+The neuropsychological assessment covered five cognitive domains, each evaluated through specific tests as follows: attention and working memory assessed using Trail Making Test Part A (TMT-A) [@bezdicek2012; @reitan2004], Prague Stroop Test – Dots (PST-D) [@bezdicek2021], Letter Number Sequencing (LNS) [@wechsler1997], WAIS Digit Span Backward (WAIS DSB) [@wechsler1997], and WAIS Corsi Block Backward (WAIS CB) [@wechsler1997]; executive function evaluated via the Tower of London (ToL) [@specific1982], Categorical Verbal Fluency (CF) [@benton1989], Trail Making Test Part B (TMT-B) [@reitan2004], subtests from the Prague Stroop Test – Words (PST-W) and Colors (PST-C) [@bezdicek2021]; language measured with the WAIS Similarities subtest [@wechsler1997] and the Boston Naming Test (BNT-60) [@kaplan1983]; memory examined using the Rey Auditory Verbal Learning Test (RAVLT) [@frydrychova2018; @rey1964] for immediate and delayed recall (including List B and recognition trials), and the Brief Visual Memory Test–Revised (BVMTR) [@havlík2020; @benedict1997], including delayed recall and forced choice recognition; visuospatial function assessed through the Judgment of Line Orientation Test (JoLO) [@benton1983], Clock Drawing Test (CLOX) [@royall1998], and the Grooved Pegboard Test (GPT) [@klove1963].
 

### 2025-04-11 texts/theart.qmd
Adjustments of the text.

@@
 Parkinson’s disease (PD) is a neurodegenerative disorder characterized by a gradual and progressive onset of motor symptoms, including rigidity, bradykinesia, and resting tremors, which eventually extend to both motor and non-motor impairments [@postuma2015].
-Beyond these hallmark features, cognitive decline is a critical aspect of the disease trajectory, culminating in Parkinson’s disease dementia (PDD) in a substantial subset of patients.
+Beyond these hallmark features, cognitive decline is a critical aspect of the disease trajectory, culminating in Parkinson’s disease dementia (PDD) in a substantial subset of patients [@meireles2012].
 
-The diagnostic criteria for PDD, first formalized in 2006, were heavily influenced by frameworks established for Alzheimer’s disease (AD) owing to the absence of PD-specific biomarkers that could facilitate a biologically grounded diagnostic system [@emre2007; @dubois2007].
-Although these criteria provided a valuable initial foundation, they lacked the specificity required to capture the distinct pathophysiological and cognitive features of PD-related dementia [@yamashita2023].
+The diagnostic criteria for PDD [@dubois2007], first formalized in 2006, were heavily influenced by frameworks established for Alzheimer’s disease (AD) owing to the absence of PD-specific biomarkers that could facilitate a biologically grounded diagnostic system.
+Although these criteria provided a valuable initial foundation, they lacked the specificity required to capture the distinct pathophysiological and cognitive features of PD-related dementia [@emre2007; @yamashita2023].
 
@@
 
-Subsequent advancements have introduced the concept of Parkinson’s disease mild cognitive impairment (PD-MCI), refining the understanding of cognitive dysfunction in PD.
-The PD-MCI criteria, published in 2012, have propelled progress in diverse areas, enabling improvements in clinical characterization, identification of genetic correlates, therapeutic interventions, clinical trial design, and the assessment of progression risk to PDD [@litvan2012; @hoogland2017; @hoogland2019; @aarsland2021].
+Subsequent advancements have introduced the concept of Parkinson’s disease mild cognitive impairment (PD-MCI), refining the understanding of cognitive dysfunction in PD [@litvan2012].
+The PD-MCI criteria, published in 2012, have propelled progress in diverse areas, enabling improvements in clinical characterization, identification of genetic correlates, therapeutic interventions, clinical trial design, and the assessment of progression risk to PDD [@hoogland2017; @hoogland2019; @aarsland2021].
 
@@
 
-The present study evaluates the diagnostic concordance between the original Level I PDD criteria, as established by the Movement Disorder Society (MDS) Task Force [@emre2007; @dubois2007] and the recently revised criteria [@kulisevsky2024] within a PD cohort selected for DBS.
-Specifically, we aimed to:
+The present study evaluates the diagnostic concordance between the original Level I PDD criteria, as established by the Movement Disorder Society (MDS) Task Force [@emre2007; @dubois2007] and the recent call for change [@kulisevsky2024] within a PD cohort selected for DBS.
 
-Compare diagnostic outcomes utilizing the original and revised Level I criteria and assess the agreement of revised Level I criteria with comprehensive neuropsychological assessment at Level II.Compare diagnostic outcomes using the original and revised Level I criteria.
+Specifically, we aimed to compare diagnostic outcomes utilizing the original and revised Level I criteria and assess the agreement of revised Level I criteria with comprehensive neuropsychological assessment at Level II.
+Moreover, we aimed to compare diagnostic outcomes using the original and revised Level I criteria.
 
@@
 
-This study retrospectively analyzed clinical data from patients diagnosed with idiopathic PD by a movement disorder specialist according to the MDS Clinical Diagnostic Criteria for PD [@postuma2015] Clinical records spanning January 2014 to December 2023 were examined.
+This study retrospectively analyzed clinical data from patients diagnosed with idiopathic PD by a movement disorder specialist according to the MDS Clinical Diagnostic Criteria for PD [@postuma2015] Clinical records spanning January 2014 to January 2024 were examined.
 All participants underwent neuropsychological evaluations conducted by a trained clinical psychologist as part of standard preoperative cognitive assessments for DBS eligibility at the General University Hospital in Prague.
@@
 
-## Operationalization of Parkinson's Disease Dementia
+## Operationalisation of Parkinson's Disease Dementia
 
@@
 
-For each of these diagnostic approaches, we applied two operationalization strategies based on deficits in Instrumental Activities of Daily Living (IADL).
+For each of these diagnostic approaches, we applied two operationalisation strategies based on deficits in Instrumental Activities of Daily Living (IADL).
 First, we utilized FAQ item 9, which corresponds to the pill questionnaire from the original criteria [@dubois2007].
@@
 
-We described continuous variables using mean, standard deviation, median, minimum, and maximum values, while categorical variables were summarized by the number of patients in each category.
+We described continuous variables using mean, standard deviation, median, minimum, and maximum values.
+Categorical variables were summarized by the number of patients in each category.
 As presented in Table 3, we reported the prevalence of PDD according to each set of diagnostic criteria.
@@
 This approach allowed for a detailed comparison of the different diagnostic frameworks.
-Table 3 displays the estimated prevalence rates of probable PDD across different sets of diagnostic criteria/To enhance clarity, the results presented in Table 3 offer a detailed summary of these operationalisations, showing how variations in diagnostic criteria lead to different estimates of PDD prevalence:
+Table 3 displays the estimated prevalence rates of probable PDD across different sets of diagnostic criteria/To enhance clarity, the results presented in Table 3 offer a detailed summary of these operationalisations, showing how variations in diagnostic criteria lead to different estimates of PDD prevalence.
 
@@
 
-To enhance clarity, the results presented in Table 3 offer a detailed summary of these operationalisations, showing how variations in diagnostic criteria lead to different estimates of PDD prevalence.
-
 # Discussion

### 2025-04-15 _manuscript/manuscript.qmd
Manuscript changes regarding methodology plus some comments.

---
lang: en
title: "Updated Criteria for the Diagnostic Procedure for Parkinson’s Disease Dementia on Level I and their Validity in Deep Brain Stimulation Cohort"
shorttitle: "Parkinson's Disease Dementia Level I Criteria"
author:
  - name: Martina Mana
    corresponding: false
    orcid: "0009-0007-4665-3946"
    role:
      - Conceptualization
      - Data curation
      - Writing - original draft
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Josef Mana
    corresponding: false
    orcid: "0000-0002-7817-3978" 
    email: "josef.mana@lf1.cuni.cz"
    role:
      - Conceptualization
      - Data curation
      - Investigation
      - Formal analysis
      - Software
      - Methodology
      - Project administration
      - Validation
      - Writing - original draft
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Tereza Uhrova
    corresponding: false
    email: "tereza.uhrova@vfn.cz"
    role:
      - Investigation
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Robert Jech
    corresponding: false
    orcid: "0000-0002-9732-8947"
    email: "jech@cesnet.cz"
    role:
      - Funding acquisition
      - Resources
      - Writing - review & editing
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Ondrej Bezdicek
    corresponding: true
    orcid: "0000-0002-5108-0181"
    email: "ondrej.bezdicek@gmail.com"
    role:
      - Investigation
      - Data curation
      - Funding acquisition
      - Conceptualization
      - Project administration
      - Supervision
      - Writing - original draft
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
abstract: ""
keywords:
  - Parkinson’s disease
  - Parkinson's disease dementia
  - level I criteria
format:
 apaquarto-pdf:
   documentmode: man
   a4paper: true
   pdf-engine: lualatex
   floatsintext: false
   keep-tex: false
bibliography: references.bib
floatsintext: false
numbered-lines: false
suppress-title-page: false
warning: false
echo: false
---

```{r}
#| label: "import"

library(targets)
library(tidyverse)

upstore <- here::here(tar_config_get("store"))
tar_source(here::here("R"))
data <- tar_read(raw_data, store = upstore)
crit <- tar_read(pdd_data, store = upstore)$criteria

```

# Introduction

Parkinson’s disease (PD) is a neurodegenerative disorder characterized by a gradual and progressive onset of motor symptoms, including rigidity, bradykinesia, and resting tremors, which eventually extend to both motor and non-motor impairments [@postuma2015]. Beyond these hallmark features, cognitive decline is a critical aspect of the disease trajectory, culminating in Parkinson’s disease dementia (PDD) in a substantial subset of patients [@meireles2012].

The diagnostic criteria for PDD [@dubois2007], first formalized in 2006, were heavily influenced by frameworks established for Alzheimer’s disease (AD) owing to the absence of PD-specific biomarkers that could facilitate a biologically grounded diagnostic system. Although these criteria provided a valuable initial foundation, they lacked the specificity required to capture the distinct pathophysiological and cognitive features of PD-related dementia [@emre2007; @yamashita2023].

An important feature of the original criteria was the provision of an algorithm that allowed for flexibility in test selection [@dubois2007]. Specifically, clinicians could choose between months reversed or seven backwards for attention assessment, lexical fluency or clock drawing for executive function evaluation, MMSE pentagons for visuospatial ability, and three-word recall for memory assessment. The availability of multiple operationalization options for PDD enhances the ability to examine the psychometric properties of the construct. Agreement across different criteria allows for the parallel computation of inter-rater reliability , which, in turn, facilitates the calculation of construct validity [@conway1995] further strengthening the diagnostic framework for PDD.

Subsequent advancements have introduced the concept of Parkinson’s disease mild cognitive impairment (PD-MCI), refining the understanding of cognitive dysfunction in PD [@litvan2012]. The PD-MCI criteria, published in 2012, have propelled progress in diverse areas, enabling improvements in clinical characterization, identification of genetic correlates, therapeutic interventions, clinical trial design, and the assessment of progression risk to PDD [@hoogland2017; @hoogland2019; @aarsland2021].

Currently, efforts are focused on refining the PDD diagnostic framework to improve its consistency and applicability in both research and clinical contexts across multiple centers [@kulisevsky2024]. Our prior investigations have underscored the importance of employing rigorous psychometric methodologies to differentiate PD-MCI from PDD, particularly given that a diagnosis of PDD remains a contraindication for deep brain stimulation (DBS) [@bezdicek2016; @deuschl2006].

The present study evaluates the diagnostic concordance between the original Level I PDD criteria, as established by the Movement Disorder Society (MDS) Task Force [@emre2007; @dubois2007] and criteria inspired by the recent call for change [@kulisevsky2024] within a PD cohort selected for DBS. **Furthermore, both sets of criteria are compared to PDD diagnosed on Level II. The study aims to address following research objectives (RO): 1) to estimate prevalence of PDD in patients considered for DBS, 2) to characterize variability in PDD diagnosis depending on PDD criteria, 3) to evaluate concordance between different sets of PDD criteria, and 4) identify factors that account for variability in PDD diagnosis between criteria.**^[**JM: It is a good practice in science to explicitly specify research objectives, questions or hypotheses to streamline the process. The idea is to create a (literal) line going theory -> objective -> estimand -> estimate -> interpretation (in light of theory presented before with maybe a special focus on mismatches between theory and results/estimate). Check whether the changes make sense to you.**] By addressing these objectives, this study seeks to validate the revised PDD criteria and evaluate their relevance in the context of DBS eligibility, thereby contributing to the refinement of cognitive assessment protocols in PD.

# Methods

## Participants

```{r}
#| label: "dates"

continue <- all(!is.na(data$assdate))
stopifnot("Some assessment dates are missing" = continue)

dates <-
  sapply(
    c("min", "max"),
    function(f) {
      date <- do.call(f, list(data$assdate))
      paste(month(date, label = T, abbr = F), year(date))
    }
  )
```

This study retrospectively analyzed clinical data from a cohort of patients with PD considered for treatment via DBS of subthalamic nucleus at the General University Hospital in Prague. All patients were diagnosed with idiopathic PD by a movement disorder specialist according to the MDS Clinical Diagnostic Criteria for PD [@postuma2015]. Clinical records spanning `r dates["min"]` to `r dates["max"]` were examined. All participants underwent neuropsychological evaluation conducted by a trained clinical psychologist as part of standard preoperative cognitive assessments for DBS eligibility at the General University Hospital in Prague.

Ethical approval for the study protocol was obtained from the Ethics Committee of the General University Hospital in Prague. Informed consent was secured from all patients prior to their neuropsychological assessments, in adherence to ethical research guidelines.

## Neuropsychological Assessment

Cognitive performance was evaluated at both Level I (abbreviated assessment) and Level II (comprehensive assessment) according to the standard MDS neuropsychology battery for Parkinson's Disease Mild Cognitive Impairment (PD-MCI) [@litvan2012, @bezdicek2017]. Level I was assessed using the Mini-Mental State Examination (MMSE) [@stepankova2014; @folstein1975] and the Montreal Cognitive Assessment (MoCA) [@kopecek2016; @nasreddine2005], both of which provide measures of global cognitive functioning. The neuropsychological assessment at Level II covered five cognitive domains, each evaluated through specific tests as follows: attention and working memory assessed using Trail Making Test Part A (TMT-A) [@bezdicek2012; @reitan2004], Prague Stroop Test – Dots (PST-D) [@bezdicek2021], Letter Number Sequencing (LNS) [@wechsler1997], WAIS Digit Span Backward (WAIS DSB) [@wechsler1997], and WAIS Corsi Block Backward (WAIS CB) [@wechsler1997]; executive function evaluated via the Tower of London (ToL) [@specific1982], Categorical Verbal Fluency (CF) [@benton1989], Trail Making Test Part B (TMT-B) [@reitan2004], subtests from the Prague Stroop Test – Words (PST-W) and Colors (PST-C) [@bezdicek2021]; language measured with the WAIS Similarities subtest [@wechsler1997] and the Boston Naming Test (BNT-60) [@kaplan1983]; memory examined using the Rey Auditory Verbal Learning Test (RAVLT) [@frydrychova2018; @rey1964] for immediate and delayed recall (including List B and recognition trials), and the Brief Visual Memory Test–Revised (BVMTR) [@havlík2020; @benedict1997], including delayed recall and forced choice recognition; visuospatial function assessed through the Judgment of Line Orientation Test (JoLO) [@benton1983], Clock Drawing Test (CLOX) [@royall1998], and the Grooved Pegboard Test (GPT) [@klove1963].

In addition to the core cognitive assessments, tasks such as the Clock Drawing Test (CDT) and Letter Fluency tasks were included to capture domain-specific impairments. The classification of Parkinson’s Disease Dementia (PDD) based on Level I criteria was determined using established scoring thresholds from Dubois et al. [@dubois2007] with corresponding MoCA equivalents.

To assess functional impairment, the Functional Activities Questionnaire (FAQ)[@bezdicek2011; @pfeffer1982] was administered. Additionally, neuropsychiatric status was evaluated using the Beck Depression Inventory-II (BDI) [@ciharova2020; @beck1996] and State-Trait Anxiety Inventory (STAI) [@spielberg1983; @mullner1980]. Psychotic symptoms were assessed through structured psychiatric interviews conducted by a trained psychiatrist.

## Theoretical and Empirical Estimands

**Following the framework of @lundberg2021, in this section we link our research objectives to theoretical (i.e., targets of inference) and empirical (i.e., observable data) estimads. Whereas, the theoretical estimand, consisting of  unit-specific quantity and the target population, represents the quantity one would compute to address a research objective under ideal conditions (e.g., data from the whole population or ideal experimental manipulations), the empirical estimand represents the quantity one can compute with the data at hand. The population of interest across our theretical estimands is a population of patients with PD worth consideration for DBS treatment.^[**JM: This will have to be stated better.**] The theoretical estimand related to our RO1 is the prevalence of PDD in patients screened for DBS and theoretical estimand related to our RO2 is the variance of prevalence estimates due to the measurement process. To estimate these quantities, we calculate the proportion of patients with probable PDD (defined below) according to each tested criterion resulting into a set of prevalence estimates. Theoretical estimand related to our RO3 is a set of population-wise contingency tables of probable PDD diagnosis assigned by different PDD criteria and metrics derived from these tables. Our empirical estimands consist of matrixes derived from pairwise Receiver Operating Characteristic (ROC) curve analyses. Finally, the theoretical estimand related to our RO4 is a set of PDD criterion components that, if observed to change, substantially change probability of a probable PDD diagnosis. We approach this RO from an exploratory point of view and evaluate the importance of each PDD criteria component by observing change in empirical estimands for RO1-3 when stratified by different oprationalization decisions.**^[**JM: Will need to work on wording here, but the idea hopefully comes across.**]

## Operationalization of Parkinson's Disease Dementia

In this study, we applied three distinct sets of diagnostic criteria for **probable**^[**JM: We should unite the terminlogy. Most importantly, I reckon, we should differentiate between "probable PDD" (what is implied by meeting all the criteria within an operationalization) and "PDD" (the latent state of patient's cognition); and to use consistently the following trifecta - "criterion," "operationalization" and their "component".**] PDD at Level I. The first set was based on the original framework [@dubois2007], which utilized the Mini-Mental State Examination (MMSE) as a global cognitive screening tool, supplemented by assessments of attention, executive function, visuospatial abilities, and memory. The second set of criteria was drawn from the recent call for change of dementia diagnostic guidelines [@kulisevsky2024], which advocate for more sensitive cognitive domain assessments in the context of Parkinson's disease (PD). This updated approach incorporated specific items from the Montreal Cognitive Assessment (MoCA) to better detect PD-related dementia. The third approach applied the Czech version of the shortened Montreal Cognitive Assessment (sMoCA) [@bezdicek2020], a time-efficient modification designed to ascertain whether equivalent cognitive impairments could be reliably identified using a reduced testing protocol. Lastly, the fourth approach followed the Level II battery protocol, which is commonly used in the evaluation of PD-MCI. **The Level II methodology, including the use of a regression-based normative scoring approach, has been detailed in a prior study [@bezdicek2017]. Refer to @tbl-criteria^[**JM: Placeholder until we make the table proper.**] to a summary of the components and scoring thresholds of each diagnostic criterion. All non-cognitive criteria of probable PDD (i.e., diagnosis of PD that developed before dementia and absence of Major Depression, delirium or other abnormalities that obscure diagnosis were established by an independent neurological and psychiatric assessments and held true for all patients in the sample).**

```{r}
#| label: "tbl-criteria"

gt::gt(crit)
```

[Insert Table 1 here]

```{r}
#| label: "criteria-check"

critnum <-
  sapply(
    c("mmse", "moca", "smoca", "lvlII"),
    function(i)
      subset(crit, group == i) |>
      nrow()
  )

continue <- !any(critnum == 0)
#stopifnot("Some criteria are missing" = continue)
```

For each of these diagnostic approaches, we applied two operationalization strategies based on deficits in Instrumental Activities of Daily Living (IADL). First, we utilized FAQ item 9, which approximates the pill questionnaire from the original criteria [@dubois2007] **employing a cut-off score of 2 points or higher**. Second, we applied the entire Functional Activities Questionnaire (FAQ) as suggested in the call for change [@kulisevsky2024]**, employing a cut-off score of 7 points or higher based on Czech normative data (cite it).**^[**JM: I am pretty sure Kulisevsky et al. did not suggest cut-off $FAQ \geq 7$ We ought to locate the Czech normative study Ondrej bases this value on.**] These methodologies resulted in a total of `r sum(critnum)` operationalizations, which were distributed across different diagnostic criteria: `r critnum["mmse"]` MMSE-based, `r critnum["moca"]` MoCA-based, `r critnum["smoca"]` sMoCA-based, and `r critnum["lvlII"]` based on the Level II battery^[**JM: Need to add Level II operationalization and allow the `stopifnot()` test code above. My job.**].

## Statistical Analyses

For sample description, we summarized continuous variables using mean, standard deviation, median, minimum, and maximum values. Categorical variables were summarized by the number of patients in each category. To address study objectives, we started by repeatedly assigning each patient the diagnosis of probable PDD based on each PDD operationalization listed in @tbl-criteria resulting in a `r sum(data$incl)` (patients) $\times$ `r sum(critnum)` (operationalizations) matrix where each cell indicates whether a patient (row) meets criteria for probable PDD according to an operationalization (column). PDD prevalence estimates were then computed as $\frac{N_{PDD}}{N_{total}}$ separately for each operationalization to address RO1.

To address RO2-4, a set of two class cross-tabulations with associated statistics was computed for each pair of operationalizations via the `confusionMatrix()` function from the R package *caret* [@kuhn2008]. For each pair of operationalizations, the analysis was repeated twice such that each variables of the pair served once as the reference and once as the predictor. Following measures were used to evaluate pairwise concordance between different operationalizations of PDD criteria: 1) Cohen's $\kappa$ with its 95% confidence interval (CI) computed via the `cohen.kappa()` function from the R package *psych* [@revelle2024]; 2) Accuracy (i.e., the proportion of correct predictions, both true positives and true negatives, among the total number of cases) with its 95% CI; 3) Sensitivity/Recall (i.e., the proportion of true positives); and 4) Specificity (i.e., the proportion of true negatives).^[Unlike Cohen's $\kappa$, Accuracy, Sensitivity and Specificity are not symmetrical, i.e., their value depend on which variable is considered reference and which is considered predictor. Consequently, we report these values twice for each pair of operationalizations. Note that the Sensitivity of a reference/predictor pair corresponds to the Positive Predictive Value if their roles were reversed. The same relationship holds true between the Specificity and the Negative Predictive Value.]

Finally, the No Information Rate (NIR) was calculated for each pair of operationalizations. NIR is the accuracy that could be obtained by always predicting the majority class and in our case it is equivalent to the complement of the PDD prevalence estimate according to the reference operationalization. Accuracy of prediction was subsequently compared to the NIR via a one-sided Exact Binomial Test as implemented by the `binom.test()` R stats function. Reference/predictor pairs associated with p < .05 were considered to show significantly better accuracy than NIR. In other word, for reference/predictor pairs associated with p < .05, we conclude that knowing the probable PDD status according to the predictor operationalization helps to estimate the probable PDD status according to the reference operationalization and the two opeartionalizations thus show substantial concordance.

Data wrangling and visualizations were done in the *tidyverse* package [@wickham2019] and tables were formatted in the *gt* package [@iannone2024]. All analyses were conducted with the R (version `r with(version, paste(major, minor, sep="."))`) software environment for statistical computing [@rsoft]. The software code supporting this article is available at [https://github.com/josefmana/DemCr1t.git](https://github.com/josefmana/DemCr1t.git).^[**JM: Do not forget to make it public before submitted!**]

# Results

```{r}
#| label: "results"

desc <- tar_read(sample_description    , store = upstore)
prev <- tar_read(prevalence_summaries  , store = upstore)
conc <- tar_read(concordance_statistics, store = upstore)
perc <- prev$table$perc
```

## Sample Description

A total of `r sum(data$incl)` patients were included. Demographic and clinical characteristics of the sample are summarized in @tbl-desc. **...**^[**JM: Give some brief impressions from the Table 2 here. E.g., evaluate sex proportion, age, (we are missing education for some reason) etc. with general PD population (e.g., from some meta-analysis). Maybe say a word or two about mean cognitive profile or its spread.**]

[Insert Table 2 here]

```{r}
#| label: "tbl-desc"

desc$gtable
```

## Prevalence Estimates

Operationalization-wise prevalence estimates are presented in @tbl-prev. On average, estimated prevalence was `r do_summary(perc, 2, "M")`% (SD = `r do_summary(perc, 2, "SD")`, Md = `r do_summary(perc, 2, "Md")`, range `r do_summary(perc, 2, "minmax")`). Notably, the prevalence estimate was substantially lower when FAQ item 9 was used as a criterion of IADL deficit (M = `r do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, "M")`%, SD = `r do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, "SD")`, range `r do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, "minmax")`) compared to using total FAQ score criterion (M = `r do_summary(perc[!grepl("FAQ 9", prev$table$IADL)], 2, "M")`%, SD = `r do_summary(perc[!grepl("FAQ 9", prev$table$IADL)], 2, "SD")`, range `r do_summary(perc[!grepl("FAQ 9", prev$table$IADL)], 2, "minmax")`) as demonstrated also in @fig-prev.^[**JM: For oneself - work on the code here, it ain't good.**]

[Insert Table 3 and Figure 1^[**JM: For thyself - add description.**] here]

```{r}
#| label: "tbl-prev"

prev$gtable
```

```{r}
#| label: "fig-prev"

prev$plot
```

## Criteria Concordance

Results of the analyses of prediction Accuracy, Cohen's $\kappa$, Sensitivity and Specificity are presented in @fig-acc, @fig-kappa, @fig-sens and @fig-spec respectively. Numerical results are available ... ^[**JM: ... as some kind of Supplementary Table, ideally html but Excel file would work as well.**]

**In this section, we need to discuss among ourselves what and how to report such that the results give enough information, the most important information and are readable as well.**

[Insert Figure 2^[**JM: For thyself - add description.**] here]

```{r}
#| label: "fig-acc"

conc$plots$Accuracy
```

# Discussion

# References

# Appendix

```{r}
#| label: "fig-kappa"

conc$plots$Kappa
```

```{r}
#| label: "fig-sens"

conc$plots$Sensitivity
```

```{r}
#| label: "fig-spec"

conc$plots$Specificity
```

### 2025-04-15 R/summarise_prevalence.R

@@ -74,12 +74,13 @@ summarise_prevalence <- function(d0, vars, descending = T) {
   tab <-
     prevs |>
     left_join(opers, by = 'type') |>
-    select(-type, -perc, -`FALSE`, -`TRUE`) |>
+    select(-type, -`FALSE`, -`TRUE`) |>
     relocate(N, .after = last_col()) |>
     relocate(Prevalence, .after = last_col())
   # Format the table:
   gtab <-
     tab |>
+    select(-perc) |>
     gt_apa_table(tit = '<b>Table 3</br>
     Estimates of prevalence.</b> Estimates of the prevalence of probable PD-D in the sample.'
     ) |>

### 2025-04-17 _manuscript/manuscript.qmd
Slight manuscript changes.

@@
 format:
- apaquarto-pdf:
+ apaquarto-docx:
    documentmode: man
@@
 The present study evaluates the diagnostic concordance between the original Level I PDD criteria, as established by the Movement Disorder Society (MDS) Task Force [@emre2007; @dubois2007] and criteria inspired by the recent call for change [@kulisevsky2024] within a PD cohort selected for DBS.
-**Furthermore, both sets of criteria are compared to PDD diagnosed on Level II.
-The study aims to address following research objectives (RO): 1) to estimate prevalence of PDD in patients considered for DBS, 2) to characterize variability in PDD diagnosis depending on PDD criteria, 3) to evaluate concordance between different sets of PDD criteria, and 4) identify factors that account for variability in PDD diagnosis between criteria.**^[**JM: It is a good practice in science to explicitly specify research objectives, questions or hypotheses to streamline the process.
-The idea is to create a (literal) line going theory -> objective -> estimand -> estimate -> interpretation (in light of theory presented before with maybe a special focus on mismatches between theory and results/estimate).
-Check whether the changes make sense to you.**] By addressing these objectives, this study seeks to validate the revised PDD criteria and evaluate their relevance in the context of DBS eligibility, thereby contributing to the refinement of cognitive assessment protocols in PD.
+Furthermore, both sets of criteria are compared to PDD diagnosed on Level II.
+The study aims to address following research objectives (RO): 1) to estimate the prevalence of PDD among patients considered for DBS (RO1), 2) to assess variability in PDD diagnosis depending on the diagnostic criteria applied (RO2), 3) to evaluate the diagnostic concordance between different sets of PDD criteria (RO3), and 4) to identify diagnostic components that contribute to variability in PDD classification across criteria (RO4).
+By addressing these objectives, this study seeks to validate the revised PDD criteria and evaluate their relevance in the context of DBS eligibility, thereby contributing to the refinement of cognitive assessment protocols in PD.
 
@@
 
-**Following the framework of @lundberg2021, in this section we link our research objectives to theoretical (i.e., targets of inference) and empirical (i.e., observable data) estimads.
-Whereas, the theoretical estimand, consisting of  unit-specific quantity and the target population, represents the quantity one would compute to address a research objective under ideal conditions (e.g., data from the whole population or ideal experimental manipulations), the empirical estimand represents the quantity one can compute with the data at hand.
-The population of interest across our theretical estimands is a population of patients with PD worth consideration for DBS treatment.^[**JM: This will have to be stated better.**] The theoretical estimand related to our RO1 is the prevalence of PDD in patients screened for DBS and theoretical estimand related to our RO2 is the variance of prevalence estimates due to the measurement process.
-To estimate these quantities, we calculate the proportion of patients with probable PDD (defined below) according to each tested criterion resulting into a set of prevalence estimates.
-Theoretical estimand related to our RO3 is a set of population-wise contingency tables of probable PDD diagnosis assigned by different PDD criteria and metrics derived from these tables.
-Our empirical estimands consist of matrixes derived from pairwise Receiver Operating Characteristic (ROC) curve analyses.
-Finally, the theoretical estimand related to our RO4 is a set of PDD criterion components that, if observed to change, substantially change probability of a probable PDD diagnosis.
-We approach this RO from an exploratory point of view and evaluate the importance of each PDD criteria component by observing change in empirical estimands for RO1-3 when stratified by different oprationalization decisions.**^[**JM: Will need to work on wording here, but the idea hopefully comes across.**]
+Following the framework proposed by Lundberg et al.
+(2021), this section connects our research objectives and their corresponding theoretical (i.e., targets of inference) and empirical (i.e., data-driven) estimands.
+The theoretical estimand refers to a unit-specific quantity defined over a target population and represents the ideal quantity that would address the research question under optimal conditions—such as access to complete population data or perfect experimental control.
+In contrast, the empirical estimand corresponds to the quantity that is actually computable using the available dataset, given real-world constraints.
+In our study, the target population for all theoretical estimands is defined as individuals diagnosed with Parkinson’s disease (PD) who are potential candidates for deep brain stimulation (DBS) treatment.
+
+The theoretical estimand corresponding to the RO1 is the true prevalence of probable PDD in this population.
+Empirically, this is estimated by calculating the proportion of patients classified as having probable PDD according to each of the diagnostic criteria under consideration.
+
+For the RO2, the theoretical estimand is the variance in prevalence attributable to the diagnostic process itself—that is, the extent to which different sets of criteria yield differing prevalence estimates for the same population.
+This variability is empirically quantified by comparing the distribution of PDD classifications across criteria within the cohort.
+
+The RO3 concerns diagnostic concordance.
+Here, the theoretical estimand is a set of population-level contingency tables comparing PDD classifications assigned by each pair of diagnostic criteria, along with derived metrics such as sensitivity, specificity, and kappa coefficients.
+The corresponding empirical estimands are represented by matrices generated through pairwise Receiver Operating Characteristic (ROC) curve analyses that evaluate the discriminatory performance of each diagnostic set.
+
+Finally, for the RO4, the theoretical estimand is defined as the set of diagnostic components whose variation systematically alters the probability of a probable PDD diagnosis.
+This aspect of the study is exploratory in nature.
+Empirically, we assess the contribution of each diagnostic feature by examining how variations in operational definitions (e.g., domain-specific thresholds, criteria for functional impairment) influence the empirical estimands derived for the first three objectives.
+This allows us to identify the diagnostic elements most responsible for inter-criterion discrepancies.
+We approach this RO from an exploratory point of view and evaluate the importance of each PDD criteria component by observing change in empirical estimands for RO1-3 when stratified by different oprationalization decisions.
 
@@
 
-In this study, we applied three distinct sets of diagnostic criteria for **probable**^[**JM: We should unite the terminlogy.
-Most importantly, I reckon, we should differentiate between "probable PDD" (what is implied by meeting all the criteria within an operationalization) and "PDD" (the latent state of patient's cognition); and to use consistently the following trifecta - "criterion," "operationalization" and their "component".**] PDD at Level I.
+In this study, we applied three distinct sets of diagnostic criteria for **probable**[^3] PDD at Level I.
 The first set was based on the original framework [@dubois2007], which utilized the Mini-Mental State Examination (MMSE) as a global cognitive screening tool, supplemented by assessments of attention, executive function, visuospatial abilities, and memory.
@@
 **The Level II methodology, including the use of a regression-based normative scoring approach, has been detailed in a prior study [@bezdicek2017].
-Refer to @tbl-criteria^[**JM: Placeholder until we make the table proper.**] to a summary of the components and scoring thresholds of each diagnostic criterion.
-All non-cognitive criteria of probable PDD (i.e., diagnosis of PD that developed before dementia and absence of Major Depression, delirium or other abnormalities that obscure diagnosis were established by an independent neurological and psychiatric assessments and held true for all patients in the sample).**
+Refer to @tbl-criteria**[^4] to a summary of the components and scoring thresholds of each diagnostic criterion.
+All non-cognitive criteria of probable PDD (i.e., diagnosis of PD that developed before dementia and absence of Major Depression, delirium or other abnormalities that obscure diagnosis were established by an independent neurological and psychiatric assessments and held true for all patients in the sample).
+
+[^3]: **JM: We should unite the terminlogy.
+Most importantly, I reckon, we should differentiate between "probable PDD" (what is implied by meeting all the criteria within an operationalization) and "PDD" (the latent state of patient's cognition); and to use consistently the following trifecta - "criterion," "operationalization" and their "component".**
+
+[^4]: **JM: Placeholder until we make the table proper.**
 
@@
 
-[Insert Table 1 here]
+\[Insert Table 1 here\]
 
@@
 For each of these diagnostic approaches, we applied two operationalization strategies based on deficits in Instrumental Activities of Daily Living (IADL).
-First, we utilized FAQ item 9, which approximates the pill questionnaire from the original criteria [@dubois2007] **employing a cut-off score of 2 points or higher**.
-Second, we applied the entire Functional Activities Questionnaire (FAQ) as suggested in the call for change [@kulisevsky2024]**, employing a cut-off score of 7 points or higher based on Czech normative data (cite it).**^[**JM: I am pretty sure Kulisevsky et al.
-did not suggest cut-off $FAQ \geq 7$ We ought to locate the Czech normative study Ondrej bases this value on.**] These methodologies resulted in a total of `r sum(critnum)` operationalizations, which were distributed across different diagnostic criteria: `r critnum["mmse"]` MMSE-based, `r critnum["moca"]` MoCA-based, `r critnum["smoca"]` sMoCA-based, and `r critnum["lvlII"]` based on the Level II battery^[**JM: Need to add Level II operationalization and allow the `stopifnot()` test code above.
-My job.**].
+First, we utilized FAQ item 9, which approximates the pill questionnaire from the original criteria [@dubois2007] employing a cut-off score of 2 points or higher.
+Second, we applied the entire Functional Activities Questionnaire (FAQ) as suggested in the call for change [@kulisevsky2024], employing a cut-off score of 7 points or higher based on Czech normative data [@bezdicek2011].
+These methodologies resulted in a total of `r sum(critnum)` operationalizations, which were distributed across different diagnostic criteria: `r critnum["mmse"]` MMSE-based, `r critnum["moca"]` MoCA-based, `r critnum["smoca"]` sMoCA-based, and `r critnum["lvlII"]` based on the Level II battery[^5].
+
+[^5]: **JM: Need to add Level II operationalization and allow the `stopifnot()` test code above. My job.**
 
@@
 For each pair of operationalizations, the analysis was repeated twice such that each variables of the pair served once as the reference and once as the predictor.
-Following measures were used to evaluate pairwise concordance between different operationalizations of PDD criteria: 1) Cohen's $\kappa$ with its 95% confidence interval (CI) computed via the `cohen.kappa()` function from the R package *psych* [@revelle2024]; 2) Accuracy (i.e., the proportion of correct predictions, both true positives and true negatives, among the total number of cases) with its 95% CI; 3) Sensitivity/Recall (i.e., the proportion of true positives); and 4) Specificity (i.e., the proportion of true negatives).^[Unlike Cohen's $\kappa$, Accuracy, Sensitivity and Specificity are not symmetrical, i.e., their value depend on which variable is considered reference and which is considered predictor.
+Following measures were used to evaluate pairwise concordance between different operationalizations of PDD criteria: 1) Cohen's $\kappa$ with its 95% confidence interval (CI) computed via the `cohen.kappa()` function from the R package *psych* [@revelle2024]; 2) Accuracy (i.e., the proportion of correct predictions, both true positives and true negatives, among the total number of cases) with its 95% CI; 3) Sensitivity/Recall (i.e., the proportion of true positives); and 4) Specificity (i.e., the proportion of true negatives).[^6]
+
+[^6]: Unlike Cohen's $\kappa$, Accuracy, Sensitivity and Specificity are not symmetrical, i.e., their value depend on which variable is considered reference and which is considered predictor.
 Consequently, we report these values twice for each pair of operationalizations.
 Note that the Sensitivity of a reference/predictor pair corresponds to the Positive Predictive Value if their roles were reversed.
-The same relationship holds true between the Specificity and the Negative Predictive Value.]
+The same relationship holds true between the Specificity and the Negative Predictive Value.
 
@@
 Accuracy of prediction was subsequently compared to the NIR via a one-sided Exact Binomial Test as implemented by the `binom.test()` R stats function.
-Reference/predictor pairs associated with p < .05 were considered to show significantly better accuracy than NIR.
-In other word, for reference/predictor pairs associated with p < .05, we conclude that knowing the probable PDD status according to the predictor operationalization helps to estimate the probable PDD status according to the reference operationalization and the two opeartionalizations thus show substantial concordance.
+Reference/predictor pairs associated with p \< .05 were considered to show significantly better accuracy than NIR.
+In other word, for reference/predictor pairs associated with p \< .05, we conclude that knowing the probable PDD status according to the predictor operationalization helps to estimate the probable PDD status according to the reference operationalization and the two opeartionalizations thus show substantial concordance.
 
@@
 All analyses were conducted with the R (version `r with(version, paste(major, minor, sep="."))`) software environment for statistical computing [@rsoft].
-The software code supporting this article is available at [https://github.com/josefmana/DemCr1t.git](https://github.com/josefmana/DemCr1t.git).^[**JM: Do not forget to make it public before submitted!**]
+The software code supporting this article is available at <https://github.com/josefmana/DemCr1t.git>.[^7]
+
+[^7]: **JM: Do not forget to make it public before submitted!**
 
@@
 
-A total of `r sum(data$incl)` patients were included.
-Demographic and clinical characteristics of the sample are summarized in @tbl-desc.
-**...**^[**JM: Give some brief impressions from the Table 2 here.
+A total of `r sum(data$incl)` patients were included. Demographic and clinical characteristics of the sample are summarized in @tbl-desc. **...**[^8]
+
+[^8]: **JM: Give some brief impressions from the Table 2 here.
 E.g., evaluate sex proportion, age, (we are missing education for some reason) etc.
 with general PD population (e.g., from some meta-analysis).
-Maybe say a word or two about mean cognitive profile or its spread.**]
+Maybe say a word or two about mean cognitive profile or its spread.**
 
-[Insert Table 2 here]
+\[Insert Table 2 here\]
 
@@
 On average, estimated prevalence was `r do_summary(perc, 2, "M")`% (SD = `r do_summary(perc, 2, "SD")`, Md = `r do_summary(perc, 2, "Md")`, range `r do_summary(perc, 2, "minmax")`).
-Notably, the prevalence estimate was substantially lower when FAQ item 9 was used as a criterion of IADL deficit (M = `r do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, "M")`%, SD = `r do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, "SD")`, range `r do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, "minmax")`) compared to using total FAQ score criterion (M = `r do_summary(perc[!grepl("FAQ 9", prev$table$IADL)], 2, "M")`%, SD = `r do_summary(perc[!grepl("FAQ 9", prev$table$IADL)], 2, "SD")`, range `r do_summary(perc[!grepl("FAQ 9", prev$table$IADL)], 2, "minmax")`) as demonstrated also in @fig-prev.^[**JM: For oneself - work on the code here, it ain't good.**]
+Notably, the prevalence estimate was substantially lower when FAQ item 9 was used as a criterion of IADL deficit (M = `r do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, "M")`%, SD = `r do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, "SD")`, range `r do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, "minmax")`) compared to using total FAQ score criterion (M = `r do_summary(perc[!grepl("FAQ 9", prev$table$IADL)], 2, "M")`%, SD = `r do_summary(perc[!grepl("FAQ 9", prev$table$IADL)], 2, "SD")`, range `r do_summary(perc[!grepl("FAQ 9", prev$table$IADL)], 2, "minmax")`) as demonstrated also in @fig-prev.[^9]
 
-[Insert Table 3 and Figure 1^[**JM: For thyself - add description.**] here]
+[^9]: **JM: For oneself - work on the code here, it ain't good.**
+
+\[Insert Table 3 and Figure 1[^10] here\]
+
+[^10]: **JM: For thyself - add description.**
 
@@
 Numerical results are available ...
-^[**JM: ...
-as some kind of Supplementary Table, ideally html but Excel file would work as well.**]
+[^11]
+
+[^11]: **JM: ... as some kind of Supplementary Table, ideally html but Excel file would work as well.**
 
@@
 
-[Insert Figure 2^[**JM: For thyself - add description.**] here]
+\[Insert Figure 2[^12] here\]
+
+[^12]: **JM: For thyself - add description.**
 

### 2025-04-17 _manuscript/manuscript.qmd
Changed manuscript back to pdf. For some reason, apaquarto-docx ain't working whereas docx does.

@@
 
-[^3]: **JM: We should unite the terminlogy.
+[^3]: **JM: We should unite the terminology.
 Most importantly, I reckon, we should differentiate between "probable PDD" (what is implied by meeting all the criteria within an operationalization) and "PDD" (the latent state of patient's cognition); and to use consistently the following trifecta - "criterion," "operationalization" and their "component".**

### 2025-04-23 R/data_paths.R
Manuscript changes - neuropsychological assessment part, slight code changes, adaptation of tables, etc.

@@ -26,19 +26,22 @@ data_paths <- function(dir = 'data-raw') {
       'item',
       'demo',
       'meta',
-      'scoring'
+      'scoring',
+      'variables'
     ),
     file = c(
       'ItemData.csv', # Item-wise data (originally 'Level1critPDD2.0.csv')
       'REDCapData.csv', # Demography (originally 'ITEMPO-ManaExportNeuropsych_DATA_2024-12-17_1821.csv')
       'MetaData.csv', # Meta-data (originally 'ITEMPO_DATA_2024-01-17_1153.csv')
-      'TestScoring.csv' # Scoring of psychological variables
+      'TestScoring.csv', # Scoring of psychological variables
+      'VariablesOfInterest.csv' # Variables to be printed in tables
     ),
     message = c(
       paste0('\nItem-wise data is missing or mislabelled!\nInsert a valid file "ItemData.csv" into the /',dir,' folder!\n'),
       paste0('\nREDCap data is missing or mislabelled!\nInsert a valid file "REDCapData.csv" into the /',dir,' folder!\n'),
       paste0('\nPatient identification data is missing or mislabelled!\nInsert a valid file "MetaData.csv" into the /',dir,' folder!\n'),
-      paste0('\nFile with test scoring data is missing or mislabelled!\nInsert a valid file "TestScoring.csv" into the /',dir,' folder!\n')
+      paste0('\nFile with test scoring data is missing or mislabelled!\nInsert a valid file "TestScoring.csv" into the /',dir,' folder!\n'),
+      paste0('\nFile with variable labels is missing or mislabelled!\nInsert a valid file "VariablesOfInterest.csv" into the /',dir,' folder!\n')
     )
   ) |>
     mutate(path = here::here(dir, file))

### 2025-04-23 _targets.R

@@ -33,11 +33,13 @@ list(
   ),
   tar_target(
     name    = sample_description,
-    command = compute_descriptives(raw_data, here::here('data-raw','VariablesOfInterest.csv'))
+    command = compute_descriptives(raw_data, here::here('data-raw','VariablesOfInterest.csv')),
+    cue     = tar_cue("always")
   ),
   tar_target(
     name    = prevalence_summaries,
-    command = summarise_prevalence(pdd_data, here::here('data-raw','VariablesOfInterest.csv'))
+    command = summarise_prevalence(pdd_data, here::here('data-raw','VariablesOfInterest.csv')),
+    cue     = tar_cue("always")
   ),
   tar_target(
     name    = concordance_statistics,

### 2025-05-07 _targets.R
JM: notes to Intro.

@@ -4,12 +4,12 @@ library(tarchetypes)
 # Set target options:
 tar_option_set(
   packages = c(
-    "here",      # for path listing
+    "here", # for path listing
     "tidyverse", # for data wrangling
-    "janitor",   # for cleaning names of diacritics
-    "psych",     # for Cohen's Kappa calculation
-    "caret",     # for confusion matrixes
-    "gt"         # for tableing
+    "janitor", # for cleaning names of diacritics
+    "psych", # for Cohen's Kappa calculation
+    "caret", # for confusion matrixes
+    "gt" # for tableing
   )
 )
 

### 2025-05-07 _manuscript/manuscript.qmd
JM: Notes to methods.

@@
 
-Parkinson’s disease (PD) is a neurodegenerative disorder characterized by a gradual and progressive onset of motor symptoms, including rigidity, bradykinesia, and resting tremors, which eventually extend to both motor and non-motor impairments [@postuma2015].
-Beyond these hallmark features, cognitive decline is a critical aspect of the disease trajectory, culminating in Parkinson’s disease dementia (PDD) in a substantial subset of patients [@meireles2012].
+Parkinson’s disease (PD) is a neurodegenerative disorder characterised by a progressive onset of motor symptoms, including rigidity, bradykinesia, postural instability and resting tremor.
+Beyond motor symptoms, patients routinely suffer from non-motor impairments [@postuma2015], such as cognitive decline.
+This decline might peak in Parkinson’s disease dementia (PDD) in a subset of patients [@meireles2012].
 
-The diagnostic criteria for PDD [@dubois2007], first formalized in 2006, were heavily influenced by frameworks established for Alzheimer’s disease (AD) owing to the absence of PD-specific biomarkers that could facilitate a biologically grounded diagnostic system.
+Despite its clinical relevance, the diagnosis of PDD remains complex.
+*The original **International Parkinson and Movement Disorder Society (MDS)** diagnostic criteria [@dubois2007], first formalized in 2006*^[**JM: I ain't sure it was the first time PDD criteria were formalised; e.g., Aarsland et al.
+- dx.doi.org/10.1001/archneur.60.3.387 - used DSM criteria for PDD already in 2003.
+Consequently I added the specification that these were MDS crtieria.**], were heavily influenced by frameworks established primarily for Alzheimer’s disease (AD).
 Although these criteria provided a valuable initial foundation, they lacked the specificity required to capture the distinct pathophysiological and cognitive features of PD-related dementia [@emre2007; @yamashita2023].
 
-An important feature of the original criteria was the provision of an algorithm that allowed for flexibility in test selection [@dubois2007].
-Specifically, clinicians could choose between months reversed or seven backwards for attention assessment, lexical fluency or clock drawing for executive function evaluation, MMSE pentagons for visuospatial ability, and three-word recall for memory assessment.
-The availability of multiple operationalization options for PDD enhances the ability to examine the psychometric properties of the construct.
+*One notable feature of the original criteria was the provision of an algorithm that allowed for flexibility in test selection [@dubois2007]*^[**JM: This came out of nowhere according to me.
+I think it would make more sense to introduce level I vs II first then introduce the algortihm for level I (i.e., Dubois's Table 2) and focus on it.**].
+Specifically, clinicians could choose between months reversed or seven backwards for attention assessment, lexical fluency or clock drawing for executive function evaluation.
+Moreover, the critaria included MMSE pentagons for evaluation of visuospatial ability, and three-word recall for memory assessment.
 Agreement across different criteria allows for the parallel computation of inter-rater reliability , which, in turn, facilitates the calculation of construct validity [@conway1995] further strengthening the diagnostic framework for PDD.
 
-Subsequent advancements have introduced the concept of Parkinson’s disease mild cognitive impairment (PD-MCI), refining the understanding of cognitive dysfunction in PD [@litvan2012].
-The PD-MCI criteria, published in 2012, have propelled progress in diverse areas, enabling improvements in clinical characterization, identification of genetic correlates, therapeutic interventions, clinical trial design, and the assessment of progression risk to PDD [@hoogland2017; @hoogland2019; @aarsland2021].
+*Subsequent steps have introduced the concept of Parkinson’s disease mild cognitive impairment (PD-MCI), refining the understanding of cognitive dysfunction in PD [@litvan2012].
+The PD-MCI criteria have propelled progress in diverse areas, enabling improvements in clinical characterization, identification of genetic correlates, therapeutic interventions, clinical trial design, and the assessment of progression risk to PDD [@hoogland2017; @hoogland2019; @aarsland2021].*^[**JM: I do not believe this section is needed for the article.
+You could keep it for yourself and use it in your dissertion but discussion history of MCI, albeit briefly, is out of scope of this manuscript in my humble correct opinion.**]
 
-Currently, efforts are focused on refining the PDD diagnostic framework to improve its consistency and applicability in both research and clinical contexts across multiple centers [@kulisevsky2024].
-Our prior investigations have underscored the importance of employing rigorous psychometric methodologies to differentiate PD-MCI from PDD, particularly given that a diagnosis of PDD remains a contraindication for deep brain stimulation (DBS) [@bezdicek2016; @deuschl2006].
+Currently, efforts are focused on refining the PDD diagnostic framework to improve its consistency and applicability in both research and clinical contexts [@kulisevsky2024].
+*Our prior investigations have underscored the importance of employing rigorous psychometric methodologies to differentiate PD-MCI from PDD [@bezdicek2016; @deuschl2006].*^[**JM: The citations do not seem to match the statement - Bezdicek (2016) is about a processing speed theory of cognitive impairment in healthy older adults, Deuschl (2006) is an RCT of DBS.
+I'd drop this sentence altogether.**]
 
-The present study evaluates the diagnostic concordance between the original Level I PDD criteria, as established by the Movement Disorder Society (MDS) Task Force [@emre2007; @dubois2007] and criteria inspired by the recent call for change [@kulisevsky2024] within a PD cohort selected for DBS.
+*According to a recent metanalysis, the epidemiological estimates of PDD prevalence among individuals with PD vary widely, ranging from 14% to 55%^[**JM: It makes sense to report some point estimate as well.**], depending on methodological criteria employed [@sousa2022].
+Moreover, factors such as patients` sex [@cereda2016], age and disease duration appear to modulate the risk of cognitive decline [@oh2016; @rana2012].*^[**JM: Good, I'd push this paragraph higher.**]
+
+The present study evaluates the diagnostic concordance between the original Level I PDD criteria, as established by the MDS Task Force [@emre2007; @dubois2007] and criteria inspired by the recent call for change [@kulisevsky2024] within a cohort of PD patients *selected for DBS*^[**JM: Not true, the cohort is of patients being considered for DBS, not selected for DBS.**].
 Furthermore, both sets of criteria are compared to PDD diagnosed on Level II.
@@
 By addressing these objectives, this study seeks to validate the revised PDD criteria and evaluate their relevance in the context of DBS eligibility, thereby contributing to the refinement of cognitive assessment protocols in PD.
+
+Suggestion: (MK) **(RO1) To estimate the prevalence of Parkinson’s disease dementia (PDD) and evaluate the diagnostic variability and concordance across different PDD criteria.
+(RO2) To identify specific diagnostic components that contribute to variability in PDD classification across the applied criteria.**^[**JM:Nice!
+Let's state it this way.**]
 
@@
 
-This study retrospectively analyzed clinical data from a cohort of patients with PD considered for treatment via DBS of subthalamic nucleus at the General University Hospital in Prague.
+This study retrospectively analyzed clinical data from a cohort of patients with PD at the General University Hospital in Prague.
 All patients were diagnosed with idiopathic PD by a movement disorder specialist according to the MDS Clinical Diagnostic Criteria for PD [@postuma2015].
 Clinical records spanning `r dates["min"]` to `r dates["max"]` were examined.
-All participants underwent neuropsychological evaluation conducted by a trained clinical psychologist as part of standard preoperative cognitive assessments for DBS eligibility at the General University Hospital in Prague.
+All participants underwent neuropsychological evaluation conducted by a trained clinical psychologist (OB) as part of standard preoperative cognitive assessments for DBS eligibility at the General University Hospital in Prague.
 
@@
 Cognitive performance was evaluated at both Level I (abbreviated assessment) and Level II (comprehensive assessment) according to the standard MDS neuropsychology battery for Parkinson's Disease Mild Cognitive Impairment (PD-MCI) [@litvan2012, @bezdicek2017].
-Level I was assessed using the Mini-Mental State Examination (MMSE) [@stepankova2014; @folstein1975] and the Montreal Cognitive Assessment (MoCA) [@kopecek2016; @nasreddine2005], both of which provide measures of global cognitive functioning.
-The neuropsychological assessment at Level II covered five cognitive domains, each evaluated through specific tests as follows: attention and working memory assessed using Trail Making Test Part A (TMT-A) [@bezdicek2012; @reitan2004], Prague Stroop Test – Dots (PST-D) [@bezdicek2021], Letter Number Sequencing (LNS) [@wechsler1997], WAIS Digit Span Backward (WAIS DSB) [@wechsler1997], and WAIS Corsi Block Backward (WAIS CB) [@wechsler1997]; executive function evaluated via the Tower of London (ToL) [@specific1982], Categorical Verbal Fluency (CF) [@benton1989], Trail Making Test Part B (TMT-B) [@reitan2004], subtests from the Prague Stroop Test – Words (PST-W) and Colors (PST-C) [@bezdicek2021]; language measured with the WAIS Similarities subtest [@wechsler1997] and the Boston Naming Test (BNT-60) [@kaplan1983]; memory examined using the Rey Auditory Verbal Learning Test (RAVLT) [@frydrychova2018; @rey1964] for immediate and delayed recall (including List B and recognition trials), and the Brief Visual Memory Test–Revised (BVMTR) [@havlík2020; @benedict1997], including delayed recall and forced choice recognition; visuospatial function assessed through the Judgment of Line Orientation Test (JoLO) [@benton1983], Clock Drawing Test (CLOX) [@royall1998], and the Grooved Pegboard Test (GPT) [@klove1963].
+Level I was assessed using the Mini-Mental State Examination (MMSE) [@stepankova2014; @folstein1975] and the Montreal Cognitive Assessment (MoCA) [@kopecek2016; @nasreddine2005].
+The neuropsychological assessment at Level II covered five cognitive domains, each evaluated through specific tests as follows: attention and working memory assessed using Trail Making Test Part A (TMT-A) [@bezdicek2012; @reitan2004], and WAIS Digit Span Backward (WAIS DSB) [@wechsler1997]; executive function evaluated via Categorical Verbal Fluency (CF) [@benton1989]^[**JM: This one instead - https://doi.org/10.14735/amcsnn2015292**], and subtest from the Prague Stroop Test – Colors (PST-C) [@bezdicek2021]; language measured with the WAIS Similarities subtest [@wechsler1997], and the Boston Naming Test (BNT-60) [@kaplan1983]; memory examined using the Rey Auditory Verbal Learning Test (RAVLT) [@frydrychova2018; @rey1964]^[**JM: Do not forget this - https://doi.org/10.1080/13825585.2013.865699**] for delayed recall, and the Brief Visual Memory Test–Revised (BVMTR) [@havlik2020; @benedict1997] for *delayed recall*^[**JM: The same label is used for RAVLT-DR and BVMT-R-DR here, we should sort it out.**], or WAIS Family Pictures subtest [@wechsler1997] *including delayed recall and forced choice recognition*^[**JM: We did not use any recognition.
+Should tidy this section such that it includes only indexes used by our study.**]; visuospatial function assessed through the Judgment of Line Orientation Test (JoLO) [@benton1983], and Clock Drawing Test (CLOX) [@royall1998].
 
-In addition to the core cognitive assessments, tasks such as the Clock Drawing Test (CDT) and Letter Fluency tasks were included to capture domain-specific impairments.
-The classification of Parkinson’s Disease Dementia (PDD) based on Level I criteria was determined using established scoring thresholds from Dubois et al.
-[@dubois2007] with corresponding MoCA equivalents.
+In addition to the core cognitive assessments, *tasks such as*^[**JM: Stating such as "such as" are usually not appropritate in Methods section of a scientific study, because all methods should be reported exactly.**] the Clock Drawing Test (CDT) and Letter Fluency tasks were included to capture domain-specific impairments.
+The classification of Parkinson’s Disease Dementia (PDD) based on Level I criteria was determined using established scoring thresholds from the original criteria [@dubois2007] with corresponding MoCA equivalents.
 
-To assess functional impairment, the Functional Activities Questionnaire (FAQ)[@bezdicek2011; @pfeffer1982] was administered.
-Additionally, neuropsychiatric status was evaluated using the Beck Depression Inventory-II (BDI) [@ciharova2020; @beck1996] and State-Trait Anxiety Inventory (STAI) [@spielberg1983; @mullner1980].
+To assess functional impairment, the Functional Activities Questionnaire (FAQ) [@bezdicek2016; @pfeffer1982] was administered.
+Additionally, neuropsychiatric status was evaluated using the Beck Depression Inventory-II (BDI-II) [@ciharova2020; @beck1996] and State-Trait Anxiety Inventory (STAI) [@spielberg1983; @mullner1980].
 Psychotic symptoms were assessed through structured psychiatric interviews conducted by a trained psychiatrist.
 
-## Theoretical and Empirical Estimands
+## Theoretical and Empirical Estimands^[**JM: Put most of it to Appendix.**]
 
-Following the framework proposed by Lundberg et al.
+*Following the framework proposed by Lundberg et al.
 (2021), this section connects our research objectives and their corresponding theoretical (i.e., targets of inference) and empirical (i.e., data-driven) estimands.
-The theoretical estimand refers to a unit-specific quantity defined over a target population and represents the ideal quantity that would address the research question under optimal conditions—such as access to complete population data or perfect experimental control.
-In contrast, the empirical estimand corresponds to the quantity that is actually computable using the available dataset, given real-world constraints.
-In our study, the target population for all theoretical estimands is defined as individuals diagnosed with Parkinson’s disease (PD) who are potential candidates for deep brain stimulation (DBS) treatment.
+The theoretical estimand refers to a unit-specific quantity defined over a target population and represents the ideal quantity that would address the research question under optimal conditions, such as access to complete population data or perfect experimental control.
+In contrast, the empirical estimand corresponds to the quantity that is actually computable using the available dataset, given real world constraints.
+In our study, the target population for all theoretical estimands is defined as individuals diagnosed with Parkinson’s disease (PD) who are potential candidates for deep brain stimulation (DBS) treatment.*
 
-The theoretical estimand corresponding to the RO1 is the true prevalence of probable PDD in this population.
-Empirically, this is estimated by calculating the proportion of patients classified as having probable PDD according to each of the diagnostic criteria under consideration.
+*The theoretical estimand corresponding to the RO1 is the true prevalence of probable PDD in this population.
+Empirically, this is estimated by calculating the proportion of patients classified as having probable PDD according to each of the diagnostic criteria under consideration.*
 
-For the RO2, the theoretical estimand is the variance in prevalence attributable to the diagnostic process itself—that is, the extent to which different sets of criteria yield differing prevalence estimates for the same population.
-This variability is empirically quantified by comparing the distribution of PDD classifications across criteria within the cohort.
+*For the RO2, the theoretical estimand is the variance in prevalence attributable to the diagnostic process itself—that is, the extent to which different sets of criteria yield differing prevalence estimates for the same population.
+This variability is empirically quantified by comparing the distribution of PDD classifications across criteria within the cohort.*
 
-The RO3 concerns diagnostic concordance.
+*The RO3 concerns diagnostic concordance.
 Here, the theoretical estimand is a set of population-level contingency tables comparing PDD classifications assigned by each pair of diagnostic criteria, along with derived metrics such as sensitivity, specificity, and kappa coefficients.
-The corresponding empirical estimands are represented by matrices generated through pairwise Receiver Operating Characteristic (ROC) curve analyses that evaluate the discriminatory performance of each diagnostic set.
+The corresponding empirical estimands are represented by matrices generated through pairwise Receiver Operating Characteristic (ROC) curve analyses that evaluate the discriminatory performance of each diagnostic set.*
 
-Finally, for the RO4, the theoretical estimand is defined as the set of diagnostic components whose variation systematically alters the probability of a probable PDD diagnosis.
+*Finally, for the RO4, the theoretical estimand is defined as the set of diagnostic components whose variation systematically alters the probability of a probable PDD diagnosis.
 This aspect of the study is exploratory in nature.
@@
 This allows us to identify the diagnostic elements most responsible for inter-criterion discrepancies.
-We approach this RO from an exploratory point of view and evaluate the importance of each PDD criteria component by observing change in empirical estimands for RO1-3 when stratified by different oprationalization decisions.
+We approach this RO from an exploratory point of view and evaluate the importance of each PDD criteria component by observing change in empirical estimands for RO1-3 when stratified by different oprationalization decisions.*
 
-## Operationalization of Parkinson's Disease Dementia
+## Operationalization of Parkinson's Disease Dementia^[**JM: Add a section with some logic (Cartezian product) to Appendix as well.**]
 
-In this study, we applied three distinct sets of diagnostic criteria for **probable**[^3] PDD at Level I.
+In this study, we applied three distinct sets of diagnostic criteria for **probable**[^1] PDD at Level I.
 The first set was based on the original framework [@dubois2007], which utilized the Mini-Mental State Examination (MMSE) as a global cognitive screening tool, supplemented by assessments of attention, executive function, visuospatial abilities, and memory.
@@
 **The Level II methodology, including the use of a regression-based normative scoring approach, has been detailed in a prior study [@bezdicek2017].
-Refer to @tbl-criteria**[^4] to a summary of the components and scoring thresholds of each diagnostic criterion.
+Refer to @tbl-criteria**[^2] to a summary of the components and scoring thresholds of each diagnostic criterion.
 All non-cognitive criteria of probable PDD (i.e., diagnosis of PD that developed before dementia and absence of Major Depression, delirium or other abnormalities that obscure diagnosis were established by an independent neurological and psychiatric assessments and held true for all patients in the sample).
+
+[^1]: **JM: We should unite the terminlogy.
+Most importantly, I reckon, we should differentiate between "probable PDD" (what is implied by meeting all the criteria within an operationalization) and "PDD" (the latent state of patient's cognition); and to use consistently the following trifecta - "criterion," "operationalization" and their "component".**
 
@@
 
-[^4]: **JM: Placeholder until we make the table proper.**
+[^2]: **JM: Placeholder until we make the table proper.**
 
@@
 continue <- !any(critnum == 0)
-#stopifnot("Some criteria are missing" = continue)
+stopifnot("Some criteria are missing" = continue)
 ```
@@
 Second, we applied the entire Functional Activities Questionnaire (FAQ) as suggested in the call for change [@kulisevsky2024], employing a cut-off score of 7 points or higher based on Czech normative data [@bezdicek2011].
-These methodologies resulted in a total of `r sum(critnum)` operationalizations, which were distributed across different diagnostic criteria: `r critnum["mmse"]` MMSE-based, `r critnum["moca"]` MoCA-based, `r critnum["smoca"]` sMoCA-based, and `r critnum["lvlII"]` based on the Level II battery[^5].
-
-[^5]: **JM: Need to add Level II operationalization and allow the `stopifnot()` test code above. My job.**
+These methodologies resulted in a total of `r sum(critnum)` operationalizations, which were distributed across different diagnostic criteria: `r critnum["mmse"]` MMSE-based, `r critnum["moca"]` MoCA-based, `r critnum["smoca"]` sMoCA-based, and `r critnum["lvlII"]` based on the Level II battery.
 
@@
 For each pair of operationalizations, the analysis was repeated twice such that each variables of the pair served once as the reference and once as the predictor.
-Following measures were used to evaluate pairwise concordance between different operationalizations of PDD criteria: 1) Cohen's $\kappa$ with its 95% confidence interval (CI) computed via the `cohen.kappa()` function from the R package *psych* [@revelle2024]; 2) Accuracy (i.e., the proportion of correct predictions, both true positives and true negatives, among the total number of cases) with its 95% CI; 3) Sensitivity/Recall (i.e., the proportion of true positives); and 4) Specificity (i.e., the proportion of true negatives).[^6]
+Following measures were used to evaluate pairwise concordance between different operationalizations of PDD criteria: 1) Cohen's $\kappa$ with its 95% confidence interval (CI) computed via the `cohen.kappa()` function from the R package *psych* [@revelle2024]; 2) Accuracy (i.e., the proportion of correct predictions, both true positives and true negatives, among the total number of cases) with its 95% CI; 3) Sensitivity/Recall (i.e., the proportion of true positives); and 4) Specificity (i.e., the proportion of true negatives).[^4]
 
-[^6]: Unlike Cohen's $\kappa$, Accuracy, Sensitivity and Specificity are not symmetrical, i.e., their value depend on which variable is considered reference and which is considered predictor.
+[^4]: Unlike Cohen's $\kappa$, Accuracy, Sensitivity and Specificity are not symmetrical, i.e., their value depend on which variable is considered reference and which is considered predictor.
 Consequently, we report these values twice for each pair of operationalizations.
@@
 All analyses were conducted with the R (version `r with(version, paste(major, minor, sep="."))`) software environment for statistical computing [@rsoft].
-The software code supporting this article is available at <https://github.com/josefmana/DemCr1t.git>.[^7]
+The software code supporting this article is available at <https://github.com/josefmana/DemCr1t.git>.[^5]
 
-[^7]: **JM: Do not forget to make it public before submitted!**
+[^5]: **JM: Do not forget to make it public before submitted!**
 
@@
 
-A total of `r sum(data$incl)` patients were included. Demographic and clinical characteristics of the sample are summarized in @tbl-desc. **...**[^8]
+A total of `r sum(data$incl)` patients were included. Demographic and clinical characteristics of the sample are summarized in @tbl-desc. **...**[^6]
 
-[^8]: **JM: Give some brief impressions from the Table 2 here.
+[^6]: **JM: Give some brief impressions from the Table 2 here.
 E.g., evaluate sex proportion, age, (we are missing education for some reason) etc.
@@
 On average, estimated prevalence was `r do_summary(perc, 2, "M")`% (SD = `r do_summary(perc, 2, "SD")`, Md = `r do_summary(perc, 2, "Md")`, range `r do_summary(perc, 2, "minmax")`).
-Notably, the prevalence estimate was substantially lower when FAQ item 9 was used as a criterion of IADL deficit (M = `r do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, "M")`%, SD = `r do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, "SD")`, range `r do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, "minmax")`) compared to using total FAQ score criterion (M = `r do_summary(perc[!grepl("FAQ 9", prev$table$IADL)], 2, "M")`%, SD = `r do_summary(perc[!grepl("FAQ 9", prev$table$IADL)], 2, "SD")`, range `r do_summary(perc[!grepl("FAQ 9", prev$table$IADL)], 2, "minmax")`) as demonstrated also in @fig-prev.[^9]
+Notably, the prevalence estimate was substantially lower when FAQ item 9 was used as a criterion of IADL deficit (M = `r do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, "M")`%, SD = `r do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, "SD")`, range `r do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, "minmax")`) compared to using total FAQ score criterion (M = `r do_summary(perc[!grepl("FAQ 9", prev$table$IADL)], 2, "M")`%, SD = `r do_summary(perc[!grepl("FAQ 9", prev$table$IADL)], 2, "SD")`, range `r do_summary(perc[!grepl("FAQ 9", prev$table$IADL)], 2, "minmax")`) as demonstrated also in @fig-prev.[^7]
 
-[^9]: **JM: For oneself - work on the code here, it ain't good.**
+[^7]: **JM: For oneself - work on the code here, it ain't good.**
 
-\[Insert Table 3 and Figure 1[^10] here\]
+\[Insert Table 3 and Figure 1[^8] here\]
 
-[^10]: **JM: For thyself - add description.**
+[^8]: **JM: For thyself - add description.**
 
@@
 Numerical results are available ...
-[^11]
+[^9]
 
-[^11]: **JM: ... as some kind of Supplementary Table, ideally html but Excel file would work as well.**
+[^9]: **JM: ... as some kind of Supplementary Table, ideally html but Excel file would work as well.**
 
@@
 
-\[Insert Figure 2[^12] here\]
+\[Insert Figure 2[^10] here\]
 
-[^12]: **JM: For thyself - add description.**
+[^10]: **JM: For thyself - add description.**
 
@@
 # Discussion
+
+This study evaluated the applicability and validity of diagnostic frameworks for diagnosing probable PDD within a cohort of patients considered for deep brain stimulation (DBS).
+Our results demonstrate that diagnostic outcomes are markedly influenced by the chosen type of operationalization, particularly in relation to the assessment of the cognitive domains and IADLs.
+
+## Variability in Prevalence Estimates
+
+One of the key findings of this study was the broad range of estimated PDD rate depending on the operationalization strategy.
+As seen in Table 3, the rate of PDD presence estimates varied from 2.00% to 16.75%, with the highest being derived from a combination of the sMoCA and the total FAQ score combination.
+When only FAQ item 9 was used to determine IADL deficits, the rate estimates were significantly lower (M = 3.09%, SD = 0.48), underscoring significant influence of functional assessment choice on diagnostic outcomes.
+
+These findings are lower in comparison with previous research demonstrating wide variability in reported PDD rate among PD patients.
+For instance, retrospective study reported a PDD rate of 19.7% [@rana2012], while a clinical investigation by Aarsland et al.
+found even higher rate of around 30% [@aarsland2011].
+A recent complex meta-analysis synthesizing global data placed the pooled rate at 26.30% [@sousa2022].
+Compared to these estimates, our study reports generally lower prevalence rates, likely reflecting differences in sample characteristics, diagnostic criteria, and methodology.
+
+Notably, our cohort consisted of patients being evaluated for DBS, a procedure typically reserved for individuals with relatively preserved cognitive function, which inherently biases against higher PDD prevalence rate.
+These comparisons emphasize that the observed lower prevalence in our sample is likely attributable to the preselection of cognitively intact individuals for DBS consideration, as well as to methodological variations such as the use of brief screening tool and differing IADL operationalizations.[^11]
+
+[^11]: **Keep it?**
+
+## Cognitive and Clinical Context [^12]
+
+The relatively low prevalence of PDD across operationalizations may also reflect the relatively preserved cognitive profile of the DBS cohort, as evident in Table 2.
+The mean MoCA (M = 24.07, SD = 3.48) and MMSE (M = 26.69, SD = 2.22) scores suggest that on average the patients were functioning at a globally intact level.
+This likely reflects the pre-selection of cognitively preserved individuals for DBS, in line with standard eligibility criteria [@bronstein2011].
+
+Memory performance, such as RAVLT and BVMTR delayed recall scores, also pointed to only mild deficits, particularly in domains central to the PDD vs.
+PD-MCI distinction.
+This aligns with previous findings that Level I criteria may overestimate dementia in patients with subtle impairments unless operational thresholds are rigorously defined [@aarsland2021; @bezdicek2016].
+
+[^12]: **Keep this section?**
+
+## Implications for DBS Eligibility and Clinical Practice
+
+Our findings also bear direct implications for clinical decision-making, especially in the context of surgical candidacy for DBS.
+Given that PDD remains a contraindication for DBS, the observed diagnostic instability could result in disparate treatment decisions based solely on which cognitive criteria are employed.
+Thus, harmonization of assessment procedures and operational cutoffs is essential to ensure equitable access to surgical treatment while maintaining diagnostic rigor.
+
+Further, our data support the use of multidimensional assessment strategies, such as Level II neuropsychological batteries or regression-based normative comparisons, as confirmatory tools in diagnostically ambiguous cases.
+These approaches may enhance diagnostic precision and mitigate the risks of both false positives and unjustified treatment exclusion.
+
+## Constraints on generalizability
+
+This study’s generalizability is limited by its retrospective design and the homogeneity of the DBS candidate cohort, which may not reflect broader PD populations with varying cognitive profiles.
+Additionally, reliance on Czech normative data may restrict international applicability, although it represents real-world clinical standards within the region [@simons2017].
+
+## Future Directions
+
+Future research should focus on prospective validation of updated Level I criteria against longitudinal functional outcomes and neurobiological markers.
+Additionally, the development of adaptive diagnostic algorithms that can integrate performance across cognitive, functional, and psychiatric domains may enhance diagnostic sensitivity while reducing the risk of over-classification.
+
+# Conclusions
+
+This study systematically investigated the application of multiple Level I diagnostic criteria for Parkinson’s disease dementia (PDD) within a cohort of patients considered for deep brain stimulation (DBS).
+The findings reveal substantial variability in prevalence estimates, strongly influenced by the choice of cognitive screening instruments and the operationalization of functional impairment.
+The divergence observed across operationalizations demonstrates the sensitivity of diagnostic outcomes to seemingly minor methodological choices.
+
+The proposed revisions to the diagnostic framework [@kulisevsky2024] criteria offer enhanced sensitivity by leveraging MoCA-based components and broader IADL assessments, the use must be cautiously calibrated to prevent over-diagnosis in populations with mild or borderline cognitive deficits.
+Conversely, overly conservative criteria, such as reliance on pill questionnaire (i.e.
+FAQ item 9 equivalence) may fail to detect meaningful functional decline and thus under-identify true cases of probable PDD.
+
+Ultimately, this study contributes to the improving landscape of PDD diagnostics by offering empirical evidence for the refinement of Level I criteria and reinforcing the value of psychometric rigor in clinical neuropsychology.
+Future work should extend this validation to longitudinal trajectories and integrate neurobiological correlates, ensuring that cognitive criteria remain both scientifically grounded and clinically actionable.
 

### 2025-05-07 _manuscript/manuscript.qmd
Moved algorithm-derivation and estimands to the Appendix.

@@
 library(tidyverse)
+library(gt)
 
@@
 crit <- tar_read(pdd_data, store = upstore)$criteria
+vars <- tar_read(variables, store = upstore)
 
@@
 
-Suggestion: (MK) **(RO1) To estimate the prevalence of Parkinson’s disease dementia (PDD) and evaluate the diagnostic variability and concordance across different PDD criteria.
+Suggestion: (MK) **(RO1) To estimate the prevalence of Parkinson’s disease dementia (PDD) and evaluate the diagnostic variability and concordance across different operationalizations of probable PDD.
 (RO2) To identify specific diagnostic components that contribute to variability in PDD classification across the applied criteria.**^[**JM:Nice!
@@
 
-## Theoretical and Empirical Estimands^[**JM: Put most of it to Appendix.**]
+## Diagnostic algorithms for probable Parkinson's Disease Dementia^[**JM: Add a section with some logic (Cartezian product) to Appendix as well.**]
 
-*Following the framework proposed by Lundberg et al.
-(2021), this section connects our research objectives and their corresponding theoretical (i.e., targets of inference) and empirical (i.e., data-driven) estimands.
-The theoretical estimand refers to a unit-specific quantity defined over a target population and represents the ideal quantity that would address the research question under optimal conditions, such as access to complete population data or perfect experimental control.
-In contrast, the empirical estimand corresponds to the quantity that is actually computable using the available dataset, given real world constraints.
-In our study, the target population for all theoretical estimands is defined as individuals diagnosed with Parkinson’s disease (PD) who are potential candidates for deep brain stimulation (DBS) treatment.*
-
-*The theoretical estimand corresponding to the RO1 is the true prevalence of probable PDD in this population.
-Empirically, this is estimated by calculating the proportion of patients classified as having probable PDD according to each of the diagnostic criteria under consideration.*
-
-*For the RO2, the theoretical estimand is the variance in prevalence attributable to the diagnostic process itself—that is, the extent to which different sets of criteria yield differing prevalence estimates for the same population.
-This variability is empirically quantified by comparing the distribution of PDD classifications across criteria within the cohort.*
-
-*The RO3 concerns diagnostic concordance.
-Here, the theoretical estimand is a set of population-level contingency tables comparing PDD classifications assigned by each pair of diagnostic criteria, along with derived metrics such as sensitivity, specificity, and kappa coefficients.
-The corresponding empirical estimands are represented by matrices generated through pairwise Receiver Operating Characteristic (ROC) curve analyses that evaluate the discriminatory performance of each diagnostic set.*
-
-*Finally, for the RO4, the theoretical estimand is defined as the set of diagnostic components whose variation systematically alters the probability of a probable PDD diagnosis.
-This aspect of the study is exploratory in nature.
-Empirically, we assess the contribution of each diagnostic feature by examining how variations in operational definitions (e.g., domain-specific thresholds, criteria for functional impairment) influence the empirical estimands derived for the first three objectives.
-This allows us to identify the diagnostic elements most responsible for inter-criterion discrepancies.
-We approach this RO from an exploratory point of view and evaluate the importance of each PDD criteria component by observing change in empirical estimands for RO1-3 when stratified by different oprationalization decisions.*
-
-## Operationalization of Parkinson's Disease Dementia^[**JM: Add a section with some logic (Cartezian product) to Appendix as well.**]
-
-In this study, we applied three distinct sets of diagnostic criteria for **probable**[^1] PDD at Level I.
+In this study, we applied three distinct sets of diagnostic criteria for probable PDD at Level I.
 The first set was based on the original framework [@dubois2007], which utilized the Mini-Mental State Examination (MMSE) as a global cognitive screening tool, supplemented by assessments of attention, executive function, visuospatial abilities, and memory.
-The second set of criteria was drawn from the recent call for change of dementia diagnostic guidelines [@kulisevsky2024], which advocate for more sensitive cognitive domain assessments in the context of Parkinson's disease (PD).
-This updated approach incorporated specific items from the Montreal Cognitive Assessment (MoCA) to better detect PD-related dementia.
-The third approach applied the Czech version of the shortened Montreal Cognitive Assessment (sMoCA) [@bezdicek2020], a time-efficient modification designed to ascertain whether equivalent cognitive impairments could be reliably identified using a reduced testing protocol.
-Lastly, the fourth approach followed the Level II battery protocol, which is commonly used in the evaluation of PD-MCI.
-**The Level II methodology, including the use of a regression-based normative scoring approach, has been detailed in a prior study [@bezdicek2017].
-Refer to @tbl-criteria**[^2] to a summary of the components and scoring thresholds of each diagnostic criterion.
-All non-cognitive criteria of probable PDD (i.e., diagnosis of PD that developed before dementia and absence of Major Depression, delirium or other abnormalities that obscure diagnosis were established by an independent neurological and psychiatric assessments and held true for all patients in the sample).
-
-[^1]: **JM: We should unite the terminlogy.
-Most importantly, I reckon, we should differentiate between "probable PDD" (what is implied by meeting all the criteria within an operationalization) and "PDD" (the latent state of patient's cognition); and to use consistently the following trifecta - "criterion," "operationalization" and their "component".**
-
-[^3]: **JM: We should unite the terminology.
-Most importantly, I reckon, we should differentiate between "probable PDD" (what is implied by meeting all the criteria within an operationalization) and "PDD" (the latent state of patient's cognition); and to use consistently the following trifecta - "criterion," "operationalization" and their "component".**
-
-[^2]: **JM: Placeholder until we make the table proper.**
+The second set of criteria was based on the recent call for change of dementia diagnostic guidelines [@kulisevsky2024], which advocate for more sensitive cognitive domain assessments in the context of PD.
+This updated approach incorporated specific items from the Montreal Cognitive Assessment (MoCA).
+The third approach applied the Czech version of the shortened Montreal Cognitive Assessment (sMoCA) [@bezdicek2020], a time-efficient modification designed to measure global cognitive performance using a reduced testing protocol that omits items providing redundant information.
+Lastly, the fourth approach followed the Level II protocol for diagnosis of PDD and Mild Cognitive Impairment in PD (PD-MCI) [@dubois2007; @litvan2012].
+The Level II methodology, including the use of a regression-based normative scoring approach, has been detailed in a prior study [@bezdicek2017].
+In this study, the threshold for cognitive impairment at Level II was set at $z \leq -1.5$.
+All non-cognitive criteria of probable PDD (i.e., diagnosis of PD that developed before dementia and absence of Major Depression, delirium or other abnormalities that obscure diagnosis were established by an independent neurological and psychiatric assessments) held true for all patients in the sample according to the psychiatric and neurological examinations.
 
@@
 #| label: "tbl-criteria"
+#| tbl-cap: "Summary of probable PDD operationalizations compared in the study."
 
-gt::gt(crit)
+table_criteria(vars)
 ```
@@
 Second, we applied the entire Functional Activities Questionnaire (FAQ) as suggested in the call for change [@kulisevsky2024], employing a cut-off score of 7 points or higher based on Czech normative data [@bezdicek2011].
-These methodologies resulted in a total of `r sum(critnum)` operationalizations, which were distributed across different diagnostic criteria: `r critnum["mmse"]` MMSE-based, `r critnum["moca"]` MoCA-based, `r critnum["smoca"]` sMoCA-based, and `r critnum["lvlII"]` based on the Level II battery.
+These methodologies resulted in a total of `r sum(critnum)` algorithms, which were distributed across different diagnostic criteria: `r critnum["mmse"]` MMSE-based, `r critnum["moca"]` MoCA-based, `r critnum["smoca"]` sMoCA-based, and `r critnum["lvlII"]` based on the Level II battery (see @tbl-criteria and Appendix for the exact specification of each algorithm).
 
@@
 
-For sample description, we summarized continuous variables using mean, standard deviation, median, minimum, and maximum values.
-Categorical variables were summarized by the number of patients in each category.
+Following the framework proposed by Lundberg et al.
+(2021), in this study we explicitly connect our research objectives and their corresponding theoretical (i.e., targets of inference) and empirical (i.e., data-driven) estimands to statistical estimates.
+The theoretical estimand refers to a unit-specific quantity defined over a target population and represents the ideal quantity that would address the research question under optimal conditions, such as access to complete population data or perfect experimental control.
+In contrast, the empirical estimand corresponds to the quantity that is actually computable using the available dataset, given real world constraints.
+The full description of the study's estimands and their relation to our research objectives is presented in the Appendix (see @tbl-estimands).
+
 To address study objectives, we started by repeatedly assigning each patient the diagnosis of probable PDD based on each PDD operationalization listed in @tbl-criteria resulting in a `r sum(data$incl)` (patients) $\times$ `r sum(critnum)` (operationalizations) matrix where each cell indicates whether a patient (row) meets criteria for probable PDD according to an operationalization (column).
 PDD prevalence estimates were then computed as $\frac{N_{PDD}}{N_{total}}$ separately for each operationalization to address RO1.
+The predictive value of age and sex was then evaluated by fitting a set of logistic regressions, one for each operationalization of probable PDD, whereby the probable PDD was predicted by age, sex and their interaction.
 
-To address RO2-4, a set of two class cross-tabulations with associated statistics was computed for each pair of operationalizations via the `confusionMatrix()` function from the R package *caret* [@kuhn2008].
+Next, a set of two class cross-tabulations with associated statistics was computed for each pair of operationalizations via the `confusionMatrix()` function from the R package *caret* [@kuhn2008].
 For each pair of operationalizations, the analysis was repeated twice such that each variables of the pair served once as the reference and once as the predictor.
@@
 
-desc <- tar_read(sample_description    , store = upstore)
-prev <- tar_read(prevalence_summaries  , store = upstore)
+desc <- tar_read(sample_description, store = upstore)
+prev <- tar_read(prevalence_summaries, store = upstore)
 conc <- tar_read(concordance_statistics, store = upstore)
@@
 #| label: "tbl-desc"
+#| tbl-cap: "Demographic, clinical and cognitive characteristics of the sample."
 
@@
 
-## Prevalence Estimates
+## PDD Rate Estimates
 
-Operationalization-wise prevalence estimates are presented in @tbl-prev.
-On average, estimated prevalence was `r do_summary(perc, 2, "M")`% (SD = `r do_summary(perc, 2, "SD")`, Md = `r do_summary(perc, 2, "Md")`, range `r do_summary(perc, 2, "minmax")`).
-Notably, the prevalence estimate was substantially lower when FAQ item 9 was used as a criterion of IADL deficit (M = `r do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, "M")`%, SD = `r do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, "SD")`, range `r do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, "minmax")`) compared to using total FAQ score criterion (M = `r do_summary(perc[!grepl("FAQ 9", prev$table$IADL)], 2, "M")`%, SD = `r do_summary(perc[!grepl("FAQ 9", prev$table$IADL)], 2, "SD")`, range `r do_summary(perc[!grepl("FAQ 9", prev$table$IADL)], 2, "minmax")`) as demonstrated also in @fig-prev.[^7]
+Operationalization-wise rate of PDD estimates are presented in @tbl-rates.
+On average, estimated PDD rate was `r do_summary(perc, 2, "M")`% (SD = `r do_summary(perc, 2, "SD")`, Md = `r do_summary(perc, 2, "Md")`, range `r do_summary(perc, 2, "minmax")`).
+Notably, the estimate was substantially lower when FAQ item 9 was used as a criterion of IADL deficit (M = `r do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, "M")`%, SD = `r do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, "SD")`, range `r do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, "minmax")`) compared to using total FAQ score criterion (M = `r do_summary(perc[!grepl("FAQ 9", prev$table$IADL)], 2, "M")`%, SD = `r do_summary(perc[!grepl("FAQ 9", prev$table$IADL)], 2, "SD")`, range `r do_summary(perc[!grepl("FAQ 9", prev$table$IADL)], 2, "minmax")`) as demonstrated also in @fig-prev.[^7]
+
+```{r}
+#| label: "tbl-rates"
+#| tbl-cap: "Estimates of the rate of probable PDD in the sample."
+
+prev$gtables$rates
+```
 
@@
 
-\[Insert Table 3 and Figure 1[^8] here\]
-
-[^8]: **JM: For thyself - add description.**
-
-```{r}
-#| label: "tbl-prev"
-
-prev$gtable
-```
+\[Insert Table 3 & Figure 1 here\]
 
@@
 #| label: "fig-prev"
+#| fig-cap: "Summary of the estimates of probable PDD rate across operationalizations under consideration.
+Vertical lines represent estimates arrived at by using sMoCA (dotted) or Level II (dashed) with FAQ item 9 (orange) or FAQ total score (blue) as criteria for probable PDD."
 
@@
 
+## Derivation of the Algorithms Set
+
+Both, the original PDD criteria [@dubois2007] and the call for their change [@kulisevsky2024] allow for several distinct combinations of test indexes to be used to define cognitive impairment.
+Consequently, in this study we derived all algorithms for probable PDD on Level I that are in line with published criteria.
+This procedure parallel the diagnostic algorithm outlined in Table 2 of @dubois2007.
+Specifically, in this study, we varied the exact specification of items 3-5 of this table (i.e., the measure of global cognitive impairment, the measure of the impact on ADLs and the measures of impaired cognition).
+
+For each set of criteria (MMSE-based, MoCA-based, sMoCA-based and Level II) we first specified measures and then the thresholds for each measure used to define probable PDD.
+Wherever more than one option was present in either the choice of measure or the choice of threshold, we created an algorithm for each choice in turn.
+The final set of algorithms tested was arrived at by computing the Cartesian product of all possibilities provided by varying measures and thresholds.
+All combinations are presented in @tbl-algos.
+
+For MMSE-based algorithms, the following sets of items served as the basis:
+
+$Global = \{MMSE < 26\}$
+
+$Attention = \{{Sevens\ backwards < 4}\}$
+
+$Executive = \{{Clock\ drawing} < 2, {Lexical\ fluency\ (S)} < 10\}$
+
+$Construction = \{Pentagons < 1\}$
+
+$Memory = \{{Three{\text-}words\ recall} < 3\}$
+
+$IADL = \{FAQ > 7, {FAQ\ (it. 9)} > 1\}$
+
+The ensuing Cartesian product $Global \times Attention \times Executive \times Construction \times Memory \times IADL$ results in $1 \times 1 \times 2 \times 1 \times 1 \times 2 = 4$ MMSE-based algorithms for probable PDD.
+
+For MoCA-based algorithms, the following sets of items served as the basis:
+
+$Global = \{MoCA < 27\}$
+
+$Attention = \{{Sevens\ backwards < 3}\}$
+
+$Executive = \{{Clock\ drawing} < 2, {Clock\ drawing} < 3, {Lexical\ fluency\ (K)} < 11\}$
+
+$Construction = \{Cube\ drawing < 1\}$
+
+$Memory = \{Five{\text-}words\ recall < 1, Five{\text-}words\ recall < 2, Five{\text-}words\ recall < 3, Five{\text-}words\ recall < 4, Five{\text-}words\ recall < 5\}$
+
+$Language = \{Abstraction < 2, Animal\ naming < 3\}$
+
+$IADL = \{FAQ > 7, {FAQ\ (it. 9)} > 1\}$
+
+Note that the additional language domain adds complexity to establishing a diagnostic algorithm because simply by adding it to the set, the number of potential algorithm doubles.
+Further complexity is added by the fact that there are so far no guidelines for selecting a diagnostic threshold for the Clock drawing and Five-words recall tests, both of which differ from their MMSE counterparts.
+Computing the Cartesian product $Global \times Attention \times Executive \times Construction \times Memory \times Language \times IADL$ yields $1 \times 1 \times 3 \times 1 \times 5 \times 2 \times 2 = 60$ distinct MoCA-based algorithms for probable PDD.
+
+For sMoCA-based algorithms, the following sets of items served as the basis:
+
+$Global = \{sMoCA < 13\}$
+
+$IADL = \{FAQ > 7, {FAQ\ (it. 9)} > 1\}$
+
+yielding $Global \times IADL$ yields $1 \times 2 = 2$ distinct sMoCA-based algorithms for probable PDD.
+
+Finally, the Level II algorithms were based on the following sets of items:
+
+$Attention = \{z(TMT\ A) < -1.5\ \cup z(WAIS\ DSB) < -1.5\}$
+
+$Executive = \{z(CF\ A) < -1.5\ \cup z(PST\ C) < -1.5\}$
+
+$Construction = \{z(JoLO) < -1.5\ \cup z(CLOXI) < -1.5\}$
+
+$Memory = \{z(RAVLT\ DR) < -1.5\ \cup z(BVMTR\ DR) < -1.5\ \cup z(WMS\text-III\ Family\ Pictures) < -1.5\}$
+
+$Language = \{z(WAIS\ Similarities) < -1.5 \ \cup z(BNT\ 60) < -1.5\}$
+
+$IADL = \{FAQ > 7, {FAQ\ (it. 9)} > 1\}$
+
+where $z()$ denotes calculation of age, sex and education adjusted z-score.
+This yields $1 \times 1 \times 1 \times 1 \times 1 \times 2 = 2$ distinct Level II algorithms for probable PDD in the current study.
+
+```{r}
+#| label: "tbl-algos"
+#| tbl-cap: "Summary of all algortihms for probable PDD used in the study."
+
+prev$gtables$algorithms
+```
+
+## Theoretical and Empirical Estimands
+
+In this study, we follow the framework proposed by Lundberg et al.
+(2021) for specifying targets of inference (i.e., the estimands) in qunatitative science to increase transparency and connect our statistical evidence to relevant theory.
+@tbl-estimands contains verbal description of the components relating to each of our proclaimed research objectives and map them to the population quantity of interest (the theoretical estimand), data-dependent quantity that could be estimated (the empirical estimand) and quantities that are reported in the study (statistical estimates).
+
+```{r}
+#| label: "tbl-estimands"
+#| tbl-cap: "Mapping between research objectives and quantities of interest in the current study."
+
+table_estimands()
+```
+
+The RO1, to estimate the prevalence of PDD and evaluate the diagnostic variability and concordance across different operationalizations of probable PDD, was further subdivided into four distinct research objectives - to estimate the rate of PDD within PD (RO1.1), to estimated variability of this rate (RO1.2), to evaluate predictive value of demographic variables for probable PDD classification (RO1.3) and to evaluate concordance between different probable PDD operationalizations and criteria (RO1.4).
+Estimates relating to RO1.1 and RO1.3 cannot be safely generalized beyond a population of PD patients that are candidates for DBS due to the systematic differences between DBS candidates pool and general PD population (such as lower age of DBS candidates compared to the general PD population).
+On the other hand, the estimates relating to RO1.4 (and to a lesser degree to RO1.2) may not be substantially influenced by the sample at hand as the primary source of their variance comes from variability in measures employed (e.g., MMSE vs MoCA to assess global cognitive performance) rather than variability in subjects' performance.
+Assuming that there is no substantial Differential Item Functioning between DBS candidates and broader PD population, the estimates relating to RO1.4 can be cautiously generalized beyond the current sample.
+
+Finally, for the RO2, the theoretical estimand is defined as the set of diagnostic components whose variation systematically alters the probability of a probable PDD diagnosis.
+This aspect of the study is exploratory in nature.
+Empirically, we assess the contribution of each diagnostic feature by examining how variations in operational definitions (e.g., domain-specific thresholds, criteria for functional impairment) influence the statistical estimates derived for the first objective.
+This allows us to identify the diagnostic elements most responsible for inter-operationalization discrepancies.
+
+## Supplementary Presentation of Results
+
 ```{r}

### 2025-05-07 R/specify_criteria.R

@@ -14,22 +14,22 @@
 specify_criteria <- function() {
   # MMSE-based criteria
   mmse <-
-    crossing(exec = c('vf_s','cloc'), iadl = c('faq_9','faq')) |>
+    crossing(exec = c("vf_s","cloc"), iadl = c("faq_9","faq")) |>
     mutate(
-      group  = 'mmse',
-      type   = paste0('mmse',seq_len(n())),
-      glob   = 'mmse',
+      group  = "mmse",
+      type   = paste0("MMSE (",seq_len(n()),")"),
+      glob   = "mmse",
       glob_t = 26,
-      atte   = 'mmse_7',
+      atte   = "mmse_7",
       atte_t = 4,
-      memo   = 'mmse_3words',
+      memo   = "mmse_3words",
       memo_t = 3,
-      cons   = 'mmse_pent',
+      cons   = "mmse_pent",
       cons_t = 1,
       lang   = NA,
       lang_t = 0,
-      exec_t = case_when(exec == 'cloc' ~ 2, exec == 'vf_s' ~ 10),
-      iadl_t = case_when(iadl == 'faq'  ~ 7, iadl == 'faq_9' ~ 1)
+      exec_t = case_when(exec == "cloc" ~ 2, exec == "vf_s" ~ 10),
+      iadl_t = case_when(iadl == "faq"  ~ 7, iadl == "faq_9" ~ 1)
     ) |>
     relocate(exec  , .after  = atte_t) |>
     relocate(exec_t, .after  = exec  ) |>
@@ -40,55 +40,55 @@ specify_criteria <- function() {
       2:3,
       function(i)
         crossing(
-          group  = 'moca',
-          glob   = 'moca_total',
+          group  = "moca",
+          glob   = "moca_total",
           glob_t = 26,
-          atte   = 'moca_7',
+          atte   = "moca_7",
           atte_t = 3,
-          exec   = c('moca_cloc', 'vf_k'),
-          memo   = 'moca_5words',
+          exec   = c("moca_cloc", "vf_k"),
+          memo   = "moca_5words",
           memo_t = 5:1,
-          cons   = 'moca_cube',
+          cons   = "moca_cube",
           cons_t = 1,
-          lang   = c('moca_anim', 'moca_abs'),
-          iadl   = c('faq', 'faq_9')
+          lang   = c("moca_anim", "moca_abs"),
[…]

### 2025-05-07 R/table_estimands.R

@@ -1,0 +1,67 @@
+#' Prepare a table summarising study
+#' estimands.
+#'
+#' Takes no parameters, can be altered
+#' only by changing the code. This function
+#' is used as a shortcut to create table
+#' verbally describing study estimands for
+#' the Appendix without cluttering the
+#' Quarto document.
+#'
+#' @returns A gt table object containing
+#' pre-specified verbal summary of the
+#' theoretical estimands, the empirical
+#' estimands and the statistical estimates
+#' regarding the study.
+#'
+#' @export
+table_estimands <- function() {
+  tibble(
+    RO = paste0("Research Objective ", c(rep(1,4), rep(2,1))),
+    Statement = c(
+      "To estimate the rate of PDD within PD.",
+      "To assess variability of PDD diagnosis depending on the algorithm applied.",
+      "To evaluate predictive information provided by demographic variables for probable PDD diagnosis.",
+      "To evaluate the diagnostic concordance between different PDD algorithms within and between PDD criteria.",
+      "To identify algorithms' components that contribute to variability in probable PDD diagnosis within and across criteria."
+    ),
+    Theoretical = c(
+      "The true rate of PDD within the population of patients with PD.",
+      "The variance of PDD rate estimates attributable to the diagnostic process itself, i.e., the extend to which different algorithms for probable PDD yield differing PDD rate estimates for the same population.",
+      "The expected change of suffering with PDD given a unit change in age or sex.",
+      "A set of population-level contingency tables comparing probable PDD classification assigned by each pair of probable PDD algorithms, along with derived metrics such as sensitivity, specificity, and inter-algorithm reliability coefficients.",
+      "A set of algorithms' components (such as diagnostic thresholds or test indexes being used to measure a certain cognitive domain) whose variation systematically alters the probability of a probable PDD diagnosis."
+    ),
+    Population = c(
+      "DBS candidates with PD.",
+      "DBS candidates/general population of patients with PD.",
+      "DBS candidates with PD.",
+      "DBS candidates/general population of patients with PD.",
+      "DBS candidates/general population of patients with PD."
+    ),
+    Empirical = c(
+      "Proportion of patients classified as having probable PDD according to each of the algorithms under consideration.",
+      "Shape of the distribution of probable PDD rate according to each of the algorithms under consideration.",
[…]

### 2025-05-07 _manuscript/manuscript.qmd
Intro change based on JM`s commentary.

--- 2025-05-07 _manuscript/manuscript.qmd
@@
 Despite its clinical relevance, the diagnosis of PDD remains complex.
-*The original **International Parkinson and Movement Disorder Society (MDS)** diagnostic criteria [@dubois2007], first formalized in 2006*^[**JM: I ain't sure it was the first time PDD criteria were formalised; e.g., Aarsland et al.
-- dx.doi.org/10.1001/archneur.60.3.387 - used DSM criteria for PDD already in 2003.
-Consequently I added the specification that these were MDS crtieria.**], were heavily influenced by frameworks established primarily for Alzheimer’s disease (AD).
+The original International Parkinson and Movement Disorder Society (MDS) diagnostic criteria [@dubois2007], first formalized in 2007, were heavily influenced by frameworks established primarily for Alzheimer’s disease (AD).
 Although these criteria provided a valuable initial foundation, they lacked the specificity required to capture the distinct pathophysiological and cognitive features of PD-related dementia [@emre2007; @yamashita2023].
 
-*One notable feature of the original criteria was the provision of an algorithm that allowed for flexibility in test selection [@dubois2007]*^[**JM: This came out of nowhere according to me.
-I think it would make more sense to introduce level I vs II first then introduce the algortihm for level I (i.e., Dubois's Table 2) and focus on it.**].
+To enhance diagnostic accuracy, the MDS introduced a two-level system for Parkinson’s disease dementia.
+Level I criteria are designed for clinical use, relying on brief cognitive assessments and clinical judgment.
+Level II criteria involve comprehensive neuropsychological testing across multiple domains, providing greater diagnostic detail for research or specialized settings [@emre2007].
+
+One notable feature of the original Level I criteria was the provision of an algorithm that allowed for flexibility in test selection [@dubois2007].
 Specifically, clinicians could choose between months reversed or seven backwards for attention assessment, lexical fluency or clock drawing for executive function evaluation.
@@
 
-*Subsequent steps have introduced the concept of Parkinson’s disease mild cognitive impairment (PD-MCI), refining the understanding of cognitive dysfunction in PD [@litvan2012].
-The PD-MCI criteria have propelled progress in diverse areas, enabling improvements in clinical characterization, identification of genetic correlates, therapeutic interventions, clinical trial design, and the assessment of progression risk to PDD [@hoogland2017; @hoogland2019; @aarsland2021].*^[**JM: I do not believe this section is needed for the article.
-You could keep it for yourself and use it in your dissertion but discussion history of MCI, albeit briefly, is out of scope of this manuscript in my humble correct opinion.**]
+According to a recent metanalysis, the epidemiological estimates of PDD prevalence among individuals with PD vary widely, ranging from 14% to 55%, depending on methodological criteria employed [@sousa2022].
+Moreover, factors such as patients` sex [@cereda2016], age and disease duration appear to modulate the risk of cognitive decline [@oh2016; @rana2012].
 
 Currently, efforts are focused on refining the PDD diagnostic framework to improve its consistency and applicability in both research and clinical contexts [@kulisevsky2024].
-*Our prior investigations have underscored the importance of employing rigorous psychometric methodologies to differentiate PD-MCI from PDD [@bezdicek2016; @deuschl2006].*^[**JM: The citations do not seem to match the statement - Bezdicek (2016) is about a processing speed theory of cognitive impairment in healthy older adults, Deuschl (2006) is an RCT of DBS.
-I'd drop this sentence altogether.**]
 
-*According to a recent metanalysis, the epidemiological estimates of PDD prevalence among individuals with PD vary widely, ranging from 14% to 55%^[**JM: It makes sense to report some point estimate as well.**], depending on methodological criteria employed [@sousa2022].
-Moreover, factors such as patients` sex [@cereda2016], age and disease duration appear to modulate the risk of cognitive decline [@oh2016; @rana2012].*^[**JM: Good, I'd push this paragraph higher.**]
-
-The present study evaluates the diagnostic concordance between the original Level I PDD criteria, as established by the MDS Task Force [@emre2007; @dubois2007] and criteria inspired by the recent call for change [@kulisevsky2024] within a cohort of PD patients *selected for DBS*^[**JM: Not true, the cohort is of patients being considered for DBS, not selected for DBS.**].
+The present study evaluates the diagnostic concordance between the original Level I PDD criteria, as established by the MDS Task Force [@emre2007; @dubois2007] and criteria inspired by the recent call for change [@kulisevsky2024] within a cohort of PD patients considered for DBS.
 Furthermore, both sets of criteria are compared to PDD diagnosed on Level II.
-The study aims to address following research objectives (RO): 1) to estimate the prevalence of PDD among patients considered for DBS (RO1), 2) to assess variability in PDD diagnosis depending on the diagnostic criteria applied (RO2), 3) to evaluate the diagnostic concordance between different sets of PDD criteria (RO3), and 4) to identify diagnostic components that contribute to variability in PDD classification across criteria (RO4).
-By addressing these objectives, this study seeks to validate the revised PDD criteria and evaluate their relevance in the context of DBS eligibility, thereby contributing to the refinement of cognitive assessment protocols in PD.
-
-Suggestion: (MK) **(RO1) To estimate the prevalence of Parkinson’s disease dementia (PDD) and evaluate the diagnostic variability and concordance across different PDD criteria.
-(RO2) To identify specific diagnostic components that contribute to variability in PDD classification across the applied criteria.**^[**JM:Nice!
-Let's state it this way.**]
+The study aims to address following research objectives (RO): (RO1) To estimate the prevalence of Parkinson’s disease dementia (PDD) and evaluate the diagnostic variability and concordance across different PDD criteria.
+(RO2) To identify specific diagnostic components that contribute to variability in PDD classification across the applied criteria.
 
@@
 Level I was assessed using the Mini-Mental State Examination (MMSE) [@stepankova2014; @folstein1975] and the Montreal Cognitive Assessment (MoCA) [@kopecek2016; @nasreddine2005].
-The neuropsychological assessment at Level II covered five cognitive domains, each evaluated through specific tests as follows: attention and working memory assessed using Trail Making Test Part A (TMT-A) [@bezdicek2012; @reitan2004], and WAIS Digit Span Backward (WAIS DSB) [@wechsler1997]; executive function evaluated via Categorical Verbal Fluency (CF) [@benton1989]^[**JM: This one instead - https://doi.org/10.14735/amcsnn2015292**], and subtest from the Prague Stroop Test – Colors (PST-C) [@bezdicek2021]; language measured with the WAIS Similarities subtest [@wechsler1997], and the Boston Naming Test (BNT-60) [@kaplan1983]; memory examined using the Rey Auditory Verbal Learning Test (RAVLT) [@frydrychova2018; @rey1964]^[**JM: Do not forget this - https://doi.org/10.1080/13825585.2013.865699**] for delayed recall, and the Brief Visual Memory Test–Revised (BVMTR) [@havlik2020; @benedict1997] for *delayed recall*^[**JM: The same label is used for RAVLT-DR and BVMT-R-DR here, we should sort it out.**], or WAIS Family Pictures subtest [@wechsler1997] *including delayed recall and forced choice recognition*^[**JM: We did not use any recognition.
-Should tidy this section such that it includes only indexes used by our study.**]; visuospatial function assessed through the Judgment of Line Orientation Test (JoLO) [@benton1983], and Clock Drawing Test (CLOX) [@royall1998].
+The neuropsychological assessment at Level II covered five cognitive domains, each evaluated through specific tests as follows: attention and working memory assessed using Trail Making Test Part A (TMT-A) [@bezdicek2012; @reitan2004], and WAIS Digit Span Backward (WAIS DSB) [@wechsler1997]; executive function evaluated via Categorical Verbal Fluency (CF) [@benton1989][^1], and subtest from the Prague Stroop Test – Colors (PST-C) [@bezdicek2021]; language measured with the WAIS Similarities subtest [@wechsler1997], and the Boston Naming Test (BNT-60) [@kaplan1983]; memory examined using the Rey Auditory Verbal Learning Test (RAVLT) [@frydrychova2018; @rey1964][^2] for delayed recall, and the Brief Visual Memory Test–Revised (BVMTR) [@havlik2020; @benedict1997] for *delayed recall*[^3], or WAIS Family Pictures subtest [@wechsler1997] *including delayed recall and forced choice recognition*[^4]; visuospatial function assessed through the Judgment of Line Orientation Test (JoLO) [@benton1983], and Clock Drawing Test (CLOX) [@royall1998].
 
-In addition to the core cognitive assessments, *tasks such as*^[**JM: Stating such as "such as" are usually not appropritate in Methods section of a scientific study, because all methods should be reported exactly.**] the Clock Drawing Test (CDT) and Letter Fluency tasks were included to capture domain-specific impairments.
+[^1]: **JM: This one instead - https://doi.org/10.14735/amcsnn2015292**
+
+[^2]: **JM: Do not forget this - https://doi.org/10.1080/13825585.2013.865699**
+
+[^3]: **JM: The same label is used for RAVLT-DR and BVMT-R-DR here, we should sort it out.**
+
+[^4]: **JM: We did not use any recognition. Should tidy this section such that it includes only indexes used by our study.**
+
+In addition to the core cognitive assessments, *tasks such as*[^5] the Clock Drawing Test (CDT) and Letter Fluency tasks were included to capture domain-specific impairments.
 The classification of Parkinson’s Disease Dementia (PDD) based on Level I criteria was determined using established scoring thresholds from the original criteria [@dubois2007] with corresponding MoCA equivalents.
+
+[^5]: **JM: Stating such as "such as" are usually not appropritate in Methods section of a scientific study, because all methods should be reported exactly.**
 
@@
 
-## Theoretical and Empirical Estimands^[**JM: Put most of it to Appendix.**]
+## Theoretical and Empirical Estimands[^6]
+
+[^6]: **JM: Put most of it to Appendix.**
 
@@
 
-## Operationalization of Parkinson's Disease Dementia^[**JM: Add a section with some logic (Cartezian product) to Appendix as well.**]
+## Operationalization of Parkinson's Disease Dementia[^7]
 
-In this study, we applied three distinct sets of diagnostic criteria for **probable**[^1] PDD at Level I.
+[^7]: **JM: Add a section with some logic (Cartezian product) to Appendix as well.**
+
+In this study, we applied three distinct sets of diagnostic criteria for **probable**[^8] PDD at Level I.
 The first set was based on the original framework [@dubois2007], which utilized the Mini-Mental State Examination (MMSE) as a global cognitive screening tool, supplemented by assessments of attention, executive function, visuospatial abilities, and memory.
@@
 **The Level II methodology, including the use of a regression-based normative scoring approach, has been detailed in a prior study [@bezdicek2017].
-Refer to @tbl-criteria**[^2] to a summary of the components and scoring thresholds of each diagnostic criterion.
+Refer to @tbl-criteria**[^9] to a summary of the components and scoring thresholds of each diagnostic criterion.
 All non-cognitive criteria of probable PDD (i.e., diagnosis of PD that developed before dementia and absence of Major Depression, delirium or other abnormalities that obscure diagnosis were established by an independent neurological and psychiatric assessments and held true for all patients in the sample).
 
-[^1]: **JM: We should unite the terminlogy.
+[^8]: **JM: We should unite the terminlogy.
 Most importantly, I reckon, we should differentiate between "probable PDD" (what is implied by meeting all the criteria within an operationalization) and "PDD" (the latent state of patient's cognition); and to use consistently the following trifecta - "criterion," "operationalization" and their "component".**
 
-[^3]: **JM: We should unite the terminology.
-Most importantly, I reckon, we should differentiate between "probable PDD" (what is implied by meeting all the criteria within an operationalization) and "PDD" (the latent state of patient's cognition); and to use consistently the following trifecta - "criterion," "operationalization" and their "component".**
-
-[^2]: **JM: Placeholder until we make the table proper.**
+[^9]: **JM: Placeholder until we make the table proper.**
 
@@
 For each pair of operationalizations, the analysis was repeated twice such that each variables of the pair served once as the reference and once as the predictor.
-Following measures were used to evaluate pairwise concordance between different operationalizations of PDD criteria: 1) Cohen's $\kappa$ with its 95% confidence interval (CI) computed via the `cohen.kappa()` function from the R package *psych* [@revelle2024]; 2) Accuracy (i.e., the proportion of correct predictions, both true positives and true negatives, among the total number of cases) with its 95% CI; 3) Sensitivity/Recall (i.e., the proportion of true positives); and 4) Specificity (i.e., the proportion of true negatives).[^4]
+Following measures were used to evaluate pairwise concordance between different operationalizations of PDD criteria: 1) Cohen's $\kappa$ with its 95% confidence interval (CI) computed via the `cohen.kappa()` function from the R package *psych* [@revelle2024]; 2) Accuracy (i.e., the proportion of correct predictions, both true positives and true negatives, among the total number of cases) with its 95% CI; 3) Sensitivity/Recall (i.e., the proportion of true positives); and 4) Specificity (i.e., the proportion of true negatives).[^10]
 
-[^4]: Unlike Cohen's $\kappa$, Accuracy, Sensitivity and Specificity are not symmetrical, i.e., their value depend on which variable is considered reference and which is considered predictor.
+[^10]: Unlike Cohen's $\kappa$, Accuracy, Sensitivity and Specificity are not symmetrical, i.e., their value depend on which variable is considered reference and which is considered predictor.
 Consequently, we report these values twice for each pair of operationalizations.
@@
 All analyses were conducted with the R (version `r with(version, paste(major, minor, sep="."))`) software environment for statistical computing [@rsoft].
-The software code supporting this article is available at <https://github.com/josefmana/DemCr1t.git>.[^5]
+The software code supporting this article is available at <https://github.com/josefmana/DemCr1t.git>.[^11]
 
-[^5]: **JM: Do not forget to make it public before submitted!**
+[^11]: **JM: Do not forget to make it public before submitted!**
 
@@
 
-A total of `r sum(data$incl)` patients were included. Demographic and clinical characteristics of the sample are summarized in @tbl-desc. **...**[^6]
+A total of `r sum(data$incl)` patients were included. Demographic and clinical characteristics of the sample are summarized in @tbl-desc. **...**[^12]
 
-[^6]: **JM: Give some brief impressions from the Table 2 here.
+[^12]: **JM: Give some brief impressions from the Table 2 here.
 E.g., evaluate sex proportion, age, (we are missing education for some reason) etc.
@@
 On average, estimated prevalence was `r do_summary(perc, 2, "M")`% (SD = `r do_summary(perc, 2, "SD")`, Md = `r do_summary(perc, 2, "Md")`, range `r do_summary(perc, 2, "minmax")`).
-Notably, the prevalence estimate was substantially lower when FAQ item 9 was used as a criterion of IADL deficit (M = `r do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, "M")`%, SD = `r do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, "SD")`, range `r do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, "minmax")`) compared to using total FAQ score criterion (M = `r do_summary(perc[!grepl("FAQ 9", prev$table$IADL)], 2, "M")`%, SD = `r do_summary(perc[!grepl("FAQ 9", prev$table$IADL)], 2, "SD")`, range `r do_summary(perc[!grepl("FAQ 9", prev$table$IADL)], 2, "minmax")`) as demonstrated also in @fig-prev.[^7]
+Notably, the prevalence estimate was substantially lower when FAQ item 9 was used as a criterion of IADL deficit (M = `r do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, "M")`%, SD = `r do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, "SD")`, range `r do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, "minmax")`) compared to using total FAQ score criterion (M = `r do_summary(perc[!grepl("FAQ 9", prev$table$IADL)], 2, "M")`%, SD = `r do_summary(perc[!grepl("FAQ 9", prev$table$IADL)], 2, "SD")`, range `r do_summary(perc[!grepl("FAQ 9", prev$table$IADL)], 2, "minmax")`) as demonstrated also in @fig-prev.[^13]
 
-[^7]: **JM: For oneself - work on the code here, it ain't good.**
+[^13]: **JM: For oneself - work on the code here, it ain't good.**
 
-\[Insert Table 3 and Figure 1[^8] here\]
+\[Insert Table 3 and Figure 1[^14] here\]
 
-[^8]: **JM: For thyself - add description.**
+[^14]: **JM: For thyself - add description.**
 
@@
 Numerical results are available ...
-[^9]
+[^15]
 
-[^9]: **JM: ... as some kind of Supplementary Table, ideally html but Excel file would work as well.**
+[^15]: **JM: ... as some kind of Supplementary Table, ideally html but Excel file would work as well.**
 
@@
 
-\[Insert Figure 2[^10] here\]
+\[Insert Figure 2[^16] here\]
 
-[^10]: **JM: For thyself - add description.**
+[^16]: **JM: For thyself - add description.**
 
@@
 Notably, our cohort consisted of patients being evaluated for DBS, a procedure typically reserved for individuals with relatively preserved cognitive function, which inherently biases against higher PDD prevalence rate.
-These comparisons emphasize that the observed lower prevalence in our sample is likely attributable to the preselection of cognitively intact individuals for DBS consideration, as well as to methodological variations such as the use of brief screening tool and differing IADL operationalizations.[^11]
+These comparisons emphasize that the observed lower prevalence in our sample is likely attributable to the preselection of cognitively intact individuals for DBS consideration, as well as to methodological variations such as the use of brief screening tool and differing IADL operationalizations.[^17]
 
-[^11]: **Keep it?**
+[^17]: **Keep it?**
 
-## Cognitive and Clinical Context [^12]
+## Cognitive and Clinical Context [^18]
+
+[^18]: **Keep this section?**
 
@@
 This aligns with previous findings that Level I criteria may overestimate dementia in patients with subtle impairments unless operational thresholds are rigorously defined [@aarsland2021; @bezdicek2016].
-
-[^12]: **Keep this section?**
 

### 2025-05-07 _manuscript/manuscript.qmd

--- 2025-05-07 _manuscript/manuscript.qmd
@@
 
-*Subsequent steps have introduced the concept of Parkinson’s disease mild cognitive impairment (PD-MCI), refining the understanding of cognitive dysfunction in PD [@litvan2012].
-The PD-MCI criteria have propelled progress in diverse areas, enabling improvements in clinical characterization, identification of genetic correlates, therapeutic interventions, clinical trial design, and the assessment of progression risk to PDD [@hoogland2017; @hoogland2019; @aarsland2021].*^[**JM: I do not believe this section is needed for the article.
-You could keep it for yourself and use it in your dissertion but discussion history of MCI, albeit briefly, is out of scope of this manuscript in my humble correct opinion.**]
+According to a recent metanalysis, the epidemiological estimates of PDD prevalence among individuals with PD vary widely, ranging from 14% to 55%, depending on methodological criteria employed [@sousa2022].
+Moreover, factors such as patients` sex [@cereda2016], age and disease duration appear to modulate the risk of cognitive decline [@oh2016; @rana2012].
 
 Currently, efforts are focused on refining the PDD diagnostic framework to improve its consistency and applicability in both research and clinical contexts [@kulisevsky2024].
-*Our prior investigations have underscored the importance of employing rigorous psychometric methodologies to differentiate PD-MCI from PDD [@bezdicek2016; @deuschl2006].*^[**JM: The citations do not seem to match the statement - Bezdicek (2016) is about a processing speed theory of cognitive impairment in healthy older adults, Deuschl (2006) is an RCT of DBS.
-I'd drop this sentence altogether.**]
 
-*According to a recent metanalysis, the epidemiological estimates of PDD prevalence among individuals with PD vary widely, ranging from 14% to 55%^[**JM: It makes sense to report some point estimate as well.**], depending on methodological criteria employed [@sousa2022].
-Moreover, factors such as patients` sex [@cereda2016], age and disease duration appear to modulate the risk of cognitive decline [@oh2016; @rana2012].*^[**JM: Good, I'd push this paragraph higher.**]
-
-The present study evaluates the diagnostic concordance between the original Level I PDD criteria, as established by the MDS Task Force [@emre2007; @dubois2007] and criteria inspired by the recent call for change [@kulisevsky2024] within a cohort of PD patients *selected for DBS*^[**JM: Not true, the cohort is of patients being considered for DBS, not selected for DBS.**].
+The present study evaluates the diagnostic concordance between the original Level I PDD criteria, as established by the MDS Task Force [@emre2007; @dubois2007] and criteria inspired by the recent call for change [@kulisevsky2024] within a cohort of PD patients considered for DBS.
 Furthermore, both sets of criteria are compared to PDD diagnosed on Level II.
-The study aims to address following research objectives (RO): 1) to estimate the prevalence of PDD among patients considered for DBS (RO1), 2) to assess variability in PDD diagnosis depending on the diagnostic criteria applied (RO2), 3) to evaluate the diagnostic concordance between different sets of PDD criteria (RO3), and 4) to identify diagnostic components that contribute to variability in PDD classification across criteria (RO4).
-By addressing these objectives, this study seeks to validate the revised PDD criteria and evaluate their relevance in the context of DBS eligibility, thereby contributing to the refinement of cognitive assessment protocols in PD.
-
-Suggestion: (MK) **(RO1) To estimate the prevalence of Parkinson’s disease dementia (PDD) and evaluate the diagnostic variability and concordance across different operationalizations of probable PDD.
-(RO2) To identify specific diagnostic components that contribute to variability in PDD classification across the applied criteria.**^[**JM:Nice!
-Let's state it this way.**]
+The study aims to address following research objectives (RO): (RO1) To estimate the prevalence of Parkinson’s disease dementia (PDD) and evaluate the diagnostic variability and concordance across different PDD criteria.
+(RO2) To identify specific diagnostic components that contribute to variability in PDD classification across the applied criteria.
 
@@
 
-## Diagnostic algorithms for probable Parkinson's Disease Dementia^[**JM: Add a section with some logic (Cartezian product) to Appendix as well.**]
+## Diagnostic algorithms for probable Parkinson's Disease Dementia
 

### 2025-05-08 _manuscript/manuscript.qmd
Almost all meths and stats done.

@@
 
-In this study, we applied three distinct sets of diagnostic criteria for probable PDD at Level I.
+In this study, we applied three distinct sets of diagnostic algorithms for probable PDD at Level I.
 The first set was based on the original framework [@dubois2007], which utilized the Mini-Mental State Examination (MMSE) as a global cognitive screening tool, supplemented by assessments of attention, executive function, visuospatial abilities, and memory.
-The second set of criteria was based on the recent call for change of dementia diagnostic guidelines [@kulisevsky2024], which advocate for more sensitive cognitive domain assessments in the context of PD.
+The second set of algorithms was based on the recent call for change of dementia diagnostic guidelines [@kulisevsky2024], which advocates for more sensitive cognitive domain assessments in the context of PD.
 This updated approach incorporated specific items from the Montreal Cognitive Assessment (MoCA).
@@
 The Level II methodology, including the use of a regression-based normative scoring approach, has been detailed in a prior study [@bezdicek2017].
-In this study, the threshold for cognitive impairment at Level II was set at $z \leq -1.5$.
-All non-cognitive criteria of probable PDD (i.e., diagnosis of PD that developed before dementia and absence of Major Depression, delirium or other abnormalities that obscure diagnosis were established by an independent neurological and psychiatric assessments) held true for all patients in the sample according to the psychiatric and neurological examinations.
-
-```{r}
-#| label: "tbl-criteria"
-#| tbl-cap: "Summary of probable PDD operationalizations compared in the study."
-
-table_criteria(vars)
-```
-
-\[Insert Table 1 here\]
+In this study, the thresholds for cognitive impairment at Level II were set at $z \leq -1.5$.
+All non-cognitive criteria of probable PDD (i.e., diagnosis of PD that developed before dementia and absence of Major Depression, delirium or other abnormalities that obscure diagnosis) held true for all patients in the sample according to the psychiatric and neurological examinations.
 
@@
 
-For each of these diagnostic approaches, we applied two operationalization strategies based on deficits in Instrumental Activities of Daily Living (IADL).
+For each of these diagnostic approaches, we applied two operationalizations of deficits in Instrumental Activities of Daily Living (IADL).
 First, we utilized FAQ item 9, which approximates the pill questionnaire from the original criteria [@dubois2007] employing a cut-off score of 2 points or higher.
 Second, we applied the entire Functional Activities Questionnaire (FAQ) as suggested in the call for change [@kulisevsky2024], employing a cut-off score of 7 points or higher based on Czech normative data [@bezdicek2011].
-These methodologies resulted in a total of `r sum(critnum)` algorithms, which were distributed across different diagnostic criteria: `r critnum["mmse"]` MMSE-based, `r critnum["moca"]` MoCA-based, `r critnum["smoca"]` sMoCA-based, and `r critnum["lvlII"]` based on the Level II battery (see @tbl-criteria and Appendix for the exact specification of each algorithm).
+These methodologies resulted in a total of `r sum(critnum)` algorithms, which were distributed across different diagnostic criteria: `r critnum["mmse"]` MMSE-based, `r critnum["moca"]` MoCA-based, `r critnum["smoca"]` sMoCA-based, and `r critnum["lvlII"]` based on the Level II battery (see @tbl-crits and Appendix @tbl-algos for the exact specification of each algorithm).
+
+```{r}
+#| label: "tbl-crits"
+#| tbl-cap: "Summary of probable PDD operationalizations compared in the study."
+
+table_criteria(vars)
+```
+
+\[Insert @tbl-crits here\]
 
@@
 
-To address study objectives, we started by repeatedly assigning each patient the diagnosis of probable PDD based on each PDD operationalization listed in @tbl-criteria resulting in a `r sum(data$incl)` (patients) $\times$ `r sum(critnum)` (operationalizations) matrix where each cell indicates whether a patient (row) meets criteria for probable PDD according to an operationalization (column).
-PDD prevalence estimates were then computed as $\frac{N_{PDD}}{N_{total}}$ separately for each operationalization to address RO1.
-The predictive value of age and sex was then evaluated by fitting a set of logistic regressions, one for each operationalization of probable PDD, whereby the probable PDD was predicted by age, sex and their interaction.
+To address study objectives, we started by repeatedly assigning each patient the diagnosis of probable PDD based on each PDD algorithm listed in @tbl-crits (see also @tbl-algos) resulting in a `r sum(data$incl)` (patients) $\times$ `r sum(critnum)` (operationalizations) matrix where each cell indicates whether a patient (row) meets criteria for probable PDD according to an algorithm (column)^[**JM: This could and should be shared most likely, as long, as we anonymize properly.
+Let's ask Oto Mestek and the NPO team how and if is it possible.**].
+PDD rate estimates were computed as $\frac{N_{PDD}}{N_{total}}$ separately for each algortihm.
+The predictive value of age and sex was then evaluated by fitting a set of logistic regressions, one for each algorithm for probable PDD, whereby the probable PDD was predicted by age, sex and their interaction.
 
-Next, a set of two class cross-tabulations with associated statistics was computed for each pair of operationalizations via the `confusionMatrix()` function from the R package *caret* [@kuhn2008].
-For each pair of operationalizations, the analysis was repeated twice such that each variables of the pair served once as the reference and once as the predictor.
-Following measures were used to evaluate pairwise concordance between different operationalizations of PDD criteria: 1) Cohen's $\kappa$ with its 95% confidence interval (CI) computed via the `cohen.kappa()` function from the R package *psych* [@revelle2024]; 2) Accuracy (i.e., the proportion of correct predictions, both true positives and true negatives, among the total number of cases) with its 95% CI; 3) Sensitivity/Recall (i.e., the proportion of true positives); and 4) Specificity (i.e., the proportion of true negatives).[^10]
+Next, a set of two class cross-tabulations with associated statistics was computed for each pair of algorithms via the `confusionMatrix()` function from the R package *caret* [@kuhn2008].
+For each pair of algorithms, the analysis was repeated twice such that each variable of the pair served once as the reference and once as the predictor.
+Following measures were used to evaluate pairwise concordance between different algorithms for probable PDD: 1) Cohen's $\kappa$ with its 95% confidence interval (CI) computed via the `cohen.kappa()` function from the R package *psych* [@revelle2024]; 2) Accuracy (i.e., the proportion of correct predictions, both true positives and true negatives, among the total number of cases) with its 95% CI; 3) Sensitivity/Recall (i.e., the proportion of true positives); and 4) Specificity (i.e., the proportion of true negatives).[^10]
 
-[^10]: Unlike Cohen's $\kappa$, Accuracy, Sensitivity and Specificity are not symmetrical, i.e., their value depend on which variable is considered reference and which is considered predictor.
-Consequently, we report these values twice for each pair of operationalizations.
+[^10]: Unlike Cohen's $\kappa$, Accuracy, Sensitivity and Specificity are not symmetrical, i.e., their value depends on which variable is considered reference and which is considered predictor.
+Consequently, we report these values twice for each pair of algorithms.
 Note that the Sensitivity of a reference/predictor pair corresponds to the Positive Predictive Value if their roles were reversed.
@@
 
-Finally, the No Information Rate (NIR) was calculated for each pair of operationalizations.
-NIR is the accuracy that could be obtained by always predicting the majority class and in our case it is equivalent to the complement of the PDD prevalence estimate according to the reference operationalization.
-Accuracy of prediction was subsequently compared to the NIR via a one-sided Exact Binomial Test as implemented by the `binom.test()` R stats function.
+Finally, the No Information Rate (NIR) was calculated for each pair of algorithms.
+NIR is the accuracy that could be obtained by always predicting the majority class and in our case it is equivalent to the complement of the PDD rate estimate according to the reference algorithm.
+Accuracy of prediction was compared to the NIR via a one-sided Exact Binomial Test as implemented by the `binom.test()` R stats function.
 Reference/predictor pairs associated with p \< .05 were considered to show significantly better accuracy than NIR.
-In other word, for reference/predictor pairs associated with p \< .05, we conclude that knowing the probable PDD status according to the predictor operationalization helps to estimate the probable PDD status according to the reference operationalization and the two opeartionalizations thus show substantial concordance.
+In other words, for reference/predictor pairs associated with p \< .05, we conclude that knowing the probable PDD status according to the predictor algorithm helps to estimate the probable PDD status according to the reference algorithm and the two algorithms thus show substantial concordance.
 
 Data wrangling and visualizations were done in the *tidyverse* package [@wickham2019] and tables were formatted in the *gt* package [@iannone2024].
-All analyses were conducted with the R (version `r with(version, paste(major, minor, sep="."))`) software environment for statistical computing [@rsoft].
+All analyses were conducted within the R (version `r with(version, paste(major, minor, sep="."))`) software environment for statistical computing [@rsoft].
 The software code supporting this article is available at <https://github.com/josefmana/DemCr1t.git>.[^11]
@@
 prev <- tar_read(prevalence_summaries, store = upstore)
+pred <- tar_read(demographic_predictors, store = upstore)
 conc <- tar_read(concordance_statistics, store = upstore)
 perc <- prev$table$perc
+
+# In-text demographic data
+sex <- subset(desc$table, variable == "Sex")$N
+age <- paste0(subset(desc$table, variable == "Age")[ , c("M", "SD")] |> paste(collapse = " (SD = "),") years of age")
+edu <- paste0(subset(desc$table, variable == "Education")[ , c("M", "SD")] |> paste(collapse = " (SD = ")," ) years of education")
+dur <- paste0(subset(desc$table, variable == "PD duration")[ , c("M", "SD")] |> paste(collapse = " (SD = "),") years of disease duration")
+updrs_off <- paste0(subset(desc$table, variable == "UPDRS III off state")[ , c("M", "SD")] |> paste(collapse = " (SD = "),") Unified Parkinson Disease Rating Scale (UPDRS), part III in medication OFF state")
+updrs_on <- paste0(subset(desc$table, variable == "UPDRS III on state")[ , c("M", "SD")] |> paste(collapse = " (SD = "),") UPDRS III in medication ON state")
+
+# In-text rate estimates
+rates_overall <- sapply(c("M", "SD", "Md", "minmax"), function(f) do_summary(perc, 2, f))
+rates_FAQ9 <- sapply(c("M", "SD", "Md", "minmax"), function(f) do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, f))
+rates_FAQtot <- sapply(c("M", "SD", "Md", "minmax"), function(f) do_summary(perc[prev$table$IADL == "FAQ > 7"], 2, f))
+
+# Minimum p-value of logistic regressions
+min_p <- subset(pred$values, quantity == "p-value")$estimate |> min() |> do_summary(3, "p")
 ```
@@
 
-A total of `r sum(data$incl)` patients were included. Demographic and clinical characteristics of the sample are summarized in @tbl-desc. **...**[^12]
-
-[^12]: **JM: Give some brief impressions from the Table 2 here.
-E.g., evaluate sex proportion, age, (we are missing education for some reason) etc.
-with general PD population (e.g., from some meta-analysis).
-Maybe say a word or two about mean cognitive profile or its spread.**
-
-\[Insert Table 2 here\]
+A total of `r sum(data$incl)` patients were included.
+The sample included `r sex` men, with an average of `r age`, `r edu`, `r dur`, `r updrs_off` and `r updrs_on`.
+Cognitive characteristics of the sample are summarized in @tbl-desc.
 
@@
 #| label: "tbl-desc"
-#| tbl-cap: "Demographic, clinical and cognitive characteristics of the sample."
+#| tbl-cap: "Cognitive characteristics of the sample."
 
@@
 
+\[Insert @tbl-desc here\]
+
 ## PDD Rate Estimates
 
-Operationalization-wise rate of PDD estimates are presented in @tbl-rates.
-On average, estimated PDD rate was `r do_summary(perc, 2, "M")`% (SD = `r do_summary(perc, 2, "SD")`, Md = `r do_summary(perc, 2, "Md")`, range `r do_summary(perc, 2, "minmax")`).
-Notably, the estimate was substantially lower when FAQ item 9 was used as a criterion of IADL deficit (M = `r do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, "M")`%, SD = `r do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, "SD")`, range `r do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, "minmax")`) compared to using total FAQ score criterion (M = `r do_summary(perc[!grepl("FAQ 9", prev$table$IADL)], 2, "M")`%, SD = `r do_summary(perc[!grepl("FAQ 9", prev$table$IADL)], 2, "SD")`, range `r do_summary(perc[!grepl("FAQ 9", prev$table$IADL)], 2, "minmax")`) as demonstrated also in @fig-prev.[^7]
+Algorithm-wise rate of PDD estimates are presented in @tbl-rates.
+On average, estimated PDD rate was `r rates_overall["M"]`% (SD = `r rates_overall["SD"]`, Md = `r rates_overall["Md"]`, range `r rates_overall["minmax"]`).
+Notably, the estimates were substantially lower when FAQ item 9 was used as a criterion of IADL deficit (M = `r rates_FAQ9["M"]`% SD = `r rates_FAQ9["SD"]`, Md = `r rates_FAQ9["Md"]`, range `r rates_FAQ9["minmax"]`) compared to using total FAQ score criterion (M = `r rates_FAQtot["M"]`% SD = `r rates_FAQtot["SD"]`, Md = `r rates_FAQtot["Md"]`, range `r rates_FAQtot["minmax"]`) as demonstrated also in @fig-prev.
+Neither age nor sex or their interaction (*p*s $\geq$ `r min_p`) reliably predicted probable PDD classification across algorithms (see @fig-data and @fig-pars).
 
-```{r}
-#| label: "tbl-rates"
-#| tbl-cap: "Estimates of the rate of probable PDD in the sample."
-
-prev$gtables$rates
-```
-
-[^7]: **JM: For oneself - work on the code here, it ain't good.**
-
-\[Insert Table 3 & Figure 1 here\]
+\[Insert @fig-prev here\]
 
@@
 
-## Criteria Concordance
+## Concordance between Algorithms
+
+```{r}
+#| label: kappas
+
+algos <- list(
+  faq_tot = subset(prev$table, IADL == "FAQ > 7")$type,
+  faq_9 = subset(prev$table, IADL == "FAQ 9 > 1")$type
+)
+
+# Check for overlap:
+continue <- !(any(algos$faq_tot %in% algos$faq_9) || any(algos$faq_9 %in% algos$faq_tot))
+stopifnot("Some algorithm(s) use both IADL operationalisations!" = continue)
+
+# Prepare combinations within IADL operationalisation:
+kappas <-
+  lapply(
+    set_names(names(algos)),
+    function(o)
+      combn(algos[[o]], 2) |>
+      t() |>
+      as_tibble() |>
+      `colnames<-`(c("predictor", "reference")) |>
+      left_join(conc$table, by = c("predictor", "reference")) |>
+      select(Kappa_raw) |>
+      pull()
+  )
+
+# Add cross-operationalisation cases
+kappas$cross <- subset(conc$table, predictor %in% algos$faq_9 & reference %in% algos$faq_tot)$Kappa_raw
+
+# Extract Kappa summaries:
+kappa_sums <-
+  sapply(
+    names(kappas),
+    function(i)
+      paste0(do_summary(kappas[[i]], 2, "M"),", SD = ", do_summary(kappas[[i]], 2, "SD"))
+  )
+```
 
 Results of the analyses of prediction Accuracy, Cohen's $\kappa$, Sensitivity and Specificity are presented in @fig-acc, @fig-kappa, @fig-sens and @fig-spec respectively.
-Numerical results are available ...
-[^15]
+**Numerical results are available at ...
+[^15].** Generally, algorithms that employed the same operationalization of IADL deficit showed substantial pairwise concordance, however, algorithms that operationalized IADL deficit differently did not.
+Whereas among algorithms with identical IADL deficit operationalization, the agreement judged by Cohen's $\kappa$ was moderately high (operationalization by FAQ total score: $\kappa$ = `r kappa_sums["faq_tot"]`; operationalization by FAQ item 9: $\kappa$ = `r kappa_sums["faq_9"]`), among algorithms that differ in IADL deficit operationalization it was low ($\kappa$ = `r kappa_sums["cross"]`).
 
@@
 
-**In this section, we need to discuss among ourselves what and how to report such that the results give enough information, the most important information and are readable as well.**
-
-\[Insert Figure 2[^16] here\]
-
-[^16]: **JM: For thyself - add description.**
+\[Insert @fig-acc here\]
 
@@
 #| label: "fig-acc"
+#| fig-cap: "Prediction accuracy matrix. The matrix depict accuracy of "
 
@@
 ```
+
+## Prediction of Level II Criteria
+
+```{r}
+#| label: "lvlII-stats"
+
+lvlII_rates <- c(
+  faq_tot = subset(prev$table, type == "Lvl.II (1)")$perc |> do_summary(2) |> paste0("%"),
+  faq_9 = subset(prev$table, type == "Lvl.II (2)")$perc |> do_summary(2) |> paste0("%")
+)
+
+smoca_stats <- c(
+  spec = subset(conc$table, predictor == "sMoCA (1)" & reference == "Lvl.II (1)")$Specificity |> do_summary(2),
+  sens = subset(conc$table, predictor == "sMoCA (1)" & reference == "Lvl.II (1)")$Sensitivity |> do_summary(2)
+)
+```
+
+For easier interpretability of our results, we next examine cases where Level II algorithms served as a reference and Level I algorithms as a predictor.
+@tbl-approx shows five Level I algorithms with the highest and five with the lowest accuracy in predicting Level II classification of probable PDD.
+
+When IADL deficit was defined by total FAQ score, the Level II estimate of PDD rate was `r lvlII_rates["faq_tot"]`.
+All five Level I algorithms that approximated the Level II classification most accurately were MoCA-based and defined Executive Function deficit by Clock drawing rather than Lexical fluency test.
+On the other hand, two out of the five Level I algorithms with the lowest accuracy were MMSE-based, whereas the remaining three were MoCA-based and defined Executive Function deficit by Lexical fluency test.
+
+When IADL deficit was defined by FAQ item 9 score, the Level II estimate of PDD rate was `r lvlII_rates["faq_9"]`.
+Overall, the difference between the most accurate and the least accurate Level I algorithms was lower than in the case of IADL deficit being defined by FAQ total score (see @tbl-approx).
+The five most accurate algorithms were all MoCA-based, defined Executive Function deficit by Clock drawing (with threshold < 2) and in majority of cases defined Language deficit by Animal naming.
+Two out of the five Level I algorithms with the lowest accuracy were MMSE-based, whereas the remaining three were MoCA-based and defined Executive Function deficit by Clock drawing (with threshold < 3) and Language deficit by Abstraction.
+
+Finally, if the predictors are sorted by their balanced accuracy (i.e., average of sensitivity and specificity) instead of raw accuracy, the results are similar with the exception that for prediction of Level II with total FAQ score algorithm for probable PDD, the highest balanced accuracy was achieved by the sMoCA algorithm with sensitivity `r smoca_stats["sens"]` and specificity `r smoca_stats["spec"]` (see @tbl-approx-bal).
+
+```{r}
+#| label: "tbl-approx"
+#| tbl-cap: "Level I algorithms for probable PDD as predictors of Level II classification as the reference."
+
+table_levelII_approximations(conc$table, algos, "Accuracy_raw")
+```
+
+\[Insert @tbl-approx here\]
 
@@
 
-Both, the original PDD criteria [@dubois2007] and the call for their change [@kulisevsky2024] allow for several distinct combinations of test indexes to be used to define cognitive impairment.
+Both, the original PDD criteria [@dubois2007] and the call for their change [@kulisevsky2024] allow for several distinct combinations of items to be used to define cognitive impairment.
 Consequently, in this study we derived all algorithms for probable PDD on Level I that are in line with published criteria.
 This procedure parallel the diagnostic algorithm outlined in Table 2 of @dubois2007.
-Specifically, in this study, we varied the exact specification of items 3-5 of this table (i.e., the measure of global cognitive impairment, the measure of the impact on ADLs and the measures of impaired cognition).
+Specifically, in this study, we varied the exact specification of items 3-5 of this table (i.e., the measure of global cognitive impairment, the measure of the impact on IADLs and the measures of impaired cognition).
 
-For each set of criteria (MMSE-based, MoCA-based, sMoCA-based and Level II) we first specified measures and then the thresholds for each measure used to define probable PDD.
-Wherever more than one option was present in either the choice of measure or the choice of threshold, we created an algorithm for each choice in turn.
-The final set of algorithms tested was arrived at by computing the Cartesian product of all possibilities provided by varying measures and thresholds.
+For each set of criteria (MMSE-based, MoCA-based, sMoCA-based and Level II), we first specified the items and then the thresholds for each item used to define probable PDD.
+If more than one option was present in either the choice of the item or the choice of the threshold, we created an algorithm for each choice in turn.
+The final set of algorithms was arrived at by computing the Cartesian product of all possibilities provided by varying items and thresholds.
 All combinations are presented in @tbl-algos.
@@
 
-$Global = \{MoCA < 27\}$
+$Global = \{MoCA < 26\}$
 
@@
 
-Note that the additional language domain adds complexity to establishing a diagnostic algorithm because simply by adding it to the set, the number of potential algorithm doubles.
-Further complexity is added by the fact that there are so far no guidelines for selecting a diagnostic threshold for the Clock drawing and Five-words recall tests, both of which differ from their MMSE counterparts.
+Note that the additional language domain adds complexity to establishing a diagnostic algorithm because simply by adding it to the set of items, the number of potential algorithms doubles.
+Further complexity is added by the fact that there are so far no guidelines for selecting a diagnostic threshold for Clock drawing and Five-words recall tests, both of which differ from their counterparts used by @dubois2007.
+Finally, although the Sevens backwards item has different thresholds in MoCA-based compared to MMSE-based algorithms, this difference is solely due to a difference in scoring whereby 3 points in MoCA correspond to 4 or 5 points in MMSE.
+The Seven backwards item threshold for MoCA-based algorithms used in this study is thus equivalent to its MMSE-based counterpart.
+
 Computing the Cartesian product $Global \times Attention \times Executive \times Construction \times Memory \times Language \times IADL$ yields $1 \times 1 \times 3 \times 1 \times 5 \times 2 \times 2 = 60$ distinct MoCA-based algorithms for probable PDD.
@@
 
-yielding $Global \times IADL$ yields $1 \times 2 = 2$ distinct sMoCA-based algorithms for probable PDD.
+yielding $Global \times IADL$, i.e., $1 \times 2 = 2$ distinct sMoCA-based algorithms for probable PDD.
 
@@
 This yields $1 \times 1 \times 1 \times 1 \times 1 \times 2 = 2$ distinct Level II algorithms for probable PDD in the current study.
+All but the BNT 60 item were evaluated using regression norms published by @bezdicek2017.
+Since the original article used BNT 30 instead of BNT 60, we approximated the deficit in BNT 60 by comparing patients' raw score to age- and education-specific normative values reported by @zemanova2016.
+Specifically, patients whose BNT 60 score fell below 5^th^ percentile of their demographic group in Table 6 of @zemanova2016 were considered to show signs of impaired performance.
+
+### Operationalization of Impaired Cognition
+
+In the original criteria, item 4 of Level I criteria, i.e., impaired cognition, was defined as follows: *"The proposed diagnostic criteria require a profile of cognitive deficits, typical of those described for PD-D, in two or more of four domains."* [@dubois2007, p.
+2316] Consequently, we defined impaired cognition as a deficit in two or more domains of four in MMSE-based criteria and as a deficit in two or more of five domains in MoCA-based criteria.
+sMoCA-based criteria omitted the "impaired cognition" item altogether because they were intended as a shorter screening alternative to classical Level I assessment.
+Finally, for the Level II criteria, we employed standard definition of impaired cognition as the *"[i]mpairment on at least two neuropsychological tests, represented by either two impaired tests in one cognitive domain or one impaired test in two different cognitive domains."* [@litvan2012, Table 1]
 
@@
 
-In this study, we follow the framework proposed by Lundberg et al.
-(2021) for specifying targets of inference (i.e., the estimands) in qunatitative science to increase transparency and connect our statistical evidence to relevant theory.
+In this study, we follow the framework proposed by @lundberg2021 for specifying targets of inference (i.e., the estimands) in qunatitative sciences to increase transparency and connect statistical evidence to relevant theory.
 @tbl-estimands contains verbal description of the components relating to each of our proclaimed research objectives and map them to the population quantity of interest (the theoretical estimand), data-dependent quantity that could be estimated (the empirical estimand) and quantities that are reported in the study (statistical estimates).
@@
 
-The RO1, to estimate the prevalence of PDD and evaluate the diagnostic variability and concordance across different operationalizations of probable PDD, was further subdivided into four distinct research objectives - to estimate the rate of PDD within PD (RO1.1), to estimated variability of this rate (RO1.2), to evaluate predictive value of demographic variables for probable PDD classification (RO1.3) and to evaluate concordance between different probable PDD operationalizations and criteria (RO1.4).
-Estimates relating to RO1.1 and RO1.3 cannot be safely generalized beyond a population of PD patients that are candidates for DBS due to the systematic differences between DBS candidates pool and general PD population (such as lower age of DBS candidates compared to the general PD population).
-On the other hand, the estimates relating to RO1.4 (and to a lesser degree to RO1.2) may not be substantially influenced by the sample at hand as the primary source of their variance comes from variability in measures employed (e.g., MMSE vs MoCA to assess global cognitive performance) rather than variability in subjects' performance.
-Assuming that there is no substantial Differential Item Functioning between DBS candidates and broader PD population, the estimates relating to RO1.4 can be cautiously generalized beyond the current sample.
+The RO1 - to estimate the prevalence of PDD and evaluate the diagnostic variability and concordance across different algorithms of probable PDD - was divided into four distinct research objectives:
+
+- to estimate the rate of PDD within PD (RO1.1),
+- to estimated variability of this rate (RO1.2),
+- to evaluate predictive value of demographic variables for probable PDD classification (RO1.3) and
+- to evaluate concordance between different probable PDD operationalizations and criteria (RO1.4).
+
+Estimates relating to RO1.1 and RO1.3 cannot be safely generalized beyond a population of PD patients that are candidates for DBS due to the systematic differences between DBS candidates pool and general PD population (such as the lower age of DBS candidates compared to the general PD population).
+On the other hand, the estimates relating to RO1.4 (and to a lesser degree to RO1.2^[Because the quantity of interest is a rate and could thus be though of as a sum of binomially distributed PDD occurences divided by the total number of patients, its variance will likely systematically vary with its mean.
+Specifically, as the rate goes from extremes to 0.5, the variance increases.
+Consequently, if our estimate of the rate was lower than the true population rate, e.g., because our sample includes younger patients compared to the general PD population, our estimate of variance would also lower than the true variance of PDD rate in the general PD population.
+Nonetheless, the between-algorithm variability may not be affected by this phenomenon as unlike variability of PDD rate, we do not have reason to assume it comes about by summing independent binomial events]) may not be substantially influenced by the sample at hand as the primary source of their variance might come from variability in measures employed (e.g., MMSE vs MoCA to assess global cognitive performance) rather than variability in patients' performance.
+Assuming that there is no substantial Differential Item Functioning for DBS candidates compared to a broader population of patients with PD, the estimates relating to RO1.4 can be cautiously generalized beyond the current sample.
 
@@
 Empirically, we assess the contribution of each diagnostic feature by examining how variations in operational definitions (e.g., domain-specific thresholds, criteria for functional impairment) influence the statistical estimates derived for the first objective.
-This allows us to identify the diagnostic elements most responsible for inter-operationalization discrepancies.
+This allows us to identify the diagnostic elements most responsible for between-algorithm discrepancies.
 
 ## Supplementary Presentation of Results
+
+```{r}
+#| label: "tbl-rates"
+#| tbl-cap: "Estimates of the rate of probable PDD in the sample."
+
+prev$gtables$rates
+```
+
+```{r}
+#| label: "tbl-approx-bal"
+#| tbl-cap: "Level I algorithms for probable PDD as predictors of Level II classification as the reference arranged by their balanced accuracy score."
+
+table_levelII_approximations(conc$table, algos, "Balanced Accuracy")
+```
+
+```{r}
+#| label: "fig-data"
+
+pred$plots$data
+```
+
+```{r}
+#| label: "fig-pars"
+
+pred$plots$parameters
+```
 

### 2025-05-08 R/describe_concordance.R

@@ -81,7 +81,7 @@ describe_concordance <- function(d0, vars) {
         function(i) ifelse(
           test = reference[i] == predictor[i],
           yes  = '1.00',
-          no   = do_summary(confmats[[i]]$overall[c('Accuracy','AccuracyUpper','AccuracyLower')], 3, 'estCI')
+          no   = do_summary(confmats[[i]]$overall[c('Accuracy','AccuracyUpper','AccuracyLower')], 2, 'estCI')
         )
       ),
       # Raw values for Kappa & Accuracy (and NIR):

### 2025-05-09 _manuscript/manuscript.qmd
Minor changes to manuscript.

--- 2025-05-07 _manuscript/manuscript.qmd
@@
 According to a recent metanalysis, the epidemiological estimates of PDD prevalence among individuals with PD vary widely, ranging from 14% to 55%, depending on methodological criteria employed [@sousa2022].
-Moreover, factors such as patients` sex [@cereda2016], age and disease duration appear to modulate the risk of cognitive decline [@oh2016; @rana2012].
+Moreover, factors such as patients\` sex [@cereda2016], age and disease duration appear to modulate the risk of cognitive decline [@oh2016; @rana2012].
 
@@
 Level I was assessed using the Mini-Mental State Examination (MMSE) [@stepankova2014; @folstein1975] and the Montreal Cognitive Assessment (MoCA) [@kopecek2016; @nasreddine2005].
-The neuropsychological assessment at Level II covered five cognitive domains, each evaluated through specific tests as follows: attention and working memory assessed using Trail Making Test Part A (TMT-A) [@bezdicek2012; @reitan2004], and WAIS Digit Span Backward (WAIS DSB) [@wechsler1997]; executive function evaluated via Categorical Verbal Fluency (CF) [@benton1989][^1], and subtest from the Prague Stroop Test – Colors (PST-C) [@bezdicek2021]; language measured with the WAIS Similarities subtest [@wechsler1997], and the Boston Naming Test (BNT-60) [@kaplan1983]; memory examined using the Rey Auditory Verbal Learning Test (RAVLT) [@frydrychova2018; @rey1964][^2] for delayed recall, and the Brief Visual Memory Test–Revised (BVMTR) [@havlik2020; @benedict1997] for *delayed recall*[^3], or WAIS Family Pictures subtest [@wechsler1997] *including delayed recall and forced choice recognition*[^4]; visuospatial function assessed through the Judgment of Line Orientation Test (JoLO) [@benton1983], and Clock Drawing Test (CLOX) [@royall1998].
+The neuropsychological assessment at Level II covered five cognitive domains, each evaluated through specific tests as follows: attention and working memory assessed using Trail Making Test Part A (TMT-A) [@bezdicek2012; @reitan2004], and WAIS Digit Span Backward (WAIS DSB) [@wechsler1997]; executive function evaluated via Categorical Verbal Fluency (CF) [@nikolai2015], and subtest from the Prague Stroop Test – Colors (PST-C) [@bezdicek2021]; language measured with the WAIS Similarities subtest [@wechsler1997], and the Boston Naming Test (BNT-60) [@kaplan1983]; memory examined using the Rey Auditory Verbal Learning Test (RAVLT) [@frydrychova2018; @rey1964; @bezdicek2013] for delayed recall, and the Brief Visual Memory Test–Revised (BVMTR) [@havlik2020; @benedict1997] for *delayed recall*[^2], or WAIS Family Pictures subtest [@wechsler1997] *including delayed recall and forced choice recognition*[^3]; visuospatial function assessed through the Judgment of Line Orientation Test (JoLO) [@benton1983], and Clock Drawing Test (CLOX) [@royall1998].
 
@@
 
-[^2]: **JM: Do not forget this - https://doi.org/10.1080/13825585.2013.865699**
+[^2]: **JM: The same label is used for RAVLT-DR and BVMT-R-DR here, we should sort it out.**
 
-[^3]: **JM: The same label is used for RAVLT-DR and BVMT-R-DR here, we should sort it out.**
+[^3]: **JM: We did not use any recognition. Should tidy this section such that it includes only indexes used by our study.**
 
-[^4]: **JM: We did not use any recognition. Should tidy this section such that it includes only indexes used by our study.**
-
-In addition to the core cognitive assessments, *tasks such as*[^5] the Clock Drawing Test (CDT) and Letter Fluency tasks were included to capture domain-specific impairments.
+In addition to the core cognitive assessments, *tasks such as*[^4] the Clock Drawing Test (CDT) and Letter Fluency tasks were included to capture domain-specific impairments.
 The classification of Parkinson’s Disease Dementia (PDD) based on Level I criteria was determined using established scoring thresholds from the original criteria [@dubois2007] with corresponding MoCA equivalents.
 
-[^5]: **JM: Stating such as "such as" are usually not appropritate in Methods section of a scientific study, because all methods should be reported exactly.**
+[^4]: **JM: Stating such as "such as" are usually not appropritate in Methods section of a scientific study, because all methods should be reported exactly.**
 
@@
 
-## Theoretical and Empirical Estimands[^6]
+## Theoretical and Empirical Estimands[^5]
 
-[^6]: **JM: Put most of it to Appendix.**
+[^5]: **JM: Put most of it to Appendix.**
 
@@
 
-## Operationalization of Parkinson's Disease Dementia[^7]
+## Operationalization of Parkinson's Disease Dementia[^6]
 
-[^7]: **JM: Add a section with some logic (Cartezian product) to Appendix as well.**
+[^6]: **JM: Add a section with some logic (Cartezian product) to Appendix as well.**
 
-In this study, we applied three distinct sets of diagnostic criteria for **probable**[^8] PDD at Level I.
+In this study, we applied three distinct sets of diagnostic criteria for **probable**[^7] PDD at Level I.
 The first set was based on the original framework [@dubois2007], which utilized the Mini-Mental State Examination (MMSE) as a global cognitive screening tool, supplemented by assessments of attention, executive function, visuospatial abilities, and memory.
@@
 **The Level II methodology, including the use of a regression-based normative scoring approach, has been detailed in a prior study [@bezdicek2017].
-Refer to @tbl-criteria**[^9] to a summary of the components and scoring thresholds of each diagnostic criterion.
+Refer to @tbl-criteria**[^8] to a summary of the components and scoring thresholds of each diagnostic criterion.
 All non-cognitive criteria of probable PDD (i.e., diagnosis of PD that developed before dementia and absence of Major Depression, delirium or other abnormalities that obscure diagnosis were established by an independent neurological and psychiatric assessments and held true for all patients in the sample).
 
-[^8]: **JM: We should unite the terminlogy.
+[^7]: **JM: We should unite the terminlogy.
 Most importantly, I reckon, we should differentiate between "probable PDD" (what is implied by meeting all the criteria within an operationalization) and "PDD" (the latent state of patient's cognition); and to use consistently the following trifecta - "criterion," "operationalization" and their "component".**
 
-[^9]: **JM: Placeholder until we make the table proper.**
+[^8]: **JM: Placeholder until we make the table proper.**
 
@@
 For each pair of operationalizations, the analysis was repeated twice such that each variables of the pair served once as the reference and once as the predictor.
-Following measures were used to evaluate pairwise concordance between different operationalizations of PDD criteria: 1) Cohen's $\kappa$ with its 95% confidence interval (CI) computed via the `cohen.kappa()` function from the R package *psych* [@revelle2024]; 2) Accuracy (i.e., the proportion of correct predictions, both true positives and true negatives, among the total number of cases) with its 95% CI; 3) Sensitivity/Recall (i.e., the proportion of true positives); and 4) Specificity (i.e., the proportion of true negatives).[^10]
+Following measures were used to evaluate pairwise concordance between different operationalizations of PDD criteria: 1) Cohen's $\kappa$ with its 95% confidence interval (CI) computed via the `cohen.kappa()` function from the R package *psych* [@revelle2024]; 2) Accuracy (i.e., the proportion of correct predictions, both true positives and true negatives, among the total number of cases) with its 95% CI; 3) Sensitivity/Recall (i.e., the proportion of true positives); and 4) Specificity (i.e., the proportion of true negatives).[^9]
 
-[^10]: Unlike Cohen's $\kappa$, Accuracy, Sensitivity and Specificity are not symmetrical, i.e., their value depend on which variable is considered reference and which is considered predictor.
+[^9]: Unlike Cohen's $\kappa$, Accuracy, Sensitivity and Specificity are not symmetrical, i.e., their value depend on which variable is considered reference and which is considered predictor.
 Consequently, we report these values twice for each pair of operationalizations.
@@
 All analyses were conducted with the R (version `r with(version, paste(major, minor, sep="."))`) software environment for statistical computing [@rsoft].
-The software code supporting this article is available at <https://github.com/josefmana/DemCr1t.git>.[^11]
+The software code supporting this article is available at <https://github.com/josefmana/DemCr1t.git>.[^10]
 
-[^11]: **JM: Do not forget to make it public before submitted!**
+[^10]: **JM: Do not forget to make it public before submitted!**
 
@@
 
-A total of `r sum(data$incl)` patients were included. Demographic and clinical characteristics of the sample are summarized in @tbl-desc. **...**[^12]
+A total of `r sum(data$incl)` patients were included. Demographic and clinical characteristics of the sample are summarized in @tbl-desc. **...**[^11]
 
-[^12]: **JM: Give some brief impressions from the Table 2 here.
+[^11]: **JM: Give some brief impressions from the Table 2 here.
 E.g., evaluate sex proportion, age, (we are missing education for some reason) etc.
@@
 On average, estimated prevalence was `r do_summary(perc, 2, "M")`% (SD = `r do_summary(perc, 2, "SD")`, Md = `r do_summary(perc, 2, "Md")`, range `r do_summary(perc, 2, "minmax")`).
-Notably, the prevalence estimate was substantially lower when FAQ item 9 was used as a criterion of IADL deficit (M = `r do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, "M")`%, SD = `r do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, "SD")`, range `r do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, "minmax")`) compared to using total FAQ score criterion (M = `r do_summary(perc[!grepl("FAQ 9", prev$table$IADL)], 2, "M")`%, SD = `r do_summary(perc[!grepl("FAQ 9", prev$table$IADL)], 2, "SD")`, range `r do_summary(perc[!grepl("FAQ 9", prev$table$IADL)], 2, "minmax")`) as demonstrated also in @fig-prev.[^13]
+Notably, the prevalence estimate was substantially lower when FAQ item 9 was used as a criterion of IADL deficit (M = `r do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, "M")`%, SD = `r do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, "SD")`, range `r do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, "minmax")`) compared to using total FAQ score criterion (M = `r do_summary(perc[!grepl("FAQ 9", prev$table$IADL)], 2, "M")`%, SD = `r do_summary(perc[!grepl("FAQ 9", prev$table$IADL)], 2, "SD")`, range `r do_summary(perc[!grepl("FAQ 9", prev$table$IADL)], 2, "minmax")`) as demonstrated also in @fig-prev.[^12]
 
-[^13]: **JM: For oneself - work on the code here, it ain't good.**
+[^12]: **JM: For oneself - work on the code here, it ain't good.**
 
-\[Insert Table 3 and Figure 1[^14] here\]
+\[Insert Table 3 and Figure 1[^13] here\]
 
-[^14]: **JM: For thyself - add description.**
+[^13]: **JM: For thyself - add description.**
 
@@
 Numerical results are available ...
-[^15]
+[^14]
 
-[^15]: **JM: ... as some kind of Supplementary Table, ideally html but Excel file would work as well.**
+[^14]: **JM: ... as some kind of Supplementary Table, ideally html but Excel file would work as well.**
 
@@
 
-\[Insert Figure 2[^16] here\]
+\[Insert Figure 2[^15] here\]
 
-[^16]: **JM: For thyself - add description.**
+[^15]: **JM: For thyself - add description.**
 
@@
 Notably, our cohort consisted of patients being evaluated for DBS, a procedure typically reserved for individuals with relatively preserved cognitive function, which inherently biases against higher PDD prevalence rate.
-These comparisons emphasize that the observed lower prevalence in our sample is likely attributable to the preselection of cognitively intact individuals for DBS consideration, as well as to methodological variations such as the use of brief screening tool and differing IADL operationalizations.[^17]
+These comparisons emphasize that the observed lower prevalence in our sample is likely attributable to the preselection of cognitively intact individuals for DBS consideration, as well as to methodological variations such as the use of brief screening tool and differing IADL operationalizations.[^16]
 
-[^17]: **Keep it?**
+[^16]: **Keep it?**
 
-## Cognitive and Clinical Context [^18]
+## Cognitive and Clinical Context [^17]
 
-[^18]: **Keep this section?**
+[^17]: **Keep this section?**
 

### 2025-05-09 _manuscript/manuscript.qmd

--- 2025-05-08 _manuscript/manuscript.qmd
@@
 
-[^11]: **JM: Do not forget to make it public before submitted!**
+[^10]: **JM: Do not forget to make it public before submitted!**
 
@@
 
-[^15]: **JM: ... as some kind of Supplementary Table, ideally html but Excel file would work as well.**
+[^14]: **JM: ... as some kind of Supplementary Table, ideally html but Excel file would work as well.**
 

### 2025-05-09 _manuscript/manuscript.qmd
Refinement of intro and discussion.

@@
 Furthermore, both sets of criteria are compared to PDD diagnosed on Level II.
-The study aims to address following research objectives (RO): (RO1) To estimate the prevalence of Parkinson’s disease dementia (PDD) and evaluate the diagnostic variability and concordance across different PDD criteria.
+The study aims to address following research objectives (RO): (RO1) To estimate the Parkinson’s disease dementia (PDD) rates and evaluate the diagnostic variability and concordance across different PDD criteria.
 (RO2) To identify specific diagnostic components that contribute to variability in PDD classification across the applied criteria.
@@
 Level I was assessed using the Mini-Mental State Examination (MMSE) [@stepankova2014; @folstein1975] and the Montreal Cognitive Assessment (MoCA) [@kopecek2016; @nasreddine2005].
-The neuropsychological assessment at Level II covered five cognitive domains, each evaluated through specific tests as follows: attention and working memory assessed using Trail Making Test Part A (TMT-A) [@bezdicek2012; @reitan2004], and WAIS Digit Span Backward (WAIS DSB) [@wechsler1997]; executive function evaluated via Categorical Verbal Fluency (CF) [@nikolai2015], and subtest from the Prague Stroop Test – Colors (PST-C) [@bezdicek2021]; language measured with the WAIS Similarities subtest [@wechsler1997], and the Boston Naming Test (BNT-60) [@kaplan1983]; memory examined using the Rey Auditory Verbal Learning Test (RAVLT) [@frydrychova2018; @rey1964; @bezdicek2013] for delayed recall, and the Brief Visual Memory Test–Revised (BVMTR) [@havlik2020; @benedict1997] for *delayed recall*[^2], or WAIS Family Pictures subtest [@wechsler1997] *including delayed recall and forced choice recognition*[^3]; visuospatial function assessed through the Judgment of Line Orientation Test (JoLO) [@benton1983], and Clock Drawing Test (CLOX) [@royall1998].
+The neuropsychological assessment at Level II covered five cognitive domains, each evaluated through specific tests as follows: attention and working memory assessed using Trail Making Test Part A (TMT-A) [@bezdicek2012; @reitan2004], and WAIS Digit Span Backward (WAIS DSB) [@wechsler1997]; executive function evaluated via Categorical Verbal Fluency (CF) [@nikolai2015], and subtest from the Prague Stroop Test – Colors (PST-C) [@bezdicek2021]; language measured with the WAIS Similarities subtest [@wechsler1997], and the Boston Naming Test (BNT-60) [@kaplan1983]; memory examined using the Rey Auditory Verbal Learning Test (RAVLT) [@frydrychova2018; @rey1964; @bezdicek2013] delayed recall, and the Brief Visual Memory Test–Revised (BVMTR) [@havlik2020; @benedict1997] delayed recall, or WAIS Family Pictures subtest [@wechsler1997] delayed recall; visuospatial function assessed through the Judgment of Line Orientation Test (JoLO) [@benton1983], and Clock Drawing Test (CLOX) [@royall1998].
 
-[^1]: **JM: This one instead - https://doi.org/10.14735/amcsnn2015292**
-
-[^2]: **JM: The same label is used for RAVLT-DR and BVMT-R-DR here, we should sort it out.**
-
-[^3]: **JM: We did not use any recognition. Should tidy this section such that it includes only indexes used by our study.**
-
-In addition to the core cognitive assessments, *tasks such as*[^4] the Clock Drawing Test (CDT) and Letter Fluency tasks were included to capture domain-specific impairments.
+In addition to the core cognitive assessments, namely the Clock Drawing Test (CDT) and Letter Fluency tasks were included to capture domain-specific impairments.
 The classification of Parkinson’s Disease Dementia (PDD) based on Level I criteria was determined using established scoring thresholds from the original criteria [@dubois2007] with corresponding MoCA equivalents.
-
-[^4]: **JM: Stating such as "such as" are usually not appropritate in Methods section of a scientific study, because all methods should be reported exactly.**
 
@@
 
-\[Insert @tbl-crits here\]
+[Insert @tbl-crits here]
 
@@
 
-To address study objectives, we started by repeatedly assigning each patient the diagnosis of probable PDD based on each PDD algorithm listed in @tbl-crits (see also @tbl-algos) resulting in a `r sum(data$incl)` (patients) $\times$ `r sum(critnum)` (operationalizations) matrix where each cell indicates whether a patient (row) meets criteria for probable PDD according to an algorithm (column)^[**JM: This could and should be shared most likely, as long, as we anonymize properly.
-Let's ask Oto Mestek and the NPO team how and if is it possible.**].
+To address study objectives, we started by repeatedly assigning each patient the diagnosis of probable PDD based on each PDD algorithm listed in @tbl-crits (see also @tbl-algos) resulting in a `r sum(data$incl)` (patients) $\times$ `r sum(critnum)` (operationalizations) matrix where each cell indicates whether a patient (row) meets criteria for probable PDD according to an algorithm (column)[^1].
 PDD rate estimates were computed as $\frac{N_{PDD}}{N_{total}}$ separately for each algortihm.
@@
 
+[^1]: **JM: This could and should be shared most likely, as long, as we anonymize properly. Let's ask Oto Mestek and the NPO team how and if is it possible.**
+
 Next, a set of two class cross-tabulations with associated statistics was computed for each pair of algorithms via the `confusionMatrix()` function from the R package *caret* [@kuhn2008].
 For each pair of algorithms, the analysis was repeated twice such that each variable of the pair served once as the reference and once as the predictor.
-Following measures were used to evaluate pairwise concordance between different algorithms for probable PDD: 1) Cohen's $\kappa$ with its 95% confidence interval (CI) computed via the `cohen.kappa()` function from the R package *psych* [@revelle2024]; 2) Accuracy (i.e., the proportion of correct predictions, both true positives and true negatives, among the total number of cases) with its 95% CI; 3) Sensitivity/Recall (i.e., the proportion of true positives); and 4) Specificity (i.e., the proportion of true negatives).[^10]
+Following measures were used to evaluate pairwise concordance between different algorithms for probable PDD: 1) Cohen's $\kappa$ with its 95% confidence interval (CI) computed via the `cohen.kappa()` function from the R package *psych* [@revelle2024]; 2) Accuracy (i.e., the proportion of correct predictions, both true positives and true negatives, among the total number of cases) with its 95% CI; 3) Sensitivity/Recall (i.e., the proportion of true positives); and 4) Specificity (i.e., the proportion of true negatives).[^2]
 
-[^10]: Unlike Cohen's $\kappa$, Accuracy, Sensitivity and Specificity are not symmetrical, i.e., their value depends on which variable is considered reference and which is considered predictor.
-Consequently, we report these values twice for each pair of algorithms.
-Note that the Sensitivity of a reference/predictor pair corresponds to the Positive Predictive Value if their roles were reversed.
-The same relationship holds true between the Specificity and the Negative Predictive Value.
+[^2]: **JM: Do not forget to make it public before submitted!**
 
@@
 All analyses were conducted within the R (version `r with(version, paste(major, minor, sep="."))`) software environment for statistical computing [@rsoft].
-The software code supporting this article is available at <https://github.com/josefmana/DemCr1t.git>.[^11]
-
-[^10]: **JM: Do not forget to make it public before submitted!**
+The software code supporting this article is available at <https://github.com/josefmana/DemCr1t.git>.\[\^11\]
 
@@
 
-\[Insert @tbl-desc here\]
+[Insert @tbl-desc here]
 
@@
 
-\[Insert @fig-prev here\]
+[Insert @fig-prev here]
 
@@
 **Numerical results are available at ...
-[^15].** Generally, algorithms that employed the same operationalization of IADL deficit showed substantial pairwise concordance, however, algorithms that operationalized IADL deficit differently did not.
+\[\^15\].** Generally, algorithms that employed the same operationalization of IADL deficit showed substantial pairwise concordance, however, algorithms that operationalized IADL deficit differently did not.
 Whereas among algorithms with identical IADL deficit operationalization, the agreement judged by Cohen's $\kappa$ was moderately high (operationalization by FAQ total score: $\kappa$ = `r kappa_sums["faq_tot"]`; operationalization by FAQ item 9: $\kappa$ = `r kappa_sums["faq_9"]`), among algorithms that differ in IADL deficit operationalization it was low ($\kappa$ = `r kappa_sums["cross"]`).
 
-[^14]: **JM: ... as some kind of Supplementary Table, ideally html but Excel file would work as well.**
-
-\[Insert @fig-acc here\]
+[Insert @fig-acc here]
 
@@
 Overall, the difference between the most accurate and the least accurate Level I algorithms was lower than in the case of IADL deficit being defined by FAQ total score (see @tbl-approx).
-The five most accurate algorithms were all MoCA-based, defined Executive Function deficit by Clock drawing (with threshold < 2) and in majority of cases defined Language deficit by Animal naming.
-Two out of the five Level I algorithms with the lowest accuracy were MMSE-based, whereas the remaining three were MoCA-based and defined Executive Function deficit by Clock drawing (with threshold < 3) and Language deficit by Abstraction.
+The five most accurate algorithms were all MoCA-based, defined Executive Function deficit by Clock drawing (with threshold \< 2) and in majority of cases defined Language deficit by Animal naming.
+Two out of the five Level I algorithms with the lowest accuracy were MMSE-based, whereas the remaining three were MoCA-based and defined Executive Function deficit by Clock drawing (with threshold \< 3) and Language deficit by Abstraction.
 
@@
 
-\[Insert @tbl-approx here\]
+[Insert @tbl-approx here]
 
@@
 
-This study evaluated the applicability and validity of diagnostic frameworks for diagnosing probable PDD within a cohort of patients considered for deep brain stimulation (DBS).
-Our results demonstrate that diagnostic outcomes are markedly influenced by the chosen type of operationalization, particularly in relation to the assessment of the cognitive domains and IADLs.
+This study comprehensively examined the diagnostic variability and validity of Level I criteria for PDD in DBS treatment candidates.
+By systematically comparing 68 diagnostic algorithms based on MMSE, MoCA, sMoCA, and Level II criteria, our findings highlight the significant influence of algorithm design.
 
@@
 
-One of the key findings of this study was the broad range of estimated PDD rate depending on the operationalization strategy.
-As seen in Table 3, the rate of PDD presence estimates varied from 2.00% to 16.75%, with the highest being derived from a combination of the sMoCA and the total FAQ score combination.
-When only FAQ item 9 was used to determine IADL deficits, the rate estimates were significantly lower (M = 3.09%, SD = 0.48), underscoring significant influence of functional assessment choice on diagnostic outcomes.
+Our results showed a wide range in estimated PDD prevalence across 68 algorithms, ranging from 2.00% to 16.75%.
+Notably, estimates were substantially lower when using FAQ item 9 alone versus the full FAQ scale.
+This discrepancy underscores the diagnostic weight of IADL assessment.
+Our PDD rates were consistently lower compared to previous studied, demonstrating wide variability in reported PDD rate among PD patients.
+For instance, retrospective study reported a PDD rate of 19.7% [@rana2012], while a clinical investigation found even higher rate of around 30% [@aarsland2011].
+A recent complex meta-analysis synthesizing global data placed the pooled rate at 26.30% [@sousa2022].
+**The discrepancy is likely attributable to our DBS specific cohort, which differs in important characteristics, such as younger age of the patients [@rana2012].
+/ Compared to these estimates, our study reports generally lower PDD rates, likely reflecting differences in sample characteristics, diagnostic criteria, and methodology.
+[^3]**
 
-These findings are lower in comparison with previous research demonstrating wide variability in reported PDD rate among PD patients.
-For instance, retrospective study reported a PDD rate of 19.7% [@rana2012], while a clinical investigation by Aarsland et al.
-found even higher rate of around 30% [@aarsland2011].
-A recent complex meta-analysis synthesizing global data placed the pooled rate at 26.30% [@sousa2022].
-Compared to these estimates, our study reports generally lower prevalence rates, likely reflecting differences in sample characteristics, diagnostic criteria, and methodology.
+[^3]: **MM to JM: Which sentence do you like better?**
 
-Notably, our cohort consisted of patients being evaluated for DBS, a procedure typically reserved for individuals with relatively preserved cognitive function, which inherently biases against higher PDD prevalence rate.
-These comparisons emphasize that the observed lower prevalence in our sample is likely attributable to the preselection of cognitively intact individuals for DBS consideration, as well as to methodological variations such as the use of brief screening tool and differing IADL operationalizations.[^16]
+## Concordance Between Diagnostic Algorithms
 
-[^16]: **Keep it?**
+Our pairwise comparison of diagnostic algorithms showed that agreement was notably stronger among those using the same IADL operationalization (κ range from 0.75, SD = 0.13 to 0.86, SD = 0.09) compared to those using different IADL definitions (κ = 0.34, SD = 0.08).
+This demonstrates that even subtle methodological differences can yield substantially divergent diagnostic outcomes.
+Such findings are critical for clinicians relying on Level I criteria for eligibility decisions, as the choice of algorithm could lead to contradictory classifications of PDD.
 
-## Cognitive and Clinical Context [^17]
+## Predictive Validity Comaprison With Level II Criteria
 
-[^17]: **Keep this section?**
+When using Level II diagnosis as the reference, MoCA-based Level I algorithms, particularly those using Clock Drawing used to assess executive function, demonstrated the highest predictive accuracy.
+This supports recent proposals to modernize PDD diagnostic frameworks [@kulisevsky2024], favoring MoCA-derived components and broader functional assessment tools.
+In contrast, MMSE-based algorithms consistently underperformed, suggesting limited sensitivity in capturing the nuanced cognitive deficits characteristic of PDD.
 
-The relatively low prevalence of PDD across operationalizations may also reflect the relatively preserved cognitive profile of the DBS cohort, as evident in Table 2.
-The mean MoCA (M = 24.07, SD = 3.48) and MMSE (M = 26.69, SD = 2.22) scores suggest that on average the patients were functioning at a globally intact level.
-This likely reflects the pre-selection of cognitively preserved individuals for DBS, in line with standard eligibility criteria [@bronstein2011].
+The advantage of MoCA-based algorithm was particularly evident when using the FAQ total score to define functional impairment in comparison with using FAQ item 9 only.
+In these cases, the best-performing algorithms achieved accuracy rates of 94–99%, together with balanced sensitivity and specificity, reinforcing the clinical utility.
 
-Memory performance, such as RAVLT and BVMTR delayed recall scores, also pointed to only mild deficits, particularly in domains central to the PDD vs.
-PD-MCI distinction.
-This aligns with previous findings that Level I criteria may overestimate dementia in patients with subtle impairments unless operational thresholds are rigorously defined [@aarsland2021; @bezdicek2016].
+## Clinical Implications
 
-## Implications for DBS Eligibility and Clinical Practice
+These findings have immediate implications for DBS candidacy assessment.
+Given that a diagnosis of PDD typically precludes surgical intervention, inconsistent application of diagnostic criteria can lead to inequitable access to the treatment.
+Our data emphasize the need for harmonized assessment protocols that balance sensitivity with specificity, particularly in high-stakes clinical decision-making.
 
-Our findings also bear direct implications for clinical decision-making, especially in the context of surgical candidacy for DBS.
-Given that PDD remains a contraindication for DBS, the observed diagnostic instability could result in disparate treatment decisions based solely on which cognitive criteria are employed.
-Thus, harmonization of assessment procedures and operational cutoffs is essential to ensure equitable access to surgical treatment while maintaining diagnostic rigor.
-
-Further, our data support the use of multidimensional assessment strategies, such as Level II neuropsychological batteries or regression-based normative comparisons, as confirmatory tools in diagnostically ambiguous cases.
-These approaches may enhance diagnostic precision and mitigate the risks of both false positives and unjustified treatment exclusion.
+Moreover, the relative success of multidomain assessments, such as Level II protocols or regression-based normative approaches, highlights the importance of psychometric rigor in diagnosing PDD.
+These approaches may serve as confirmatory tools in cases where Level I assessments yield ambiguous results.
 
@@
 
-This study systematically investigated the application of multiple Level I diagnostic criteria for Parkinson’s disease dementia (PDD) within a cohort of patients considered for deep brain stimulation (DBS).
-The findings reveal substantial variability in prevalence estimates, strongly influenced by the choice of cognitive screening instruments and the operationalization of functional impairment.
-The divergence observed across operationalizations demonstrates the sensitivity of diagnostic outcomes to seemingly minor methodological choices.
+This study systematically investigated the application of multiple Level I diagnostic criteria for PDD within a cohort of patients considered for DBS.
+The findings reveal substantial variability in PDD rates estimates, strongly influenced by the choice of cognitive screening instruments and the operationalization of functional impairment.
+The divergence observed across algorithms demonstrates the sensitivity of diagnostic outcomes to seemingly minor methodological choices.
 
-The proposed revisions to the diagnostic framework [@kulisevsky2024] criteria offer enhanced sensitivity by leveraging MoCA-based components and broader IADL assessments, the use must be cautiously calibrated to prevent over-diagnosis in populations with mild or borderline cognitive deficits.
+The proposed revisions to the diagnostic framework [@kulisevsky2024] criteria offer enhanced sensitivity by leveraging MoCA-based components and broader IADL assessments.
 Conversely, overly conservative criteria, such as reliance on pill questionnaire (i.e.
-FAQ item 9 equivalence) may fail to detect meaningful functional decline and thus under-identify true cases of probable PDD.
+FAQ item 9 equivalence) may fail to detect meaningful functional decline and thus under-identify true cases of PDD presence.
 
 Ultimately, this study contributes to the improving landscape of PDD diagnostics by offering empirical evidence for the refinement of Level I criteria and reinforcing the value of psychometric rigor in clinical neuropsychology.
-Future work should extend this validation to longitudinal trajectories and integrate neurobiological correlates, ensuring that cognitive criteria remain both scientifically grounded and clinically actionable.
+Future research should validate these operationalizations longitudinally and explore integration with biomarkers to further enhance diagnostic reliability.
 
@@
 sMoCA-based criteria omitted the "impaired cognition" item altogether because they were intended as a shorter screening alternative to classical Level I assessment.
-Finally, for the Level II criteria, we employed standard definition of impaired cognition as the *"[i]mpairment on at least two neuropsychological tests, represented by either two impaired tests in one cognitive domain or one impaired test in two different cognitive domains."* [@litvan2012, Table 1]
+Finally, for the Level II criteria, we employed standard definition of impaired cognition as the *"\[i\]mpairment on at least two neuropsychological tests, represented by either two impaired tests in one cognitive domain or one impaired test in two different cognitive domains."* [@litvan2012, Table 1]
 
@@
 
-The RO1 - to estimate the prevalence of PDD and evaluate the diagnostic variability and concordance across different algorithms of probable PDD - was divided into four distinct research objectives:
+The RO1 - to estimate the PDD rate and evaluate the diagnostic variability and concordance across different algorithms of probable PDD - was divided into four distinct research objectives:
 
@@
 Estimates relating to RO1.1 and RO1.3 cannot be safely generalized beyond a population of PD patients that are candidates for DBS due to the systematic differences between DBS candidates pool and general PD population (such as the lower age of DBS candidates compared to the general PD population).
-On the other hand, the estimates relating to RO1.4 (and to a lesser degree to RO1.2^[Because the quantity of interest is a rate and could thus be though of as a sum of binomially distributed PDD occurences divided by the total number of patients, its variance will likely systematically vary with its mean.
+On the other hand, the estimates relating to RO1.4 (and to a lesser degree to RO1.2[^3]) may not be substantially influenced by the sample at hand as the primary source of their variance might come from variability in measures employed (e.g., MMSE vs MoCA to assess global cognitive performance) rather than variability in patients' performance.
+Assuming that there is no substantial Differential Item Functioning for DBS candidates compared to a broader population of patients with PD, the estimates relating to RO1.4 can be cautiously generalized beyond the current sample.
+
+[^3]: Because the quantity of interest is a rate and could thus be though of as a sum of binomially distributed PDD occurences divided by the total number of patients, its variance will likely systematically vary with its mean.
 Specifically, as the rate goes from extremes to 0.5, the variance increases.
 Consequently, if our estimate of the rate was lower than the true population rate, e.g., because our sample includes younger patients compared to the general PD population, our estimate of variance would also lower than the true variance of PDD rate in the general PD population.
-Nonetheless, the between-algorithm variability may not be affected by this phenomenon as unlike variability of PDD rate, we do not have reason to assume it comes about by summing independent binomial events]) may not be substantially influenced by the sample at hand as the primary source of their variance might come from variability in measures employed (e.g., MMSE vs MoCA to assess global cognitive performance) rather than variability in patients' performance.
-Assuming that there is no substantial Differential Item Functioning for DBS candidates compared to a broader population of patients with PD, the estimates relating to RO1.4 can be cautiously generalized beyond the current sample.
+Nonetheless, the between-algorithm variability may not be affected by this phenomenon as unlike variability of PDD rate, we do not have reason to assume it comes about by summing independent binomial events
 

### 2025-05-12 _manuscript/manuscript.qmd
Last few changes in the text, now up to JM to check the text for  possible baloney :)

@@
 
-Parkinson’s disease (PD) is a neurodegenerative disorder characterised by a progressive onset of motor symptoms, including rigidity, bradykinesia, postural instability and resting tremor.
-Beyond motor symptoms, patients routinely suffer from non-motor impairments [@postuma2015], such as cognitive decline.
-This decline might peak in Parkinson’s disease dementia (PDD) in a subset of patients [@meireles2012].
+Parkinson’s disease (PD) is a neurodegenerative disorder defined by a gradual onset of motor impairment, including rigidity, bradykinesia, postural instability and resting tremor.
+Additionally, patients frequently suffer from a range of non-motor impairments [@postuma2015], particularly cognitive decline.
+This state might result in Parkinson’s disease dementia (PDD) in a subset of patients [@meireles2012].
 
 Despite its clinical relevance, the diagnosis of PDD remains complex.
-The original International Parkinson and Movement Disorder Society (MDS) diagnostic criteria [@dubois2007], first formalized in 2007, were heavily influenced by frameworks established primarily for Alzheimer’s disease (AD).
-Although these criteria provided a valuable initial foundation, they lacked the specificity required to capture the distinct pathophysiological and cognitive features of PD-related dementia [@emre2007; @yamashita2023].
+The original criteria formalized in 2007 by the International Parkinson and Movement Disorder Society (MDS) [@dubois2007] were heavily influenced by frameworks established primarily for Alzheimer’s disease (AD).
+Although these criteria provided a valuable base, they lacked the specificity to detect the specific cognitive aspects of PDD [@emre2007; @yamashita2023].
 
-To enhance diagnostic accuracy, the MDS introduced a two-level system for Parkinson’s disease dementia.
-Level I criteria are designed for clinical use, relying on brief cognitive assessments and clinical judgment.
-Level II criteria involve comprehensive neuropsychological testing across multiple domains, providing greater diagnostic detail for research or specialized settings [@emre2007].
+In order to increase diagnostic accuracy, the MDS also developed a levelled system for PDD detection.
+Level I criteria are designed primarily for clinical use, consisting of brief cognitive assessments and relying on clinical judgement.
+In contrast, Level II criteria use comprehensive neuropsychological testing across cognitive domains to provide greater diagnostic specificity [@emre2007].
 
-One notable feature of the original Level I criteria was the provision of an algorithm that allowed for flexibility in test selection [@dubois2007].
-Specifically, clinicians could choose between months reversed or seven backwards for attention assessment, lexical fluency or clock drawing for executive function evaluation.
-Moreover, the critaria included MMSE pentagons for evaluation of visuospatial ability, and three-word recall for memory assessment.
-Agreement across different criteria allows for the parallel computation of inter-rater reliability , which, in turn, facilitates the calculation of construct validity [@conway1995] further strengthening the diagnostic framework for PDD.
+One specific feature of the original Level I criteria was the presence of an algorithm allowing flexibility in test selection.
+The criteria included MMSE pentagons for evaluation of visuospatial ability, three-word recall for memory assessment,  months reversed or seven backwards for attention assessment and lexical fluency or clock drawing for executive function assessment [@dubois2007].
 
-According to a recent metanalysis, the epidemiological estimates of PDD prevalence among individuals with PD vary widely, ranging from 14% to 55%, depending on methodological criteria employed [@sousa2022].
-Moreover, factors such as patients\` sex [@cereda2016], age and disease duration appear to modulate the risk of cognitive decline [@oh2016; @rana2012].
-
+According to a recent meta-nalysis, the PDD prevalence estimates vary widely, ranging from 14% up to 55%, depending on methodological criteria employed [@sousa2022].
+Moreover, factors such as patients` sex [@cereda2016], age and disease duration appear to modulate the risk of cognitive decline [@oh2016; @rana2012].
 Currently, efforts are focused on refining the PDD diagnostic framework to improve its consistency and applicability in both research and clinical contexts [@kulisevsky2024].
@@
 Furthermore, both sets of criteria are compared to PDD diagnosed on Level II.
-The study aims to address following research objectives (RO): (RO1) To estimate the Parkinson’s disease dementia (PDD) rates and evaluate the diagnostic variability and concordance across different PDD criteria.
-(RO2) To identify specific diagnostic components that contribute to variability in PDD classification across the applied criteria.
+The study aims to address the following research objectives (RO): (RO1) To estimate the Parkinson’s disease dementia (PDD) rates and evaluate the diagnostic variability and concordance across different PDD criteria.
+(RO2) To identify specific diagnostic components contributing to PDD classification variability across the applied criteria.
 
@@
 
-This study comprehensively examined the diagnostic variability and validity of Level I criteria for PDD in DBS treatment candidates.
-By systematically comparing 68 diagnostic algorithms based on MMSE, MoCA, sMoCA, and Level II criteria, our findings highlight the significant influence of algorithm design.
+This study examined the diagnostic variability and validity of Level I criteria for PDD, specifically in candidates for DBS treatment.
+By systematically comparing 68 diagnostic algorithms based on MMSE, MoCA, sMoCA, and Level II criteria, our findings highlight the significance of algorithm design.
 
@@
 
-Our results showed a wide range in estimated PDD prevalence across 68 algorithms, ranging from 2.00% to 16.75%.
-Notably, estimates were substantially lower when using FAQ item 9 alone versus the full FAQ scale.
-This discrepancy underscores the diagnostic weight of IADL assessment.
-Our PDD rates were consistently lower compared to previous studied, demonstrating wide variability in reported PDD rate among PD patients.
-For instance, retrospective study reported a PDD rate of 19.7% [@rana2012], while a clinical investigation found even higher rate of around 30% [@aarsland2011].
-A recent complex meta-analysis synthesizing global data placed the pooled rate at 26.30% [@sousa2022].
-**The discrepancy is likely attributable to our DBS specific cohort, which differs in important characteristics, such as younger age of the patients [@rana2012].
+Our results showed a wide range in estimated PDD prevalence across the total of 68 algorithms, ranging from 2.00% to 16.75%.
+Estimates reached lower rates when using solely FAQ item 9 in comparison with the full FAQ scale.
+This discrepancy highlights the diagnostic importance of IADL assessment.
+Our overall PDD rates were consistently lower compared to previous studies regarding PDD among PD patients, demonstrating wide variability based on various criteria used.
+For instance, a retrospective study reported a PDD rate of 19.7% [@rana2012], while other clinical investigation found even higher rates, reaching up to 30% [@aarsland2011].
+A recent complex meta-analysis synthesizing global data placed the rate at 26.30% [@sousa2022].
+**The discrepancy is likely attributable to our DBS-specific cohort, which differs in important characteristics, such as the younger age of the patients [@rana2012].
 / Compared to these estimates, our study reports generally lower PDD rates, likely reflecting differences in sample characteristics, diagnostic criteria, and methodology.
@@
 Our pairwise comparison of diagnostic algorithms showed that agreement was notably stronger among those using the same IADL operationalization (κ range from 0.75, SD = 0.13 to 0.86, SD = 0.09) compared to those using different IADL definitions (κ = 0.34, SD = 0.08).
-This demonstrates that even subtle methodological differences can yield substantially divergent diagnostic outcomes.
+This demonstrates that even slight methodological differences can yield divergent diagnostic outcomes.
 Such findings are critical for clinicians relying on Level I criteria for eligibility decisions, as the choice of algorithm could lead to contradictory classifications of PDD.
 
-## Predictive Validity Comaprison With Level II Criteria
+## Predictive Validity Comparison With Level II Criteria
 
-When using Level II diagnosis as the reference, MoCA-based Level I algorithms, particularly those using Clock Drawing used to assess executive function, demonstrated the highest predictive accuracy.
+When using Level II diagnosis as the reference, MoCA-based Level I algorithms, particularly those using Clock Drawing to assess executive function, demonstrated the highest predictive accuracy.
 This supports recent proposals to modernize PDD diagnostic frameworks [@kulisevsky2024], favoring MoCA-derived components and broader functional assessment tools.
@@
 
-The advantage of MoCA-based algorithm was particularly evident when using the FAQ total score to define functional impairment in comparison with using FAQ item 9 only.
-In these cases, the best-performing algorithms achieved accuracy rates of 94–99%, together with balanced sensitivity and specificity, reinforcing the clinical utility.
+The advantage of MoCA-based algorithm was evident when using the FAQ total score to define functional impairment in comparison with using solely FAQ item 9.
+In these cases, the best-performing algorithms achieved accuracy rates of 94–99%, together with high sensitivity and specificity.
 
@@
 
-These findings have immediate implications for DBS candidacy assessment.
-Given that a diagnosis of PDD typically precludes surgical intervention, inconsistent application of diagnostic criteria can lead to inequitable access to the treatment.
-Our data emphasize the need for harmonized assessment protocols that balance sensitivity with specificity, particularly in high-stakes clinical decision-making.
+Our findings have immediate implications for DBS candidacy assessment.
+Given that a diagnosis of PDD typically excludes surgical intervention, inconsistent application of diagnostic criteria can lead to unbalanced access to the treatment.
 
-Moreover, the relative success of multidomain assessments, such as Level II protocols or regression-based normative approaches, highlights the importance of psychometric rigor in diagnosing PDD.
-These approaches may serve as confirmatory tools in cases where Level I assessments yield ambiguous results.
+Moreover, the relative success of multidomain assessments, such as Level II protocols, highlights the importance of psychometric precision.
+This approach may serve as a confirmatory tool in cases where Level I assessments provide ambiguous results.
 
@@
 This study’s generalizability is limited by its retrospective design and the homogeneity of the DBS candidate cohort, which may not reflect broader PD populations with varying cognitive profiles.
-Additionally, reliance on Czech normative data may restrict international applicability, although it represents real-world clinical standards within the region [@simons2017].
+Moreover, use of the Czech normative data may limit international applicability, although our results represent clinical standards in the region [@simons2017].
 
@@
 
-Future research should focus on prospective validation of updated Level I criteria against longitudinal functional outcomes and neurobiological markers.
-Additionally, the development of adaptive diagnostic algorithms that can integrate performance across cognitive, functional, and psychiatric domains may enhance diagnostic sensitivity while reducing the risk of over-classification.
+Future research should focus on the validation of updated Level I criteria using longitudinal data and neurobiological markers.
+Furthermore, the development of adaptive diagnostic algorithms that can combine performance across cognitive and functional domains may enhance diagnostic sensitivity while reducing the risk of over-classification of PDD.
 
@@
 
-This study systematically investigated the application of multiple Level I diagnostic criteria for PDD within a cohort of patients considered for DBS.
-The findings reveal substantial variability in PDD rates estimates, strongly influenced by the choice of cognitive screening instruments and the operationalization of functional impairment.
-The divergence observed across algorithms demonstrates the sensitivity of diagnostic outcomes to seemingly minor methodological choices.
+This study systematically investigated the application of multiple Level I diagnostic criteria for PDD within a cohort of patients considered for DBS tretment.
+Our results show variability in PDD rate estimates, strongly influenced by the choice of cognitive screening instruments and the operationalization of functional impairment.
+The divergence observed across algorithms demonstrates the sensitivity of diagnostic outcomes to seemingly negligible methodological choices.
 
-The proposed revisions to the diagnostic framework [@kulisevsky2024] criteria offer enhanced sensitivity by leveraging MoCA-based components and broader IADL assessments.
-Conversely, overly conservative criteria, such as reliance on pill questionnaire (i.e.
-FAQ item 9 equivalence) may fail to detect meaningful functional decline and thus under-identify true cases of PDD presence.
+The proposed changes of the current diagnostic criteria [@kulisevsky2024] offer heightened sensitivity by proposing the use of MoCA-based components and broader IADL assessments.
+Overly conservative criteria, such as reliance on pill questionnaire (i.e.
+FAQ item 9 equivalence), may fail to detect meaningful functional decline and thus under-identify true cases of PDD.
 
-Ultimately, this study contributes to the improving landscape of PDD diagnostics by offering empirical evidence for the refinement of Level I criteria and reinforcing the value of psychometric rigor in clinical neuropsychology.
-Future research should validate these operationalizations longitudinally and explore integration with biomarkers to further enhance diagnostic reliability.
+Conclusively, this study contributes to the improvement of PDD diagnostics by offering evidence for the clarification of Level I criteria and emphasizing the value of psychometric precision in clinical neuropsychology.
+Future research should validate the algorithms longitudinally and explore possible integration of biomarkers to further improve diagnostic reliability.
 

### 2025-05-13 _manuscript/manuscript.qmd
JM: intro notes and formatting for the apaquarto to work.

@@
 format:
+ apaquarto-docx: 
+    toc: false
  apaquarto-pdf:
+    # Can be jou (journal), man (manuscript), stu (student), or doc (document)
     documentmode: man
-   a4paper: true
-   pdf-engine: lualatex
-   floatsintext: false
-   keep-tex: false
+    keep-tex: true
 bibliography: references.bib

### 2025-05-13 _manuscript/text.qmd
JM: intro notes and formatting for the apaquarto to work.

---
lang: en
title: "Updated Criteria for the Diagnostic Procedure for Parkinson’s Disease Dementia on Level I and their Validity in Deep Brain Stimulation Cohort"
shorttitle: "Parkinson's Disease Dementia Level I Criteria"
author:
  - name: Martina Mana
    corresponding: false
    orcid: "0009-0007-4665-3946"
    role:
      - Conceptualization
      - Data curation
      - Writing - original draft
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Josef Mana
    corresponding: false
    orcid: "0000-0002-7817-3978" 
    email: "josef.mana@lf1.cuni.cz"
    role:
      - Conceptualization
      - Data curation
      - Investigation
      - Formal analysis
      - Software
      - Methodology
      - Project administration
      - Validation
      - Writing - original draft
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Tereza Uhrova
    corresponding: false
    email: "tereza.uhrova@vfn.cz"
    role:
      - Investigation
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Robert Jech
    corresponding: false
    orcid: "0000-0002-9732-8947"
    email: "jech@cesnet.cz"
    role:
      - Funding acquisition
      - Resources
      - Writing - review & editing
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Ondrej Bezdicek
    corresponding: true
    orcid: "0000-0002-5108-0181"
    email: "ondrej.bezdicek@gmail.com"
    role:
      - Investigation
      - Data curation
      - Funding acquisition
      - Conceptualization
      - Project administration
      - Supervision
      - Writing - original draft
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
abstract: ""
keywords:
  - Parkinson’s disease
  - Parkinson's disease dementia
  - level I criteria
format:
 apaquarto-pdf:
    # Can be jou (journal), man (manuscript), stu (student), or doc (document)
    documentmode: man
    keep-tex: true
bibliography: bibliography.bib
floatsintext: false
numbered-lines: false
suppress-title-page: false
warning: false
echo: false
---

```{r}
#| label: "import"

library(targets)
library(tidyverse)
library(gt)

upstore <- here::here(tar_config_get("store"))
tar_source(here::here("R"))
data <- tar_read(raw_data, store = upstore)
crit <- tar_read(pdd_data, store = upstore)$criteria
vars <- tar_read(variables, store = upstore)
```

# Introduction

Parkinson’s disease (PD) is a neurodegenerative disorder defined by a gradual onset of motor impairment, including rigidity, bradykinesia, postural instability and resting tremor. Additionally, patients frequently suffer from a range of non-motor impairments [@postuma2015], particularly cognitive decline. This state might result in Parkinson’s disease dementia (PDD) in a subset of patients [@meireles2012].

Despite its clinical relevance, the diagnosis of PDD remains complex. The original criteria formalized in 2007 by the International Parkinson and Movement Disorder Society (MDS) [@dubois2007] were heavily influenced by frameworks established primarily for Alzheimer’s disease (AD). **Although these criteria provided a valuable base, they lacked the specificity to detect the specific cognitive aspects of PDD [@emre2007; @yamashita2023].**^[**JM: Are we sure the references provide evidence supporting our statement that MDS criteria lacked specificity for cognitive aspects of PDD? Emre et al. (2007) is one of the series of PDD-criteria-creating articles together with Dubois et al. (2007) and it does not say the criteria are not specific. Yamashita et al. (2023) is not concerned with criteria at all, instead they use any definition of PDD that was present in their source studies to evaluate predictive value of biomarkers for PDD. I do not think our argument here is valid based on the provided evidence.**]

In order to increase diagnostic accuracy, the MDS also developed a levelled system for PDD detection. Level I criteria are designed primarily for clinical use, consisting of brief cognitive assessments **and relying on clinical judgement**^[**JM: Is this a statement directly from the source? If not, I would drop the clinical judgement part as all diagnoses are likely to rely on clinical judgement anyway.**]. In contrast, Level II criteria use comprehensive neuropsychological testing across cognitive domains to provide greater diagnostic specificity [@emre2007].

One specific feature of the original Level I criteria was the presence of an algorithm allowing flexibility in test selection. The criteria included (Mini-Mental State Examination) MMSE pentagons for evaluation of visuospatial ability, three-word recall for memory assessment,  months reversed or seven backwards for attention assessment and lexical fluency or clock drawing for executive function assessment [@dubois2007].^[**JM: I would introduce the algorithm in a little bit of more detail. E.g., "The original criteria included a straightforward algorithm for diagnosis of probable PDD consisting of eight items that needed to be fullfilled simultaneously. ... **]

According to a recent meta-nalysis, the estimates of PDD rate within PD vary widely, ranging from 14% up to 55%, depending on methodological criteria employed [@sousa2022]. Moreover, factors such as patients` sex [@cereda2016], age and disease duration appear to modulate the risk of cognitive decline [@oh2016; @rana2012]. Currently, efforts are focused on refining the PDD diagnostic framework to improve its consistency and applicability in both research and clinical contexts [@kulisevsky2024].^[**JM: This shall be even higher, the second paragraphs is a good place for this one according to my humble correct opinion.**]

**The present study evaluates the diagnostic concordance**^[**JM: There is not much reasoning substantiating these goals in the introduction. We would be better off restructuring the introduction with a more straightforward logic that goes: "This is PDD" -> "PDD is important" -> "This is how PDD is measured" -> "This is why the PDD measurement is imperfect" -> "This is what can help us to make sense of the PDD measurement" -> "Ergo, that's why we do what we do in this study."**] between the original Level I criteria for PDD, as established by the MDS Task Force [@emre2007; @dubois2007] and criteria inspired by the recent call for change [@kulisevsky2024]. Furthermore, both sets of criteria are compared to PDD diagnosed on Level II. The study aims to address the following research objectives (RO): (RO1) To estimate the Parkinson’s disease dementia (PDD) rate and evaluate the diagnostic variability and concordance across different PDD criteria. (RO2) To identify specific diagnostic components contributing to PDD classification variability across criteria.

# Methods

## Participants

```{r}
#| label: "dates"

continue <- all(!is.na(data$assdate))
stopifnot("Some assessment dates are missing" = continue)

dates <-
  sapply(
    c("min", "max"),
    function(f) {
      date <- do.call(f, list(data$assdate))
      paste(month(date, label = T, abbr = F), year(date))
    }
  )
```

This study retrospectively analyzed clinical data from a cohort of patients with PD at the General University Hospital in Prague. All patients were diagnosed with idiopathic PD by a movement disorder specialist according to the MDS Clinical Diagnostic Criteria for PD [@postuma2015]. Clinical records spanning `r dates["min"]` to `r dates["max"]` were examined. All participants underwent neuropsychological evaluation conducted by a trained clinical psychologist (OB) as part of standard preoperative cognitive assessments for DBS eligibility at the General University Hospital in Prague.

Ethical approval for the study protocol was obtained from the Ethics Committee of the General University Hospital in Prague. Informed consent was secured from all patients prior to their neuropsychological assessments, in adherence to ethical research guidelines.

## Neuropsychological Assessment

Cognitive performance was evaluated at both Level I (abbreviated assessment) and Level II (comprehensive assessment) according to the standard MDS neuropsychology battery for Parkinson's Disease Mild Cognitive Impairment (PD-MCI) [@litvan2012, @bezdicek2017]. Level I was assessed using the Mini-Mental State Examination (MMSE) [@stepankova2014; @folstein1975] and the Montreal Cognitive Assessment (MoCA) [@kopecek2016; @nasreddine2005]. The neuropsychological assessment at Level II covered five cognitive domains, each evaluated through specific tests as follows: attention and working memory assessed using Trail Making Test Part A (TMT-A) [@bezdicek2012; @reitan2004], and WAIS Digit Span Backward (WAIS DSB) [@wechsler1997]; executive function evaluated via Categorical Verbal Fluency (CF) [@nikolai2015], and subtest from the Prague Stroop Test – Colors (PST-C) [@bezdicek2021]; language measured with the WAIS Similarities subtest [@wechsler1997], and the Boston Naming Test (BNT-60) [@zemanova2016; @kaplan1983]; memory examined using the Rey Auditory Verbal Learning Test (RAVLT) [@frydrychova2018; @bezdicek2013; @rey1964] delayed recall, and the Brief Visual Memory Test–Revised (BVMTR) [@havlik2020; @benedict1997] delayed recall, or WAIS Family Pictures subtest [@wechsler1997] delayed recall; visuospatial function assessed through the Judgment of Line Orientation Test (JoLO) [@benton1983], and Clock Drawing Test (CLOX) [@royall1998].

# References

### 2025-05-13 _manuscript/text.qmd
Notes to discussion. Finishing touches to re-formatting. apaquarto-docx appears to work now.

@@
     keep-tex: true
+ apaquarto-docx: 
+    toc: false
 bibliography: bibliography.bib
@@
 The original criteria formalized in 2007 by the International Parkinson and Movement Disorder Society (MDS) [@dubois2007] were heavily influenced by frameworks established primarily for Alzheimer’s disease (AD).
-**Although these criteria provided a valuable base, they lacked the specificity to detect the specific cognitive aspects of PDD [@emre2007; @yamashita2023].**^[**JM: Are we sure the references provide evidence supporting our statement that MDS criteria lacked specificity for cognitive aspects of PDD?
+**Although these criteria provided a valuable base, they lacked the specificity to detect the specific cognitive aspects of PDD [@emre2007; @yamashita2023].**[^1]
+
+[^1]: **JM: Are we sure the references provide evidence supporting our statement that MDS criteria lacked specificity for cognitive aspects of PDD?
 Emre et al.
@@
 (2023) is not concerned with criteria at all, instead they use any definition of PDD that was present in their source studies to evaluate predictive value of biomarkers for PDD.
-I do not think our argument here is valid based on the provided evidence.**]
+I do not think our argument here is valid based on the provided evidence.**
 
 In order to increase diagnostic accuracy, the MDS also developed a levelled system for PDD detection.
-Level I criteria are designed primarily for clinical use, consisting of brief cognitive assessments **and relying on clinical judgement**^[**JM: Is this a statement directly from the source?
-If not, I would drop the clinical judgement part as all diagnoses are likely to rely on clinical judgement anyway.**].
+Level I criteria are designed primarily for clinical use, consisting of brief cognitive assessments **and relying on clinical judgement**[^2].
 In contrast, Level II criteria use comprehensive neuropsychological testing across cognitive domains to provide greater diagnostic specificity [@emre2007].
 
+[^2]: **JM: Is this a statement directly from the source? If not, I would drop the clinical judgement part as all diagnoses are likely to rely on clinical judgement anyway.**
+
 One specific feature of the original Level I criteria was the presence of an algorithm allowing flexibility in test selection.
-The criteria included (Mini-Mental State Examination) MMSE pentagons for evaluation of visuospatial ability, three-word recall for memory assessment,  months reversed or seven backwards for attention assessment and lexical fluency or clock drawing for executive function assessment [@dubois2007].^[**JM: I would introduce the algorithm in a little bit of more detail.
+The criteria included (Mini-Mental State Examination) MMSE pentagons for evaluation of visuospatial ability, three-word recall for memory assessment, months reversed or seven backwards for attention assessment and lexical fluency or clock drawing for executive function assessment [@dubois2007].[^3]
+
+[^3]: **JM: I would introduce the algorithm in a little bit of more detail.
 E.g., "The original criteria included a straightforward algorithm for diagnosis of probable PDD consisting of eight items that needed to be fullfilled simultaneously.
-...
-**]
+...**
 
 According to a recent meta-nalysis, the estimates of PDD rate within PD vary widely, ranging from 14% up to 55%, depending on methodological criteria employed [@sousa2022].
-Moreover, factors such as patients` sex [@cereda2016], age and disease duration appear to modulate the risk of cognitive decline [@oh2016; @rana2012].
-Currently, efforts are focused on refining the PDD diagnostic framework to improve its consistency and applicability in both research and clinical contexts [@kulisevsky2024].^[**JM: This shall be even higher, the second paragraphs is a good place for this one according to my humble correct opinion.**]
+Moreover, factors such as patients\` sex [@cereda2016], age and disease duration appear to modulate the risk of cognitive decline [@oh2016; @rana2012].
+Currently, efforts are focused on refining the PDD diagnostic framework to improve its consistency and applicability in both research and clinical contexts [@kulisevsky2024].[^4]
 
-**The present study evaluates the diagnostic concordance**^[**JM: There is not much reasoning substantiating these goals in the introduction.
-We would be better off restructuring the introduction with a more straightforward logic that goes: "This is PDD" -> "PDD is important" -> "This is how PDD is measured" -> "This is why the PDD measurement is imperfect" -> "This is what can help us to make sense of the PDD measurement" -> "Ergo, that's why we do what we do in this study."**] between the original Level I criteria for PDD, as established by the MDS Task Force [@emre2007; @dubois2007] and criteria inspired by the recent call for change [@kulisevsky2024].
+[^4]: **JM: This shall be even higher, the second paragraphs is a good place for this one according to my humble correct opinion.**
+
+**The present study evaluates the diagnostic concordance**[^5] between the original Level I criteria for PDD, as established by the MDS Task Force [@emre2007; @dubois2007] and criteria inspired by the recent call for change [@kulisevsky2024].
 Furthermore, both sets of criteria are compared to PDD diagnosed on Level II.
@@
 (RO2) To identify specific diagnostic components contributing to PDD classification variability across criteria.
+
+[^5]: **JM: There is not much reasoning substantiating these goals in the introduction.
+We would be better off restructuring the introduction with a more straightforward logic that goes: "This is PDD" -\> "PDD is important" -\> "This is how PDD is measured" -\> "This is why the PDD measurement is imperfect" -\> "This is what can help us to make sense of the PDD measurement" -\> "Ergo, that's why we do what we do in this study."**
 
@@
 ```{r}
-#| label: "dates"
+#| label: dates
 
@@
 
+To assess functional impairment, the Functional Activities Questionnaire (FAQ) [@bezdicek2016; @pfeffer1982] was administered.
+Additionally, neuropsychiatric status was evaluated using the Beck Depression Inventory-II (BDI-II) [@ciharova2020; @beck1996] and State-Trait Anxiety Inventory (STAI) [@spielberg1983; @mullner1980].
+Psychotic symptoms were assessed through structured psychiatric interviews conducted by a trained psychiatrist.
+
+## Diagnostic algorithms for probable Parkinson's Disease Dementia
+
+In this study, we applied three distinct sets of diagnostic algorithms for probable PDD at Level I.
+The first set was based on the original framework [@dubois2007], which utilized the Mini-Mental State Examination (MMSE) as a global cognitive screening tool, supplemented by assessments of attention, executive function, visuospatial abilities, and memory.
+The second set of algorithms was based on the recent call for change of dementia diagnostic guidelines [@kulisevsky2024], which advocates for more sensitive cognitive domain assessments in the context of PD.
+This updated approach incorporated specific items from the Montreal Cognitive Assessment (MoCA).
+The third approach applied the Czech version of the shortened Montreal Cognitive Assessment (sMoCA) [@bezdicek2020], a time-efficient modification designed to measure global cognitive performance using a reduced testing protocol that omits items providing redundant information.
+Lastly, the fourth approach followed the Level II protocol for diagnosis of PDD and Mild Cognitive Impairment in PD (PD-MCI) [@dubois2007; @litvan2012].
+The Level II methodology, including the use of a regression-based normative scoring approach, has been detailed in a prior study [@bezdicek2017].
+In this study, the thresholds for cognitive impairment at Level II were set at $z \leq -1.5$.
+All non-cognitive criteria of probable PDD (i.e., diagnosis of PD that developed before dementia and absence of Major Depression, delirium or other abnormalities that obscure diagnosis) held true for all patients in the sample according to the psychiatric and neurological examinations.
+
+```{r}
+#| label: crit-check
+
+critnum <-
+  sapply(
+    c("mmse", "moca", "smoca", "lvlII"),
+    function(i)
+      subset(crit, group == i) |>
+      nrow()
+  )
+
+continue <- !any(critnum == 0)
+stopifnot("Some criteria are missing" = continue)
+```
+
+For each of these diagnostic approaches, we applied two operationalizations of deficits in Instrumental Activities of Daily Living (IADL).
+First, we utilized FAQ item 9, which approximates the pill questionnaire from the original criteria [@dubois2007] employing a cut-off score of 2 points or higher.
+Second, we applied the entire Functional Activities Questionnaire (FAQ) as suggested in the call for change [@kulisevsky2024], employing a cut-off score of 7 points or higher based on Czech normative data [@bezdicek2011].
+These methodologies resulted in a total of `r sum(critnum)` algorithms, which were distributed across different diagnostic criteria: `r critnum["mmse"]` MMSE-based, `r critnum["moca"]` MoCA-based, `r critnum["smoca"]` sMoCA-based, and `r critnum["lvlII"]` based on the Level II battery (see @tbl-crits and Appendix @tbl-algos for the exact specification of each algorithm).
+
+```{r}
+#| label: tbl-crits
+#| tbl-cap: "Summary of probable PDD operationalizations compared in the study."
+
+table_criteria(vars)
+```
+
+\[Insert Table 1 here\]
+
+## Statistical Analyses
+
+Following the framework proposed by @lundberg2021, in this study we explicitly connect our research objectives and their corresponding theoretical (i.e., targets of inference) and empirical (i.e., data-driven) estimands to statistical estimates.
+The theoretical estimand refers to a unit-specific quantity defined over a target population and represents the ideal quantity that would address the research question under optimal conditions, such as access to complete population data or perfect experimental control.
+In contrast, the empirical estimand corresponds to the quantity that is actually computable using the available dataset, given real world constraints.
+The full description of the study's estimands and their relation to our research objectives is presented in the Appendix (see @tbl-estimands).
+
+To address study objectives, we started by repeatedly assigning each patient the diagnosis of probable PDD based on each PDD algorithm listed in @tbl-crits (see also @tbl-algos) resulting in a `r sum(data$incl)` (patients) $\times$ `r sum(critnum)` (algorithms) matrix where each cell indicates whether a patient (row) meets criteria for probable PDD according to an algorithm (column)[^6].
+PDD rate estimates were computed as $\frac{N_{PDD}}{N_{total}}$ separately for each algortihm.
+The predictive value of age and sex was then evaluated by fitting a set of logistic regressions, one for each algorithm for probable PDD, whereby the probable PDD was predicted by age, sex and their interaction.
+
+[^6]: **JM: This could and should be shared most likely, as long, as we anonymize properly. Let's ask Oto Mestek and the NPO team how and if is it possible.**
+
+Next, a set of two class cross-tabulations with associated statistics was computed for each pair of algorithms via the `confusionMatrix()` function from the R package *caret* [@kuhn2008].
+For each pair of algorithms, the analysis was repeated twice such that each variable of the pair served once as the reference and once as the predictor.
+Following measures were used to evaluate pairwise concordance between different algorithms for probable PDD: 1) Cohen's $\kappa$ with its 95% confidence interval (CI) computed via the `cohen.kappa()` function from the R package *psych* [@revelle2024]; 2) Accuracy (i.e., the proportion of correct predictions, both true positives and true negatives, among the total number of cases) with its 95% CI; 3) Sensitivity/Recall (i.e., the proportion of true positives); and 4) Specificity (i.e., the proportion of true negatives).
+
+Finally, the No Information Rate (NIR) was calculated for each pair of algorithms.
+NIR is the accuracy that could be obtained by always predicting the majority class and in our case it is equivalent to the complement of the PDD rate estimate according to the reference algorithm.
+Accuracy of prediction was compared to the NIR via a one-sided Exact Binomial Test as implemented by the `binom.test()` R stats function.
+Reference/predictor pairs associated with p \< .05 were considered to show significantly better accuracy than NIR.
+In other words, for reference/predictor pairs associated with p \< .05, we conclude that knowing the probable PDD status according to the predictor algorithm helps to estimate the probable PDD status according to the reference algorithm and the two algorithms thus show substantial concordance.
+
+Data wrangling and visualizations were done in the *tidyverse* package [@wickham2019] and tables were formatted in the *gt* package [@iannone2024].
+All analyses were conducted within the R (version `r with(version, paste(major, minor, sep="."))`) software environment for statistical computing [@rsoft].
+The software code supporting this article is available at <https://github.com/josefmana/DemCr1t.git>.[^7]
+
+[^7]: **JM: Do not forget to make it public before submitted!**
+
+# Results
+
+```{r}
+#| label: results
+
+desc <- tar_read(sample_description, store = upstore)
+prev <- tar_read(prevalence_summaries, store = upstore)
+pred <- tar_read(demographic_predictors, store = upstore)
+conc <- tar_read(concordance_statistics, store = upstore)
+perc <- prev$table$perc
+
+# In-text demographic data
+sex <- subset(desc$table, variable == "Sex")$N
+age <- paste0(subset(desc$table, variable == "Age")[ , c("M", "SD")] |> paste(collapse = " (SD = "),") years of age")
+edu <- paste0(subset(desc$table, variable == "Education")[ , c("M", "SD")] |> paste(collapse = " (SD = ")," ) years of education")
+dur <- paste0(subset(desc$table, variable == "PD duration")[ , c("M", "SD")] |> paste(collapse = " (SD = "),") years of disease duration")
+updrs_off <- paste0(subset(desc$table, variable == "UPDRS III off state")[ , c("M", "SD")] |> paste(collapse = " (SD = "),") Unified Parkinson Disease Rating Scale (UPDRS), part III in medication OFF state")
+updrs_on <- paste0(subset(desc$table, variable == "UPDRS III on state")[ , c("M", "SD")] |> paste(collapse = " (SD = "),") UPDRS III in medication ON state")
+
+# In-text rate estimates
+rates_overall <- sapply(c("M", "SD", "Md", "minmax"), function(f) do_summary(perc, 2, f))
+rates_FAQ9 <- sapply(c("M", "SD", "Md", "minmax"), function(f) do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, f))
+rates_FAQtot <- sapply(c("M", "SD", "Md", "minmax"), function(f) do_summary(perc[prev$table$IADL == "FAQ > 7"], 2, f))
+
+# Minimum p-value of logistic regressions
+min_p <- subset(pred$values, quantity == "p-value")$estimate |> min() |> do_summary(3, "p")
+```
+
+## Sample Description
+
+A total of `r sum(data$incl)` patients were included.
+The sample included `r sex` men, with an average of `r age`, `r edu`, `r dur`, `r updrs_off` and `r updrs_on`.
+Cognitive characteristics of the sample are summarized in @tbl-desc.
+
+```{r}
+#| label: tbl-desc
+#| tbl-cap: "Cognitive characteristics of the sample."
+
+desc$gtable
+```
+
+\[Insert Table 2 here\]
+
+## PDD Rate Estimates
+
+Algorithm-wise rate of PDD estimates are presented in @tbl-rates.
+On average, estimated PDD rate was `r rates_overall["M"]`% (SD = `r rates_overall["SD"]`, Md = `r rates_overall["Md"]`, range `r rates_overall["minmax"]`).
+Notably, the estimates were substantially lower when FAQ item 9 was used as a criterion of IADL deficit (M = `r rates_FAQ9["M"]`% SD = `r rates_FAQ9["SD"]`, Md = `r rates_FAQ9["Md"]`, range `r rates_FAQ9["minmax"]`) compared to using total FAQ score criterion (M = `r rates_FAQtot["M"]`% SD = `r rates_FAQtot["SD"]`, Md = `r rates_FAQtot["Md"]`, range `r rates_FAQtot["minmax"]`) as demonstrated also in @fig-rates.
+Neither age nor sex or their interaction (*p*s $\geq$ `r min_p`) reliably predicted probable PDD classification across algorithms (see @fig-data and @fig-pars).
+
+\[Insert Figure 1 here\]
+
+```{r}
+#| label: fig-rates
+#| fig-cap: Summary of the estimates of probable PDD rate.
+#| apa-note: Vertical lines represent estimates arrived at by using sMoCA (dotted) or Level II (dashed) with FAQ item 9 (orange) or FAQ total score (blue) as criteria for probable PDD.
+
+prev$plot
+```
+
+## Concordance between Algorithms
+
+```{r}
+#| label: kappas
+
+algos <- list(
+  faq_tot = subset(prev$table, IADL == "FAQ > 7")$type,
+  faq_9 = subset(prev$table, IADL == "FAQ 9 > 1")$type
+)
+
+# Check for overlap:
+continue <- !(any(algos$faq_tot %in% algos$faq_9) || any(algos$faq_9 %in% algos$faq_tot))
+stopifnot("Some algorithm(s) use both IADL operationalisations!" = continue)
+
+# Prepare combinations within IADL operationalisation:
+kappas <-
+  lapply(
+    set_names(names(algos)),
+    function(o)
+      combn(algos[[o]], 2) |>
+      t() |>
+      as_tibble() |>
+      `colnames<-`(c("predictor", "reference")) |>
+      left_join(conc$table, by = c("predictor", "reference")) |>
+      select(Kappa_raw) |>
+      pull()
+  )
+
+# Add cross-operationalisation cases
+kappas$cross <- subset(conc$table, predictor %in% algos$faq_9 & reference %in% algos$faq_tot)$Kappa_raw
+
+# Extract Kappa summaries:
+kappa_sums <-
+  sapply(
+    names(kappas),
+    function(i)
+      paste0(do_summary(kappas[[i]], 2, "M"),", SD = ", do_summary(kappas[[i]], 2, "SD"))
+  )
+```
+
+Results of the analyses of prediction Accuracy, Cohen's $\kappa$, Sensitivity and Specificity are presented in @fig-acc, @fig-kappa, @fig-sens and @fig-spec respectively.
+**Numerical results are available at ...** [^8].
+Generally, algorithms that employed the same operationalization of IADL deficit showed substantial pairwise concordance, however, algorithms that operationalized IADL deficit differently did not.
+Whereas among algorithms with identical IADL deficit operationalization, the agreement judged by Cohen's $\kappa$ was moderately high (operationalization by FAQ total score: $\kappa$ = `r kappa_sums["faq_tot"]`; operationalization by FAQ item 9: $\kappa$ = `r kappa_sums["faq_9"]`), among algorithms that differ in IADL deficit operationalization it was low ($\kappa$ = `r kappa_sums["cross"]`).
+
+[^8]: **JM: Should provide an excel sheet or html table the readers could go through by themselves.**
+
+\[Insert Figure 2 here\]
+
+```{r}
+#| label: fig-acc
+#| fig-cap: Prediction accuracy matrix.
+#| apa-note: The matrix depicts classification accuracy of algorithms for PDD depicted on x-axis in predicting outcomes based on algorithms on the y-axis.
+Algorithms printed in blue defined IADL deficit by FAQ total score, algorithms printed in black defined IADL deficit by FAQ item 9 response.
+Cases with asterisk indicate predictive accuracy statistically significantly higher than the No Information Rate.
+
+conc$plots$Accuracy
+```
+
+## Prediction of Level II Criteria
+
+```{r}
+#| label: lvlII-stats
+
+lvlII_rates <- c(
+  faq_tot = subset(prev$table, type == "Lvl.II (1)")$perc |> do_summary(2) |> paste0("%"),
+  faq_9 = subset(prev$table, type == "Lvl.II (2)")$perc |> do_summary(2) |> paste0("%")
+)
+
+smoca_stats <- c(
+  spec = subset(conc$table, predictor == "sMoCA (1)" & reference == "Lvl.II (1)")$Specificity |> do_summary(2),
+  sens = subset(conc$table, predictor == "sMoCA (1)" & reference == "Lvl.II (1)")$Sensitivity |> do_summary(2)
+)
+```
+
+For easier interpretability of our results, we next examine cases where Level II algorithms served as a reference and Level I algorithms as a predictor.
+@tbl-approx shows five Level I algorithms with the highest and five with the lowest accuracy in predicting Level II classification of probable PDD.
+
+When IADL deficit was defined by total FAQ score, the Level II estimate of PDD rate was `r lvlII_rates["faq_tot"]`.
+All five Level I algorithms that approximated the Level II classification most accurately were MoCA-based and defined Executive Function deficit by Clock drawing rather than Lexical fluency test.
+On the other hand, two out of the five Level I algorithms with the lowest accuracy were MMSE-based, whereas the remaining three were MoCA-based and defined Executive Function deficit by Lexical fluency test.
+
+When IADL deficit was defined by FAQ item 9 score, the Level II estimate of PDD rate was `r lvlII_rates["faq_9"]`.
+Overall, the difference between the most accurate and the least accurate Level I algorithms was lower than in the case of IADL deficit being defined by FAQ total score (see @tbl-approx).
+The five most accurate algorithms were all MoCA-based, defined Executive Function deficit by Clock drawing (with threshold \< 2) and in majority of cases defined Language deficit by Animal naming.
+Two out of the five Level I algorithms with the lowest accuracy were MMSE-based, whereas the remaining three were MoCA-based and defined Executive Function deficit by Clock drawing (with threshold \< 3) and Language deficit by Abstraction.
+
+Finally, if the predictors are sorted by their balanced accuracy (i.e., average of sensitivity and specificity) instead of raw accuracy, the results are similar with the exception that for prediction of Level II with total FAQ score algorithm for probable PDD, the highest balanced accuracy was achieved by the sMoCA algorithm with sensitivity `r smoca_stats["sens"]` and specificity `r smoca_stats["spec"]` (see @tbl-approx-bal).
+
+```{r}
+#| label: tbl-approx
+#| tbl-cap: "Level I algorithms for probable PDD as predictors of Level II classification as the reference."
+
+table_levelII_approximations(conc$table, algos, "Accuracy_raw")
+```
+
+# Discussion
+
+This study examined the diagnostic variability and validity of Level I criteria for PDD, **specifically in candidates for DBS treatment**[^9].
+By systematically comparing 68 diagnostic algorithms based on MMSE, MoCA, sMoCA, and Level II criteria, our findings highlight the significance of algorithm design.
+
+[^9]: **JM: All-right, I do not think it needs to be here either.**
+
+## Variability in PDD Rate Estimates
+
+Our results showed a wide range in estimated PDD rate across algorithms, ranging from 2.00% to 16.75%.
+Estimates reached lower rates when using solely FAQ item 9 (as an approximation of the pill questionnaire suggested by @dubois2007) in comparison with the full FAQ scale.
+This discrepancy highlights the diagnostic importance of IADL assessment.
+Our overall PDD rates were consistently lower compared to previous studies regarding PDD among PD patients, **demonstrating wide variability based on various criteria used**[^10].
+For instance, a retrospective study reported a PDD rate of 19.7% [@rana2012], while other clinical investigation found even higher rates, reaching up to 30% [@aarsland2005][^11].
+A recent complex meta-analysis synthesizing global data placed the expected PDD rate in PD at 26.30% [@sousa2022].
+**The discrepancy is likely attributable to our DBS-specific cohort, which differs in important characteristics, such as the younger age of the patients [@rana2012].
+/ Compared to these estimates, our study reports generally lower PDD rates, likely reflecting differences in sample characteristics, diagnostic criteria, and methodology.** [^12] [^13]
+
+[^10]: **JM: What is demonstrating this variability? Our findings or previous ones?**
+
+[^11]: **JM: There was aarsland2011 originally but I did not find such a study. Is this reference correct?**
+
+[^12]: **Because the quantity of interest is a rate and could thus be though of as a sum of binomially distributed PDD occurences divided by the total number of patients, its variance will likely systematically vary with its mean.
+Specifically, as the rate goes from extremes to 0.5, the variance increases.
+Consequently, if our estimate of the rate was lower than the true population rate, e.g., because our sample includes younger patients compared to the general PD population, our estimate of variance would also lower than the true variance of PDD rate in the general PD population.
+Nonetheless, the between-algorithm variability may not be affected by this phenomenon as unlike variability of PDD rate, we do not have reason to assume it comes about by summing independent binomial events**
+
+[^13]: \*\*JM: I prefer the second one with added sentence like: "Specifically, our cohort included candidates for DBS who are generally younger than the general PD population and age was repeatedly shown to be a strong predictor of PDD (some citations)." We should also incorporate here a statement about missing evidence of age being predictor in our sample.\*
+
+## Concordance Between Diagnostic Algorithms
+
+Our pairwise comparisons of diagnostic algorithms showed that agreement was notably stronger among those using the same IADL operationalization **(**$\kappa$ range from 0.75, SD = 0.13 to 0.86, SD = 0.09)[^14] compared to those using different IADL definitions **(**$\kappa$ = 0.34, SD = 0.08)[^15].
+**Moreover, the agreement was slightly higher between algorithms that defined IADL deficit by FAQ item 9 compared to algorithms that defined it using the full FAQ scale.
+One possible explanation of this difference follows from the observation that algorithms using the full scale definition yielded higher PDD rate estimates.
+Because there was higher probability of being diagnosed with IADL deficit based on the full FAQ scale, there was also a bigger room for disagreement in the Cognitive Impairment status when different indexes weer used (e.g., by defining executive deficit via clock drawing vs.
+lexical fluency).**
+
+[^14]: **JM: There is no need to repeat these numbers in this section.**
+
+[^15]: **JM: Same as above.**
+
+**Overall, algorithms using the same definition of IADL deficit showed degree of concordance equivalent to moderate (when using FAQ total score to define IADL deficit) or strong (when using FAQ item 9 to define IADL deficit) level of agreement between raters in an inter-rater reliability analysis [@mchugh2012].
+On the other hand, the concordance between algorithms using different IADL deficit definitions was equivalent to minimal agreement.** This demonstrates that even slight methodological differences can yield divergent diagnostic outcomes.
+Such findings are critical for clinicians relying on Level I criteria for eligibility decisions, as the choice of algorithm could lead to contradictory classifications of PDD.
+
+## Predictive Validity Comparison With Level II Criteria
+
+When using Level II diagnosis as the reference, MoCA-based Level I algorithms, particularly those using Clock Drawing to assess executive function, demonstrated the highest predictive accuracy.
+This supports recent proposals to modernize PDD diagnostic frameworks [@kulisevsky2024], favoring MoCA-derived components and broader functional assessment tools.
+In contrast, MMSE-based algorithms consistently underperformed, suggesting limited sensitivity in capturing the nuanced cognitive deficits characteristic of PDD.
+
+The advantage of MoCA-based algorithm was evident when using the FAQ total score to define functional impairment in comparison with using solely FAQ item 9.
+In these cases, the best-performing algorithms achieved accuracy rates of 94–99%, together with high sensitivity and specificity.
+**JM: Some sMoCA praise should be added here ...**
+
+## Clinical Implications^[**JM: Is this section needed? It does not seem to say much to me. I would't be afraid to drop it.**]
+
+Our findings have immediate implications for DBS candidacy assessment.
+Given that a diagnosis of PDD typically excludes surgical intervention, inconsistent application of diagnostic criteria can lead to unbalanced access to the treatment.
+
+Moreover, the relative success of multidomain assessments, such as Level II protocols, highlights the importance of psychometric precision.
+This approach may serve as a confirmatory tool in cases where Level I assessments provide ambiguous results.
+
+## Limitations and Future Directions
+
+This study’s generalizability is limited by its retrospective design and the homogeneity of the patient cohort, which may not reflect broader PD populations with varying cognitive profiles.
+**Specifically, younger age of the sample and possible lack of patients with high risk phenotypes for PDD compared to general population of people with PD might have been responsible for comparatively lower PDD rate estimates in our study.
+Nonetheless, the analysis of the concordance between different algorithms might not be affected by the sample to the same degree allowing for a broader generalization.
+Future studies could replicate our results on different cohorts of PD patients.
+To make this process easier, the code used to generate our results is publicly available and easily applicable to other data sets with a similar data structure.**
+
+**JM: The fact that no biomarkers were used and the IADL was at best estimated by FAQ which is somewhat archaic (comp.
+to the e-IADL thing which could be also cited and is more in line with Kulisevsky) should be mentioned.
+Also, this is a great space for an advert for NOT/DOT because the self-report of IADLs can be viewed as a limitation worth of addressing, especially since our results imply possible central role for IADL deficit definition in further work on PDD definition to achieve good reliability.**
+
+Future research should focus on the validation of updated Level I criteria using longitudinal data and neurobiological markers.
+Furthermore, the development of adaptive diagnostic algorithms that can combine performance across cognitive and functional domains may enhance diagnostic sensitivity while reducing the risk of **over-classification of PDD**^[**JM: What's an over-classification of PDD?**].
+
+# Conclusions
+
+This study systematically investigated the application of multiple Level I diagnostic criteria for PDD within a cohort of patients considered for DBS treatment.
+Our results show variability in PDD rate estimates, strongly influenced by the choice of cognitive screening instruments and the operationalization of functional impairment.
+The divergence observed across algorithms demonstrates the sensitivity of diagnostic outcomes to seemingly negligible methodological choices.^[**JM: This sounds as a great opening to the discussion, rather than conclusions :-)**]
+
+The **proposed changes**^[**JM: They did not propose anything though and from what I gather, the authors stress this a lot - that their article in not a proposal of new criteria, only a position paper.**] of the current diagnostic criteria [@kulisevsky2024] offer heightened sensitivity by proposing the use of MoCA-based components and broader IADL assessments.
+Conservative criteria, such as reliance on pill questionnaire (i.e.
+FAQ item 9 equivalence), may fail to detect meaningful functional decline and thus under-identify true cases of PDD.
+**On the other hand, such criteria show slightly higher agreement between their different forms.**
+
+Conclusively, this study contributes to the improvement of PDD diagnostics by offering **evidence for the clarification of Level I criteria**^[**JM: What is meant by clarification of Level I criteria?
+Also, how does our study offer evidence for such thing?**] and **emphasizing the value of psychometric precision in clinical neuropsychology**^[**JM: This is independent of our study.**].
+Future research should validate the algorithms longitudinally and explore possible integration of biomarkers to further improve diagnostic reliability.
+
 # References
+
+# Appendix
+
+## Derivation of the Algorithms Set
+
+Both, the original PDD criteria [@dubois2007] and the call for their change [@kulisevsky2024] allow for several distinct combinations of items to be used to define cognitive impairment.
+Consequently, in this study we derived all algorithms for probable PDD on Level I that are in line with published criteria.
+This procedure parallel the diagnostic algorithm outlined in Table 2 of @dubois2007.
+Specifically, in this study, we varied the exact specification of items 3-5 of this table (i.e., the measure of global cognitive impairment, the measure of the impact on IADLs and the measures of impaired cognition).
+
+For each set of criteria (MMSE-based, MoCA-based, sMoCA-based and Level II), we first specified the items and then the thresholds for each item used to define probable PDD.
+If more than one option was present in either the choice of the item or the choice of the threshold, we created an algorithm for each choice in turn.
+The final set of algorithms was arrived at by computing the Cartesian product of all possibilities provided by varying items and thresholds.
+All combinations are presented in @tbl-algos.
+
+For MMSE-based algorithms, the following sets of items served as the basis:
+
+$Global = \{MMSE < 26\}$
+
+$Attention = \{{Sevens\ backwards < 4}\}$
+
+$Executive = \{{Clock\ drawing} < 2, {Lexical\ fluency\ (S)} < 10\}$
+
+$Construction = \{Pentagons < 1\}$
+
+$Memory = \{{Three{\text-}words\ recall} < 3\}$
+
+$IADL = \{FAQ > 7, {FAQ\ (it. 9)} > 1\}$
+
+The ensuing Cartesian product $Global \times Attention \times Executive \times Construction \times Memory \times IADL$ results in $1 \times 1 \times 2 \times 1 \times 1 \times 2 = 4$ MMSE-based algorithms for probable PDD.
+
+For MoCA-based algorithms, the following sets of items served as the basis:
+
+$Global = \{MoCA < 26\}$
+
+$Attention = \{{Sevens\ backwards < 3}\}$
+
+$Executive = \{{Clock\ drawing} < 2, {Clock\ drawing} < 3, {Lexical\ fluency\ (K)} < 11\}$
+
+$Construction = \{Cube\ drawing < 1\}$
+
+$Memory = \{Five{\text-}words\ recall < 1, Five{\text-}words\ recall < 2, Five{\text-}words\ recall < 3, Five{\text-}words\ recall < 4, Five{\text-}words\ recall < 5\}$
+
+$Language = \{Abstraction < 2, Animal\ naming < 3\}$
+
+$IADL = \{FAQ > 7, {FAQ\ (it. 9)} > 1\}$
+
+Note that the additional language domain adds complexity to establishing a diagnostic algorithm because simply by adding it to the set of items, the number of potential algorithms doubles.
+Further complexity is added by the fact that there are so far no guidelines for selecting a diagnostic threshold for Clock drawing and Five-words recall tests, both of which differ from their counterparts used by @dubois2007.
+Finally, although the Sevens backwards item has different thresholds in MoCA-based compared to MMSE-based algorithms, this difference is solely due to a difference in scoring whereby 3 points in MoCA correspond to 4 or 5 points in MMSE.
+The Seven backwards item threshold for MoCA-based algorithms used in this study is thus equivalent to its MMSE-based counterpart.
+
+Computing the Cartesian product $Global \times Attention \times Executive \times Construction \times Memory \times Language \times IADL$ yields $1 \times 1 \times 3 \times 1 \times 5 \times 2 \times 2 = 60$ distinct MoCA-based algorithms for probable PDD.
+
+For sMoCA-based algorithms, the following sets of items served as the basis:
+
+$Global = \{sMoCA < 13\}$
+
+$IADL = \{FAQ > 7, {FAQ\ (it. 9)} > 1\}$
+
+yielding $Global \times IADL$, i.e., $1 \times 2 = 2$ distinct sMoCA-based algorithms for probable PDD.
+
+Finally, the Level II algorithms were based on the following sets of items:
+
+$Attention = \{z(TMT\ A) < -1.5\ \cup z(WAIS\ DSB) < -1.5\}$
+
+$Executive = \{z(CF\ A) < -1.5\ \cup z(PST\ C) < -1.5\}$
+
+$Construction = \{z(JoLO) < -1.5\ \cup z(CLOXI) < -1.5\}$
+
+$Memory = \{z(RAVLT\ DR) < -1.5\ \cup z(BVMTR\ DR) < -1.5\ \cup z(WMS\text-III\ Family\ Pictures) < -1.5\}$
+
+$Language = \{z(WAIS\ Similarities) < -1.5 \ \cup z(BNT\ 60) < -1.5\}$
+
+$IADL = \{FAQ > 7, {FAQ\ (it. 9)} > 1\}$
+
+where $z()$ denotes calculation of age, sex and education adjusted z-score.
+This yields $1 \times 1 \times 1 \times 1 \times 1 \times 2 = 2$ distinct Level II algorithms for probable PDD in the current study.
+All but the BNT 60 item were evaluated using regression norms published by @bezdicek2017.
+Since the original article used BNT 30 instead of BNT 60, we approximated the deficit in BNT 60 by comparing patients' raw score to age- and education-specific normative values reported by @zemanova2016.
+Specifically, patients whose BNT 60 score fell below 5^th^ percentile of their demographic group in Table 6 of @zemanova2016 were considered to show signs of impaired performance.
+
+### Operationalization of Impaired Cognition
+
+In the original criteria, item 4 of Level I criteria, i.e., impaired cognition, was defined as follows: *"The proposed diagnostic criteria require a profile of cognitive deficits, typical of those described for PD-D, in two or more of four domains."* [@dubois2007, p.
+2316] Consequently, we defined impaired cognition as a deficit in two or more domains of four in MMSE-based criteria and as a deficit in two or more of five domains in MoCA-based criteria.
+sMoCA-based criteria omitted the "impaired cognition" item altogether because they were intended as a shorter screening alternative to classical Level I assessment.
+Finally, for the Level II criteria, we employed standard definition of impaired cognition as the *"\[i\]mpairment on at least two neuropsychological tests, represented by either two impaired tests in one cognitive domain or one impaired test in two different cognitive domains."* [@litvan2012, Table 1]
+
+```{r}
+#| label: tbl-algos
+#| tbl-cap: Summary of all algortihms for probable PDD used in the study.
+
+prev$gtables$algorithms
+```
+
+## Theoretical and Empirical Estimands
+
+In this study, we follow the framework proposed by @lundberg2021 for specifying targets of inference (i.e., the estimands) in qunatitative sciences to increase transparency and connect statistical evidence to relevant theory.
+@tbl-estimands contains verbal description of the components relating to each of our proclaimed research objectives and map them to the population quantity of interest (the theoretical estimand), data-dependent quantity that could be estimated (the empirical estimand) and quantities that are reported in the study (statistical estimates).
+
+```{r}
+#| label: tbl-estimands
+#| tbl-cap: Mapping between research objectives and quantities of interest in the current study.
+
+table_estimands()
+```
+
+The RO1 - to estimate the PDD rate and evaluate the diagnostic variability and concordance across different algorithms of probable PDD - was divided into four distinct research objectives:
+
+-   to estimate the rate of PDD within PD (RO1.1),
+-   to estimated variability of this rate (RO1.2),
+-   to evaluate predictive value of demographic variables for probable PDD classification (RO1.3) and
+-   to evaluate concordance between different probable PDD operationalizations and criteria (RO1.4).
+
+Estimates relating to RO1.1 and RO1.3 cannot be safely generalized beyond a population of PD patients that are candidates for DBS due to the systematic differences between DBS candidates pool and general PD population (such as the lower age of DBS candidates compared to the general PD population).
+On the other hand, the estimates relating to RO1.4 (and to a lesser degree to RO1.2[^16]) may not be substantially influenced by the sample at hand as the primary source of their variance might come from variability in measures employed (e.g., MMSE vs MoCA to assess global cognitive performance) rather than variability in patients' performance.
+Assuming that there is no substantial Differential Item Functioning for DBS candidates compared to a broader population of patients with PD, the estimates relating to RO1.4 can be cautiously generalized beyond the current sample.
+
+[^16]: Because the quantity of interest is a rate and could thus be though of as a sum of binomially distributed PDD occurences divided by the total number of patients, its variance will likely systematically vary with its mean.
+Specifically, as the rate goes from extremes to 0.5, the variance increases.
+Consequently, if our estimate of the rate was lower than the true population rate, e.g., because our sample includes younger patients compared to the general PD population, our estimate of variance would also lower than the true variance of PDD rate in the general PD population.
+Nonetheless, the between-algorithm variability may not be affected by this phenomenon as unlike variability of PDD rate, we do not have reason to assume it comes about by summing independent binomial events
+
+Finally, for the RO2, the theoretical estimand is defined as the set of diagnostic components whose variation systematically alters the probability of a probable PDD diagnosis.
+This aspect of the study is exploratory in nature.
+Empirically, we assess the contribution of each diagnostic feature by examining how variations in operational definitions (e.g., domain-specific thresholds, criteria for functional impairment) influence the statistical estimates derived for the first objective.
+This allows us to identify the diagnostic elements most responsible for between-algorithm discrepancies.
+
+## Supplementary Presentation of Results
+
+```{r}
+#| label: tbl-rates
+#| tbl-cap: Estimates of the rate of probable PDD in the sample.
+
+prev$gtables$rates
+```
+
+```{r}
+#| label: tbl-approx-bal
+#| tbl-cap: Level I algorithms for probable PDD as predictors of Level II classification as the reference arranged by their balanced accuracy score.
+
+table_levelII_approximations(conc$table, algos, "Balanced Accuracy")
+```
+
+```{r}
+#| label: fig-data
+#| fig-cap: Representation of study data.
+#| apa-note: The figure shows whether patients (x-axis) ordered from the youngest (left) to the oldest (right) were classified as probable PDD by each tested algorithm (y-axis) ordered from the one with the lowest (bottom) to the highest (top) PDD rate estimate.
+Patients printed in red are women, patients printed in blue are men.
+Red cells indicate probable PDD diagnosis, grey cells indicate non-PDD diagnosis and white cells indicate missing diagnosis.
+
+pred$plots$data
+```
+
+```{r}
+#| label: fig-pars
+#| fig-cap: Summary of logistic regressions parameters prediction probable PDD by age and sex.
+#| apa-note: Histograms represent odd ratio (OR) estimates and p-values associated with age, sex, and their interaction as predictors of each of the 68 probable PDD classification.
+In the case of parameters for sex, values higher than 20 were omitted for clarity.
+Vertical lines indicate OR = 1 and p = .05.
+
+pred$plots$parameters
+```
+
+```{r}
+#| label: fig-kappa
+#| fig-cap: Cohen's κ matrix.
+#| apa-note: The matrix depicts Cohen's κ measuring agreement between algorithms for PDD.
+Algorithms printed in red defined IADL deficit by FAQ total score, algorithms printed in black defined IADL deficit by FAQ item 9 response.
+
+conc$plots$Kappa
+```
+
+```{r}
+#| label: fig-sens
+#| fig-cap: Sensitivity matrix.
+#| apa-note: The matrix depicts sensitivity of algorithms for PDD depicted on x-axis in predicting outcomes based on algorithms on the y-axis.
+Algorithms printed in blue defined IADL deficit by FAQ total score, algorithms printed in black defined IADL deficit by FAQ item 9 response.
+
+conc$plots$Sensitivity
+```
+
+```{r}
+#| label: fig-spec
+#| fig-cap: Specificity matrix.
+#| apa-note: The matrix depicts specificity of algorithms for PDD depicted on x-axis in predicting outcomes based on algorithms on the y-axis.
+Algorithms printed in blue defined IADL deficit by FAQ total score, algorithms printed in black defined IADL deficit by FAQ item 9 response.
+
+conc$plots$Specificity
+```

### 2025-05-13 _manuscript/manuscript.qmd
Final introduction refinement (JM approved).

@@
 
-Parkinson’s disease (PD) is a neurodegenerative disorder defined by a gradual onset of motor impairment, including rigidity, bradykinesia, postural instability and resting tremor.
+Parkinson’s disease (PD) is a neurodegenerative disorder.
+It is primarily defined by a gradual onset of motor impairment, including rigidity, bradykinesia, postural instability and resting tremor.
 Additionally, patients frequently suffer from a range of non-motor impairments [@postuma2015], particularly cognitive decline.
@@
 
-Despite its clinical relevance, the diagnosis of PDD remains complex.
-The original criteria formalized in 2007 by the International Parkinson and Movement Disorder Society (MDS) [@dubois2007] were heavily influenced by frameworks established primarily for Alzheimer’s disease (AD).
-Although these criteria provided a valuable base, they lacked the specificity to detect the specific cognitive aspects of PDD [@emre2007; @yamashita2023].
+According to a recent meta-analysis, approximately one-quarter of PD patients is likely to be diagnosed with PDD [@sousa2022].
+However, reported PDD rate estimates vary widely, ranging from 14% up to 55%, depending on methodological criteria employed [@sousa2022].
+Moreover, factors such as patients' sex [@cereda2016], age and disease duration appear to modulate the risk of cognitive decline and PDD [@oh2016; @rana2012].
 
-In order to increase diagnostic accuracy, the MDS also developed a levelled system for PDD detection.
-Level I criteria are designed primarily for clinical use, consisting of brief cognitive assessments and relying on clinical judgement.
-In contrast, Level II criteria use comprehensive neuropsychological testing across cognitive domains to provide greater diagnostic specificity [@emre2007].
+Despite the clinical relevance of PDD, its diagnosis remains complex.
+A milestone in research of PDD was the publication of diagnostic criteria established in 2007 by the International Parkinson and Movement Disorder Society (MDS) [@dubois2007].
+In these citeria, the MDS introduced a two-levelled system for PDD detection.
+Level I consists of brief cognitive assessments, while Level II involves comprehensive neuropsychological testing across cognitive domains [@emre2007].
 
-One specific feature of the original Level I criteria was the presence of an algorithm allowing flexibility in test selection.
-The criteria included MMSE pentagons for evaluation of visuospatial ability, three-word recall for memory assessment,  months reversed or seven backwards for attention assessment and lexical fluency or clock drawing for executive function assessment [@dubois2007].
+The original Level I algorithm included eight conditions that had to be satisfied simultaneously in order to diagnose probable PDD.
+These included: 1) diagnosis of PD proposed by the Queen Square Brain Bank; 2) PD onset prior to the PDD emergence; 3) evidence of global cognitive impairment (MMSE score \< 26 points); 4) cognitive deficit interference with the IADL (assessed by the pill questionnaire or caregiver interview); 5) impairment in at least two cognitive domains, namely memory, attention, visuo-constructive abilities and executive functions; 6) there was absence of Major Depressive Disorder; 7) absence of delirium; and 8) exclusion of other abnormalities and potential causes of dementia [@dubois2007].
 
-According to a recent meta-nalysis, the PDD prevalence estimates vary widely, ranging from 14% up to 55%, depending on methodological criteria employed [@sousa2022].
-Moreover, factors such as patients` sex [@cereda2016], age and disease duration appear to modulate the risk of cognitive decline [@oh2016; @rana2012].
-Currently, efforts are focused on refining the PDD diagnostic framework to improve its consistency and applicability in both research and clinical contexts [@kulisevsky2024].
+Currently, efforts are focused on refining this PDD diagnostic framework.
+A recent call for a change pinpoints limitations regarding the original criteria and suggest various updates to enhance their utility [@kulisevsky2024].
+Proposed suggestions include replacement of Mini Mental State Examination (MMSE) by Montreal Cognitive Assessment (MoCA), which is more sensitive to PD specific cognitive impairment; expansion of instrumental activities of daily living (IADL) evaluation; inclusion of language assessment; recognition of anxiety as one the neuropsychiatric symptoms relevant in PDD; and intergration of biomarkers.
 
-The present study evaluates the diagnostic concordance between the original Level I PDD criteria, as established by the MDS Task Force [@emre2007; @dubois2007] and criteria inspired by the recent call for change [@kulisevsky2024] within a cohort of PD patients considered for DBS.
-Furthermore, both sets of criteria are compared to PDD diagnosed on Level II.
-The study aims to address the following research objectives (RO): (RO1) To estimate the Parkinson’s disease dementia (PDD) rates and evaluate the diagnostic variability and concordance across different PDD criteria.
+In the light of these proposals, the current study aims to evaluate the diagnostic concordance between the original MDS Level I PDD criteria [@emre2007; @dubois2007] and a modified framework based on the recent call for change [@kulisevsky2024].
+Furthermore, both Level I diagnostic approaches are compared to PDD diagnosed on Level II.
+The study aims to address the following research objectives (RO): (RO1) To estimate the PDD rate and evaluate the diagnostic variability and concordance across different PDD criteria.
 (RO2) To identify specific diagnostic components contributing to PDD classification variability across the applied criteria.
@@
 Clinical records spanning `r dates["min"]` to `r dates["max"]` were examined.
-All participants underwent neuropsychological evaluation conducted by a trained clinical psychologist (OB) as part of standard preoperative cognitive assessments for DBS eligibility at the General University Hospital in Prague.
+All participants were candidates for for Deep Brain Stimulation (DBS) treatment and underwent neuropsychological evaluation conducted by a trained clinical psychologist (OB) as part of standard preoperative assessments forDBS eligibility at the General University Hospital in Prague.
 
@@
 **The discrepancy is likely attributable to our DBS-specific cohort, which differs in important characteristics, such as the younger age of the patients [@rana2012].
-/ Compared to these estimates, our study reports generally lower PDD rates, likely reflecting differences in sample characteristics, diagnostic criteria, and methodology.
-[^3]**
+/ Compared to these estimates, our study reports generally lower PDD rates, likely reflecting differences in sample characteristics, diagnostic criteria, and methodology.** [^3]
 
-[^3]: **MM to JM: Which sentence do you like better?**
+[^3]: **Because the quantity of interest is a rate and could thus be though of as a sum of binomially distributed PDD occurences divided by the total number of patients, its variance will likely systematically vary with its mean.
+Specifically, as the rate goes from extremes to 0.5, the variance increases.
+Consequently, if our estimate of the rate was lower than the true population rate, e.g., because our sample includes younger patients compared to the general PD population, our estimate of variance would also lower than the true variance of PDD rate in the general PD population.
+Nonetheless, the between-algorithm variability may not be affected by this phenomenon as unlike variability of PDD rate, we do not have reason to assume it comes about by summing independent binomial events**
 
@@
 Estimates relating to RO1.1 and RO1.3 cannot be safely generalized beyond a population of PD patients that are candidates for DBS due to the systematic differences between DBS candidates pool and general PD population (such as the lower age of DBS candidates compared to the general PD population).
-On the other hand, the estimates relating to RO1.4 (and to a lesser degree to RO1.2[^3]) may not be substantially influenced by the sample at hand as the primary source of their variance might come from variability in measures employed (e.g., MMSE vs MoCA to assess global cognitive performance) rather than variability in patients' performance.
+On the other hand, the estimates relating to RO1.4 (and to a lesser degree to RO1.2[^4]) may not be substantially influenced by the sample at hand as the primary source of their variance might come from variability in measures employed (e.g., MMSE vs MoCA to assess global cognitive performance) rather than variability in patients' performance.
 Assuming that there is no substantial Differential Item Functioning for DBS candidates compared to a broader population of patients with PD, the estimates relating to RO1.4 can be cautiously generalized beyond the current sample.
 
-[^3]: Because the quantity of interest is a rate and could thus be though of as a sum of binomially distributed PDD occurences divided by the total number of patients, its variance will likely systematically vary with its mean.
+[^4]: Because the quantity of interest is a rate and could thus be though of as a sum of binomially distributed PDD occurences divided by the total number of patients, its variance will likely systematically vary with its mean.
 Specifically, as the rate goes from extremes to 0.5, the variance increases.

### 2025-05-13 _manuscript/manuscript.qmd
The merge of manuscript.qmd (main) and text.qmd (josefswritings).

@@
 format:
- apaquarto-docx: 
-    toc: false
  apaquarto-pdf:
@@
     keep-tex: true
-bibliography: references.bib
+ apaquarto-docx: 
+    toc: false
+bibliography: bibliography.bib
 floatsintext: false
@@
 Level I was assessed using the Mini-Mental State Examination (MMSE) [@stepankova2014; @folstein1975] and the Montreal Cognitive Assessment (MoCA) [@kopecek2016; @nasreddine2005].
-The neuropsychological assessment at Level II covered five cognitive domains, each evaluated through specific tests as follows: attention and working memory assessed using Trail Making Test Part A (TMT-A) [@bezdicek2012; @reitan2004], and WAIS Digit Span Backward (WAIS DSB) [@wechsler1997]; executive function evaluated via Categorical Verbal Fluency (CF) [@nikolai2015], and subtest from the Prague Stroop Test – Colors (PST-C) [@bezdicek2021]; language measured with the WAIS Similarities subtest [@wechsler1997], and the Boston Naming Test (BNT-60) [@kaplan1983]; memory examined using the Rey Auditory Verbal Learning Test (RAVLT) [@frydrychova2018; @rey1964; @bezdicek2013] delayed recall, and the Brief Visual Memory Test–Revised (BVMTR) [@havlik2020; @benedict1997] delayed recall, or WAIS Family Pictures subtest [@wechsler1997] delayed recall; visuospatial function assessed through the Judgment of Line Orientation Test (JoLO) [@benton1983], and Clock Drawing Test (CLOX) [@royall1998].
-
-In addition to the core cognitive assessments, namely the Clock Drawing Test (CDT) and Letter Fluency tasks were included to capture domain-specific impairments.
-The classification of Parkinson’s Disease Dementia (PDD) based on Level I criteria was determined using established scoring thresholds from the original criteria [@dubois2007] with corresponding MoCA equivalents.
+The neuropsychological assessment at Level II covered five cognitive domains, each evaluated through specific tests as follows: attention and working memory assessed using Trail Making Test Part A (TMT-A) [@bezdicek2012; @reitan2004], and WAIS Digit Span Backward (WAIS DSB) [@wechsler1997]; executive function evaluated via Categorical Verbal Fluency (CF) [@nikolai2015], and subtest from the Prague Stroop Test – Colors (PST-C) [@bezdicek2021]; language measured with the WAIS Similarities subtest [@wechsler1997], and the Boston Naming Test (BNT-60) [@zemanova2016; @kaplan1983]; memory examined using the Rey Auditory Verbal Learning Test (RAVLT) [@frydrychova2018; @bezdicek2013; @rey1964] delayed recall, and the Brief Visual Memory Test–Revised (BVMTR) [@havlik2020; @benedict1997] delayed recall, or WAIS Family Pictures subtest [@wechsler1997] delayed recall; visuospatial function assessed through the Judgment of Line Orientation Test (JoLO) [@benton1983], and Clock Drawing Test (CLOX) [@royall1998].
 
@@
 ```{r}
-#| label: "criteria-check"
+#| label: crit-check
 
@@
 ```{r}
-#| label: "tbl-crits"
+#| label: tbl-crits
 #| tbl-cap: "Summary of probable PDD operationalizations compared in the study."
@@
 
-[Insert @tbl-crits here]
+\[Insert Table 1 here\]
 
@@
 
-Following the framework proposed by Lundberg et al.
-(2021), in this study we explicitly connect our research objectives and their corresponding theoretical (i.e., targets of inference) and empirical (i.e., data-driven) estimands to statistical estimates.
+Following the framework proposed by @lundberg2021, in this study we explicitly connect our research objectives and their corresponding theoretical (i.e., targets of inference) and empirical (i.e., data-driven) estimands to statistical estimates.
 The theoretical estimand refers to a unit-specific quantity defined over a target population and represents the ideal quantity that would address the research question under optimal conditions, such as access to complete population data or perfect experimental control.
@@
 
-To address study objectives, we started by repeatedly assigning each patient the diagnosis of probable PDD based on each PDD algorithm listed in @tbl-crits (see also @tbl-algos) resulting in a `r sum(data$incl)` (patients) $\times$ `r sum(critnum)` (operationalizations) matrix where each cell indicates whether a patient (row) meets criteria for probable PDD according to an algorithm (column)[^1].
+To address study objectives, we started by repeatedly assigning each patient the diagnosis of probable PDD based on each PDD algorithm listed in @tbl-crits (see also @tbl-algos) resulting in a `r sum(data$incl)` (patients) $\times$ `r sum(critnum)` (algorithms) matrix where each cell indicates whether a patient (row) meets criteria for probable PDD according to an algorithm (column)[^6].
 PDD rate estimates were computed as $\frac{N_{PDD}}{N_{total}}$ separately for each algortihm.
@@
 
-[^1]: **JM: This could and should be shared most likely, as long, as we anonymize properly. Let's ask Oto Mestek and the NPO team how and if is it possible.**
+[^6]: **JM: This could and should be shared most likely, as long, as we anonymize properly. Let's ask Oto Mestek and the NPO team how and if is it possible.**
 
@@
 For each pair of algorithms, the analysis was repeated twice such that each variable of the pair served once as the reference and once as the predictor.
-Following measures were used to evaluate pairwise concordance between different algorithms for probable PDD: 1) Cohen's $\kappa$ with its 95% confidence interval (CI) computed via the `cohen.kappa()` function from the R package *psych* [@revelle2024]; 2) Accuracy (i.e., the proportion of correct predictions, both true positives and true negatives, among the total number of cases) with its 95% CI; 3) Sensitivity/Recall (i.e., the proportion of true positives); and 4) Specificity (i.e., the proportion of true negatives).[^2]
-
-[^2]: **JM: Do not forget to make it public before submitted!**
+Following measures were used to evaluate pairwise concordance between different algorithms for probable PDD: 1) Cohen's $\kappa$ with its 95% confidence interval (CI) computed via the `cohen.kappa()` function from the R package *psych* [@revelle2024]; 2) Accuracy (i.e., the proportion of correct predictions, both true positives and true negatives, among the total number of cases) with its 95% CI; 3) Sensitivity/Recall (i.e., the proportion of true positives); and 4) Specificity (i.e., the proportion of true negatives).
 
@@
 All analyses were conducted within the R (version `r with(version, paste(major, minor, sep="."))`) software environment for statistical computing [@rsoft].
-The software code supporting this article is available at <https://github.com/josefmana/DemCr1t.git>.\[\^11\]
+The software code supporting this article is available at <https://github.com/josefmana/DemCr1t.git>.[^7]
+
+[^7]: **JM: Do not forget to make it public before submitted!**
 
@@
 ```{r}
-#| label: "results"
+#| label: results
 
@@
 ```{r}
-#| label: "tbl-desc"
+#| label: tbl-desc
 #| tbl-cap: "Cognitive characteristics of the sample."
@@
 
-[Insert @tbl-desc here]
+\[Insert Table 2 here\]
 
@@
 On average, estimated PDD rate was `r rates_overall["M"]`% (SD = `r rates_overall["SD"]`, Md = `r rates_overall["Md"]`, range `r rates_overall["minmax"]`).
-Notably, the estimates were substantially lower when FAQ item 9 was used as a criterion of IADL deficit (M = `r rates_FAQ9["M"]`% SD = `r rates_FAQ9["SD"]`, Md = `r rates_FAQ9["Md"]`, range `r rates_FAQ9["minmax"]`) compared to using total FAQ score criterion (M = `r rates_FAQtot["M"]`% SD = `r rates_FAQtot["SD"]`, Md = `r rates_FAQtot["Md"]`, range `r rates_FAQtot["minmax"]`) as demonstrated also in @fig-prev.
+Notably, the estimates were substantially lower when FAQ item 9 was used as a criterion of IADL deficit (M = `r rates_FAQ9["M"]`% SD = `r rates_FAQ9["SD"]`, Md = `r rates_FAQ9["Md"]`, range `r rates_FAQ9["minmax"]`) compared to using total FAQ score criterion (M = `r rates_FAQtot["M"]`% SD = `r rates_FAQtot["SD"]`, Md = `r rates_FAQtot["Md"]`, range `r rates_FAQtot["minmax"]`) as demonstrated also in @fig-rates.
 Neither age nor sex or their interaction (*p*s $\geq$ `r min_p`) reliably predicted probable PDD classification across algorithms (see @fig-data and @fig-pars).
 
-[Insert @fig-prev here]
+\[Insert Figure 1 here\]
 
 ```{r}
-#| label: "fig-prev"
-#| fig-cap: "Summary of the estimates of probable PDD rate across operationalizations under consideration.
-Vertical lines represent estimates arrived at by using sMoCA (dotted) or Level II (dashed) with FAQ item 9 (orange) or FAQ total score (blue) as criteria for probable PDD."
+#| label: fig-rates
+#| fig-cap: Summary of the estimates of probable PDD rate.
+#| apa-note: Vertical lines represent estimates arrived at by using sMoCA (dotted) or Level II (dashed) with FAQ item 9 (orange) or FAQ total score (blue) as criteria for probable PDD.
 
@@
 Results of the analyses of prediction Accuracy, Cohen's $\kappa$, Sensitivity and Specificity are presented in @fig-acc, @fig-kappa, @fig-sens and @fig-spec respectively.
-**Numerical results are available at ...
-\[\^15\].** Generally, algorithms that employed the same operationalization of IADL deficit showed substantial pairwise concordance, however, algorithms that operationalized IADL deficit differently did not.
+**Numerical results are available at ...** [^8].
+Generally, algorithms that employed the same operationalization of IADL deficit showed substantial pairwise concordance, however, algorithms that operationalized IADL deficit differently did not.
 Whereas among algorithms with identical IADL deficit operationalization, the agreement judged by Cohen's $\kappa$ was moderately high (operationalization by FAQ total score: $\kappa$ = `r kappa_sums["faq_tot"]`; operationalization by FAQ item 9: $\kappa$ = `r kappa_sums["faq_9"]`), among algorithms that differ in IADL deficit operationalization it was low ($\kappa$ = `r kappa_sums["cross"]`).
 
-[Insert @fig-acc here]
+[^8]: **JM: Should provide an excel sheet or html table the readers could go through by themselves.**
+
+\[Insert Figure 2 here\]
 
 ```{r}
-#| label: "fig-acc"
-#| fig-cap: "Prediction accuracy matrix. The matrix depict accuracy of "
+#| label: fig-acc
+#| fig-cap: Prediction accuracy matrix.
+#| apa-note: The matrix depicts classification accuracy of algorithms for PDD depicted on x-axis in predicting outcomes based on algorithms on the y-axis.
+Algorithms printed in blue defined IADL deficit by FAQ total score, algorithms printed in black defined IADL deficit by FAQ item 9 response.
+Cases with asterisk indicate predictive accuracy statistically significantly higher than the No Information Rate.
 
@@
 ```{r}
-#| label: "lvlII-stats"
+#| label: lvlII-stats
 
@@
 ```{r}
-#| label: "tbl-approx"
+#| label: tbl-approx
 #| tbl-cap: "Level I algorithms for probable PDD as predictors of Level II classification as the reference."
@@
 
-[Insert @tbl-approx here]
-
 # Discussion
 
-This study examined the diagnostic variability and validity of Level I criteria for PDD, specifically in candidates for DBS treatment.
+This study examined the diagnostic variability and validity of Level I criteria for PDD, **specifically in candidates for DBS treatment**[^9].
 By systematically comparing 68 diagnostic algorithms based on MMSE, MoCA, sMoCA, and Level II criteria, our findings highlight the significance of algorithm design.
 
-## Variability in Prevalence Estimates
+[^9]: **JM: All-right, I do not think it needs to be here either.**
 
-Our results showed a wide range in estimated PDD prevalence across the total of 68 algorithms, ranging from 2.00% to 16.75%.
-Estimates reached lower rates when using solely FAQ item 9 in comparison with the full FAQ scale.
+## Variability in PDD Rate Estimates
+
+Our results showed a wide range in estimated PDD rate across algorithms, ranging from 2.00% to 16.75%.
+Estimates reached lower rates when using solely FAQ item 9 (as an approximation of the pill questionnaire suggested by @dubois2007) in comparison with the full FAQ scale.
 This discrepancy highlights the diagnostic importance of IADL assessment.
-Our overall PDD rates were consistently lower compared to previous studies regarding PDD among PD patients, demonstrating wide variability based on various criteria used.
-For instance, a retrospective study reported a PDD rate of 19.7% [@rana2012], while other clinical investigation found even higher rates, reaching up to 30% [@aarsland2011].
-A recent complex meta-analysis synthesizing global data placed the rate at 26.30% [@sousa2022].
+Our overall PDD rates were consistently lower compared to previous studies regarding PDD among PD patients, **demonstrating wide variability based on various criteria used**[^10].
+For instance, a retrospective study reported a PDD rate of 19.7% [@rana2012], while other clinical investigation found even higher rates, reaching up to 30% [@aarsland2005][^11].
+A recent complex meta-analysis synthesizing global data placed the expected PDD rate in PD at 26.30% [@sousa2022].
 **The discrepancy is likely attributable to our DBS-specific cohort, which differs in important characteristics, such as the younger age of the patients [@rana2012].
-/ Compared to these estimates, our study reports generally lower PDD rates, likely reflecting differences in sample characteristics, diagnostic criteria, and methodology.** [^3]
+/ Compared to these estimates, our study reports generally lower PDD rates, likely reflecting differences in sample characteristics, diagnostic criteria, and methodology.** [^12] [^13]
 
-[^3]: **Because the quantity of interest is a rate and could thus be though of as a sum of binomially distributed PDD occurences divided by the total number of patients, its variance will likely systematically vary with its mean.
+[^10]: **JM: What is demonstrating this variability? Our findings or previous ones?**
+
+[^11]: **JM: There was aarsland2011 originally but I did not find such a study. Is this reference correct?**
+
+[^12]: **Because the quantity of interest is a rate and could thus be though of as a sum of binomially distributed PDD occurences divided by the total number of patients, its variance will likely systematically vary with its mean.
 Specifically, as the rate goes from extremes to 0.5, the variance increases.
@@
 
+[^13]: \*\*JM: I prefer the second one with added sentence like: "Specifically, our cohort included candidates for DBS who are generally younger than the general PD population and age was repeatedly shown to be a strong predictor of PDD (some citations)." We should also incorporate here a statement about missing evidence of age being predictor in our sample.\*
+
 ## Concordance Between Diagnostic Algorithms
 
-Our pairwise comparison of diagnostic algorithms showed that agreement was notably stronger among those using the same IADL operationalization (κ range from 0.75, SD = 0.13 to 0.86, SD = 0.09) compared to those using different IADL definitions (κ = 0.34, SD = 0.08).
-This demonstrates that even slight methodological differences can yield divergent diagnostic outcomes.
+Our pairwise comparisons of diagnostic algorithms showed that agreement was notably stronger among those using the same IADL operationalization **(**$\kappa$ range from 0.75, SD = 0.13 to 0.86, SD = 0.09)[^14] compared to those using different IADL definitions **(**$\kappa$ = 0.34, SD = 0.08)[^15].
+**Moreover, the agreement was slightly higher between algorithms that defined IADL deficit by FAQ item 9 compared to algorithms that defined it using the full FAQ scale.
+One possible explanation of this difference follows from the observation that algorithms using the full scale definition yielded higher PDD rate estimates.
+Because there was higher probability of being diagnosed with IADL deficit based on the full FAQ scale, there was also a bigger room for disagreement in the Cognitive Impairment status when different indexes weer used (e.g., by defining executive deficit via clock drawing vs.
+lexical fluency).**
+
+[^14]: **JM: There is no need to repeat these numbers in this section.**
+
+[^15]: **JM: Same as above.**
+
+**Overall, algorithms using the same definition of IADL deficit showed degree of concordance equivalent to moderate (when using FAQ total score to define IADL deficit) or strong (when using FAQ item 9 to define IADL deficit) level of agreement between raters in an inter-rater reliability analysis [@mchugh2012].
+On the other hand, the concordance between algorithms using different IADL deficit definitions was equivalent to minimal agreement.** This demonstrates that even slight methodological differences can yield divergent diagnostic outcomes.
 Such findings are critical for clinicians relying on Level I criteria for eligibility decisions, as the choice of algorithm could lead to contradictory classifications of PDD.
@@
 In these cases, the best-performing algorithms achieved accuracy rates of 94–99%, together with high sensitivity and specificity.
+**JM: Some sMoCA praise should be added here ...**
 
-## Clinical Implications
+## Clinical Implications^[**JM: Is this section needed? It does not seem to say much to me. I would't be afraid to drop it.**]
 
@@
 
-## Constraints on generalizability
+## Limitations and Future Directions
 
-This study’s generalizability is limited by its retrospective design and the homogeneity of the DBS candidate cohort, which may not reflect broader PD populations with varying cognitive profiles.
-Moreover, use of the Czech normative data may limit international applicability, although our results represent clinical standards in the region [@simons2017].
+This study’s generalizability is limited by its retrospective design and the homogeneity of the patient cohort, which may not reflect broader PD populations with varying cognitive profiles.
+**Specifically, younger age of the sample and possible lack of patients with high risk phenotypes for PDD compared to general population of people with PD might have been responsible for comparatively lower PDD rate estimates in our study.
+Nonetheless, the analysis of the concordance between different algorithms might not be affected by the sample to the same degree allowing for a broader generalization.
+Future studies could replicate our results on different cohorts of PD patients.
+To make this process easier, the code used to generate our results is publicly available and easily applicable to other data sets with a similar data structure.**
 
-## Future Directions
+**JM: The fact that no biomarkers were used and the IADL was at best estimated by FAQ which is somewhat archaic (comp.
+to the e-IADL thing which could be also cited and is more in line with Kulisevsky) should be mentioned.
+Also, this is a great space for an advert for NOT/DOT because the self-report of IADLs can be viewed as a limitation worth of addressing, especially since our results imply possible central role for IADL deficit definition in further work on PDD definition to achieve good reliability.**
 
 Future research should focus on the validation of updated Level I criteria using longitudinal data and neurobiological markers.
-Furthermore, the development of adaptive diagnostic algorithms that can combine performance across cognitive and functional domains may enhance diagnostic sensitivity while reducing the risk of over-classification of PDD.
+Furthermore, the development of adaptive diagnostic algorithms that can combine performance across cognitive and functional domains may enhance diagnostic sensitivity while reducing the risk of **over-classification of PDD**^[**JM: What's an over-classification of PDD?**].
 
@@
 
-This study systematically investigated the application of multiple Level I diagnostic criteria for PDD within a cohort of patients considered for DBS tretment.
+This study systematically investigated the application of multiple Level I diagnostic criteria for PDD within a cohort of patients considered for DBS treatment.
 Our results show variability in PDD rate estimates, strongly influenced by the choice of cognitive screening instruments and the operationalization of functional impairment.
-The divergence observed across algorithms demonstrates the sensitivity of diagnostic outcomes to seemingly negligible methodological choices.
+The divergence observed across algorithms demonstrates the sensitivity of diagnostic outcomes to seemingly negligible methodological choices.^[**JM: This sounds as a great opening to the discussion, rather than conclusions :-)**]
 
-The proposed changes of the current diagnostic criteria [@kulisevsky2024] offer heightened sensitivity by proposing the use of MoCA-based components and broader IADL assessments.
-Overly conservative criteria, such as reliance on pill questionnaire (i.e.
+The **proposed changes**^[**JM: They did not propose anything though and from what I gather, the authors stress this a lot - that their article in not a proposal of new criteria, only a position paper.**] of the current diagnostic criteria [@kulisevsky2024] offer heightened sensitivity by proposing the use of MoCA-based components and broader IADL assessments.
+Conservative criteria, such as reliance on pill questionnaire (i.e.
 FAQ item 9 equivalence), may fail to detect meaningful functional decline and thus under-identify true cases of PDD.
+**On the other hand, such criteria show slightly higher agreement between their different forms.**
 
-Conclusively, this study contributes to the improvement of PDD diagnostics by offering evidence for the clarification of Level I criteria and emphasizing the value of psychometric precision in clinical neuropsychology.
+Conclusively, this study contributes to the improvement of PDD diagnostics by offering **evidence for the clarification of Level I criteria**^[**JM: What is meant by clarification of Level I criteria?
+Also, how does our study offer evidence for such thing?**] and **emphasizing the value of psychometric precision in clinical neuropsychology**^[**JM: This is independent of our study.**].
 Future research should validate the algorithms longitudinally and explore possible integration of biomarkers to further improve diagnostic reliability.
@@
 ```{r}
-#| label: "tbl-algos"
-#| tbl-cap: "Summary of all algortihms for probable PDD used in the study."
+#| label: tbl-algos
+#| tbl-cap: Summary of all algortihms for probable PDD used in the study.
 
@@
 ```{r}
-#| label: "tbl-estimands"
-#| tbl-cap: "Mapping between research objectives and quantities of interest in the current study."
+#| label: tbl-estimands
+#| tbl-cap: Mapping between research objectives and quantities of interest in the current study.
 
@@
 Estimates relating to RO1.1 and RO1.3 cannot be safely generalized beyond a population of PD patients that are candidates for DBS due to the systematic differences between DBS candidates pool and general PD population (such as the lower age of DBS candidates compared to the general PD population).
-On the other hand, the estimates relating to RO1.4 (and to a lesser degree to RO1.2[^4]) may not be substantially influenced by the sample at hand as the primary source of their variance might come from variability in measures employed (e.g., MMSE vs MoCA to assess global cognitive performance) rather than variability in patients' performance.
+On the other hand, the estimates relating to RO1.4 (and to a lesser degree to RO1.2[^16]) may not be substantially influenced by the sample at hand as the primary source of their variance might come from variability in measures employed (e.g., MMSE vs MoCA to assess global cognitive performance) rather than variability in patients' performance.
 Assuming that there is no substantial Differential Item Functioning for DBS candidates compared to a broader population of patients with PD, the estimates relating to RO1.4 can be cautiously generalized beyond the current sample.
 
-[^4]: Because the quantity of interest is a rate and could thus be though of as a sum of binomially distributed PDD occurences divided by the total number of patients, its variance will likely systematically vary with its mean.
+[^16]: Because the quantity of interest is a rate and could thus be though of as a sum of binomially distributed PDD occurences divided by the total number of patients, its variance will likely systematically vary with its mean.
 Specifically, as the rate goes from extremes to 0.5, the variance increases.
@@
 ```{r}
-#| label: "tbl-rates"
-#| tbl-cap: "Estimates of the rate of probable PDD in the sample."
+#| label: tbl-rates
+#| tbl-cap: Estimates of the rate of probable PDD in the sample.
 
@@
 ```{r}
-#| label: "tbl-approx-bal"
-#| tbl-cap: "Level I algorithms for probable PDD as predictors of Level II classification as the reference arranged by their balanced accuracy score."
+#| label: tbl-approx-bal
+#| tbl-cap: Level I algorithms for probable PDD as predictors of Level II classification as the reference arranged by their balanced accuracy score.
 
@@
 ```{r}
-#| label: "fig-data"
+#| label: fig-data
+#| fig-cap: Representation of study data.
+#| apa-note: The figure shows whether patients (x-axis) ordered from the youngest (left) to the oldest (right) were classified as probable PDD by each tested algorithm (y-axis) ordered from the one with the lowest (bottom) to the highest (top) PDD rate estimate.
+Patients printed in red are women, patients printed in blue are men.
+Red cells indicate probable PDD diagnosis, grey cells indicate non-PDD diagnosis and white cells indicate missing diagnosis.
 
@@
 ```{r}
-#| label: "fig-pars"
+#| label: fig-pars
+#| fig-cap: Summary of logistic regressions parameters prediction probable PDD by age and sex.
+#| apa-note: Histograms represent odd ratio (OR) estimates and p-values associated with age, sex, and their interaction as predictors of each of the 68 probable PDD classification.
+In the case of parameters for sex, values higher than 20 were omitted for clarity.
+Vertical lines indicate OR = 1 and p = .05.
 
@@
 ```{r}
-#| label: "fig-kappa"
+#| label: fig-kappa
+#| fig-cap: Cohen's κ matrix.
+#| apa-note: The matrix depicts Cohen's κ measuring agreement between algorithms for PDD.
+Algorithms printed in red defined IADL deficit by FAQ total score, algorithms printed in black defined IADL deficit by FAQ item 9 response.
 
@@
 ```{r}
-#| label: "fig-sens"
+#| label: fig-sens
+#| fig-cap: Sensitivity matrix.
+#| apa-note: The matrix depicts sensitivity of algorithms for PDD depicted on x-axis in predicting outcomes based on algorithms on the y-axis.
+Algorithms printed in blue defined IADL deficit by FAQ total score, algorithms printed in black defined IADL deficit by FAQ item 9 response.
 
@@
 ```{r}
-#| label: "fig-spec"
+#| label: fig-spec
+#| fig-cap: Specificity matrix.
+#| apa-note: The matrix depicts specificity of algorithms for PDD depicted on x-axis in predicting outcomes based on algorithms on the y-axis.
+Algorithms printed in blue defined IADL deficit by FAQ total score, algorithms printed in black defined IADL deficit by FAQ item 9 response.
 

### 2025-05-14 _manuscript/manuscript.qmd
CHANGEEESSS!!!

@@
 
-Parkinson’s disease (PD) is a neurodegenerative disorder.
-It is primarily defined by a gradual onset of motor impairment, including rigidity, bradykinesia, postural instability and resting tremor.
-Additionally, patients frequently suffer from a range of non-motor impairments [@postuma2015], particularly cognitive decline.
-This state might result in Parkinson’s disease dementia (PDD) in a subset of patients [@meireles2012].
+Parkinson’s disease (PD) is a neurodegenerative disorder typically characterized by a progressive onset of motor symptoms, including rigidity, bradykinesia, postural instability and resting tremor.
+Moreover, patients suffer from a range of non-motor impairments [@postuma2015], particularly cognitive decline.
+This factor might result in Parkinson’s disease dementia (PDD) in a subset of patients [@meireles2012].
 
@@
 
-Ethical approval for the study protocol was obtained from the Ethics Committee of the General University Hospital in Prague.
-Informed consent was secured from all patients prior to their neuropsychological assessments, in adherence to ethical research guidelines.
-
 ## Neuropsychological Assessment
 
-Cognitive performance was evaluated at both Level I (abbreviated assessment) and Level II (comprehensive assessment) according to the standard MDS neuropsychology battery for Parkinson's Disease Mild Cognitive Impairment (PD-MCI) [@litvan2012, @bezdicek2017].
-Level I was assessed using the Mini-Mental State Examination (MMSE) [@stepankova2014; @folstein1975] and the Montreal Cognitive Assessment (MoCA) [@kopecek2016; @nasreddine2005].
-The neuropsychological assessment at Level II covered five cognitive domains, each evaluated through specific tests as follows: attention and working memory assessed using Trail Making Test Part A (TMT-A) [@bezdicek2012; @reitan2004], and WAIS Digit Span Backward (WAIS DSB) [@wechsler1997]; executive function evaluated via Categorical Verbal Fluency (CF) [@nikolai2015], and subtest from the Prague Stroop Test – Colors (PST-C) [@bezdicek2021]; language measured with the WAIS Similarities subtest [@wechsler1997], and the Boston Naming Test (BNT-60) [@zemanova2016; @kaplan1983]; memory examined using the Rey Auditory Verbal Learning Test (RAVLT) [@frydrychova2018; @bezdicek2013; @rey1964] delayed recall, and the Brief Visual Memory Test–Revised (BVMTR) [@havlik2020; @benedict1997] delayed recall, or WAIS Family Pictures subtest [@wechsler1997] delayed recall; visuospatial function assessed through the Judgment of Line Orientation Test (JoLO) [@benton1983], and Clock Drawing Test (CLOX) [@royall1998].
+Cognitive performance was evaluated at both Level I and Level II according to the standard MDS battery for Parkinson's Disease Mild Cognitive Impairment (PD-MCI) [@litvan2012, @bezdicek2017].
+Cognitive performance at Level I was assessed by the Mini-Mental State Examination (MMSE) [@stepankova2014; @folstein1975] and the Montreal Cognitive Assessment (MoCA) [@kopecek2016; @nasreddine2005].
+The comprehensive neuropsychological assessment at Level II evaluated five cognitive domains through specific test: attention and working memory assessed by Trail Making Test Part A (TMT-A) [@bezdicek2012; @reitan2004], and WAIS Digit Span Backward (WAIS DSB) [@wechsler1997], executive functions by Categorical Verbal Fluency (CF) [@nikolai2015], and subtest from the Prague Stroop Test – Colors (PST-C) [@bezdicek2021], language by the WAIS Similarities subtest [@wechsler1997], and the Boston Naming Test (BNT-60) [@zemanova2016; @kaplan1983], memory by the Rey Auditory Verbal Learning Test (RAVLT) [@frydrychova2018; @bezdicek2013; @rey1964] delayed recall, and the Brief Visual Memory Test–Revised (BVMTR) [@havlik2020; @benedict1997] delayed recall, or WAIS Family Pictures subtest [@wechsler1997] delayed recall, visuospatial function assessed by the Judgment of Line Orientation Test (JoLO) [@benton1983], and Clock Drawing Test (CLOX) [@royall1998].
 
-To assess functional impairment, the Functional Activities Questionnaire (FAQ) [@bezdicek2016; @pfeffer1982] was administered.
-Additionally, neuropsychiatric status was evaluated using the Beck Depression Inventory-II (BDI-II) [@ciharova2020; @beck1996] and State-Trait Anxiety Inventory (STAI) [@spielberg1983; @mullner1980].
+The Functional Activities Questionnaire (FAQ) [@bezdicek2016; @pfeffer1982] was administered to assess functional impairment.
+The Beck Depression Inventory-II (BDI-II) [@ciharova2020; @beck1996] and State-Trait Anxiety Inventory (STAI) [@spielberg1983; @mullner1980] were used to assess neuropsychiatric status.
 Psychotic symptoms were assessed through structured psychiatric interviews conducted by a trained psychiatrist.
@@
 
-To address study objectives, we started by repeatedly assigning each patient the diagnosis of probable PDD based on each PDD algorithm listed in @tbl-crits (see also @tbl-algos) resulting in a `r sum(data$incl)` (patients) $\times$ `r sum(critnum)` (algorithms) matrix where each cell indicates whether a patient (row) meets criteria for probable PDD according to an algorithm (column)[^6].
+To address study objectives, we started by repeatedly assigning each patient the diagnosis of probable PDD based on each PDD algorithm listed in @tbl-crits (see also @tbl-algos) resulting in a `r sum(data$incl)` (patients) $\times$ `r sum(critnum)` (algorithms) matrix where each cell indicates whether a patient (row) meets criteria for probable PDD according to an algorithm (column)[^1].
 PDD rate estimates were computed as $\frac{N_{PDD}}{N_{total}}$ separately for each algortihm.
@@
 
-[^6]: **JM: This could and should be shared most likely, as long, as we anonymize properly. Let's ask Oto Mestek and the NPO team how and if is it possible.**
+[^1]: **JM: This could and should be shared most likely, as long, as we anonymize properly. Let's ask Oto Mestek and the NPO team how and if is it possible.**
 
@@
 All analyses were conducted within the R (version `r with(version, paste(major, minor, sep="."))`) software environment for statistical computing [@rsoft].
-The software code supporting this article is available at <https://github.com/josefmana/DemCr1t.git>.[^7]
+The software code supporting this article is available at <https://github.com/josefmana/DemCr1t.git>.[^2]
 
-[^7]: **JM: Do not forget to make it public before submitted!**
+[^2]: **JM: Do not forget to make it public before submitted!**
 
@@
 Results of the analyses of prediction Accuracy, Cohen's $\kappa$, Sensitivity and Specificity are presented in @fig-acc, @fig-kappa, @fig-sens and @fig-spec respectively.
-**Numerical results are available at ...** [^8].
+**Numerical results are available at ...** [^3].
 Generally, algorithms that employed the same operationalization of IADL deficit showed substantial pairwise concordance, however, algorithms that operationalized IADL deficit differently did not.
@@
 
-[^8]: **JM: Should provide an excel sheet or html table the readers could go through by themselves.**
+[^3]: **JM: Should provide an excel sheet or html table the readers could go through by themselves.**
 
@@
 
-This study examined the diagnostic variability and validity of Level I criteria for PDD, **specifically in candidates for DBS treatment**[^9].
-By systematically comparing 68 diagnostic algorithms based on MMSE, MoCA, sMoCA, and Level II criteria, our findings highlight the significance of algorithm design.
-
-[^9]: **JM: All-right, I do not think it needs to be here either.**
+This study systematically investigated the application of multiple Level I diagnostic criteria for PDD.
+Our results show variability in PDD rate estimates, strongly influenced by the choice of cognitive screening instruments (MMSE, MoCA and sMoCA) and the operationalization of functional impairment.
+The divergence observed across algorithms demonstrates the sensitivity of diagnostic outcomes to seemingly negligible methodological choices.
 
@@
 Estimates reached lower rates when using solely FAQ item 9 (as an approximation of the pill questionnaire suggested by @dubois2007) in comparison with the full FAQ scale.
-This discrepancy highlights the diagnostic importance of IADL assessment.
-Our overall PDD rates were consistently lower compared to previous studies regarding PDD among PD patients, **demonstrating wide variability based on various criteria used**[^10].
-For instance, a retrospective study reported a PDD rate of 19.7% [@rana2012], while other clinical investigation found even higher rates, reaching up to 30% [@aarsland2005][^11].
+This discrepancy highlights the diagnostic importance of how IADLs are assessed.
+
+Our overall PDD rates were consistently lower than previous studies regarding PDD among PD patients, demonstrating wide variability based on various criteria used.
+For instance, a retrospective study reported a PDD rate of 19.7% [@rana2012], while other clinical investigation found even higher rate, reaching up to 30% [@aarsland2005].
 A recent complex meta-analysis synthesizing global data placed the expected PDD rate in PD at 26.30% [@sousa2022].
-**The discrepancy is likely attributable to our DBS-specific cohort, which differs in important characteristics, such as the younger age of the patients [@rana2012].
-/ Compared to these estimates, our study reports generally lower PDD rates, likely reflecting differences in sample characteristics, diagnostic criteria, and methodology.** [^12] [^13]
-
-[^10]: **JM: What is demonstrating this variability? Our findings or previous ones?**
-
-[^11]: **JM: There was aarsland2011 originally but I did not find such a study. Is this reference correct?**
-
-[^12]: **Because the quantity of interest is a rate and could thus be though of as a sum of binomially distributed PDD occurences divided by the total number of patients, its variance will likely systematically vary with its mean.
-Specifically, as the rate goes from extremes to 0.5, the variance increases.
-Consequently, if our estimate of the rate was lower than the true population rate, e.g., because our sample includes younger patients compared to the general PD population, our estimate of variance would also lower than the true variance of PDD rate in the general PD population.
-Nonetheless, the between-algorithm variability may not be affected by this phenomenon as unlike variability of PDD rate, we do not have reason to assume it comes about by summing independent binomial events**
-
-[^13]: \*\*JM: I prefer the second one with added sentence like: "Specifically, our cohort included candidates for DBS who are generally younger than the general PD population and age was repeatedly shown to be a strong predictor of PDD (some citations)." We should also incorporate here a statement about missing evidence of age being predictor in our sample.\*
+Compared to these estimates, our study reports generally lower PDD rates, likely reflecting differences in diagnostic criteria, methodology and sample characteristics.
+Specifically, age was repeatedly shown to be a strong predictor of PDD across studies[@rana2012; @sousa2022].
+In our sample, we did not observe age-related PDD rate within our cohort, likely due to low age variability.
 
@@
 
-Our pairwise comparisons of diagnostic algorithms showed that agreement was notably stronger among those using the same IADL operationalization **(**$\kappa$ range from 0.75, SD = 0.13 to 0.86, SD = 0.09)[^14] compared to those using different IADL definitions **(**$\kappa$ = 0.34, SD = 0.08)[^15].
-**Moreover, the agreement was slightly higher between algorithms that defined IADL deficit by FAQ item 9 compared to algorithms that defined it using the full FAQ scale.
-One possible explanation of this difference follows from the observation that algorithms using the full scale definition yielded higher PDD rate estimates.
-Because there was higher probability of being diagnosed with IADL deficit based on the full FAQ scale, there was also a bigger room for disagreement in the Cognitive Impairment status when different indexes weer used (e.g., by defining executive deficit via clock drawing vs.
-lexical fluency).**
+Pairwise comparisons of diagnostic algorithms showed that agreement was notably stronger among those using the same IADL operationalization compared to those using different IADL definitions.
+Moreover, the agreement was slightly higher between algorithms that defined IADL deficit by FAQ item 9 compared to algorithms that defined it using the full FAQ scale.
+One possible explanation of this difference follows from the observation that algorithms using the full-scale definition yielded higher PDD rate estimates.
+Because there was a higher probability of being diagnosed with IADL deficit based on the full FAQ scale, there was also a bigger room for disagreement in the Cognitive Impairment status when different indexes were used (e.g.
+by defining executive deficit via clock drawing vs.
+lexical fluency).
 
-[^14]: **JM: There is no need to repeat these numbers in this section.**
-
-[^15]: **JM: Same as above.**
-
-**Overall, algorithms using the same definition of IADL deficit showed degree of concordance equivalent to moderate (when using FAQ total score to define IADL deficit) or strong (when using FAQ item 9 to define IADL deficit) level of agreement between raters in an inter-rater reliability analysis [@mchugh2012].
-On the other hand, the concordance between algorithms using different IADL deficit definitions was equivalent to minimal agreement.** This demonstrates that even slight methodological differences can yield divergent diagnostic outcomes.
+Overall, when the same IADL definitions were used across algorithms, we observed concordance levels varying from moderate (using FAQ total score) to strong (using FAQ item 9), consistent with inter-rater reliability analysis [@mchugh2012].
+Contrarily, the concordance between algorithms using different IADL deficit definitions was equivalent to minimal agreement.
+This demonstrates that even slight methodological differences can yield divergent diagnostic outcomes.
 Such findings are critical for clinicians relying on Level I criteria for eligibility decisions, as the choice of algorithm could lead to contradictory classifications of PDD.
@@
 
-When using Level II diagnosis as the reference, MoCA-based Level I algorithms, particularly those using Clock Drawing to assess executive function, demonstrated the highest predictive accuracy.
-This supports recent proposals to modernize PDD diagnostic frameworks [@kulisevsky2024], favoring MoCA-derived components and broader functional assessment tools.
-In contrast, MMSE-based algorithms consistently underperformed, suggesting limited sensitivity in capturing the nuanced cognitive deficits characteristic of PDD.
+Using Level II diagnosis as the reference, MoCA-based Level I algorithms, particularly those using Clock Drawing to assess executive function, demonstrated the highest predictive accuracy.
+This supports recent proposals to modernize PDD diagnostic frameworks [@kulisevsky2024], favouring MoCA-derived components and broader functional assessment tools.
+In contrast, MMSE-based algorithms consistently underperformed, suggesting limited sensitivity in capturing cognitive deficits typical in PDD.
 
-The advantage of MoCA-based algorithm was evident when using the FAQ total score to define functional impairment in comparison with using solely FAQ item 9.
-In these cases, the best-performing algorithms achieved accuracy rates of 94–99%, together with high sensitivity and specificity.
-**JM: Some sMoCA praise should be added here ...**
-
-## Clinical Implications^[**JM: Is this section needed? It does not seem to say much to me. I would't be afraid to drop it.**]
-
-Our findings have immediate implications for DBS candidacy assessment.
-Given that a diagnosis of PDD typically excludes surgical intervention, inconsistent application of diagnostic criteria can lead to unbalanced access to the treatment.
-
-Moreover, the relative success of multidomain assessments, such as Level II protocols, highlights the importance of psychometric precision.
-This approach may serve as a confirmatory tool in cases where Level I assessments provide ambiguous results.
+The advantage of MoCA-based algorithm was evident in combination with using the FAQ total score, reaching accuracy rates of 94–99%, along with high sensitivity and specificity.
+Furthermore, in the algorithm using sMoCA, the raw accuracy was moderate, however, the balanced accuracy (i.e.
+combined sensitivity and specificity) was robust.
+These findings imply that simplified tools such as sMoCA may be a promising tool for heightening efficiency of the assessment while maintaining diagnostic accuracy.
 
@@
 This study’s generalizability is limited by its retrospective design and the homogeneity of the patient cohort, which may not reflect broader PD populations with varying cognitive profiles.
-**Specifically, younger age of the sample and possible lack of patients with high risk phenotypes for PDD compared to general population of people with PD might have been responsible for comparatively lower PDD rate estimates in our study.
-Nonetheless, the analysis of the concordance between different algorithms might not be affected by the sample to the same degree allowing for a broader generalization.
-Future studies could replicate our results on different cohorts of PD patients.
-To make this process easier, the code used to generate our results is publicly available and easily applicable to other data sets with a similar data structure.**
+Specifically, the younger age of the sample and possible lack of high-risk phenotypes for PDD compared to the general population of people with PD might have been responsible for comparatively lower PDD rate estimates in our study.
 
-**JM: The fact that no biomarkers were used and the IADL was at best estimated by FAQ which is somewhat archaic (comp.
-to the e-IADL thing which could be also cited and is more in line with Kulisevsky) should be mentioned.
-Also, this is a great space for an advert for NOT/DOT because the self-report of IADLs can be viewed as a limitation worth of addressing, especially since our results imply possible central role for IADL deficit definition in further work on PDD definition to achieve good reliability.**
+Nevertheless, the analysis of the algorithm concordance might not be affected by the sample to the same degree allowing for a broader generalization.
+Future studies should aim to replicate our results on larger and different cohorts of PD patients.
+To make this process easier, the code used to generate our results is publicly available and easily applicable to similarly structured data.
 
-Future research should focus on the validation of updated Level I criteria using longitudinal data and neurobiological markers.
-Furthermore, the development of adaptive diagnostic algorithms that can combine performance across cognitive and functional domains may enhance diagnostic sensitivity while reducing the risk of **over-classification of PDD**^[**JM: What's an over-classification of PDD?**].
+Furthermore, another limitation is the use of the FAQ questionnaire for IADL assessment.
+FAQ is highly subjective and informant-reliant, which may result in biased data.
+We recommend the use of more objective, modernized tools for IADl assessment, such as performance-based methods[@schmitter-edgecombe2020] or questionnaire adaptation including questions regarding gadgets use and digital literacy [@postema2024].
+Our results emphasise the importance of IADL deficit for the PDD diagnosis, therefore, we advocate for using more reliable tools with high ecological validity.
 
@@
 
-This study systematically investigated the application of multiple Level I diagnostic criteria for PDD within a cohort of patients considered for DBS treatment.
-Our results show variability in PDD rate estimates, strongly influenced by the choice of cognitive screening instruments and the operationalization of functional impairment.
-The divergence observed across algorithms demonstrates the sensitivity of diagnostic outcomes to seemingly negligible methodological choices.^[**JM: This sounds as a great opening to the discussion, rather than conclusions :-)**]
+Our study highlights the variability in PDD classification across Level I diagnostic algorithms, heavily influenced by IADL operationalization and the choice of cognitive screening tools.
+The findings support the call for a change of the current diagnostic criteria [@kulisevsky2024], favouring the use of MoCA-based components and comprehensive IADL assessments.
 
-The **proposed changes**^[**JM: They did not propose anything though and from what I gather, the authors stress this a lot - that their article in not a proposal of new criteria, only a position paper.**] of the current diagnostic criteria [@kulisevsky2024] offer heightened sensitivity by proposing the use of MoCA-based components and broader IADL assessments.
 Conservative criteria, such as reliance on pill questionnaire (i.e.
-FAQ item 9 equivalence), may fail to detect meaningful functional decline and thus under-identify true cases of PDD.
-**On the other hand, such criteria show slightly higher agreement between their different forms.**
+FAQ item 9 equivalence), may fail to detect functional decline and thus under-identify true cases of PDD.
+Importantly, concordance across algorithms rises significantly, reaching moderate to high values, when the same definition of IADL is used (either FAQ total or FAQ item 9).
+Moreover, when using MoCA-based algorithms instead of MMSE-based ones, we can observe better approximations to the Level II battery.
 
-Conclusively, this study contributes to the improvement of PDD diagnostics by offering **evidence for the clarification of Level I criteria**^[**JM: What is meant by clarification of Level I criteria?
-Also, how does our study offer evidence for such thing?**] and **emphasizing the value of psychometric precision in clinical neuropsychology**^[**JM: This is independent of our study.**].
-Future research should validate the algorithms longitudinally and explore possible integration of biomarkers to further improve diagnostic reliability.
+Future research should validate these strategies longitudinally and explore the possible integration of biomarkers, neuropsychiatric variables (e.g.
+anxiety profile) and modernised IADL methods to further improve diagnostic reliability, accuracy and consistency of PDD diagnostics.
 
@@
 Estimates relating to RO1.1 and RO1.3 cannot be safely generalized beyond a population of PD patients that are candidates for DBS due to the systematic differences between DBS candidates pool and general PD population (such as the lower age of DBS candidates compared to the general PD population).
-On the other hand, the estimates relating to RO1.4 (and to a lesser degree to RO1.2[^16]) may not be substantially influenced by the sample at hand as the primary source of their variance might come from variability in measures employed (e.g., MMSE vs MoCA to assess global cognitive performance) rather than variability in patients' performance.
+On the other hand, the estimates relating to RO1.4 (and to a lesser degree to RO1.2[^4]) may not be substantially influenced by the sample at hand as the primary source of their variance might come from variability in measures employed (e.g., MMSE vs MoCA to assess global cognitive performance) rather than variability in patients' performance.
 Assuming that there is no substantial Differential Item Functioning for DBS candidates compared to a broader population of patients with PD, the estimates relating to RO1.4 can be cautiously generalized beyond the current sample.
 
-[^16]: Because the quantity of interest is a rate and could thus be though of as a sum of binomially distributed PDD occurences divided by the total number of patients, its variance will likely systematically vary with its mean.
+[^4]: Because the quantity of interest is a rate and could thus be though of as a sum of binomially distributed PDD occurences divided by the total number of patients, its variance will likely systematically vary with its mean.
 Specifically, as the rate goes from extremes to 0.5, the variance increases.

### 2025-05-14 _manuscript/manuscript.qmd
Changes for the final crunching up of the article by Martaska.

@@
 
-To address study objectives, we started by repeatedly assigning each patient the diagnosis of probable PDD based on each PDD algorithm listed in @tbl-crits (see also @tbl-algos) resulting in a `r sum(data$incl)` (patients) $\times$ `r sum(critnum)` (algorithms) matrix where each cell indicates whether a patient (row) meets criteria for probable PDD according to an algorithm (column)[^1].
+To address study objectives, we started by repeatedly assigning each patient the diagnosis of probable PDD based on each PDD algorithm listed in @tbl-crits (see also @tbl-algos) resulting in a `r sum(data$incl)` (patients) $\times$ `r sum(critnum)` (algorithms) matrix where each cell indicates whether a patient (row) meets criteria for probable PDD according to an algorithm (column).
 PDD rate estimates were computed as $\frac{N_{PDD}}{N_{total}}$ separately for each algortihm.
 The predictive value of age and sex was then evaluated by fitting a set of logistic regressions, one for each algorithm for probable PDD, whereby the probable PDD was predicted by age, sex and their interaction.
-
-[^1]: **JM: This could and should be shared most likely, as long, as we anonymize properly. Let's ask Oto Mestek and the NPO team how and if is it possible.**
 
@@
   )
+
+# Add Kappa summaries of algos with same everything but different IADL:
+kappa_iadl <-
+  data.frame(
+    predictor = c("Lvl.II (1)", "MMSE (1)", "MMSE (3)", paste0("MoCA (",seq(1,59,2),")"), "sMoCA (1)"),
+    reference = c("Lvl.II (2)", "MMSE (2)", "MMSE (4)", paste0("MoCA (",seq(2,60,2),")"), "sMoCA (2)")
+  ) |>
+    left_join(conc$table) |>
+    select(Kappa_raw) |>
+    pull()
+
+kappa_iadl_sums <- paste0(do_summary(x = kappa_iadl, 2, "M"),", SD = ", do_summary(x = kappa_iadl, 2, "SD"))
 ```
@@
 Generally, algorithms that employed the same operationalization of IADL deficit showed substantial pairwise concordance, however, algorithms that operationalized IADL deficit differently did not.
-Whereas among algorithms with identical IADL deficit operationalization, the agreement judged by Cohen's $\kappa$ was moderately high (operationalization by FAQ total score: $\kappa$ = `r kappa_sums["faq_tot"]`; operationalization by FAQ item 9: $\kappa$ = `r kappa_sums["faq_9"]`), among algorithms that differ in IADL deficit operationalization it was low ($\kappa$ = `r kappa_sums["cross"]`).
+Whereas among algorithms with identical IADL deficit operationalization, the agreement judged by Cohen's $\kappa$ was moderately high (operationalization by FAQ total score: $\kappa$ = `r kappa_sums["faq_tot"]`; operationalization by FAQ item 9: $\kappa$ = `r kappa_sums["faq_9"]`), among algorithms that differ in IADL deficit operationalization but are otherwise identical it was low: $\kappa$ = `r kappa_iadl_sums`.
 
@@
 This study systematically investigated the application of multiple Level I diagnostic criteria for PDD.
-Our results show variability in PDD rate estimates, strongly influenced by the choice of cognitive screening instruments (MMSE, MoCA and sMoCA) and the operationalization of functional impairment.
+Our results show variability in PDD rate estimates, strongly influenced by the choice of cognitive screening instrument (MMSE, MoCA and sMoCA) and the operationalization of functional impairment.
 The divergence observed across algorithms demonstrates the sensitivity of diagnostic outcomes to seemingly negligible methodological choices.
@@
 
-Our results showed a wide range in estimated PDD rate across algorithms, ranging from 2.00% to 16.75%.
+```{r}
+#| label: rates-minmax
+
+rate_min <- sub("-.*", "", rates_overall["minmax"])
+rate_max <- sub(".*-", "", rates_overall["minmax"])
+```
+
+Our results showed a wide range in estimated PDD rate across algorithms, ranging from `r rate_min`% to `r rate_max`%.
 Estimates reached lower rates when using solely FAQ item 9 (as an approximation of the pill questionnaire suggested by @dubois2007) in comparison with the full FAQ scale.
@@
 For instance, a retrospective study reported a PDD rate of 19.7% [@rana2012], while other clinical investigation found even higher rate, reaching up to 30% [@aarsland2005].
-A recent complex meta-analysis synthesizing global data placed the expected PDD rate in PD at 26.30% [@sousa2022].
+A recent meta-analysis synthesizing global data placed the expected PDD rate in PD at 26.30% [@sousa2022].
 Compared to these estimates, our study reports generally lower PDD rates, likely reflecting differences in diagnostic criteria, methodology and sample characteristics.
-Specifically, age was repeatedly shown to be a strong predictor of PDD across studies[@rana2012; @sousa2022].
-In our sample, we did not observe age-related PDD rate within our cohort, likely due to low age variability.
+Specifically, our sample was younger compared to other PD cohorts and age was repeatedly shown to be a strong predictor of PDD across studies [@rana2012; @sousa2022].
+
+Interestingly, we did not observe any reliable age-related differences in PDD rate within our cohort.
+This lack of age-dependency may, however, be also stemming from the relatively younger age of our cohort, because previous reports indicate that association between age and PDD is not linear but increases with age and may not reach substantial values before the age of 70 [@rana2012].
 
@@
 
-The advantage of MoCA-based algorithm was evident in combination with using the FAQ total score, reaching accuracy rates of 94–99%, along with high sensitivity and specificity.
+The advantage of MoCA-based algorithm was evident in combination with using the FAQ total score, reaching accuracy rates of 94–99%.
 Furthermore, in the algorithm using sMoCA, the raw accuracy was moderate, however, the balanced accuracy (i.e.
-combined sensitivity and specificity) was robust.
+combined sensitivity and specificity) was high.
 These findings imply that simplified tools such as sMoCA may be a promising tool for heightening efficiency of the assessment while maintaining diagnostic accuracy.
@@
 
-This study’s generalizability is limited by its retrospective design and the homogeneity of the patient cohort, which may not reflect broader PD populations with varying cognitive profiles.
+This study’s generalizability is limited by its retrospective design and the homogeneity of the patient cohort, which does not reflect broader PD populations with varying cognitive profiles.
 Specifically, the younger age of the sample and possible lack of high-risk phenotypes for PDD compared to the general population of people with PD might have been responsible for comparatively lower PDD rate estimates in our study.
@@
 Furthermore, another limitation is the use of the FAQ questionnaire for IADL assessment.
-FAQ is highly subjective and informant-reliant, which may result in biased data.
-We recommend the use of more objective, modernized tools for IADl assessment, such as performance-based methods[@schmitter-edgecombe2020] or questionnaire adaptation including questions regarding gadgets use and digital literacy [@postema2024].
-Our results emphasise the importance of IADL deficit for the PDD diagnosis, therefore, we advocate for using more reliable tools with high ecological validity.
+FAQ is subjective and informant-reliant, which may result in biased data.
+For future research, we suggest the use of more objective, modernized tools for IADl assessment, such as performance-based methods [@schmitter-edgecombe2020] or questionnaire adaptation including questions regarding gadgets use and digital literacy [@postema2024].
+Our results emphasize the importance of IADL deficit for the PDD diagnosis, therefore, we suggest exploring more reliable tools with high ecological validity.
+
+Finally, whereas our study systematically investigated how varying definition of global deficit, impaired cognition and IADL deficit affect probable PDD  classification, it did not explore associations of PDD diagnosis with its neuropsychiatric (e.g.
+anxiety profile) and biomarker correlates.
+Future studies would benefit from focusing on these aspects of PDD.
 
@@
 Moreover, when using MoCA-based algorithms instead of MMSE-based ones, we can observe better approximations to the Level II battery.
-
-Future research should validate these strategies longitudinally and explore the possible integration of biomarkers, neuropsychiatric variables (e.g.
-anxiety profile) and modernised IADL methods to further improve diagnostic reliability, accuracy and consistency of PDD diagnostics.
 

### 2025-05-14 _manuscript/manuscript.qmd
Adding citation

@@
 lang: en
-title: "Updated Criteria for the Diagnostic Procedure for Parkinson’s Disease Dementia on Level I and their Validity in Deep Brain Stimulation Cohort"
+title: "Evaluation of Diagnostic Concordance Between Algorithms for Parkinson’s Disease Dementia"
 shorttitle: "Parkinson's Disease Dementia Level I Criteria"
@@
 format:
- apaquarto-pdf:
-    # Can be jou (journal), man (manuscript), stu (student), or doc (document)
-    documentmode: man
-    keep-tex: true
+ # apaquarto-pdf:
+   # documentmode: man
+   # keep-tex: true
  apaquarto-docx: 
@@
 All analyses were conducted within the R (version `r with(version, paste(major, minor, sep="."))`) software environment for statistical computing [@rsoft].
-The software code supporting this article is available at <https://github.com/josefmana/DemCr1t.git>.[^2]
+The software code supporting this article is available at <https://github.com/josefmana/DemCr1t.git>.[^1]
 
-[^2]: **JM: Do not forget to make it public before submitted!**
+[^1]: **JM: Do not forget to make it public before submitted!**
 
@@
 Results of the analyses of prediction Accuracy, Cohen's $\kappa$, Sensitivity and Specificity are presented in @fig-acc, @fig-kappa, @fig-sens and @fig-spec respectively.
-**Numerical results are available at ...** [^3].
+**Numerical results are available at ...** [^2].
 Generally, algorithms that employed the same operationalization of IADL deficit showed substantial pairwise concordance, however, algorithms that operationalized IADL deficit differently did not.
@@
 
-[^3]: **JM: Should provide an excel sheet or html table the readers could go through by themselves.**
+[^2]: **JM: Should provide an excel sheet or html table the readers could go through by themselves.**
 
@@
 Estimates relating to RO1.1 and RO1.3 cannot be safely generalized beyond a population of PD patients that are candidates for DBS due to the systematic differences between DBS candidates pool and general PD population (such as the lower age of DBS candidates compared to the general PD population).
-On the other hand, the estimates relating to RO1.4 (and to a lesser degree to RO1.2[^4]) may not be substantially influenced by the sample at hand as the primary source of their variance might come from variability in measures employed (e.g., MMSE vs MoCA to assess global cognitive performance) rather than variability in patients' performance.
+On the other hand, the estimates relating to RO1.4 (and to a lesser degree to RO1.2[^3]) may not be substantially influenced by the sample at hand as the primary source of their variance might come from variability in measures employed (e.g., MMSE vs MoCA to assess global cognitive performance) rather than variability in patients' performance.
 Assuming that there is no substantial Differential Item Functioning for DBS candidates compared to a broader population of patients with PD, the estimates relating to RO1.4 can be cautiously generalized beyond the current sample.
 
-[^4]: Because the quantity of interest is a rate and could thus be though of as a sum of binomially distributed PDD occurences divided by the total number of patients, its variance will likely systematically vary with its mean.
+[^3]: Because the quantity of interest is a rate and could thus be though of as a sum of binomially distributed PDD occurences divided by the total number of patients, its variance will likely systematically vary with its mean.
 Specifically, as the rate goes from extremes to 0.5, the variance increases.

### 2025-05-15 _manuscript/manuscript.qmd
Added financial support

@@
     toc: false
+author-note:
+  disclosures:
+    financial-support: "Supported by The project National Institute for Neurological Research (Programme EXCELES, ID Project No.
+LX22NPO5107) - Funded by the European Union – Next Generation EU; Charles University: Cooperatio Program in Neuroscience; General University Hospital in Prague project MH CZ-DRO-VFN64165."
 bibliography: bibliography.bib
@@
 Clinical records spanning `r dates["min"]` to `r dates["max"]` were examined.
-All participants were candidates for for Deep Brain Stimulation (DBS) treatment and underwent neuropsychological evaluation conducted by a trained clinical psychologist (OB) as part of standard preoperative assessments forDBS eligibility at the General University Hospital in Prague.
+All participants were candidates for Deep Brain Stimulation (DBS) treatment and underwent neuropsychological evaluation conducted by a trained clinical psychologist (OB) as part of standard preoperative assessments for DBS eligibility at the General University Hospital in Prague.
 

### 2025-06-25 _manuscript/manuscript.qmd
Put manuscript rendering directly inside the _targets script.

@@
 vars <- tar_read(variables, store = upstore)
-
 ```
@@
 
-dates <-
-  sapply(
-    c("min", "max"),
-    function(f) {
+dates <- sapply(
+  c("min", "max"), function(f) {
     date <- do.call(f, list(data$assdate))
@@
 ```{r}
-#| label: crit-check
+#| label: "crit-check"
 
-critnum <-
-  sapply(
-    c("mmse", "moca", "smoca", "lvlII"),
-    function(i)
+critnum <- sapply(
+  c("mmse", "moca", "smoca", "lvlII"), function(i) {
     subset(crit, group == i) |>
       nrow()
+  }
 )
@@
 ```{r}
-#| label: tbl-crits
+#| label: "tbl-crits"
 #| tbl-cap: "Summary of probable PDD operationalizations compared in the study."
@@
 ```{r}
-#| label: results
+#| label: "results"
 
 desc <- tar_read(sample_description, store = upstore)
-prev <- tar_read(prevalence_summaries, store = upstore)
+rates <- tar_read(rate_summaries, store = upstore)
 pred <- tar_read(demographic_predictors, store = upstore)
 conc <- tar_read(concordance_statistics, store = upstore)
-perc <- prev$table$perc
+perc <- rates$table$perc
 
@@
 rates_overall <- sapply(c("M", "SD", "Md", "minmax"), function(f) do_summary(perc, 2, f))
-rates_FAQ9 <- sapply(c("M", "SD", "Md", "minmax"), function(f) do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, f))
-rates_FAQtot <- sapply(c("M", "SD", "Md", "minmax"), function(f) do_summary(perc[prev$table$IADL == "FAQ > 7"], 2, f))
+rates_FAQ9 <- sapply(c("M", "SD", "Md", "minmax"), function(f) do_summary(perc[grepl("FAQ 9", rates$table$IADL)], 2, f))
+rates_FAQtot <- sapply(c("M", "SD", "Md", "minmax"), function(f) do_summary(perc[rates$table$IADL == "FAQ > 7"], 2, f))
 
-# Minimum p-value of logistic regressions
-min_p <- subset(pred$values, quantity == "p-value")$estimate |> min() |> do_summary(3, "p")
+# Minimum p-value of logistic regressions:
+min_p <-
+  subset(pred$values, quantity == "p-value")$estimate |>
+  min() |>
+  do_summary(3, "p")
+
+# List of algorithms:
+algos <- list(
+  faq_tot = subset(rates$table, IADL == "FAQ > 7")$type,
+  faq_9 = subset(rates$table, IADL == "FAQ 9 > 1")$type
+)
+
+continue <- !(any(algos$faq_tot %in% algos$faq_9) || any(algos$faq_9 %in% algos$faq_tot))
+stopifnot("Some algorithm(s) use both IADL operationalisations!" = continue)
 ```
@@
 ```{r}
-#| label: tbl-desc
+#| label: "tbl-desc"
 #| tbl-cap: "Cognitive characteristics of the sample."
@@
 ```{r}
-#| label: fig-rates
-#| fig-cap: Summary of the estimates of probable PDD rate.
-#| apa-note: Vertical lines represent estimates arrived at by using sMoCA (dotted) or Level II (dashed) with FAQ item 9 (orange) or FAQ total score (blue) as criteria for probable PDD.
+#| label: "fig-rates"
+#| fig-cap: "Summary of the estimates of probable PDD rate."
+#| apa-note: "Vertical lines represent estimates arrived at by using sMoCA (dotted) or Level II (dashed) with FAQ item 9 (orange) or FAQ total score (blue) as criteria for probable PDD."
 
-prev$plot
+rates$plot
 ```
@@
 ```{r}
-#| label: kappas
-
-algos <- list(
-  faq_tot = subset(prev$table, IADL == "FAQ > 7")$type,
-  faq_9 = subset(prev$table, IADL == "FAQ 9 > 1")$type
-)
-
-# Check for overlap:
-continue <- !(any(algos$faq_tot %in% algos$faq_9) || any(algos$faq_9 %in% algos$faq_tot))
-stopifnot("Some algorithm(s) use both IADL operationalisations!" = continue)
+#| label: "kappas"
 
 # Prepare combinations within IADL operationalisation:
-kappas <-
-  lapply(
-    set_names(names(algos)),
-    function(o)
+kappas <- lapply(
+  set_names(names(algos)), function(o) {
     combn(algos[[o]], 2) |>
@@
       pull()
+  }
 )
@@
 ```{r}
-#| label: fig-acc
-#| fig-cap: Prediction accuracy matrix.
-#| apa-note: The matrix depicts classification accuracy of algorithms for PDD depicted on x-axis in predicting outcomes based on algorithms on the y-axis.
+#| label: "fig-acc"
+#| fig-cap: "Prediction accuracy matrix."
+#| apa-note: "The matrix depicts classification accuracy of algorithms for PDD depicted on x-axis in predicting outcomes based on algorithms on the y-axis.
 Algorithms printed in blue defined IADL deficit by FAQ total score, algorithms printed in black defined IADL deficit by FAQ item 9 response.
-Cases with asterisk indicate predictive accuracy statistically significantly higher than the No Information Rate.
+Cases with asterisk indicate predictive accuracy statistically significantly higher than the No Information Rate."
 
@@
 ```{r}
-#| label: lvlII-stats
+#| label: "lvlII-stats"
 
 lvlII_rates <- c(
-  faq_tot = subset(prev$table, type == "Lvl.II (1)")$perc |> do_summary(2) |> paste0("%"),
-  faq_9 = subset(prev$table, type == "Lvl.II (2)")$perc |> do_summary(2) |> paste0("%")
+  faq_tot = subset(rates$table, type == "Lvl.II (1)")$perc |> do_summary(2) |> paste0("%"),
+  faq_9 = subset(rates$table, type == "Lvl.II (2)")$perc |> do_summary(2) |> paste0("%")
 )
@@
 ```{r}
-#| label: tbl-approx
+#| label: "tbl-approx"
 #| tbl-cap: "Level I algorithms for probable PDD as predictors of Level II classification as the reference."
@@
 ```{r}
-#| label: rates-minmax
+#| label: "rates-minmax"
 
@@
 ```{r}
-#| label: tbl-algos
-#| tbl-cap: Summary of all algortihms for probable PDD used in the study.
+#| label: "tbl-algos"
+#| tbl-cap: "Summary of all algortihms for probable PDD used in the study."
 
-prev$gtables$algorithms
+rates$gtables$algorithms
 ```
@@
 ```{r}
-#| label: tbl-estimands
-#| tbl-cap: Mapping between research objectives and quantities of interest in the current study.
+#| label: "tbl-estimands"
+#| tbl-cap: "Mapping between research objectives and quantities of interest in the current study."
 
@@
 ```{r}
-#| label: tbl-rates
-#| tbl-cap: Estimates of the rate of probable PDD in the sample.
+#| label: "tbl-rates"
+#| tbl-cap: "Estimates of the rate of probable PDD in the sample."
 
-prev$gtables$rates
+rates$gtables$rates
 ```
@@
 ```{r}
-#| label: tbl-approx-bal
-#| tbl-cap: Level I algorithms for probable PDD as predictors of Level II classification as the reference arranged by their balanced accuracy score.
+#| label: "tbl-approx-bal"
+#| tbl-cap: "Level I algorithms for probable PDD as predictors of Level II classification as the reference arranged by their balanced accuracy score."
 
@@
 ```{r}
-#| label: fig-data
-#| fig-cap: Representation of study data.
-#| apa-note: The figure shows whether patients (x-axis) ordered from the youngest (left) to the oldest (right) were classified as probable PDD by each tested algorithm (y-axis) ordered from the one with the lowest (bottom) to the highest (top) PDD rate estimate.
+#| label: "fig-data"
+#| fig-cap: "Representation of study data."
+#| apa-note: "The figure shows whether patients (x-axis) ordered from the youngest (left) to the oldest (right) were classified as probable PDD by each tested algorithm (y-axis) ordered from the one with the lowest (bottom) to the highest (top) PDD rate estimate.
 Patients printed in red are women, patients printed in blue are men.
-Red cells indicate probable PDD diagnosis, grey cells indicate non-PDD diagnosis and white cells indicate missing diagnosis.
+Red cells indicate probable PDD diagnosis, grey cells indicate non-PDD diagnosis and white cells indicate missing diagnosis."
 
@@
 ```{r}
-#| label: fig-pars
-#| fig-cap: Summary of logistic regressions parameters prediction probable PDD by age and sex.
-#| apa-note: Histograms represent odd ratio (OR) estimates and p-values associated with age, sex, and their interaction as predictors of each of the 68 probable PDD classification.
+#| label: "fig-pars"
+#| fig-cap: "Summary of logistic regressions parameters prediction probable PDD by age and sex."
+#| apa-note: "Histograms represent odd ratio (OR) estimates and p-values associated with age, sex, and their interaction as predictors of each of the 68 probable PDD classification.
 In the case of parameters for sex, values higher than 20 were omitted for clarity.
-Vertical lines indicate OR = 1 and p = .05.
+Vertical lines indicate OR = 1 and p = .05."
 
@@
 ```{r}
-#| label: fig-kappa
-#| fig-cap: Cohen's κ matrix.
-#| apa-note: The matrix depicts Cohen's κ measuring agreement between algorithms for PDD.
-Algorithms printed in red defined IADL deficit by FAQ total score, algorithms printed in black defined IADL deficit by FAQ item 9 response.
+#| label: "fig-kappa"
+#| fig-cap: "Cohen's κ matrix."
+#| apa-note: "The matrix depicts Cohen's κ measuring agreement between algorithms for PDD.
+Algorithms printed in red defined IADL deficit by FAQ total score, algorithms printed in black defined IADL deficit by FAQ item 9 response."
 
@@
 ```{r}
-#| label: fig-sens
-#| fig-cap: Sensitivity matrix.
-#| apa-note: The matrix depicts sensitivity of algorithms for PDD depicted on x-axis in predicting outcomes based on algorithms on the y-axis.
-Algorithms printed in blue defined IADL deficit by FAQ total score, algorithms printed in black defined IADL deficit by FAQ item 9 response.
+#| label: "fig-sens"
+#| fig-cap: "Sensitivity matrix."
+#| apa-note: "The matrix depicts sensitivity of algorithms for PDD depicted on x-axis in predicting outcomes based on algorithms on the y-axis.
+Algorithms printed in blue defined IADL deficit by FAQ total score, algorithms printed in black defined IADL deficit by FAQ item 9 response."
 
@@
 ```{r}
-#| label: fig-spec
-#| fig-cap: Specificity matrix.
-#| apa-note: The matrix depicts specificity of algorithms for PDD depicted on x-axis in predicting outcomes based on algorithms on the y-axis.
-Algorithms printed in blue defined IADL deficit by FAQ total score, algorithms printed in black defined IADL deficit by FAQ item 9 response.
+#| label: "fig-spec"
+#| fig-cap: "Specificity matrix."
+#| apa-note: "The matrix depicts specificity of algorithms for PDD depicted on x-axis in predicting outcomes based on algorithms on the y-axis.
+Algorithms printed in blue defined IADL deficit by FAQ total score, algorithms printed in black defined IADL deficit by FAQ item 9 response."
 

### 2025-06-25 _targets.R

@@ -21,36 +21,40 @@ list(
   tar_files(
     name = inputs, # Monitor input data files
     command = data_paths("data-raw"),
-    format  = "file"
+    format = "file"
   ),
   tar_target(
-    name = raw_data,
-    command = prepare_data(inputs)
-  ),
-  tar_target(
-    name    = pdd_data,
-    command = diagnose_pdd_sample(raw_data)
-  ),
-  tar_target(
-    name = variables,
+    name = variables, # List variables for future use
     command = here::here("data-raw","VariablesOfInterest.csv"),
     format = "file"
   ),
   tar_target(
-    name = sample_description,
+    name = raw_data, # Pre-process input data
+    command = prepare_data(inputs)
+  ),
+  tar_target(
+    name = pdd_data, # Assign probable PDD
+    command = diagnose_pdd_sample(raw_data)
+  ),
+  tar_target(
+    name = sample_description, # Describe the sample
     command = compute_descriptives(raw_data, variables)
   ),
   tar_target(
-    name = prevalence_summaries,
+    name = rate_summaries, # Summarise PDD rates
     command = summarise_rates(pdd_data, variables)
   ),
   tar_target(
-    name = demographic_predictors,
+    name = demographic_predictors, # Regress probable PDD on demographics
     command = regress_pdd_on_demographics(raw_data, pdd_data$PDD)
   ),
   tar_target(
-    name = concordance_statistics,
+    name = concordance_statistics, # Describe concordance between different PDD algorithms
     command = describe_concordance(pdd_data)
   ),
-  NULL
+  tar_quarto(
+    name = manuscript, # Prepare the manuscript
+    path = here::here("_manuscript", "manuscript.qmd"),
+    quiet = FALSE
+  )
 )

### 2025-09-23 _manuscript/manuscript.qmd
Tidying the code.

@@
 
-dates <- sapply(
-  c("min", "max"), function(f) {
+dates <- sapply(c("min", "max"), function(f) {
   date <- do.call(f, list(data$assdate))
-    paste(month(date, label = T, abbr = F), year(date))
-  }
-)
+  paste(month(date, label = TRUE, abbr = FALSE), year(date))
+})
 ```
@@
 
-critnum <- sapply(
-  c("mmse", "moca", "smoca", "lvlII"), function(i) {
+critnum <- sapply(c("mmse", "moca", "smoca", "lvlII"), function(i) {
   subset(crit, group == i) |>
     nrow()
-  }
-)
+})
 
@@
 # In-text rate estimates
-rates_overall <- sapply(c("M", "SD", "Md", "minmax"), function(f) do_summary(perc, 2, f))
-rates_FAQ9 <- sapply(c("M", "SD", "Md", "minmax"), function(f) do_summary(perc[grepl("FAQ 9", rates$table$IADL)], 2, f))
-rates_FAQtot <- sapply(c("M", "SD", "Md", "minmax"), function(f) do_summary(perc[rates$table$IADL == "FAQ > 7"], 2, f))
+rates_overall <- sapply(c("M", "SD", "Md", "minmax"), \(f) do_summary(perc, 2, f))
+rates_FAQ9 <- sapply(c("M", "SD", "Md", "minmax"), \(f) do_summary(perc[grepl("FAQ 9", rates$table$IADL)], 2, f))
+rates_FAQtot <- sapply(c("M", "SD", "Md", "minmax"), \(f) do_summary(perc[rates$table$IADL == "FAQ > 7"], 2, f))
 
 # Minimum p-value of logistic regressions:
-min_p <-
-  subset(pred$values, quantity == "p-value")$estimate |>
+min_p <- subset(pred$values, quantity == "p-value")$estimate |>
   min() |>
@@
 # Prepare combinations within IADL operationalisation:
-kappas <- lapply(
-  set_names(names(algos)), function(o) {
+kappas <- lapply(set_names(names(algos)), function(o) {
   combn(algos[[o]], 2) |>
@@
       `colnames<-`(c("predictor", "reference")) |>
-      left_join(conc$table, by = c("predictor", "reference")) |>
-      select(Kappa_raw) |>
-      pull()
-  }
-)
+      left_join(conc$table, by = join_by(predictor, reference)) |>
+      pull(Kappa_raw)
+})
 
@@
 # Extract Kappa summaries:
-kappa_sums <-
-  sapply(
-    names(kappas),
-    function(i)
+kappa_sums <- sapply(names(kappas), function(i) {
   paste0(do_summary(kappas[[i]], 2, "M"),", SD = ", do_summary(kappas[[i]], 2, "SD"))
-  )
+})
 
 # Add Kappa summaries of algos with same everything but different IADL:
-kappa_iadl <-
-  data.frame(
+kappa_iadl <- data.frame(
   predictor = c("Lvl.II (1)", "MMSE (1)", "MMSE (3)", paste0("MoCA (",seq(1,59,2),")"), "sMoCA (1)"),
@@
   left_join(conc$table) |>
-    select(Kappa_raw) |>
-    pull()
+  pull(Kappa_raw)
 

### 2025-09-23 R/assign_cognitive_impairment.R

@@ -42,64 +42,58 @@
 #' @export
 assign_cognitive_impairment <- function(d0, calc, map, output = "both") {
   # Read map if it is a file
-  if(is.character(map)) {
+  if (is.character(map)) {
     m <- readr::read_delim(map, delim = ";", col_types = cols())
   } else {
     m <- map
   }
   # Read regression calculator file:
-  r <-
-    readxl::read_excel(calc, sheet = "equations", skip = 1) |>
+  r <- readxl::read_excel(calc, sheet = "equations", skip = 1) |>
     rename("calc_lab" = "...1") |>
     right_join(m, by = "calc_lab")
   # Prepare thresholds for BNT-60:
-  bnt_thresh <-
-    data.frame(
-      age_bottom = c(0, 0, 60, 60),
-      age_top    = c(60, 60, Inf, Inf),
-      edu_bottom = c(0, 12, 0, 12),
-      edu_top    = c(12, Inf, 12, Inf),
-      threshold  = c(49, 52, 50, 53)
-    )
+  bnt_thresh <- data.frame(
+    age_bottom = c(0, 0, 60, 60),
+    age_top    = c(60, 60, Inf, Inf),
+    edu_bottom = c(0, 12, 0, 12),
+    edu_top    = c(12, Inf, 12, Inf),
+    threshold  = c(49, 52, 50, 53)
+  )
   # Add CI-related values to data:
-  d1 <-
-    d0 |>
+  d1 <- d0 |>
     mutate(
-      gender = as.numeric(sex)-1, # to get man = 1, women = 0
+      gender = as.numeric(sex) - 1, # to get man = 1, women = 0
       across(
         .cols  = all_of(r$data_lab),
-        .fns   = ~compute_z_score(r, .x, r$calc_lab[r$data_lab == cur_column()], age, gender, edu_years),
+        .fns   = \(x) compute_z_score(r, x, r$calc_lab[r$data_lab == cur_column()], age, gender, edu_years),
         .names = "z_{.col}"
       ),
       across(
         .cols  = starts_with("z_"),
-        .fns   = ~if_else(.x <= -1.5, 1, 0),
+        .fns   = \(x) if_else(x <= -1.5, 1, 0),
         .names = "mci_{.col}"
       ),
-      mci_z_bnt_60 = # Re-score BNT because the calculator is for BNT-30 not -60
-        sapply(
-          seq_len(length(mci_z_bnt_60)),
-          function(i)
-            ifelse(
[…]

### 2025-09-23 R/check_compatibility.R

@@ -24,33 +24,26 @@ check_compatibility <- function(d1, d2) {
     NULL
   )
   # Prepare a table comparing common variables:
-  tvar <-
-    sapply(
-      d1$id,
-      function(i)
-        sapply(
-          commonvars,
-          function(j) ifelse(
-            test = !(i %in% d2$id),
-            yes  = NA,
-            no   = with(d1, get(j)[id == i & incl == 1]) == with(d2, get(j)[id == i])
-          )
-        )
-    ) |>
+  tvar <- sapply(d1$id, function(i) {
+    sapply(commonvars, function(j) {
+      ifelse(
+        test = !(i %in% d2$id),
+        yes  = NA,
+        no   = with(d1, get(j)[id == i & incl == 1]) == with(d2, get(j)[id == i])
+      )
+    })
+  }) |>
     t()
   # Evaluate mismatches:
-  tdisc <- lapply(
-    set_names(commonvars),
-    function(i) left_join(
+  tdisc <- map_dfr(set_names(commonvars), function(i) {
+    left_join(
       d1[d1$id %in% rownames(tvar[!tvar[ ,i], ]) & d1$incl == 1, c("id", i)],
       d2[d2$id %in% rownames(tvar[!tvar[ ,i], ]), c("id", i)],
-      by = "id",
+      by = join_by(id),
       suffix = c("_ItemData", "_REDCap")
     )
-  ) |>
-    reduce(full_join, by = "id")
+  })
   # Return mismatches as a data frame
-  d1 |>
-    select(id, surname, firstname, assdate) |>
+  d1 |> select(id, surname, firstname, assdate) |>
     right_join(tdisc, by = "id")
 }

### 2025-09-23 R/check_names.R

@@ -20,36 +20,29 @@
 #' data <- prepare_data(p, check.names = FALSE)
 #' }
 check_names <- function(d, nms) {
-  nms <-
-    nms |>
+  nms <- nms |>
     filter(study_id %in% d$id) |>
     column_to_rownames("study_id") |>
     select(jmeno, prijmeni) |>
-    mutate_all(~make_clean_names(., allow_dupes = T))
-  tnam <-
-    sapply(
-      rownames(nms),
-      function(i) c(
-        forname = nms[i, "jmeno"] == subset(d, id == i & incl == 1)$firstname,
-        surname = nms[i, "prijmeni"] == subset(d, id == i & incl == 1)$surname
-      )
-    ) |>
+    mutate_all(\(x) janitor::make_clean_names(x, allow_dupes = TRUE))
+  tnam <- sapply(rownames(nms), function(i) {
+    c(forname = nms[i, "jmeno"] == subset(d, id == i & incl == 1)$firstname,
+      surname = nms[i, "prijmeni"] == subset(d, id == i & incl == 1)$surname
+    )
+  }) |>
     t()
   discid <- rownames(tnam[(!tnam[ ,1] | !tnam[ ,2]), ])
-  disctab <-
-    left_join(
-      nms[discid, ] |> rownames_to_column("id"),
-      d[d$id %in% discid, c("id", "firstname", "surname")],
-      by = "id"
-    ) |>
-    mutate(
-      reason = if_else(
-        condition = id == "IPN143",
-        true = "married",
-        false = "typo"
-      )
-    )
-  if(nrow(disctab) > 0) {
+  disctab <- left_join(
+    nms[discid, ] |> rownames_to_column("id"),
+    d[d$id %in% discid, c("id", "firstname", "surname")],
+    by = join_by(id)
+  ) |>
+    mutate(reason = if_else(
+      condition = id == "IPN143",
+      true = "married",
+      false = "typo"
+    ))
+  if (nrow(disctab) > 0) {
     cat("Please, check whether the reasons of name discrepancies
 of patients below are valid. If not, revise the data.\n\n")
     print(disctab)

### 2025-09-23 R/check_ranges.R

@@ -58,16 +58,15 @@ check_ranges <- function(d) {
   )
   #zerotrunc <- c("vf_k", "vf_s", "tmt_a", "pst_d", "tmt_b", "pst_w", "pst_c", "cf", "gp_r", "gp_l")
   zerotrunc <- c("tmt_a", "pst_c", "vf_animals")
-  for(i in zerotrunc) mistakes[[i]] <- subset(d, get(i) < 0)
-  for(i in paste0("faq_", seq_len(10))) mistakes[[i]] <- subset(d, !(get(i) %in% seq0(3)))
-  stop <- F
+  for (i in zerotrunc) mistakes[[i]] <- subset(d, get(i) < 0)
+  for (i in paste0("faq_", seq_len(10))) mistakes[[i]] <- subset(d, !(get(i) %in% seq0(3)))
+  stop <- FALSE
   for(i in names(mistakes)) {
-    mistakes[[i]] <-
-      mistakes[[i]] |>
+    mistakes[[i]] <- mistakes[[i]] |>
       select(all_of(c("id", "surname", "firstname", "assdate", i))) |>
       filter(!is.na(get(i)))
     if(nrow(mistakes[[i]]) > 0) {
-      stop <- T
+      stop <- TRUE
     }
   }
   # Return the results:

### 2025-09-23 R/compute_descriptives.R

@@ -35,60 +35,51 @@ compute_descriptives <- function(d0, vois) {
     v <- vois
   }
   # Prepare a demography table:
-  demtab <-
-    sapply(
-      seq_len(nrow(v)),
-      function(i) c(
-        variable = pull(v[i, 2]),
-        group    = pull(v[i, 4]),
-        N  = case_when(
-          pull(v[i, 3]) == "binary" ~ do_summary(d0[ , pull(v[i, 1])], 0, "Nperc"),
-          pull(v[i, 3]) == "nominal" ~ do_summary(d0[ , pull(v[i, 1])], 0, "Nslash"),
-          pull(v[i, 3]) == "continuous" ~ do_summary(d0[ , pull(v[i, 1])], 0, "N")
-        ),
-        sapply(
-          c("Md","minmax","M","SD"),
-          function(f) ifelse(
-            test = pull(v[i, 3]) == "continuous",
-            yes = ifelse(
-              test = pull(v[i, 4]) %in% c("Attention and Working Memory", "Executive Function", "Language", "Memory", "Visuospatial Function"),
-              yes = do_summary(pull(d0[ , pull(v[i, 1])]), 2, f),
-              no = do_summary(pull(d0[ , pull(v[i, 1])]), ifelse(f %in% c("M","SD"), 2, 0), f)
-            ),
-            no   = "-"
-          )
+  demtab <- map_df(seq_len(nrow(v)), function(i) {
+    c(
+      variable = pull(v[i, 2]),
+      group = pull(v[i, 4]),
+      N = case_when(
+        pull(v[i, 3]) == "binary" ~ do_summary(d0[ , pull(v[i, 1])], 0, "Nperc"),
+        pull(v[i, 3]) == "nominal" ~ do_summary(d0[ , pull(v[i, 1])], 0, "Nslash"),
+        pull(v[i, 3]) == "continuous" ~ do_summary(d0[ , pull(v[i, 1])], 0, "N")
+      ),
+      sapply(c("Md","minmax","M","SD"), function(f) {
+        ifelse(
+          test = pull(v[i, 3]) == "continuous",
+          yes = ifelse(
+            test = pull(v[i, 4]) %in% c("Attention and Working Memory", "Executive Function", "Language", "Memory", "Visuospatial Function"),
+            yes = do_summary(pull(d0[ , pull(v[i, 1])]), 2, f),
+            no = do_summary(pull(d0[ , pull(v[i, 1])]), ifelse(f %in% c("M","SD"), 2, 0), f)
+          ),
[…]

### 2025-09-23 R/compute_z_score.R

@@ -27,13 +27,11 @@
 #' }
 #' @export
 compute_z_score <- function(calc, x, nam, AGE, GEN, EDU) {
-  with(
-    calc, {
-      pars <- as.numeric(c(Constant[calc_lab == nam], age[calc_lab == nam], gender[calc_lab == nam], education[calc_lab == nam]))
-      data <- as.matrix(cbind(rep(1, length(x)), AGE, GEN, EDU))
-      x_bar <- data %*% pars
-      z <- c(sign[calc_lab == nam] * (x - x_bar) / RMSE[calc_lab == nam])
-      z
-    }
-  )
+  with(calc, {
+    pars <- as.numeric(c(Constant[calc_lab == nam], age[calc_lab == nam], gender[calc_lab == nam], education[calc_lab == nam]))
+    data <- as.matrix(cbind(rep(1, length(x)), AGE, GEN, EDU))
+    x_bar <- data %*% pars
+    z <- c(sign[calc_lab == nam] * (x - x_bar) / RMSE[calc_lab == nam])
+    z
+  })
 }

### 2025-09-23 R/data_paths.R

@@ -33,11 +33,11 @@ data_paths <- function(dir = "data-raw") {
       "VariablesOfInterest.csv" # Variables to be printed in tables
     ),
     message = c(
-      paste0("\nItem-wise data is missing or mislabelled!\nInsert a valid file 'ItemData.csv' into the /", dir, " folder!\n"),
-      paste0("\nREDCap data is missing or mislabelled!\nInsert a valid file 'REDCapData.csv' into the /", dir, " folder!\n"),
-      paste0("\nPatient identification data is missing or mislabelled!\nInsert a valid file 'MetaData.csv' into the /", dir, " folder!\n"),
-      paste0("\nFile with test scoring data is missing or mislabelled!\nInsert a valid file 'TestScoring.csv' into the /", dir, " folder!\n"),
-      paste0("\nFile with variable labels is missing or mislabelled!\nInsert a valid file 'VariablesOfInterest.csv' into the /", dir, " folder!\n")
+      glue::glue("\nItem-wise data is missing or mislabelled!\nInsert a valid file 'ItemData.csv' into the /{dir} folder!\n"),
+      glue::glue("\nREDCap data is missing or mislabelled!\nInsert a valid file 'REDCapData.csv' into the /{dir} folder!\n"),
+      glue::glue("\nPatient identification data is missing or mislabelled!\nInsert a valid file 'MetaData.csv' into the /{dir} folder!\n"),
+      glue::glue("\nFile with test scoring data is missing or mislabelled!\nInsert a valid file 'TestScoring.csv' into the /{dir} folder!\n"),
+      glue::glue("\nFile with variable labels is missing or mislabelled!\nInsert a valid file 'VariablesOfInterest.csv' into the /{dir} folder!\n")
     )
   ) |>
     mutate(path = here::here(dir, file))

### 2025-09-23 R/describe_concordance.R

@@ -25,160 +25,139 @@
 #' concmats <- describe_concordance(pdd)
 #' }
 #' @export
-describe_concordance <- function(d0, vars) {
+describe_concordance <- function(d0) {
   # Get pairs of operationalisations:
   p <- crossing(predictor = d0$criteria$type, reference = d0$criteria$type)
   # Calculate Kappas:
-  k <- lapply(
-    seq_len(nrow(p)),
-    function(i) {
-      if(p[i, 1] == p[i, 2]) {
-        list(kappa = 1)
-      } else {
-        d0$PDD |>
-          filter(type %in% p[i, ]) |>
-          pivot_wider(names_from = type, values_from = PDD, id_cols = id) |>
-          select(-id) |>
-          as.matrix() |>
-          cohen.kappa()
-      }
+  k <- lapply(seq_len(nrow(p)), function(i) {
+    if (p[i, 1] == p[i, 2]) {
+      list(kappa = 1)
+    } else {
+      d0$PDD |>
+        filter(type %in% p[i, ]) |>
+        pivot_wider(names_from = type, values_from = PDD, id_cols = id) |>
+        select(-id) |>
+        as.matrix() |>
+        psych::cohen.kappa()
     }
-  )
+  })
   # Calculate confusion matrixes:
-  confmats <- lapply(
-    seq_len(nrow(p)),
-    function(i) {
-      if(p[i, 1] == p[i, 2]) {
-        NULL
-      } else {
-        d0$PDD |>
-          filter(type %in% p[i, ]) |>
-          pivot_wider(names_from = type, values_from = PDD, id_cols = id) |>
-          select(all_of(unlist(p[i, ], use.names = FALSE))) |>
-          mutate_all(~2-as.numeric(.x)) |> # re-scoring such that prevalence is calculated for PDD == 1
-          table() |>
-          confusionMatrix()
-      }
+  confmats <- lapply(seq_len(nrow(p)), function(i) {
+    if (p[i, 1] == p[i, 2]) {
+      NULL
+    } else {
+      d0$PDD |>
+        filter(type %in% p[i, ]) |>
+        pivot_wider(names_from = type, values_from = PDD, id_cols = id) |>
+        select(all_of(unlist(p[i, ], use.names = FALSE))) |>
+        mutate_all(\(x) 2 - as.numeric(x)) |> # re-scoring such that prevalence is calculated for PDD == 1
+        table() |>
+        caret::confusionMatrix()
     }
[…]

### 2025-09-23 R/diagnose_pdd_case.R

@@ -24,27 +24,26 @@
 #' @export
 diagnose_pdd_case <- function(x, c) {
   c[ , c(is.na(c))] <- "drsii" # DRS-2 as a dummy variable.
-  x |>
-    mutate(
-      impaired_orig_att = if_else(x[[c$atte]] < c$atte_t, TRUE, FALSE),
-      impaired_orig_exe = if_else(x[[c$exec]] < c$exec_t, TRUE, FALSE),
-      impaired_orig_con = if_else(x[[c$cons]] < c$cons_t, TRUE, FALSE),
-      impaired_orig_mem = if_else(x[[c$memo]] < c$memo_t, TRUE, FALSE),
-      impaired_new_lan  = if_else(x[[c$lang]] < c$lang_t, TRUE, FALSE),
-      # PDD criteria proper:
-      crit1 = TRUE, # 1. Parkinson’s disease
-      crit2 = TRUE, # 2. Parkinson’s disease developed before dementia
-      crit3 = if_else(x[[c$glob]] < c$glob_t, TRUE, FALSE), # 3. Global deficit
-      crit4 = if_else(x[[c$iadl]] > c$iadl_t, TRUE, FALSE), # 4. Dementia has Impact on ADLs
-      crit5 = case_when( # 5. Impaired cognition (for Yes, at least of 2 tests abnormal)
-        c$group == "mmse" ~if_else(rowSums(across(starts_with("impaired_orig"))) > 1, TRUE, FALSE),
-        c$group == "moca" ~if_else(rowSums(across(starts_with("impaired"))) > 1, TRUE, FALSE),
-        c$group %in% c("smoca","lvlII") ~ crit3
-      ),
-      crit6 = TRUE, # 6. Absence of Major Depression
-      crit7 = TRUE, # 7. Absence of Major Depression
-      crit8 = TRUE, # 8. Absence of other abnormalities that obscure diagnosis
-      PDD   = if_else(rowSums(across(starts_with("crit"))) == 8, TRUE, FALSE)
-    ) |>
+  x |> mutate(
+    impaired_orig_att = if_else(x[[c$atte]] < c$atte_t, TRUE, FALSE),
+    impaired_orig_exe = if_else(x[[c$exec]] < c$exec_t, TRUE, FALSE),
+    impaired_orig_con = if_else(x[[c$cons]] < c$cons_t, TRUE, FALSE),
[…]

### 2025-09-23 R/diagnose_pdd_sample.R

@@ -30,12 +30,10 @@
 #' @export
 diagnose_pdd_sample <- function(d0) {
   crit <- specify_criteria()
-  pdd  <- lapply(seq_len(nrow(crit)), function(i) {
+  pdd <- map_dfr(seq_len(nrow(crit)), function(i) {
     diagnose_pdd_case(d0, crit[i, ]) |>
       mutate(type = crit$type[i], .after = id)
-  }) |>
-    reduce(full_join) |>
-    suppressMessages()
+  })
   # Return both criteria specification and PDD diagnoses:
   list(criteria = crit, PDD = pdd)
 }

### 2025-09-23 R/do_summary.R

@@ -71,7 +71,7 @@ do_summary <- function(x, dec, sum = NULL) {
     ifelse(x < .001, "< .001", paste0("= ", zerolead(x, dec)))
   } else if(sum == "Nperc") {
     t <- table(x)
-    paste0(t[2]," (",rprint(100*prop.table(t)[2], dec), "%)")
+    paste0(t[2]," (",rprint(100 * prop.table(t)[2], dec), "%)")
   } else if(sum == "Nslash") {
     paste(table(x), collapse = "/")
   } else if(sum == "estCI") {

### 2025-09-23 R/gt_apa_table.R

@@ -16,8 +16,7 @@
 #'
 #' @export
 gt_apa_table <- function(x, grp = NULL, nms = NULL, tit = "") {
-  x |>
-    gt(groupname_col = grp, rowname_col = nms) |>
+  x |> gt(groupname_col = grp, rowname_col = nms) |>
     tab_options(
       table.border.top.color = "white",
       heading.title.font.size = px(16),

### 2025-09-23 R/import_item_data.R

@@ -11,14 +11,13 @@
 #'
 #' @examples
 #' \dontrun{
-#' p    <- data_paths("data-raw")
+#' p <- data_paths("data-raw")
 #' data <- import_item_data(p[1])
 #' }
 #'
 #' @export
 import_item_data <- function(path) {
-  df <-
-    read_delim(path, delim = ";", col_types = cols()) |>
+  df <- readr::read_delim(path, delim = ";", col_types = cols()) |>
     rename(
       id = IPN,
       birth = born_NA_RC,
@@ -35,16 +34,16 @@ import_item_data <- function(path) {
         case_when(sex == "F" ~ 0, sex == "M" ~ 1),
         levels = 0:1,
         labels = c("female", "male"),
-        ordered = F
+        ordered = FALSE
       ),
       hand = factor(
         case_when(hand == "R" ~ 0, hand == "L" ~ 1),
         levels = 0:1,
         labels = c("right", "left"),
-        ordered = F
+        ordered = FALSE
       ),
-      across(ends_with("name"), ~make_clean_names(.x, allow_dupes = TRUE)),
-      across(all_of(c("assdate", "birth")), ~as.Date(.x, tryFormats = "%d.%m.%Y")),
+      across(ends_with("name"), \(x) janitor::make_clean_names(x, allow_dupes = TRUE)),
+      across(all_of(c("assdate", "birth")), \(x) as.Date(x, tryFormats = "%d.%m.%Y")),
       incl = 1 # As a baseline, include everyone
     )
   # Check MMSE variables:
@@ -68,7 +67,7 @@ import_item_data <- function(path) {
       print(mistakes[[i]][ , c("surname", "firstname", "id", "assdate", i)])
     }
   }
-  if(stop) stop("There seem to be typos in the data, check the data printed above to locate and repair them.")
+  if (stop) stop("There seem to be typos in the data.\nCheck the data printed above to locate and repair them.")
   # !Duplicated cases, rows selected manually:
   # IPN138: keep the first assessment because it is the "screening" in REDCap.
   # IPN347: keep the first assessment because the second one was just one year later & the first one is REDCap's "screening".

### 2025-09-23 R/import_redcap_data.R

@@ -12,14 +12,13 @@
 #' @examples
 #' \dontrun{
 #' p <- data_paths("data-raw")
-#' scor <- readr::read_delim(p[4], delim = ";")
+#' scor <- readr::read_delim(p[4], delim = ";", col_types = cols(rev = "c"))
 #' data <- import_redcap_data(p[2])
 #' }
 #'
 #' @export
 import_redcap_data <- function(path, scoring) {
-  df <-
-    read_csv(path, show_col_types = FALSE) |>
+  df <- readr::read_csv(path, show_col_types = FALSE) |>
     select(-contains("dbs"), -contains("post")) |>
     filter(grepl("screening", redcap_event_name)) |>
     rename(
@@ -40,7 +39,7 @@ import_redcap_data <- function(path, scoring) {
   }
   # Extract correct FAQ scores:
   faq_its <- unlist(strsplit(with(scoring, item[scale == "faq"]), ","))
-  for(i in faq_its) {
+  for (i in faq_its) {
     df[ , paste0("faq_", i)] <- NA
     for (j in seq_len(nrow(df))) {
       if (is.na(df[j, paste0("faq_uvod_", i)])) next
@@ -49,26 +48,24 @@ import_redcap_data <- function(path, scoring) {
     }
   }
   # Reverse item scores where applicable:
-  with(
-    scoring,
+  with(scoring, {
     for (i in scale[complete.cases(rev)]) {
       for (j in unlist(strsplit(rev[scale == i], ","))) {
         df[ , paste(i, j, sep = "_")] <<- (max[scale == i] + min[scale == i]) - df[ , paste(i, j, sep = "_")]
       }
     }
-  )
+  })
   # Check whether MoCA verbal fluency and vf_k are the same:
-  vf_fail <-
-    df |>
+  vf_fail <- df |>
     filter(vf_k != moca_fluence_k) |>
     select(id, neuropsy_years, vf_k, moca_fluence_k)
-  stop <- F
+  stop <- FALSE
   if (nrow(vf_fail) > 0) {
-    stop <- T
+    stop <- TRUE
     cat("\nSee the problematic cases below:\n\n")
     print(vf_fail)
   }
-  if(stop) cat("
+  if (stop) cat("
   There are some incongruities in verbal fluency data between MoCA and Level II.
   Using the Level II data in these cases to keep it consistent with the rest of data.\n\n"
   )
[…]

### 2025-09-23 R/prepare_data.R

@@ -24,25 +24,27 @@ prepare_data <- function(p, check.names = TRUE) {
   Scoring <- readr::read_delim(p[4], delim = ";", col_types = cols(rev = "c"))
   REDCap <- import_redcap_data(p[2], Scoring)
   # Check names nd compatibility:
-  if(check.names) {
-    check_names(d = ItemData, nms = read_csv(p[3], show_col_types = FALSE))
+  if (check.names) {
+    check_names(d = ItemData, nms = readr::read_csv(p[3], show_col_types = FALSE))
   }
   disc <- check_compatibility(d1 = ItemData, d2 = REDCap)
   if (nrow(disc) > 1) {
     discfile <- here("temp", "discrepancies.csv")
-    cat(paste0("\n!!! There were some discrepancies, REDCap data were used !!!\nCheck the ", discfile," file\nto reconcile any incompatibilities.\n"))
-    if (!dir.exists("temp")) dir.create("temp")
-    write_csv(disc, discfile, na = "")
+    cat(glue::glue("\nThere were some discrepancies, REDCap data were used.\nCheck the {discfile} file to reconcile any incompatibilities.\n"))
+    if (!dir.exists("temp")) {
+      dir.create("temp")
+    }
+    readr::write_csv(disc, discfile, na = "")
   } else {
-    if (dir.exists("temp")) unlink("temp", recursive = TRUE)
+    if (dir.exists("temp")) {
+      unlink("temp", recursive = TRUE)
+    }
   }
   # Prepare a full data set:
-  df <-
-    ItemData |>
+  df <- ItemData |>
     filter(incl == 1) |>
     left_join(REDCap, by = "id", suffix = c("_iw", "_rc"))
-  cv <-
-    names(df)[grepl("_iw", names(df))] |>
+  cv <- names(df)[grepl("_iw", names(df))] |>
     sub(x = _, "_iw", "")
   for (i in cv) {
     df[ , i] <- NA
@@ -71,5 +73,5 @@ prepare_data <- function(p, check.names = TRUE) {
   map_file  <- here::here("data-raw", "CalculatorMapping.csv")
   ci_lvlII <- assign_cognitive_impairment(df, calc_file, map_file, "both")
   # Return the data:
-  left_join(df, ci_lvlII, by = "id")
+  df |> left_join(ci_lvlII, by = "id")
 }

### 2025-09-23 R/regress_pdd_on_demographics.R

@@ -31,22 +31,18 @@
 #' @export
 regress_pdd_on_demographics <- function(d0, d1) {
   # Extract orders for visualisation:
-  ord_id <-
-    d0 |>
+  ord_id <- d0 |>
     arrange(age) |>
     filter(!is.na(age)) |>
-    select(id) |>
-    pull()
-  ord_type <-
-    d1 |>
+    pull(id)
+  ord_type <- d1 |>
     select(PDD, type) |>
     table() |>
     prop.table() |>
     as_tibble() |>
     filter(PDD == TRUE) |>
     arrange(n) |>
-    select(type) |>
-    pull()
+    pull(type)
   # Prepare an anonimisation mapping:
   anon <- left_join(
     data.frame(id = ord_id),
@@ -57,26 +53,22 @@ regress_pdd_on_demographics <- function(d0, d1) {
     by ="id"
   )
   # Prepare data for analysis:
-  df <-
-    d1 |>
+  df <- d1 |>
     select(id, type, PDD) |>
-    left_join(d0 |> select(id, age, sex), by = "id") |>
+    left_join(d0 |> select(id, age, sex), by = join_by(id)) |>
     mutate(
       PDD = if_else(PDD, 1, 0),
       sid = factor(
-        sapply(
-          seq_len(length(id)),
-          function(i)
-            anon$sid[anon$id == id[i]]
-        ),
+        sapply(seq_along(id), function(i) {
+          anon$sid[anon$id == id[i]]
+        }),
         levels = anon$sid,
         ordered = TRUE
       ),
-      type = factor(type, levels = ord_type, ordered = T)
+      type = factor(type, levels = ord_type, ordered = TRUE)
     )
   # Visualise raw data:
-  plt_raw <-
-    df |>
+  plt_raw <- df |>
     filter(!is.na(sid)) |>
     ggplot() +
     aes(x = sid, y = type, fill = factor(PDD)) +
@@ -95,24 +87,19 @@ regress_pdd_on_demographics <- function(d0, d1) {
       )
     )
   # Get regression fits:
-  fits <-
-    lapply(
-      set_names(unique(df$type)),
-      function(i) glm(
-        formula = PDD ~ age * sex,
-        data = subset(df, type == i),
-        family = binomial(link = "logit")
-      )
+  fits <- lapply(set_names(unique(df$type)), function(i) {
+    glm(
+      formula = PDD ~ age * sex,
[…]

### 2025-09-23 R/specify_criteria.R

@@ -12,8 +12,7 @@
 #' @export
 specify_criteria <- function() {
   # MMSE-based criteria
-  mmse <-
-    crossing(exec = c("vf_s","cloc"), iadl = c("faq_9","faq")) |>
+  mmse <- crossing(exec = c("vf_s","cloc"), iadl = c("faq_9","faq")) |>
     mutate(
       group = "mmse",
       type = paste0("MMSE (",seq_len(n()),")"),
@@ -30,64 +29,55 @@ specify_criteria <- function() {
       exec_t = case_when(exec == "cloc" ~ 2, exec == "vf_s" ~ 10),
       iadl_t = case_when(iadl == "faq"  ~ 7, iadl == "faq_9" ~ 1)
     ) |>
-    relocate(exec , .after = atte_t) |>
+    relocate(exec, .after = atte_t) |>
     relocate(exec_t, .after = exec) |>
     relocate(iadl, .before = iadl_t)
   # MoCA-based criteria
-  moca <-
-    lapply(
-      2:3,
-      function(i)
-        crossing(
-          group = "moca",
-          glob = "moca_total",
-          glob_t = 26,
-          atte = "moca_7",
-          atte_t = 3,
-          exec = c("moca_cloc", "vf_k"),
-          memo = "moca_5words",
-          memo_t = 5:1,
-          cons = "moca_cube",
-          cons_t = 1,
-          lang = c("moca_anim", "moca_abs"),
-          iadl = c("faq", "faq_9")
-        ) |>
-        mutate(
-          exec_t = case_when(exec == "vf_k" ~ 11, exec == "moca_cloc" ~ i),
-          iadl_t = case_when(iadl == "faq" ~ 7, iadl == "faq_9" ~ 1),
-          lang_t = case_when(lang == "moca_abs" ~ 2 , lang == "moca_anim" ~ 3)
-        )
+  moca <- map_dfr(2:3, function(i) {
+    crossing(
+      group = "moca",
+      glob = "moca_total",
+      glob_t = 26,
+      atte = "moca_7",
+      atte_t = 3,
+      exec = c("moca_cloc", "vf_k"),
+      memo = "moca_5words",
+      memo_t = 5:1,
+      cons = "moca_cube",
+      cons_t = 1,
+      lang = c("moca_anim", "moca_abs"),
+      iadl = c("faq", "faq_9")
     ) |>
[…]

### 2025-09-23 R/summarise_rates.R

@@ -43,9 +43,8 @@ summarise_rates <- function(d0, vars, descending = TRUE) {
   } else {
     v <- vars
   }
-  # Prepare and arrange the table with prevalences:
-  prevs <-
-    d0$PDD |>
+  # Prepare and arrange the table with PDD rates (previously called 'prevalences'):
+  prevs <- d0$PDD |>
     select(type, PDD) |>
     table() |>
     as_tibble() |>
@@ -55,23 +54,21 @@ summarise_rates <- function(d0, vars, descending = TRUE) {
       perc = 100 * `TRUE`/N,
       Rate = paste0(`TRUE`, " (", do_summary(perc, 2), "%)")
     )
-  # Prepare operationalisation labels:
-  opers <-
-    d0$criteria |>
+  # Prepare algorithm (previously called 'operationalisation') labels:
+  opers <- d0$criteria |>
     mutate(
-      Global = sapply(seq_len(length(type)), function(i) paste0(v[v[ , 1] == glob[i], 2]," < ", glob_t[i])),
-      Attention = sapply(seq_len(length(type)), function(i) paste0(v[v[ , 1] == atte[i], 2]," < ", atte_t[i])),
-      Executive = sapply(seq_len(length(type)), function(i) paste0(v[v[ , 1] == exec[i], 2]," < ", exec_t[i])),
-      Construction = sapply(seq_len(length(type)), function(i) paste0(v[v[ , 1] == cons[i], 2]," < ", cons_t[i])),
-      Memory = sapply(seq_len(length(type)), function(i) paste0(v[v[ , 1] == memo[i], 2]," < ", memo_t[i])),
-      Language = sapply(seq_len(length(type)), function(i) paste0(v[v[ , 1] == lang[i], 2]," < ", lang_t[i])),
-      IADL = sapply(seq_len(length(type)), function(i) paste0(v[v[ , 1] == iadl[i], 2]," > ", iadl_t[i]))
+      Global = sapply(seq_along(type), \(i) paste0(v[v[ , 1] == glob[i], 2]," < ", glob_t[i])),
+      Attention = sapply(seq_along(type), \(i) paste0(v[v[ , 1] == atte[i], 2]," < ", atte_t[i])),
+      Executive = sapply(seq_along(type), \(i) paste0(v[v[ , 1] == exec[i], 2]," < ", exec_t[i])),
[…]

### 2025-09-23 R/table_criteria.R

@@ -22,16 +22,14 @@ table_criteria <- function(vars) {
     v <- vars
   }
   v <- subset(v, group %in% c("Attention and Working Memory", "Executive Function", "Language", "Memory", "Visuospatial Function"))
-  note <-
-    paste0(
-      "Note. MMSE: Mini-Mental State Examination; MoCA: Montreal Cognitive Assessment; sMoCA: short version of the MoCA; ",
-      paste0(v$label,": ",v$note) |> paste(collapse = "; "),
-      ". The OR operator implies that exactly one of the criteria listed is utilized within a single operationalization;
+  note <- paste0(
+    "Note. MMSE: Mini-Mental State Examination; MoCA: Montreal Cognitive Assessment; sMoCA: short version of the MoCA; ",
+    paste0(v$label,": ",v$note) |> paste(collapse = "; "),
+    ". The OR operator implies that exactly one of the criteria listed is utilized within a single operationalization;
       the & operator implies that both criteria are used at the same time within a single operationalization;
       each threshold value within the set brackets {} was used to define probable PDD once in combination with all the
       other criteria on the same row."
-
-    )
+  )
   # Do the table:
   tibble(
     Type = c(

### 2025-09-23 R/table_levelII_approximations.R

@@ -24,18 +24,15 @@
 #'
 #' @export
 table_levelII_approximations <- function(values, struct, crit) {
-  lapply(
-    1:2,
-    function(ii)
-      values |>
+  map_dfr(1:2, function(ii) {
+    values |>
       filter(reference == paste0("Lvl.II (", ii, ")")) |>
       filter(predictor %in% struct[[ii]]) |>
       filter(!grepl("Lvl.II", predictor)) |>
       arrange(desc(get(crit))) |>
       select(reference, predictor, Kappa_raw, Accuracy_raw, AccuracyPValue, Sensitivity, Specificity) |>
       mutate(order = seq_len(length(reference)))
-  ) |>
-    do.call(rbind.data.frame, args = _) |>
+  }) |>
     pivot_wider(
       names_from = reference,
       id_cols = order,
@@ -46,8 +43,8 @@ table_levelII_approximations <- function(values, struct, crit) {
         order %in% 1:5 ~ "Top five",
         order %in% (length(order)-4):length(order) ~ "Bottom five"
       ),
-      across(contains("PValue"), ~do_summary(.x, 3, "p")),
-      across(where(is.numeric), ~do_summary(.x, 2))
+      across(contains("PValue"), \(x) do_summary(x, 3, "p")),
+      across(where(is.numeric), \(x) do_summary(x, 2))
     ) |>
     filter(!is.na(part)) |>
     select(part, ends_with("(1)"), ends_with("(2)")) |>

### 2025-09-23 _targets.R

@@ -1,17 +1,10 @@
+library(tidyverse) |> suppressPackageStartupMessages()
+library(gt)
 library(targets)
 library(tarchetypes)
 
 # Set target options:
-tar_option_set(
-  packages = c(
-    "here", # for path listing
-    "tidyverse", # for data wrangling
-    "janitor", # for cleaning names of diacritics
-    "psych", # for Cohen's Kappa calculation
-    "caret", # for confusion matrixes
-    "gt" # for tableing
-  )
-)
+tar_option_set()
 
 # Load all in-house functions:
 tar_source()

### 2025-10-04 _targets.R
R2/9 vignette prefinal

@@ -18,7 +18,7 @@ list(
   ),
   tar_target(
     name = variables, # List variables for future use
-    command = here::here("data-raw","VariablesOfInterest.csv"),
+    command = here::here("data-raw", "VariablesOfInterest.csv"),
     format = "file"
   ),
   tar_target(

### 2025-10-08 _manuscript/manuscript.qmd
Added option to regress PDD on covariates, cleaned the code for the original text.

@@
 rates <- tar_read(rate_summaries, store = upstore)
+algos <- tar_read(algorithms, store = upstore)
 pred <- tar_read(demographic_predictors, store = upstore)
 conc <- tar_read(concordance_statistics, store = upstore)
+kappa <- tar_read(kappa_summmaries, store = upstore)
 perc <- rates$table$perc
@@
   do_summary(3, "p")
-
-# List of algorithms:
-algos <- list(
-  faq_tot = subset(rates$table, IADL == "FAQ > 7")$type,
-  faq_9 = subset(rates$table, IADL == "FAQ 9 > 1")$type
-)
-
-continue <- !(any(algos$faq_tot %in% algos$faq_9) || any(algos$faq_9 %in% algos$faq_tot))
-stopifnot("Some algorithm(s) use both IADL operationalisations!" = continue)
 ```
@@
 
-```{r}
-#| label: "kappas"
-
-# Prepare combinations within IADL operationalisation:
-kappas <- lapply(set_names(names(algos)), function(o) {
-  combn(algos[[o]], 2) |>
-      t() |>
-      as_tibble() |>
-      `colnames<-`(c("predictor", "reference")) |>
-      left_join(conc$table, by = join_by(predictor, reference)) |>
-      pull(Kappa_raw)
-})
-
-# Add cross-operationalisation cases
-kappas$cross <- subset(conc$table, predictor %in% algos$faq_9 & reference %in% algos$faq_tot)$Kappa_raw
-
-# Extract Kappa summaries:
-kappa_sums <- sapply(names(kappas), function(i) {
-  paste0(do_summary(kappas[[i]], 2, "M"),", SD = ", do_summary(kappas[[i]], 2, "SD"))
-})
-
-# Add Kappa summaries of algos with same everything but different IADL:
-kappa_iadl <- data.frame(
-  predictor = c("Lvl.II (1)", "MMSE (1)", "MMSE (3)", paste0("MoCA (",seq(1,59,2),")"), "sMoCA (1)"),
-    reference = c("Lvl.II (2)", "MMSE (2)", "MMSE (4)", paste0("MoCA (",seq(2,60,2),")"), "sMoCA (2)")
-) |>
-  left_join(conc$table) |>
-  pull(Kappa_raw)
-
-kappa_iadl_sums <- paste0(do_summary(x = kappa_iadl, 2, "M"),", SD = ", do_summary(x = kappa_iadl, 2, "SD"))
-```
-
 Results of the analyses of prediction Accuracy, Cohen's $\kappa$, Sensitivity and Specificity are presented in @fig-acc, @fig-kappa, @fig-sens and @fig-spec respectively.
@@
 Generally, algorithms that employed the same operationalization of IADL deficit showed substantial pairwise concordance, however, algorithms that operationalized IADL deficit differently did not.
-Whereas among algorithms with identical IADL deficit operationalization, the agreement judged by Cohen's $\kappa$ was moderately high (operationalization by FAQ total score: $\kappa$ = `r kappa_sums["faq_tot"]`; operationalization by FAQ item 9: $\kappa$ = `r kappa_sums["faq_9"]`), among algorithms that differ in IADL deficit operationalization but are otherwise identical it was low: $\kappa$ = `r kappa_iadl_sums`.
+Whereas among algorithms with identical IADL deficit operationalization, the agreement judged by Cohen's $\kappa$ was moderately high (operationalization by FAQ total score: $\kappa$ = `r kappa$sums["faq_tot"]`; operationalization by FAQ item 9: $\kappa$ = `r kappa$sums["faq_9"]`), among algorithms that differ in IADL deficit operationalization but are otherwise identical it was low: $\kappa$ = `r kappa$sums["iadl"]`.
 

### 2025-10-08 R/assign_cognitive_impairment.R

@@ -7,7 +7,7 @@
 #'
 #' @param d0 A list of data, typically generated by \code{prepare_data}, containing test scores.
 #' @param calc A character string specifying the file path to the regression calculator
-#' (.xlsx format).
+#'    (.xlsx format).
 #' @param map A data frame, tibble, or matrix with:
 #'   \itemize{
 #'     \item Column 1: Label used in the calculator file

### 2025-10-08 R/check_compatibility.R

@@ -7,8 +7,8 @@
 #' @param d2 A data frame generated by \code{import_redcap_data}, containing REDCap data.
 #'
 #' @returns Nothing. If discrepancies are found, an informative message is printed and the
-#' function stops execution (e.g., via \code{stop()}). This function is intended to be used
-#' inside \code{prepare_data()} as a consistency check.
+#'    function stops execution (e.g., via \code{stop()}). This function is intended to be used
+#'    inside \code{prepare_data()} as a consistency check.
 #'
 #' @examples
 #' \dontrun{

### 2025-10-08 R/check_names.R

@@ -7,7 +7,7 @@
 #' @param nms A data frame or vector containing the correct patient names (e.g., from REDCap).
 #'
 #' @returns Nothing. Prints a message listing any discrepancies found. Intended to be called
-#' within the \code{prepare_data()} function, where the check can optionally be silenced.
+#'    within the \code{prepare_data()} function, where the check can optionally be silenced.
 #'
 #' @examples
 #' \dontrun{

### 2025-10-08 R/check_ranges.R

@@ -6,7 +6,7 @@
 #' @param d A data frame prepared via the \code{prepare_data} function.
 #'
 #' @returns Nothing. Prints a message listing any out-of-range values and may
-#' terminate execution if invalid data are detected.
+#'    terminate execution if invalid data are detected.
 #'
 #' @examples
 #' \dontrun{

### 2025-10-08 R/data_paths.R

@@ -6,7 +6,7 @@
 #' for the \code{prepare_data} function.
 #'
 #' @param dir A character string specifying the input directory where raw data files reside.
-#' Defaults to \code{"data-raw"}.
+#'    Defaults to \code{"data-raw"}.
 #'
 #' @returns A named character vector with file paths to all required raw data files.
 #'

### 2025-10-08 R/describe_concordance.R

@@ -6,11 +6,11 @@
 #' Value (NPV), and Accuracy. Visual summaries are also generated.
 #'
 #' @param d0 A list containing PDD diagnostic results, as produced by
-#' \code{diagnose_pdd_sample}.
+#'    \code{diagnose_pdd_sample}.
 #'
 #' @returns A list with two components:
 #' \describe{
-#'   \item{\code{stats}}{A data frame summarising pairwise concordance statistics
+#'   \item{\code{table}}{A data frame summarising pairwise concordance statistics
 #'   (e.g., Kappa, Accuracy, Sensitivity, Specificity, PPV, NPV).}
 #'   \item{\code{plots}}{A list of ggplot2-based visualisations, including matrices
 #'   of Cohen's Kappa, Accuracy (with significance tests against the No Information Rate),

### 2025-10-08 R/diagnose_pdd_case.R

@@ -8,7 +8,7 @@
 #'
 #' @param x A data frame or matrix containing patient-level variables used for diagnosis.
 #' @param c A named vector (or single-row data frame) specifying the criteria
-#' to be applied for PDD diagnosis.
+#'    to be applied for PDD diagnosis.
 #'
 #' @returns A logical vector indicating PDD diagnosis for each patient
 #' (TRUE = diagnosed, FALSE = not diagnosed).

### 2025-10-08 R/gt_apa_table.R

@@ -6,11 +6,11 @@
 #'
 #' @param x A tibble or data frame to be formatted as an APA-style table.
 #' @param grp An optional column name (as a string or symbol) to use for grouping
-#' rows via \code{gt::gt()}'s \code{groupname_col} parameter. Defaults to \code{NULL}.
+#'    rows via \code{gt::gt()}'s \code{groupname_col} parameter. Defaults to \code{NULL}.
 #' @param nms An optional column name (as a string or symbol) to use for row names
-#' via \code{gt::gt()}'s \code{rowname_col} parameter. Defaults to \code{NULL}.
+#'    via \code{gt::gt()}'s \code{rowname_col} parameter. Defaults to \code{NULL}.
 #' @param tit A character string specifying the table title. Will be converted to
-#' HTML format. Defaults to \code{"No Name"}.
+#'    HTML format. Defaults to \code{"No Name"}.
 #'
 #' @returns A \code{gt} table object formatted according to APA style.
 #'

### 2025-10-08 R/import_item_data.R

@@ -5,7 +5,7 @@
 #' This function is intended to be called within the larger \code{prepare_data} workflow.
 #'
 #' @param paths A character string specifying the path to the data file. Typically generated
-#' by the \code{data_paths} function.
+#'    by the \code{data_paths} function.
 #'
 #' @returns A tibble containing item-level data.
 #'

### 2025-10-08 R/import_redcap_data.R

@@ -5,7 +5,7 @@
 #' within the broader \code{prepare_data} workflow.
 #'
 #' @param paths A character string specifying the path to the REDCap data file.
-#' Typically generated by the \code{data_paths} function.
+#'    Typically generated by the \code{data_paths} function.
 #'
 #' @returns A tibble containing the imported and processed REDCap data.
 #'

### 2025-10-08 R/list_algos.r

@@ -1,0 +1,42 @@
+#' Prepare a list of algorithms stratified by IADL mesure
+#'
+#' Using the table with PDD rates estimates per diagnostic algorithm,
+#' this function extracts a list of algorithms using different IADL
+#' criteria for PDD. If there was an algorithm that uses both criteria
+#' at the same time, the function prints an error.
+#'
+#' @param rates A tibble containing raw summaries of PDD rates as
+#'    generated by \code{summarise_rates}.
+#'
+#' @returns A list with two components:
+#' \describe{
+#'   \item{\code{faq_tot}}{A character vector containing names of
+#'   algorithms that defined IADL deficit by FAQ total score > 7.}
+#'   \item{\code{faq_9}}{A character vector containing names of
+#'   algorithms that defined IADL deficit by FAQ item 9 > 1.}
+#' }
+#'
+#'@examples
+#' \dontrun{
+#' p <- data_paths("data-raw")
+#' data <- prepare_data(p)
+#' pdd  <- diagnose_pdd_sample(data)
+#' vars <- here::here("data-raw", "VariablesOfInterest.csv")
+#'
+#' rates <- summarise_rates(pdd, vars)
+#' algos <- list_algos(rates$table)
+#' }
+#'
+#' @export
+list_algos <- function(rates) {
+  # List of algorithms:
+  algs <- list(
+    faq_tot = subset(rates, IADL == "FAQ > 7")$type,
+    faq_9 = subset(rates, IADL == "FAQ 9 > 1")$type
+  )
+  # Sanity check:
+  continue <- !(any(algs$faq_tot %in% algs$faq_9) || any(algs$faq_9 %in% algs$faq_tot))
+  stopifnot("Some algorithm(s) use both IADL operationalisations!" = continue)
+  # Return:
+  algs
+}

### 2025-10-08 R/prepare_data.R

@@ -8,8 +8,8 @@
 #' @param p A named character vector of file paths generated by \code{data_paths}.
 #'
 #' @returns A tibble containing cleaned and pre-processed variables ready for analysis.
-#' The function also performs validation checks and prints warnings or stops execution
-#' if critical issues are detected (e.g., mismatched names, out-of-range values).
+#'    The function also performs validation checks and prints warnings or stops execution
+#'    if critical issues are detected (e.g., mismatched names, out-of-range values).
 #'
 #' @examples
 #' \dontrun{

### 2025-10-08 R/regress_pdd_on_demographics.R

@@ -7,7 +7,9 @@
 #'
 #' @param d0 A tibble of demographic data generated by \code{prepare_data}.
 #' @param d1 A tibble of PDD diagnoses (typically the \code{PDD} element from
-#' the list returned by \code{diagnose_pdd_sample}).
+#'    the list returned by \code{diagnose_pdd_sample}).
+#' @param form A character containing linear model formula. Defaults to
+#'    "PDD ~ age * sex" but could be altered to include covariates.
 #'
 #' @returns A list with the following components:
 #' \describe{
@@ -29,7 +31,7 @@
 #' }
 #'
 #' @export
-regress_pdd_on_demographics <- function(d0, d1) {
+regress_pdd_on_demographics <- function(d0, d1, form = "PDD ~ age * sex") {
   # Extract orders for visualisation:
   ord_id <- d0 |>
     arrange(age) |>
@@ -55,7 +57,7 @@ regress_pdd_on_demographics <- function(d0, d1) {
   # Prepare data for analysis:
   df <- d1 |>
     select(id, type, PDD) |>
-    left_join(d0 |> select(id, age, sex), by = join_by(id)) |>
+    left_join(d0, by = join_by(id)) |>
     mutate(
       PDD = if_else(PDD, 1, 0),
       sid = factor(
@@ -76,20 +78,18 @@ regress_pdd_on_demographics <- function(d0, d1) {
     scale_fill_manual(values = c("grey89", "red3"), na.value = "white") +
     labs(x = NULL, y = NULL) +
     theme(legend.position = "none") +
-    theme(
-      axis.text.x = element_text(
-        angle = 90,
-        vjust = .5,
-        colour = case_when(
-          df$sex == "male" ~ "steelblue",
-          df$sex == "female" ~ "red2"
-        )
+    theme(axis.text.x = element_text(
+      angle = 90,
+      vjust = .5,
+      colour = case_when(
+        df$sex == "male" ~ "steelblue",
+        df$sex == "female" ~ "red2"
       )
-    )
+    ))
   # Get regression fits:
   fits <- lapply(set_names(unique(df$type)), function(i) {
     glm(
-      formula = PDD ~ age * sex,
+      formula = as.formula(form),
       data = subset(df, type == i),
       family = binomial(link = "logit")
     )
@@ -102,7 +102,7 @@ regress_pdd_on_demographics <- function(d0, d1) {
   }) |>
     t()|>
     as_tibble(rownames = "type") |>
[…]

### 2025-10-08 R/specify_criteria.R

@@ -6,8 +6,8 @@
 #' \code{diagnose_pdd_sample}.
 #'
 #' @returns A tibble where each row represents one unique operationalisation
-#' of PDD diagnostic criteria. Columns indicate the type of criteria, individual
-#' test indexes, thresholds, and other relevant parameters.
+#'    of PDD diagnostic criteria. Columns indicate the type of criteria, individual
+#'    test indexes, thresholds, and other relevant parameters.
 #'
 #' @export
 specify_criteria <- function() {

### 2025-10-08 R/summarise_kappa.r

@@ -1,0 +1,54 @@
+#' Prepare a list of algorithms stratified by IADL mesure
+#'
+#' Using the table with PDD rates estimates per diagnostic algorithm,
+#' this function extracts a list of algorithms using different IADL
+#' criteria for PDD. If there was an algorithm that uses both criteria
+#' at the same time, the function prints an error.
+#'
+#' @param algorithms A tibble containing raw summaries of PDD rates as
+#'    generated by \code{summarise_rates}.
+#' @param concordance description
+#'
+#' @returns A list with two components:
+#' \describe{
+#'   \item{\code{faq_tot}}{A numeric vector containing all Cohen's kappas
+#'   of algorithm pairs with IADL deficit defined as FAQ total > 7.}
+#'   \item{\code{faq_9}}{A numeric vector containing all Cohen's kappas
+#'   of algorithm pairs with IADL deficit defined as FAQ item 9 > 1.}
+#'   \item{\code{cross}}{A numeric vector containing all Cohen's kappas
+#'   of algorithm pairs with different IADL deficit definitions.}
+#'   \item{\code{iadl}}{A numeric vector containing all Cohen's kappas
+#'   of algorithm pairs which differ only in IADL deficit definition.}
+#'   \item{\code{sums}}{A named character vector containing mean and
+#'   standard deviation of each vector of Kappas contained in this list.}
+#' }
+#'
+#' @export
+summarise_kappa <- function(algorithms, concordance) {
+  # Prepare combinations within IADL operationalisation:
+  k <- lapply(set_names(names(algorithms)), function(o) {
+    combn(algorithms[[o]], 2) |>
+      t() |>
+      as_tibble() |>
+      `colnames<-`(c("predictor", "reference")) |>
+      left_join(concordance, by = join_by(predictor, reference)) |>
+      pull(Kappa_raw)
+  })
+  # Add cross-operationalisation cases:
+  k$cross <- subset(concordance, predictor %in% algorithms$faq_9 & reference %in% algorithms$faq_tot)$Kappa_raw
+  # Add Kappa summaries of algorithms with same everything but different IADL:
+  k$iadl <- data.frame(
[…]

### 2025-10-08 R/summarise_rates.R

@@ -5,12 +5,12 @@
 #'
 #' @param d0 A list with PDD data generated via \code{diagnose_pdd_sample}.
 #' @param vars A data.frame, tibble, or matrix with variable names in the first column,
-#' variable labels in the second column, type of variable (continuous, binary, or nominal)
-#' in the third column, and optionally a fourth column denoting group and a fifth column
-#' mapping each label to its description in the table’s note.
-#' Alternatively, a path to a CSV file (semicolon-delimited) containing such a table.
+#'    variable labels in the second column, type of variable (continuous, binary, or nominal)
+#'    in the third column, and optionally a fourth column denoting group and a fifth column
+#'    mapping each label to its description in the table’s note.
+#'    Alternatively, a path to a CSV file (semicolon-delimited) containing such a table.
 #' @param descending A logical indicating whether operationalisations with the highest
-#' estimated PDD rates should be listed first (default is \code{TRUE}).
+#'    estimated PDD rates should be listed first (default is \code{TRUE}).
 #'
 #' @returns A list with:
 #' \describe{

### 2025-10-08 R/table_criteria.R

@@ -6,10 +6,10 @@
 #' The table content is hardcoded and can only be changed by modifying the function code.
 #'
 #' @param vars A data.frame, tibble, or matrix with variable names in the first column,
-#' variable labels in the second column, variable type (continuous, binary, or nominal)
-#' in the third column, an optional fourth column indicating group, and a fifth column
-#' mapping each label to a note or description. Alternatively, a path to a semicolon-delimited
-#' CSV file containing such a table.
+#'    variable labels in the second column, variable type (continuous, binary, or nominal)
+#'    in the third column, an optional fourth column indicating group, and a fifth column
+#'    mapping each label to a note or description. Alternatively, a path to a semicolon-delimited
+#'    CSV file containing such a table.
 #'
 #' @returns A \code{gt} table object containing the summary of criteria used in the study.
 #'

### 2025-10-08 R/table_estimands.R

@@ -7,7 +7,8 @@
 #' modified by editing the function's code directly.
 #'
 #' @returns A \code{gt} table object summarising the theoretical estimands,
-#' empirical estimands, and the corresponding statistical estimates used in the study.
+#'    empirical estimands, and the corresponding statistical estimates used
+#'    in the study.
 #'
 #' @export
 table_estimands <- function() {

### 2025-10-08 R/table_levelII_approximations.R

@@ -7,7 +7,7 @@
 #' for use in the manuscript to keep Quarto documents uncluttered.
 #'
 #' @param values A tibble of concordance statistics generated by
-#' \code{describe_concordance}.
+#'    \code{describe_concordance}.
 #' @param struct A list with two elements:
 #' \describe{
 #'   \item{[[1]]}{Character vector of Level I algorithm names derived using FAQ total > 7

### 2025-10-08 _targets.R

@@ -38,12 +38,24 @@ list(
     command = summarise_rates(pdd_data, variables)
   ),
   tar_target(
+    name = algorithms, # List all algorithms used in the study
+    command = list_algos(rate_summaries$table)
+  ),
+  tar_target(
     name = demographic_predictors, # Regress probable PDD on demographics
     command = regress_pdd_on_demographics(raw_data, pdd_data$PDD)
   ),
   tar_target(
+    name = demographic_predictors_adjusted, # Regress probable PDD on demographics adjusting for neuropsychiatry
+    command = regress_pdd_on_demographics(raw_data, pdd_data$PDD, form = "PDD ~ age * sex + bdi + stai_1 + stai_2")
+  ),
+  tar_target(
     name = concordance_statistics, # Describe concordance between different PDD algorithms
     command = describe_concordance(pdd_data)
+  ),
+  tar_target(
+    name = kappa_summmaries, # Etxract summaries of Cohen's kappa for the manuscript
+    command = summarise_kappa(algorithms, concordance_statistics$table)
   ),
   tar_quarto(
     name = manuscript, # Prepare the manuscript

### 2025-10-09 _manuscript/manuscript.qmd
The first build of the package.

@@
 tar_source(here::here("R"))
-data <- tar_read(raw_data, store = upstore)
-crit <- tar_read(pdd_data, store = upstore)$criteria
-vars <- tar_read(variables, store = upstore)
+
+load_target <- function(x) {
+  tar_read_raw(x, store = upstore)
+}
+
+data <- load_target("raw_data")
+crit <- load_target("pdd_data")$criteria
+vars <- load_target("variables")
 ```
@@
 
-desc <- tar_read(sample_description, store = upstore)
-rates <- tar_read(rate_summaries, store = upstore)
-algos <- tar_read(algorithms, store = upstore)
-pred <- tar_read(demographic_predictors, store = upstore)
-conc <- tar_read(concordance_statistics, store = upstore)
-kappa <- tar_read(kappa_summmaries, store = upstore)
+desc  <- load_target("sample_description")
+rates <- load_target("rate_summaries")
+algos <- load_target("algorithms")
+pred  <- load_target("demographic_predictors")
+conc  <- load_target("concordance_statistics")
+kappa <- load_target("kappa_summmaries")
 perc  <- rates$table$perc

### 2025-10-09 R/assign_cognitive_impairment.R

@@ -43,14 +43,14 @@
 assign_cognitive_impairment <- function(d0, calc, map, output = "both") {
   # Read map if it is a file
   if (is.character(map)) {
-    m <- readr::read_delim(map, delim = ";", col_types = cols())
+    m <- readr::read_delim(map, delim = ";", col_types = readr::cols())
   } else {
     m <- map
   }
   # Read regression calculator file:
   r <- readxl::read_excel(calc, sheet = "equations", skip = 1) |>
-    rename("calc_lab" = "...1") |>
-    right_join(m, by = "calc_lab")
+    dplyr::rename("calc_lab" = "...1") |>
+    dplyr::right_join(m, by = "calc_lab")
   # Prepare thresholds for BNT-60:
   bnt_thresh <- data.frame(
     age_bottom = c(0, 0, 60, 60),
@@ -61,16 +61,16 @@ assign_cognitive_impairment <- function(d0, calc, map, output = "both") {
   )
   # Add CI-related values to data:
   d1 <- d0 |>
-    mutate(
+    dplyr::mutate(
       gender = as.numeric(sex) - 1, # to get man = 1, women = 0
-      across(
-        .cols  = all_of(r$data_lab),
-        .fns   = \(x) compute_z_score(r, x, r$calc_lab[r$data_lab == cur_column()], age, gender, edu_years),
+      dplyr::across(
+        .cols = tidyselect::all_of(r$data_lab),
+        .fns = \(x) compute_z_score(r, x, r$calc_lab[r$data_lab == dplyr::cur_column()], age, gender, edu_years),
         .names = "z_{.col}"
       ),
-      across(
-        .cols  = starts_with("z_"),
-        .fns   = \(x) if_else(x <= -1.5, 1, 0),
+      dplyr::across(
+        .cols = tidyselect::starts_with("z_"),
+        .fns = \(x) dplyr::if_else(x <= -1.5, 1, 0),
         .names = "mci_{.col}"
       ),
       # Re-score BNT because the calculator is for BNT-30 not -60
@@ -78,7 +78,7 @@ assign_cognitive_impairment <- function(d0, calc, map, output = "both") {
         ifelse(
           test = is.na(age[i]) || is.na(edu_years[i]), # To ensure correct format.
           yes = NA,
-          no = if_else(
+          no = dplyr::if_else(
[…]

### 2025-10-09 R/check_compatibility.R

@@ -20,30 +20,29 @@ check_compatibility <- function(d1, d2) {
   commonvars <- c(
     "sex", "mmse", "faq", "faq_9", "smoca_total", # "vf_k", is not checked because it comes from MoCA in ItemData but from Level II in REDCap data.
     paste0("moca_", c("7", "cube", "5words", "anim", "total")),
-    "bdi", "stai_1", "stai_2", # These variables are not used in analyses directly but will likely used to describe the sample.
-    NULL
+    "bdi", "stai_1", "stai_2" # These variables are not used in analyses directly but will likely used to describe the sample.
   )
   # Prepare a table comparing common variables:
   tvar <- sapply(d1$id, function(i) {
     sapply(commonvars, function(j) {
       ifelse(
         test = !(i %in% d2$id),
-        yes  = NA,
-        no   = with(d1, get(j)[id == i & incl == 1]) == with(d2, get(j)[id == i])
+        yes = NA,
+        no = with(d1, get(j)[id == i & incl == 1]) == with(d2, get(j)[id == i])
       )
     })
   }) |>
     t()
   # Evaluate mismatches:
-  tdisc <- map_dfr(set_names(commonvars), function(i) {
-    left_join(
+  tdisc <- purrr::map_dfr(rlang::set_names(commonvars), function(i) {
+    dplyr::left_join(
       d1[d1$id %in% rownames(tvar[!tvar[ ,i], ]) & d1$incl == 1, c("id", i)],
       d2[d2$id %in% rownames(tvar[!tvar[ ,i], ]), c("id", i)],
-      by = join_by(id),
+      by = dplyr::join_by(id),
       suffix = c("_ItemData", "_REDCap")
     )
   })
   # Return mismatches as a data frame
-  d1 |> select(id, surname, firstname, assdate) |>
-    right_join(tdisc, by = "id")
+  d1 |> dplyr::select(id, surname, firstname, assdate) |>
+    dplyr::right_join(tdisc, by = "id")
 }

### 2025-10-09 R/check_names.R

@@ -21,10 +21,10 @@
 #' }
 check_names <- function(d, nms) {
   nms <- nms |>
-    filter(study_id %in% d$id) |>
-    column_to_rownames("study_id") |>
-    select(jmeno, prijmeni) |>
-    mutate_all(\(x) janitor::make_clean_names(x, allow_dupes = TRUE))
+    dplyr::filter(study_id %in% d$id) |>
+    tibble::column_to_rownames("study_id") |>
+    dplyr::select(jmeno, prijmeni) |>
+    dplyr::mutate_all(\(x) janitor::make_clean_names(x, allow_dupes = TRUE))
   tnam <- sapply(rownames(nms), function(i) {
     c(forname = nms[i, "jmeno"] == subset(d, id == i & incl == 1)$firstname,
       surname = nms[i, "prijmeni"] == subset(d, id == i & incl == 1)$surname
@@ -32,12 +32,12 @@ check_names <- function(d, nms) {
   }) |>
     t()
   discid <- rownames(tnam[(!tnam[ ,1] | !tnam[ ,2]), ])
-  disctab <- left_join(
-    nms[discid, ] |> rownames_to_column("id"),
+  disctab <- dplyr::left_join(
+    nms[discid, ] |> tibble::rownames_to_column("id"),
     d[d$id %in% discid, c("id", "firstname", "surname")],
-    by = join_by(id)
+    by = dplyr::join_by(id)
   ) |>
-    mutate(reason = if_else(
+    dplyr::mutate(reason = dplyr::if_else(
       condition = id == "IPN143",
       true = "married",
       false = "typo"

### 2025-10-09 R/check_ranges.R

@@ -58,14 +58,18 @@ check_ranges <- function(d) {
   )
   #zerotrunc <- c("vf_k", "vf_s", "tmt_a", "pst_d", "tmt_b", "pst_w", "pst_c", "cf", "gp_r", "gp_l")
   zerotrunc <- c("tmt_a", "pst_c", "vf_animals")
-  for (i in zerotrunc) mistakes[[i]] <- subset(d, get(i) < 0)
-  for (i in paste0("faq_", seq_len(10))) mistakes[[i]] <- subset(d, !(get(i) %in% seq0(3)))
+  for (i in zerotrunc) {
+    mistakes[[i]] <- subset(d, get(i) < 0)
+  }
+  for (i in paste0("faq_", seq_len(10))) {
+    mistakes[[i]] <- subset(d, !(get(i) %in% seq0(3)))
+  }
   stop <- FALSE
-  for(i in names(mistakes)) {
+  for (i in names(mistakes)) {
     mistakes[[i]] <- mistakes[[i]] |>
-      select(all_of(c("id", "surname", "firstname", "assdate", i))) |>
-      filter(!is.na(get(i)))
-    if(nrow(mistakes[[i]]) > 0) {
+      dplyr::select(tidyselect::all_of(c("id", "surname", "firstname", "assdate", i))) |>
+      dplyr::filter(!is.na(get(i)))
+    if (nrow(mistakes[[i]]) > 0) {
       stop <- TRUE
     }
   }

### 2025-10-09 R/compute_descriptives.R

@@ -30,27 +30,27 @@
 compute_descriptives <- function(d0, vois) {
   # Get variables of interest:
   if (is.character(vois)) {
-    v <- readr::read_delim(vois, delim = ";", col_types = cols())
+    v <- readr::read_delim(vois, delim = ";", col_types = readr::cols())
   } else {
     v <- vois
   }
   # Prepare a demography table:
-  demtab <- map_df(seq_len(nrow(v)), function(i) {
+  demtab <- purrr::map_df(seq_len(nrow(v)), function(i) {
     c(
-      variable = pull(v[i, 2]),
-      group = pull(v[i, 4]),
-      N = case_when(
-        pull(v[i, 3]) == "binary" ~ do_summary(d0[ , pull(v[i, 1])], 0, "Nperc"),
-        pull(v[i, 3]) == "nominal" ~ do_summary(d0[ , pull(v[i, 1])], 0, "Nslash"),
-        pull(v[i, 3]) == "continuous" ~ do_summary(d0[ , pull(v[i, 1])], 0, "N")
+      variable = dplyr::pull(v[i, 2]),
+      group = dplyr::pull(v[i, 4]),
+      N = dplyr::case_when(
+        dplyr::pull(v[i, 3]) == "binary" ~ do_summary(d0[ , dplyr::pull(v[i, 1])], 0, "Nperc"),
+        dplyr::pull(v[i, 3]) == "nominal" ~ do_summary(d0[ , dplyr::pull(v[i, 1])], 0, "Nslash"),
+        dplyr::pull(v[i, 3]) == "continuous" ~ do_summary(d0[ , dplyr::pull(v[i, 1])], 0, "N")
       ),
       sapply(c("Md","minmax","M","SD"), function(f) {
         ifelse(
-          test = pull(v[i, 3]) == "continuous",
+          test = dplyr::pull(v[i, 3]) == "continuous",
           yes = ifelse(
-            test = pull(v[i, 4]) %in% c("Attention and Working Memory", "Executive Function", "Language", "Memory", "Visuospatial Function"),
-            yes = do_summary(pull(d0[ , pull(v[i, 1])]), 2, f),
-            no = do_summary(pull(d0[ , pull(v[i, 1])]), ifelse(f %in% c("M","SD"), 2, 0), f)
+            test = dplyr::pull(v[i, 4]) %in% c("Attention and Working Memory", "Executive Function", "Language", "Memory", "Visuospatial Function"),
[…]

### 2025-10-09 R/data.r

@@ -1,0 +1,48 @@
+#' Table of Pairwise Concordance Measures
+#'
+#' A comprehensive table containing all pairwise comparisons between 68
+#' diagnostic algorithms for Parkinson’s Disease Dementia (PDD) evaluated in
+#' the study. The table serves as the basis for Figures 2, A3, A4, and A5, as
+#' well as for other summary results presented throughout the article.
+#'
+#' Each algorithm is used once as a reference and once as a predictor, meaning
+#' that each algorithm pair appears twice. Some metrics, such as Cohen’s kappa,
+#' are symmetrical, whereas others (e.g., accuracy, sensitivity, or specificity)
+#' are not.
+#'
+#' @seealso
+#' * [psych::cohen.kappa()] — computes and documents Cohen’s kappa and related
+#'   metrics.
+#' * [caret::confusionMatrix()] — computes and documents the remaining metrics.
+#'
+#' @format ## `concords`
+#' A tibble with 7,744 rows and 20 columns:
+#' \describe{
+#'   \item{predictor}{Character; name of the algorithm used as a predictor.}
+#'   \item{reference}{Character; name of the algorithm used as a reference.}
+#'   \item{Kappa}{Character; Cohen’s kappa with its 95% confidence interval.}
+#'   \item{Accuracy}{Character; accuracy estimate with its 95% confidence interval.}
+#'   \item{Kappa_raw}{Numeric; Cohen’s kappa estimate.}
+#'   \item{McnemarPValue}{Numeric; p-value from McNemar’s chi-squared test for
+#'     symmetry between rows and columns.}
+#'   \item{Accuracy_raw}{Numeric; accuracy estimate.}
+#'   \item{NoInformationRate_raw}{Numeric; no-information rate, i.e., the accuracy
+#'     if all patients were classified as non-PDD.}
+#'   \item{AccuracyPValue}{Numeric; p-value from the Exact Binomial Test comparing
+#'     accuracy to the no-information rate.}
+#'   \item{Sensitivity}{Numeric; sensitivity estimate.}
+#'   \item{Specificity}{Numeric; specificity estimate.}
+#'   \item{Pos Pred Value}{Numeric; positive predictive value.}
+#'   \item{Neg Pred Value}{Numeric; negative predictive value.}
+#'   \item{Precision}{Numeric; precision estimate.}
+#'   \item{Recall}{Numeric; recall estimate.}
[…]

### 2025-10-09 R/data_paths.R

@@ -40,7 +40,7 @@ data_paths <- function(dir = "data-raw") {
       glue::glue("\nFile with variable labels is missing or mislabelled!\nInsert a valid file 'VariablesOfInterest.csv' into the /{dir} folder!\n")
     )
   ) |>
-    mutate(path = here::here(dir, file))
+    dplyr::mutate(path = here::here(dir, file))
   proceed <- TRUE
   for (i in seq_len(nrow(dfs))) {
     if (!file.exists(dfs[i, "path"])) {
@@ -48,7 +48,6 @@ data_paths <- function(dir = "data-raw") {
       proceed <- FALSE
     }
   }
-  message("\n")
   stopifnot("\nThe pipeline was terminated due to one or more data files being absent.
 Provide valid data files before marching on." = proceed)
   dfs$path

### 2025-10-09 R/describe_concordance.R

@@ -27,16 +27,15 @@
 #' @export
 describe_concordance <- function(d0) {
   # Get pairs of operationalisations:
-  p <- crossing(predictor = d0$criteria$type, reference = d0$criteria$type)
+  p <- tidyr::crossing(predictor = d0$criteria$type, reference = d0$criteria$type)
   # Calculate Kappas:
   k <- lapply(seq_len(nrow(p)), function(i) {
     if (p[i, 1] == p[i, 2]) {
       list(kappa = 1)
     } else {
-      d0$PDD |>
-        filter(type %in% p[i, ]) |>
-        pivot_wider(names_from = type, values_from = PDD, id_cols = id) |>
-        select(-id) |>
+      d0$PDD |> dplyr::filter(type %in% p[i, ]) |>
+        tidyr::pivot_wider(names_from = type, values_from = PDD, id_cols = id) |>
+        dplyr::select(-id) |>
         as.matrix() |>
         psych::cohen.kappa()
     }
@@ -46,10 +45,9 @@ describe_concordance <- function(d0) {
     if (p[i, 1] == p[i, 2]) {
       NULL
     } else {
-      d0$PDD |>
-        filter(type %in% p[i, ]) |>
-        pivot_wider(names_from = type, values_from = PDD, id_cols = id) |>
-        select(all_of(unlist(p[i, ], use.names = FALSE))) |>
+      d0$PDD |> dplyr::filter(type %in% p[i, ]) |>
+        tidyr::pivot_wider(names_from = type, values_from = PDD, id_cols = id) |>
+        dplyr::select(tidyselect::all_of(unlist(p[i, ], use.names = FALSE))) |>
         mutate_all(\(x) 2 - as.numeric(x)) |> # re-scoring such that prevalence is calculated for PDD == 1
         table() |>
         caret::confusionMatrix()
@@ -65,8 +63,8 @@ describe_concordance <- function(d0) {
       Kappa = sapply(seq_along(reference), function(i) {
         ifelse(
           test = reference[i] == predictor[i],
-          yes  = "-",
-          no   = do_summary(k[[i]]$confid[1, ], 2, "estCI")
+          yes = "-",
+          no = do_summary(k[[i]]$confid[1, ], 2, "estCI")
         )
       }),
       Accuracy = sapply(seq_along(reference), function(i) {
@@ -113,15 +111,15 @@ describe_concordance <- function(d0) {
         )
       }),
       # All the other metrics:
[…]

### 2025-10-09 R/diagnose_pdd_case.R

@@ -24,26 +24,26 @@
 #' @export
 diagnose_pdd_case <- function(x, c) {
   c[ , c(is.na(c))] <- "drsii" # DRS-2 as a dummy variable.
-  x |> mutate(
-    impaired_orig_att = if_else(x[[c$atte]] < c$atte_t, TRUE, FALSE),
-    impaired_orig_exe = if_else(x[[c$exec]] < c$exec_t, TRUE, FALSE),
-    impaired_orig_con = if_else(x[[c$cons]] < c$cons_t, TRUE, FALSE),
-    impaired_orig_mem = if_else(x[[c$memo]] < c$memo_t, TRUE, FALSE),
-    impaired_new_lan  = if_else(x[[c$lang]] < c$lang_t, TRUE, FALSE),
+  x |> dplyr::mutate(
+    impaired_orig_att = dplyr::if_else(x[[c$atte]] < c$atte_t, TRUE, FALSE),
+    impaired_orig_exe = dplyr::if_else(x[[c$exec]] < c$exec_t, TRUE, FALSE),
+    impaired_orig_con = dplyr::if_else(x[[c$cons]] < c$cons_t, TRUE, FALSE),
+    impaired_orig_mem = dplyr::if_else(x[[c$memo]] < c$memo_t, TRUE, FALSE),
+    impaired_new_lan  = dplyr::if_else(x[[c$lang]] < c$lang_t, TRUE, FALSE),
     # PDD criteria proper:
     crit1 = TRUE, # 1. Parkinson’s disease
     crit2 = TRUE, # 2. Parkinson’s disease developed before dementia
-    crit3 = if_else(x[[c$glob]] < c$glob_t, TRUE, FALSE), # 3. Global deficit
-    crit4 = if_else(x[[c$iadl]] > c$iadl_t, TRUE, FALSE), # 4. Dementia has Impact on ADLs
+    crit3 = dplyr::if_else(x[[c$glob]] < c$glob_t, TRUE, FALSE), # 3. Global deficit
+    crit4 = dplyr::if_else(x[[c$iadl]] > c$iadl_t, TRUE, FALSE), # 4. Dementia has Impact on ADLs
     crit5 = case_when( # 5. Impaired cognition (for Yes, at least of 2 tests abnormal)
-      c$group == "mmse" ~ if_else(rowSums(across(starts_with("impaired_orig"))) > 1, TRUE, FALSE),
-      c$group == "moca" ~ if_else(rowSums(across(starts_with("impaired"))) > 1, TRUE, FALSE),
[…]

### 2025-10-09 R/diagnose_pdd_sample.R

@@ -30,9 +30,9 @@
 #' @export
 diagnose_pdd_sample <- function(d0) {
   crit <- specify_criteria()
-  pdd <- map_dfr(seq_len(nrow(crit)), function(i) {
+  pdd <- purrr::map_dfr(seq_len(nrow(crit)), function(i) {
     diagnose_pdd_case(d0, crit[i, ]) |>
-      mutate(type = crit$type[i], .after = id)
+      dplyr::mutate(type = crit$type[i], .after = id)
   })
   # Return both criteria specification and PDD diagnoses:
   list(criteria = crit, PDD = pdd)

### 2025-10-09 R/do_summary.R

@@ -10,19 +10,19 @@
 #' \describe{
 #'   \item{NULL (default)}{Simple printing of the vector}
 #'   \item{"N"}{Number of observations}
-#'   \item{"msd"}{Mean ± standard deviation}
+#'   \item{"msd"}{Mean +/- standard deviation}
 #'   \item{"M"}{Mean}
 #'   \item{"SD"}{Standard deviation}
 #'   \item{"Md"}{Median}
 #'   \item{"IQR"}{Interquartile range}
-#'   \item{"minmax"}{Minimum–maximum range}
+#'   \item{"minmax"}{Minimum-maximum range}
 #'   \item{"Min"}{Minimum}
 #'   \item{"Max"}{Maximum}
 #'   \item{"p"}{p-value (formatted for tables)}
 #'   \item{"ptext"}{p-value (formatted for in-text reporting)}
 #'   \item{"Nperc"}{Count and percentage for binary variables}
 #'   \item{"Nslash"}{Counts separated by slashes for nominal variables with any number of categories}
-#'   \item{"estCI"}{Estimate with confidence interval, e.g., "estimate [CI]"}
+#'   \item{"estCI"}{Estimate with confidence interval, e.g., "estimate `[CI]`"}
 #' }
 #'
 #' @returns A character string or numeric summary as specified by \code{sum}.
@@ -60,7 +60,7 @@ do_summary <- function(x, dec, sum = NULL) {
   } else if(sum == "N") {
     as.character(sum(!is.na(x)))
   } else if(sum == "msd") {
-    paste0(rprint(M(x),dec)," ± ",rprint(SD(x), dec))
+    paste0(rprint(M(x),dec)," \u00B1 ",rprint(SD(x), dec))
   } else if (sum == "IQR") {
     rprint(IQRna(x),dec)
   } else if(sum == "minmax") {

### 2025-10-09 R/gt_apa_table.R

@@ -16,39 +16,39 @@
 #'
 #' @export
 gt_apa_table <- function(x, grp = NULL, nms = NULL, tit = "") {
-  x |> gt(groupname_col = grp, rowname_col = nms) |>
-    tab_options(
+  x |> gt::gt(groupname_col = grp, rowname_col = nms) |>
+    gt::tab_options(
       table.border.top.color = "white",
-      heading.title.font.size = px(16),
+      heading.title.font.size = gt::px(16),
       column_labels.border.top.width = 3,
       column_labels.border.top.color = "black",
       column_labels.border.bottom.width = 3,
       column_labels.border.bottom.color = "black",
       table_body.border.bottom.color = "black",
       table.border.bottom.color = "white",
-      table.width = pct(100),
+      table.width = gt::pct(100),
       table.background.color = "white"
     ) |>
-    cols_align(align = "center") |>
-    tab_style(
+    gt::cols_align(align = "center") |>
+    gt::tab_style(
       style = list(
-        cell_borders(
+        gt::cell_borders(
           sides = c("top", "bottom"),
           color = "white",
-          weight = px(1)
+          weight = gt::px(1)
         ),
-        cell_text(
+        gt::cell_text(
           align = "center"
         ),
-        cell_fill(color = "white", alpha = NULL)
+        gt::cell_fill(color = "white", alpha = NULL)
       ),
-      locations = cells_body(
-        columns = everything(),
-        rows = everything()
+      locations = gt::cells_body(
+        columns = tidyselect::everything(),
+        rows = tidyselect::everything()
       )
     ) |>
     tab_header( # title setup
-      title = html(tit)
+      title = gt::html(tit)
     ) |>
-    opt_align_table_header(align = "left")
+    gt::opt_align_table_header(align = "left")
 }

### 2025-10-09 R/import_item_data.R

@@ -4,7 +4,7 @@
 #' (e.g., in MMSE scores), and prepares the data for further processing.
 #' This function is intended to be called within the larger \code{prepare_data} workflow.
 #'
-#' @param paths A character string specifying the path to the data file. Typically generated
+#' @param path A character string specifying the path to the data file. Typically generated
 #'    by the \code{data_paths} function.
 #'
 #' @returns A tibble containing item-level data.
@@ -17,8 +17,8 @@
 #'
 #' @export
 import_item_data <- function(path) {
-  df <- readr::read_delim(path, delim = ";", col_types = cols()) |>
-    rename(
+  df <- readr::read_delim(path, delim = ";", col_types = readr::cols()) |>
+    dplyr::rename(
       id = IPN,
       birth = born_NA_RC,
       sex = gender_NA_RC,
@@ -27,23 +27,23 @@ import_item_data <- function(path) {
       faq = FAQ_seb,
       bdi = `BDI-II`
     ) |>
-    rename_all(tolower) |>
-    mutate(
+    dplyr::rename_all(tolower) |>
+    dplyr::mutate(
       cloc = clox_num + clox_hands,
       sex = factor(
-        case_when(sex == "F" ~ 0, sex == "M" ~ 1),
+        dplyr::case_when(sex == "F" ~ 0, sex == "M" ~ 1),
         levels = 0:1,
         labels = c("female", "male"),
         ordered = FALSE
       ),
       hand = factor(
-        case_when(hand == "R" ~ 0, hand == "L" ~ 1),
+        dplyr::case_when(hand == "R" ~ 0, hand == "L" ~ 1),
         levels = 0:1,
         labels = c("right", "left"),
         ordered = FALSE
       ),
-      across(ends_with("name"), \(x) janitor::make_clean_names(x, allow_dupes = TRUE)),
-      across(all_of(c("assdate", "birth")), \(x) as.Date(x, tryFormats = "%d.%m.%Y")),
+      dplyr::across(tidyselect::ends_with("name"), \(x) janitor::make_clean_names(x, allow_dupes = TRUE)),
+      dplyr::across(tidyselect::all_of(c("assdate", "birth")), \(x) as.Date(x, tryFormats = "%d.%m.%Y")),
       incl = 1 # As a baseline, include everyone
     )
   # Check MMSE variables:

### 2025-10-09 R/import_redcap_data.R

@@ -4,24 +4,26 @@
 #' This function prepares the data in a suitable format and is intended to be used
 #' within the broader \code{prepare_data} workflow.
 #'
-#' @param paths A character string specifying the path to the REDCap data file.
+#' @param path A character string specifying the path to the REDCap data file.
 #'    Typically generated by the \code{data_paths} function.
+#' @param scoring A tibble containing scoring conventions for variables of
+#'    interest.
 #'
 #' @returns A tibble containing the imported and processed REDCap data.
 #'
 #' @examples
 #' \dontrun{
 #' p <- data_paths("data-raw")
-#' scor <- readr::read_delim(p[4], delim = ";", col_types = cols(rev = "c"))
+#' scor <- readr::read_delim(p[4], delim = ";", col_types = readr::cols(rev = "c"))
 #' data <- import_redcap_data(p[2])
 #' }
 #'
 #' @export
 import_redcap_data <- function(path, scoring) {
   df <- readr::read_csv(path, show_col_types = FALSE) |>
-    select(-contains("dbs"), -contains("post")) |>
-    filter(grepl("screening", redcap_event_name)) |>
-    rename(
+    dplyr::select(-tidyselect::contains("dbs"), -tidyselect::contains("post")) |>
+    dplyr::filter(grepl("screening", redcap_event_name)) |>
+    dplyr::rename(
       id = study_id,
       pd_dur = rok_vzniku_pn,
       ledd = levodopa_equivalent,
@@ -57,8 +59,8 @@ import_redcap_data <- function(path, scoring) {
   })
   # Check whether MoCA verbal fluency and vf_k are the same:
   vf_fail <- df |>
-    filter(vf_k != moca_fluence_k) |>
-    select(id, neuropsy_years, vf_k, moca_fluence_k)
+    dplyr::filter(vf_k != moca_fluence_k) |>
+    dplyr::select(id, neuropsy_years, vf_k, moca_fluence_k)
   stop <- FALSE
   if (nrow(vf_fail) > 0) {
     stop <- TRUE
@@ -74,16 +76,16 @@ import_redcap_data <- function(path, scoring) {
   #Check the data printed above to locate these inconsistencies and repair them."
   #)
   df |>
[…]

### 2025-10-09 R/make_dendrogram.r

@@ -18,26 +18,26 @@
 make_dendrogram <- function(tab) {
   # Prepare data for plot edges:
   data <- tab |>
-    mutate(
-      Global = case_when(
-        grepl("sMoCA", type) ~ str_replace(Global, "total score ", ""),
-        grepl("MMSE", type) ~ str_replace(Global, "Total score", "MMSE"),
-        grepl("MoCA", type) ~ str_replace(Global, "Total score", "MoCA")
+    dplyr::mutate(
+      Global = dplyr::case_when(
+        grepl("sMoCA", type) ~ stringr::str_replace(Global, "total score ", ""),
+        grepl("MMSE", type) ~ stringr::str_replace(Global, "Total score", "MMSE"),
+        grepl("MoCA", type) ~ stringr::str_replace(Global, "Total score", "MoCA")
       ),
-      Attention = case_when(
-        grepl("MMSE", type) ~ str_replace(Attention, "Sevens", "MMSE 7s"),
-        grepl("MoCA", type) ~ str_replace(Attention, "Sevens", "MoCA 7s"),
+      Attention = dplyr::case_when(
+        grepl("MMSE", type) ~ stringr::str_replace(Attention, "Sevens", "MMSE 7s"),
+        grepl("MoCA", type) ~ stringr::str_replace(Attention, "Sevens", "MoCA 7s"),
         .default = Attention
       ),
-      Executive = case_when(
-        grepl("MoCA", type) ~ str_replace(Executive, "Clock drawing", "MoCA clock"),
+      Executive = dplyr::case_when(
+        grepl("MoCA", type) ~ stringr::str_replace(Executive, "Clock drawing", "MoCA clock"),
         .default = Executive
       ),
-      across(
+      dplyr::across(
         .cols = c(Attention, Executive, Construction, Memory, Language),
-        .fns = \(x) if_else(grepl("OR", x), NA, x)
+        .fns = \(x) dplyr::if_else(grepl("OR", x), NA, x)
       ),
-      across(everything(), \(x) if_else(x == "-", NA, x)),
+      dplyr::across(tidyselect::everything(), \(x) dplyr::if_else(x == "-", NA, x)),
       lvl0 = "",
       lvl1 = IADL,
       lvl2 = glue::glue("{lvl1}.{Global}"),
[…]

### 2025-10-09 R/prepare_data.R

@@ -6,6 +6,8 @@
 #' from raw sources.
 #'
 #' @param p A named character vector of file paths generated by \code{data_paths}.
+#' @param check.names A logical indicating whether \code{check_names} should be
+#'    run to compare names in meta data and item data.
 #'
 #' @returns A tibble containing cleaned and pre-processed variables ready for analysis.
 #'    The function also performs validation checks and prints warnings or stops execution
@@ -20,16 +22,16 @@
 #' @export
 prepare_data <- function(p, check.names = TRUE) {
   # Read data sets:
-  ItemData <- import_item_data(p[1])
-  Scoring <- readr::read_delim(p[4], delim = ";", col_types = cols(rev = "c"))
+  Scoring <- readr::read_delim(p[4], delim = ";", col_types = readr::cols(rev = "c"))
   REDCap <- import_redcap_data(p[2], Scoring)
-  # Check names nd compatibility:
+  ItemData <- import_item_data(p[1])
+  # Check names and compatibility:
   if (check.names) {
     check_names(d = ItemData, nms = readr::read_csv(p[3], show_col_types = FALSE))
   }
   disc <- check_compatibility(d1 = ItemData, d2 = REDCap)
   if (nrow(disc) > 1) {
-    discfile <- here("temp", "discrepancies.csv")
+    discfile <- here::here("temp", "discrepancies.csv")
     cat(glue::glue("\nThere were some discrepancies, REDCap data were used.\nCheck the {discfile} file to reconcile any incompatibilities.\n"))
     if (!dir.exists("temp")) {
       dir.create("temp")
@@ -42,8 +44,8 @@ prepare_data <- function(p, check.names = TRUE) {
   }
   # Prepare a full data set:
   df <- ItemData |>
-    filter(incl == 1) |>
-    left_join(REDCap, by = "id", suffix = c("_iw", "_rc"))
+    dplyr::filter(incl == 1) |>
+    dplyr::left_join(REDCap, by = dplyr::join_by(id), suffix = c("_iw", "_rc"))
   cv <- names(df)[grepl("_iw", names(df))] |>
     sub(x = _, "_iw", "")
   for (i in cv) {
@@ -51,8 +53,8 @@ prepare_data <- function(p, check.names = TRUE) {
     for (j in seq_len(nrow(df))) {
       df[j , i] <- ifelse(
         test = !is.na(df[j , paste0(i, "_rc")]),
[…]

### 2025-10-09 R/regress_pdd_on_demographics.R

@@ -34,19 +34,19 @@
 regress_pdd_on_demographics <- function(d0, d1, form = "PDD ~ age * sex") {
   # Extract orders for visualisation:
   ord_id <- d0 |>
-    arrange(age) |>
-    filter(!is.na(age)) |>
-    pull(id)
+    dplyr::arrange(age) |>
+    dplyr::filter(!is.na(age)) |>
+    dplyr::pull(id)
   ord_type <- d1 |>
-    select(PDD, type) |>
+    dplyr::select(PDD, type) |>
     table() |>
     prop.table() |>
-    as_tibble() |>
-    filter(PDD == TRUE) |>
-    arrange(n) |>
-    pull(type)
+    tibble::as_tibble() |>
+    dplyr::filter(PDD == TRUE) |>
+    dplyr::arrange(n) |>
+    dplyr::pull(type)
   # Prepare an anonimisation mapping:
-  anon <- left_join(
+  anon <- dplyr::left_join(
     data.frame(id = ord_id),
     cbind.data.frame(
       id = unique(d0$id),
@@ -56,10 +56,10 @@ regress_pdd_on_demographics <- function(d0, d1, form = "PDD ~ age * sex") {
   )
   # Prepare data for analysis:
   df <- d1 |>
-    select(id, type, PDD) |>
-    left_join(d0, by = join_by(id)) |>
-    mutate(
-      PDD = if_else(PDD, 1, 0),
+    dplyr::select(id, type, PDD) |>
+    dplyr::left_join(d0, by = join_by(id)) |>
+    dplyr::mutate(
+      PDD = dplyr::if_else(PDD, 1, 0),
       sid = factor(
         sapply(seq_along(id), function(i) {
           anon$sid[anon$id == id[i]]
@@ -71,23 +71,23 @@ regress_pdd_on_demographics <- function(d0, d1, form = "PDD ~ age * sex") {
     )
   # Visualise raw data:
   plt_raw <- df |>
-    filter(!is.na(sid)) |>
-    ggplot() +
-    aes(x = sid, y = type, fill = factor(PDD)) +
-    geom_tile() +
-    scale_fill_manual(values = c("grey89", "red3"), na.value = "white") +
-    labs(x = NULL, y = NULL) +
-    theme(legend.position = "none") +
-    theme(axis.text.x = element_text(
+    dplyr::filter(!is.na(sid)) |>
+    ggplot2::ggplot() +
[…]

### 2025-10-09 R/specify_criteria.R

@@ -12,10 +12,10 @@
 #' @export
 specify_criteria <- function() {
   # MMSE-based criteria
-  mmse <- crossing(exec = c("vf_s","cloc"), iadl = c("faq_9","faq")) |>
-    mutate(
+  mmse <- tidyr::crossing(exec = c("vf_s","cloc"), iadl = c("faq_9","faq")) |>
+    dplyr::mutate(
       group = "mmse",
-      type = paste0("MMSE (",seq_len(n()),")"),
+      type = paste0("MMSE (",seq_len(dplyr::n()),")"),
       glob = "mmse",
       glob_t = 26,
       atte = "mmse_7",
@@ -26,15 +26,15 @@ specify_criteria <- function() {
       cons_t = 1,
       lang = NA,
       lang_t = 0,
-      exec_t = case_when(exec == "cloc" ~ 2, exec == "vf_s" ~ 10),
-      iadl_t = case_when(iadl == "faq"  ~ 7, iadl == "faq_9" ~ 1)
+      exec_t = dplyr::case_when(exec == "cloc" ~ 2, exec == "vf_s" ~ 10),
+      iadl_t = dplyr::case_when(iadl == "faq"  ~ 7, iadl == "faq_9" ~ 1)
     ) |>
-    relocate(exec, .after = atte_t) |>
-    relocate(exec_t, .after = exec) |>
-    relocate(iadl, .before = iadl_t)
+    dplyr::relocate(exec, .after = atte_t) |>
+    dplyr::relocate(exec_t, .after = exec) |>
+    dplyr::relocate(iadl, .before = iadl_t)
   # MoCA-based criteria
-  moca <- map_dfr(2:3, function(i) {
-    crossing(
+  moca <- purrr::map_dfr(2:3, function(i) {
+    tidyr::crossing(
       group = "moca",
       glob = "moca_total",
       glob_t = 26,
@@ -49,35 +49,35 @@ specify_criteria <- function() {
       iadl = c("faq", "faq_9")
     ) |>
       mutate(
-        exec_t = case_when(exec == "vf_k" ~ 11, exec == "moca_cloc" ~ i),
-        iadl_t = case_when(iadl == "faq" ~ 7, iadl == "faq_9" ~ 1),
-        lang_t = case_when(lang == "moca_abs" ~ 2 , lang == "moca_anim" ~ 3)
[…]

### 2025-10-09 R/summarise_kappa.r

@@ -26,13 +26,13 @@
 #' @export
 summarise_kappa <- function(algorithms, concordance) {
   # Prepare combinations within IADL operationalisation:
-  k <- lapply(set_names(names(algorithms)), function(o) {
+  k <- lapply(rlang::set_names(names(algorithms)), function(o) {
     combn(algorithms[[o]], 2) |>
       t() |>
-      as_tibble() |>
+      tibble::as_tibble() |>
       `colnames<-`(c("predictor", "reference")) |>
-      left_join(concordance, by = join_by(predictor, reference)) |>
-      pull(Kappa_raw)
+      dplyr::left_join(concordance, by = dplyr::join_by(predictor, reference)) |>
+      dplyr::pull(Kappa_raw)
   })
   # Add cross-operationalisation cases:
   k$cross <- subset(concordance, predictor %in% algorithms$faq_9 & reference %in% algorithms$faq_tot)$Kappa_raw
@@ -41,8 +41,8 @@ summarise_kappa <- function(algorithms, concordance) {
     predictor = c("Lvl.II (1)", "MMSE (1)", "MMSE (3)", paste0("MoCA (", seq(1, 59, 2),")"), "sMoCA (1)"),
     reference = c("Lvl.II (2)", "MMSE (2)", "MMSE (4)", paste0("MoCA (", seq(2, 60, 2),")"), "sMoCA (2)")
   ) |>
-    left_join(concordance, by = join_by(predictor, reference)) |>
-    pull(Kappa_raw)
+    dplyr::left_join(concordance, by = dplyr::join_by(predictor, reference)) |>
+    dplyr::pull(Kappa_raw)
   # Extract Kappa summaries:
   k$sums <- sapply(names(k), function(i) {
     paste0(do_summary(k[[i]], 2, "M"),", SD = ", do_summary(k[[i]], 2, "SD"))

### 2025-10-09 R/summarise_rates.R

@@ -18,8 +18,8 @@
 #'   \item{\code{plot}}{A \code{ggplot2} object visualising estimated PDD rates.}
 #'   \item{\code{gtable}}{A named list of gt tables:
 #'      \describe{
-#'        \item{\code{gtab_rates}{An APA-style \code{gt} table summarising the results.}}
-#'        \item{\code{gtab_algos}{An APA-style \code{gt} table listing algorithms used.}}
+#'        \item{\code{gtab_rates}}{An APA-style \code{gt} table summarising the results.}
+#'        \item{\code{gtab_algos}}{An APA-style \code{gt} table listing algorithms used.}
 #'        }}
 #' }
 #'
@@ -39,24 +39,24 @@
 summarise_rates <- function(d0, vars, descending = TRUE) {
   # Get variables mapping:
   if (is.character(vars)) {
-    v <- readr::read_delim(vars, delim = ";", col_types = cols())
+    v <- readr::read_delim(vars, delim = ";", col_types = readr::cols())
   } else {
     v <- vars
   }
   # Prepare and arrange the table with PDD rates (previously called 'prevalences'):
   prevs <- d0$PDD |>
-    select(type, PDD) |>
+    dplyr::select(type, PDD) |>
     table() |>
-    as_tibble() |>
-    pivot_wider(names_from = PDD, values_from = n) |>
-    mutate(
-      N = rowSums(across(all_of(c("FALSE", "TRUE")))),
-      perc = 100 * `TRUE`/N,
+    tibble::as_tibble() |>
+    tidyr::pivot_wider(names_from = PDD, values_from = n) |>
+    dplyr::mutate(
+      N = rowSums(dplyr::across(tidyselect::all_of(c("FALSE", "TRUE")))),
+      perc = 100 * `TRUE` / N,
       Rate = paste0(`TRUE`, " (", do_summary(perc, 2), "%)")
     )
   # Prepare algorithm (previously called 'operationalisation') labels:
   opers <- d0$criteria |>
-    mutate(
+    dplyr::mutate(
       Global = sapply(seq_along(type), \(i) paste0(v[v[ , 1] == glob[i], 2]," < ", glob_t[i])),
       Attention = sapply(seq_along(type), \(i) paste0(v[v[ , 1] == atte[i], 2]," < ", atte_t[i])),
[…]

### 2025-10-09 R/table_criteria.R

@@ -17,7 +17,7 @@
 table_criteria <- function(vars) {
   # Prepare the note:
   if (is.character(vars)) {
-    v <- readr::read_delim(vars, delim = ";", col_types = cols())
+    v <- readr::read_delim(vars, delim = ";", col_types = readr::cols())
   } else {
     v <- vars
   }
@@ -31,7 +31,7 @@ table_criteria <- function(vars) {
       other criteria on the same row."
   )
   # Do the table:
-  tibble(
+  tibble::tibble(
     Type = c(
       rep("MMSE-based", 5),
       rep("MoCA-based", 6),
@@ -51,19 +51,19 @@ table_criteria <- function(vars) {
       "TMT A & WAIS DSB", "CF A & PST C", "JoLO & CLOXI", "RAVLT DR & (BVMTR DR OR WMS-III Family Pictures DR)", "WAIS Similarities & BNT 60"
     )
   ) |>
-    pivot_wider(names_from = Domain, values_from = Indexes) |>
+    tidyr::pivot_wider(names_from = Domain, values_from = Indexes) |>
     gt_apa_table() |>
-    sub_missing(missing_text = "-") |>
-    tab_spanner(
-      columns = all_of(c("Attention", "Executive Function", "Construction", "Memory", "Language")),
+    gt::sub_missing(missing_text = "-") |>
+    gt::tab_spanner(
+      columns = tidyselect::all_of(c("Attention", "Executive Function", "Construction", "Memory", "Language")),
       label = "Impaired Cognition"
     ) |>
-    tab_source_note(note) |>
-    tab_footnote(
-      locations = cells_body(rows = 4, columns = Memory),
+    gt::tab_source_note(note) |>
+    gt::tab_footnote(
+      locations = gt::cells_body(rows = 4, columns = Memory),
       footnote = "The visual memory was evaluated based on WMS-III Family Pictures or BVMTR depending on which
       test was used in the assessment. This lead to no missing values because each patient underwent assessment
       via one of these tests."
     ) |>
-    opt_footnote_marks(marks = "letters")
+    gt::opt_footnote_marks(marks = "letters")
 }

### 2025-10-09 R/table_estimands.R

@@ -12,7 +12,7 @@
 #'
 #' @export
 table_estimands <- function() {
-  tibble(
+  tibble::tibble(
     RO = paste0("Research Objective ", c(rep(1,4), rep(2,1))),
     Statement = c(
       "To estimate the rate of PDD within PD.",
@@ -54,8 +54,8 @@ table_estimands <- function() {
       grp = "RO",
       nms = "Statement"
     ) |>
-    tab_spanner(columns = c("Theoretical", "Population"), label = "Theoretical Estimand", gather = FALSE) |>
-    cols_label(
+    gt::tab_spanner(columns = c("Theoretical", "Population"), label = "Theoretical Estimand", gather = FALSE) |>
+    gt::cols_label(
       Theoretical ~ "Quantity",
       Empirical ~ "Empirical Estimand",
       Estimate ~ "Statistical Estimate"

### 2025-10-09 R/table_levelII_approximations.R

@@ -10,9 +10,9 @@
 #'    \code{describe_concordance}.
 #' @param struct A list with two elements:
 #' \describe{
-#'   \item{[[1]]}{Character vector of Level I algorithm names derived using FAQ total > 7
+#'   \item{`[[1]]`}{Character vector of Level I algorithm names derived using FAQ total > 7
 #'   as the IADL criterion.}
-#'   \item{[[2]]}{Character vector of Level I algorithm names derived using FAQ item 9 > 1
+#'   \item{`[[2]]`}{Character vector of Level I algorithm names derived using FAQ item 9 > 1
 #'   as the IADL criterion. Note: order matters.}
 #' }
 #' @param crit A character string indicating which Level II algorithm is used
@@ -24,55 +24,55 @@
 #'
 #' @export
 table_levelII_approximations <- function(values, struct, crit) {
-  map_dfr(1:2, function(ii) {
+  purrr::map_dfr(1:2, function(ii) {
     values |>
-      filter(reference == paste0("Lvl.II (", ii, ")")) |>
-      filter(predictor %in% struct[[ii]]) |>
-      filter(!grepl("Lvl.II", predictor)) |>
-      arrange(desc(get(crit))) |>
-      select(reference, predictor, Kappa_raw, Accuracy_raw, AccuracyPValue, Sensitivity, Specificity) |>
-      mutate(order = seq_len(length(reference)))
+      dplyr::filter(reference == paste0("Lvl.II (", ii, ")")) |>
+      dplyr::filter(predictor %in% struct[[ii]]) |>
+      dplyr::filter(!grepl("Lvl.II", predictor)) |>
+      dplyr::arrange(dplyr::desc(get(crit))) |>
+      dplyr::select(reference, predictor, Kappa_raw, Accuracy_raw, AccuracyPValue, Sensitivity, Specificity) |>
+      dplyr::mutate(order = seq_len(length(reference)))
   }) |>
-    pivot_wider(
+    tidyr::pivot_wider(
       names_from = reference,
       id_cols = order,
       values_from = -c(reference, order)
     ) |>
-    mutate(
-      part = case_when(
+    dplyr::mutate(
+      part = dplyr::case_when(
         order %in% 1:5 ~ "Top five",
         order %in% (length(order)-4):length(order) ~ "Bottom five"
       ),
-      across(contains("PValue"), \(x) do_summary(x, 3, "p")),
-      across(where(is.numeric), \(x) do_summary(x, 2))
[…]

### 2025-10-09 _targets_packages.R

@@ -1,15 +1,0 @@
-# Generated by targets::tar_renv(): do not edit by hand
-library(bslib)
-library(caret)
-library(crew)
-library(gt)
-library(here)
-library(janitor)
-library(markdown)
-library(psych)
-library(rstudioapi)
-library(shiny)
-library(shinyWidgets)
-library(shinybusy)
-library(tidyverse)
-library(visNetwork)

### 2025-10-09 _manuscript/manuscript.qmd
Aligning the quarto version with current manuscript on GDrive for easier manipulation during review.

--- 2025-10-08 _manuscript/manuscript.qmd
@@
 tar_source(here::here("R"))
+
 data <- tar_read(raw_data, store = upstore)
@@
 According to a recent meta-analysis, approximately one-quarter of PD patients is likely to be diagnosed with PDD [@sousa2022].
-However, reported PDD rate estimates vary widely, ranging from 14% up to 55%, depending on methodological criteria employed [@sousa2022].
+However, reported PDD rate estimates vary widely, ranging from 14% up to 55%, depending on the methodological criteria employed [@sousa2022].
 Moreover, factors such as patients' sex [@cereda2016], age and disease duration appear to modulate the risk of cognitive decline and PDD [@oh2016; @rana2012].
@@
 A milestone in research of PDD was the publication of diagnostic criteria established in 2007 by the International Parkinson and Movement Disorder Society (MDS) [@dubois2007].
-In these citeria, the MDS introduced a two-levelled system for PDD detection.
+In these criteria, the MDS introduced a two-levelled system for PDD detection.
 Level I consists of brief cognitive assessments, while Level II involves comprehensive neuropsychological testing across cognitive domains [@emre2007].
@@
 The original Level I algorithm included eight conditions that had to be satisfied simultaneously in order to diagnose probable PDD.
-These included: 1) diagnosis of PD proposed by the Queen Square Brain Bank; 2) PD onset prior to the PDD emergence; 3) evidence of global cognitive impairment (MMSE score \< 26 points); 4) cognitive deficit interference with the IADL (assessed by the pill questionnaire or caregiver interview); 5) impairment in at least two cognitive domains, namely memory, attention, visuo-constructive abilities and executive functions; 6) there was absence of Major Depressive Disorder; 7) absence of delirium; and 8) exclusion of other abnormalities and potential causes of dementia [@dubois2007].
+These included: 1) diagnosis of PD proposed by the Queen Square Brain Bank; 2) PD onset prior to the PDD emergence; 3) evidence of global cognitive impairment (MMSE score \< 26 points); 4) cognitive deficit interference with the IADL (assessed by the pill questionnaire or caregiver interview); 5) impairment in at least two cognitive domains, namely memory, attention, visuo-constructive abilities and executive function; 6) there was absence of Major Depressive Disorder; 7) absence of delirium; and 8) exclusion of other abnormalities and potential causes of dementia [@dubois2007].
 
@@
 A recent call for a change pinpoints limitations regarding the original criteria and suggest various updates to enhance their utility [@kulisevsky2024].
-Proposed suggestions include replacement of Mini Mental State Examination (MMSE) by Montreal Cognitive Assessment (MoCA), which is more sensitive to PD specific cognitive impairment; expansion of instrumental activities of daily living (IADL) evaluation; inclusion of language assessment; recognition of anxiety as one the neuropsychiatric symptoms relevant in PDD; and intergration of biomarkers.
+Proposed suggestions include replacement of Mini Mental State Examination (MMSE) by Montreal Cognitive Assessment (MoCA), which is more sensitive to PD specific cognitive impairment; expansion of instrumental activities of daily living (IADL) evaluation; inclusion of language assessment; recognition of anxiety as one **of** the neuropsychiatric symptoms relevant in PDD; and intergration of biomarkers.
 
@@
 
-Cognitive performance was evaluated at both Level I and Level II according to the standard MDS battery for Parkinson's Disease Mild Cognitive Impairment (PD-MCI) [@litvan2012, @bezdicek2017].
+Cognitive performance was evaluated at both Level I and Level II according to the standard MDS battery for Parkinson’s Disease Mild Cognitive Impairment (PD-MCI) [@litvan2012, @bezdicek2017].
 Cognitive performance at Level I was assessed by the Mini-Mental State Examination (MMSE) [@stepankova2014; @folstein1975] and the Montreal Cognitive Assessment (MoCA) [@kopecek2016; @nasreddine2005].
-The comprehensive neuropsychological assessment at Level II evaluated five cognitive domains through specific test: attention and working memory assessed by Trail Making Test Part A (TMT-A) [@bezdicek2012; @reitan2004], and WAIS Digit Span Backward (WAIS DSB) [@wechsler1997], executive functions by Categorical Verbal Fluency (CF) [@nikolai2015], and subtest from the Prague Stroop Test – Colors (PST-C) [@bezdicek2021], language by the WAIS Similarities subtest [@wechsler1997], and the Boston Naming Test (BNT-60) [@zemanova2016; @kaplan1983], memory by the Rey Auditory Verbal Learning Test (RAVLT) [@frydrychova2018; @bezdicek2013; @rey1964] delayed recall, and the Brief Visual Memory Test–Revised (BVMTR) [@havlik2020; @benedict1997] delayed recall, or WAIS Family Pictures subtest [@wechsler1997] delayed recall, visuospatial function assessed by the Judgment of Line Orientation Test (JoLO) [@benton1983], and Clock Drawing Test (CLOX) [@royall1998].
+The comprehensive neuropsychological assessment at Level II evaluated five cognitive domains through specific tests: attention and working memory assessed by Trail Making Test Part A (TMT-A) [@bezdicek2012; @reitan2004], and WAIS Digit Span Backward (WAIS DSB) [@wechsler1997], executive function by Categorical Verbal Fluency (CF) [@nikolai2015], and subtest from the Prague Stroop Test – Colors (PST-C) [@bezdicek2021], language by the WAIS Similarities subtest [@wechsler1997], and the Boston Naming Test (BNT-60) [@zemanova2016; @kaplan1983], memory by the Rey Auditory Verbal Learning Test (RAVLT) [@frydrychova2018; @bezdicek2013; @rey1964] delayed recall, and the Brief Visual Memory Test–Revised (BVMTR) [@havlik2020; @benedict1997] delayed recall, or WAIS Family Pictures subtest [@wechsler1997] delayed recall, visuospatial function assessed by the Judgment of Line Orientation Test (JoLO) [@benton1983], and Clock Drawing Test (CLOX) [@royall1998].
 
@@
 The Beck Depression Inventory-II (BDI-II) [@ciharova2020; @beck1996] and State-Trait Anxiety Inventory (STAI) [@spielberg1983; @mullner1980] were used to assess neuropsychiatric status.
-Psychotic symptoms were assessed through structured psychiatric interviews conducted by a trained psychiatrist.
+**BDI-II and STAI are a part of the standard methods in our pre-DBS protocol, having well-established psychometric properties in PD populations [@leentjens2000; @dissanayaka2010].
+Other psychiatric scales data, such as the Neuropsychiatric Inventory, were not available for our retrospective dataset.
+However, all patients underwent a comprehensive evaluation by a trained neuropsychiatrist, excluding the presence of major depression, apathy, psychosis, and visual hallucinations, ensuring that confounding neuropsychiatric conditions were not present in our sample.**
 
@@
 The third approach applied the Czech version of the shortened Montreal Cognitive Assessment (sMoCA) [@bezdicek2020], a time-efficient modification designed to measure global cognitive performance using a reduced testing protocol that omits items providing redundant information.
+**The sMoCA has been validated in Czech PD cohorts [@bezdicek2020] and shown to be sensitive to cognitive deficits while lowering patient burden [@roalf2017a].
+Inclusion of the sMoCA in our study was motivated by its clinical utility in pre-surgical settings, where time restrictions and patients’ fatigue often limit the feasibility of longer assessments.
+Moreover, its inclusion allowed us to compare diagnostic performance across different screening lengths.**
+
 Lastly, the fourth approach followed the Level II protocol for diagnosis of PDD and Mild Cognitive Impairment in PD (PD-MCI) [@dubois2007; @litvan2012].
@@
 
-Following the framework proposed by @lundberg2021, in this study we explicitly connect our research objectives and their corresponding theoretical (i.e., targets of inference) and empirical (i.e., data-driven) estimands to statistical estimates.
+Following the framework proposed @lundberg2021, in this study we explicitly connect our research objectives and their corresponding theoretical (i.e., targets of inference) and empirical (i.e., data-driven) estimands to statistical estimates.
 The theoretical estimand refers to a unit-specific quantity defined over a target population and represents the ideal quantity that would address the research question under optimal conditions, such as access to complete population data or perfect experimental control.
 In contrast, the empirical estimand corresponds to the quantity that is actually computable using the available dataset, given real world constraints.
-The full description of the study's estimands and their relation to our research objectives is presented in the Appendix (see @tbl-estimands).
+The full description of the study’s estimands and their relation to our research objectives is presented in the Appendix (see @tbl-estimands).
 
 To address study objectives, we started by repeatedly assigning each patient the diagnosis of probable PDD based on each PDD algorithm listed in @tbl-crits (see also @tbl-algos) resulting in a `r sum(data$incl)` (patients) $\times$ `r sum(critnum)` (algorithms) matrix where each cell indicates whether a patient (row) meets criteria for probable PDD according to an algorithm (column).
-PDD rate estimates were computed as $\frac{N_{PDD}}{N_{total}}$ separately for each algortihm.
+PDD rate estimates were computed as $\frac{N_{PDD}}{N_{total}}$ separately for each algorithm.
 The predictive value of age and sex was then evaluated by fitting a set of logistic regressions, one for each algorithm for probable PDD, whereby the probable PDD was predicted by age, sex and their interaction.
@@
 
+# Function for writing M (SD):
+msd_txt <- function(x, txt = "", tab = desc$table) {
+  paste0(
+    subset(tab, variable == x)[ , c("M", "SD")] |> paste(collapse = " (SD = "),
+    ")",
+    txt
+  )
+}
+
 # In-text demographic data
 sex <- subset(desc$table, variable == "Sex")$N
-age <- paste0(subset(desc$table, variable == "Age")[ , c("M", "SD")] |> paste(collapse = " (SD = "),") years of age")
-edu <- paste0(subset(desc$table, variable == "Education")[ , c("M", "SD")] |> paste(collapse = " (SD = ")," ) years of education")
-dur <- paste0(subset(desc$table, variable == "PD duration")[ , c("M", "SD")] |> paste(collapse = " (SD = "),") years of disease duration")
-updrs_off <- paste0(subset(desc$table, variable == "UPDRS III off state")[ , c("M", "SD")] |> paste(collapse = " (SD = "),") Unified Parkinson Disease Rating Scale (UPDRS), part III in medication OFF state")
-updrs_on <- paste0(subset(desc$table, variable == "UPDRS III on state")[ , c("M", "SD")] |> paste(collapse = " (SD = "),") UPDRS III in medication ON state")
+age <- msd_txt("Age", " years of age")
+edu <- msd_txt("Education", " years of education")
+dur <- msd_txt("PD duration", " years of disease duration")
+updrs_off <- msd_txt("UPDRS III off state", " Unified Parkinson Disease Rating Scale (UPDRS), part III in medication OFF state")
+updrs_on <- msd_txt("UPDRS III on state", " Unified Parkinson Disease Rating Scale (UPDRS), part III in medication ON state")
+bdi <- msd_txt("BDI")
+staix1 <- msd_txt("STAI X1")
+staix2 <- msd_txt("STAI X2")
 
@@
 The sample included `r sex` men, with an average of `r age`, `r edu`, `r dur`, `r updrs_off` and `r updrs_on`.
-Cognitive characteristics of the sample are summarized in @tbl-desc.
+**Descriptive statistics for neuropsychiatric symptoms indicated mild depressive and anxiety symptoms in our cohort, with the average BDI-II of `r bdi`, average STAI X1 of `r staix1`, and average STAI X2 of `r staix2`.
+However, according to the psychiatric assessment, none of the patients with probable PDD was suffering major depressive disorder, delirium or other neuropsychiatric abnormalities that would disqualify the diagnosis.** Cognitive characteristics of the sample are summarized in @tbl-desc.
 
@@
 This lack of age-dependency may, however, be also stemming from the relatively younger age of our cohort, because previous reports indicate that association between age and PDD is not linear but increases with age and may not reach substantial values before the age of 70 [@rana2012].
+**In our sample, the younger age and preserved cognitive level might limit age-related effects.
+This fact could bias algorithm results toward higher specificity but lower sensitivity [@larvin2023].
+Previous studies report that DBS candidates tend to be, on average, younger and cognitively less impaired than the general PD population [@bronstein2011], which could bias diagnostic algorithm performance toward overestimating specificity.
+Therefore, our findings may not be fully generalizable to older or more cognitively impaired people with PD.**
 
@@
 These findings imply that simplified tools such as sMoCA may be a promising tool for heightening efficiency of the assessment while maintaining diagnostic accuracy.
+**This underscores the clinical utility of sMoCA screening tool.
+Although raw accuracy was moderate, balanced accuracy results ensure avoidance of bias from class imbalance.
+From a practical point of view, it suggests that even brief screening tools can achieve reliable classification, which means an advantage for routine clinical settings and time-restricted assessments.**
 
@@
 Nevertheless, the analysis of the algorithm concordance might not be affected by the sample to the same degree allowing for a broader generalization.
-Future studies should aim to replicate our results on larger and different cohorts of PD patients.
+**However, the use of a DBS cohort offered certain methodological advantages.
+For example, all patients underwent standardized and comprehensive neuropsychological testing, which resulted in well-characterized and high-quality dataset, allowing us to systematically evaluate multiple diagnostic algorithms.** Future studies should aim to replicate our results on larger and different cohorts **with more heterogenous sample**.
 To make this process easier, the code used to generate our results is publicly available and easily applicable to similarly structured data.
@@
 FAQ is subjective and informant-reliant, which may result in biased data.
-For future research, we suggest the use of more objective, modernized tools for IADl assessment, such as performance-based methods [@schmitter-edgecombe2020] or questionnaire adaptation including questions regarding gadgets use and digital literacy [@postema2024].
+**At the same time, the questionnaire may vary across cultures, therefore, the results may not be transferable.
+For example, activities such as financial management or driving, may not be equally practiced across cultures or contexts, which might complicate interpretation of impairment levels.
+The FAQ scores may vary across cultures, and diagnostic thresholds may not be transferable across countries [@odonald2025].
+This may influence both the sensitivity and ecological validity of the functional criteria used.
+Another limitation is that the FAQ scores may correlate with neuropsychological results of certain domains, specifically executive functions or memory [@vanrentergem2020; @moheb2017].
+This overlap could artificially increase concordance between cognitive and functional criteria.
+For future research, we suggest using a PD-specific questionnaire, such as the Penn Parkinson's Daily Activities Questionnaire-15 [@brennan2016] or generally more objective methods, preferably performance-based ones.
+Our results underscore the importance of IADL measurement for the PDD diagnosis.
+Therefore, we recommend using more reliable tools with high ecological validity.** For future research, we suggest the use of more objective, modernized tools for IADl assessment, such as performance-based methods [@schmitter-edgecombe2020] or questionnaire adaptation including questions regarding gadgets use and digital literacy [@postema2024].
 Our results emphasize the importance of IADL deficit for the PDD diagnosis, therefore, we suggest exploring more reliable tools with high ecological validity.
 
-Finally, whereas our study systematically investigated how varying definition of global deficit, impaired cognition and IADL deficit affect probable PDD classification, it did not explore associations of PDD diagnosis with its neuropsychiatric (e.g.
+**Additionally, the current study did not explore associations of PDD diagnosis with its neuropsychiatric correlates and biomarkers.
+We recommend examining models where psychiatric symptoms and biomarkers are treated as covariates to better capture their impact on PDD diagnosis.**
+
+??Finally, whereas our study systematically investigated how varying definition of global deficit, impaired cognition and IADL deficit affect probable PDD classification, it did not explore associations of PDD diagnosis with its neuropsychiatric (e.g.
 anxiety profile) and biomarker correlates.
-Future studies would benefit from focusing on these aspects of PDD.
+Future studies would benefit from focusing on these aspects of PDD.??^[**JM: Why was the previous paragraph added instead of this one?
+This looks perfectly functional.
+The same goes largely for the IADL discussion, the original was better (in terms of resources and logical coherence) at most places.**]
 
@@
 Specifically, as the rate goes from extremes to 0.5, the variance increases.
-Consequently, if our estimate of the rate was lower than the true population rate, e.g., because our sample includes younger patients compared to the general PD population, our estimate of variance would also lower than the true variance of PDD rate in the general PD population.
-Nonetheless, the between-algorithm variability may not be affected by this phenomenon as unlike variability of PDD rate, we do not have reason to assume it comes about by summing independent binomial events
+Consequently, if our estimate of the rate was lower than the true population rate, e.g., because our sample includes younger patients compared to the general PD population, our estimate of variance would also be lower than the true variance of PDD rate in the general PD population.
+Nonetheless, the between-algorithm variability may not be affected by this phenomenon as unlike variability of PDD rate, we do not have reason to assume it comes about by summing independent binomial events.
 

### 2025-10-10 R/compute_descriptives.R
Documentation proofread.

@@ -1,23 +1,27 @@
-#' Compute descriptive statistics for selected sample variables
+#' Describe the Sample
 #'
 #' Computes descriptive summaries for a specified set of variables in a dataset.
 #' The output includes a raw summary tibble and an APA-style formatted gt table.
 #'
 #' @param d0 A tibble containing the data, as generated by \code{prepare_data}.
-#' @param vois A data frame, tibble, or matrix with the following columns:
-#'   (1) variable name,
-#'   (2) variable label,
-#'   (3) variable type (one of "continuous", "binary", "nominal"),
-#'   (4) optional: variable group,
-#'   (5) optional: label-to-description mapping for table notes,
-#'   (6) optional: score type description to append in parentheses.
-#'   Alternatively, a path to a CSV file (semicolon-delimited) containing this information.
+#' @param vois A data frame, tibble, or matrix with the following columns (in this order):
+#'
+#' 1. variable name,
+#' 2. variable label,
+#' 3. variable type (one of "continuous", "binary", "nominal"),
+#' 4. optional: variable group,
+#' 5. optional: label-to-description mapping for table notes,
+#' 6. optional: score type description to append in parentheses.
+#'
+#' Alternatively, a path to a CSV file (semicolon-delimited) containing this information.
 #'
 #' @returns A list containing:
 #'   \describe{
 #'     \item{\code{table}}{A tibble with raw descriptive statistics.}
-#'     \item{\code{gtable}}{An APA-style formatted gt table summarizing selected variables.}
+#'     \item{\code{gtable}}{An APA-style formatted {gt} table summarizing selected variables.}
 #'   }
+#'
+#' @seealso [prepare_data()] prepares `d0`.
 #'
 #' @examples
 #' \dontrun{

### 2025-10-10 R/specify_algorithms.R

@@ -1,17 +1,20 @@
-#' Specify criteria operationalisations for PDD diagnosis
+#' Specify Algorithms for Probable PDD
 #'
 #' Computes a Cartesian product of possible diagnostic components to define
-#' all combinations of operationalised criteria for probable Parkinson’s disease
-#' dementia (PDD). This function is a helper and is typically called within
+#' all combinations of algorithms for probable Parkinson’s disease dementia (PDD).
+#' This function is a helper with no inputs and is typically called within
 #' \code{diagnose_pdd_sample}.
 #'
-#' @returns A tibble where each row represents one unique operationalisation
-#'    of PDD diagnostic criteria. Columns indicate the type of criteria, individual
-#'    test indexes, thresholds, and other relevant parameters.
+#' @returns A tibble where each row represents one unique diagnostic algorithm
+#'    for probable PDD. Columns indicate the type of algorithms, individual test
+#'    indexes, thresholds, and other relevant parameters.
+#'
+#' @seealso [diagnose_pdd_sample()] combines the algorithms with data to diagnose
+#'    single patients.
 #'
 #' @export
-specify_criteria <- function() {
-  # MMSE-based criteria
+specify_algorithms <- function() {
+  # MMSE-based algorithms
   mmse <- tidyr::crossing(exec = c("vf_s","cloc"), iadl = c("faq_9","faq")) |>
     dplyr::mutate(
       group = "mmse",
@@ -32,7 +35,7 @@ specify_criteria <- function() {
     dplyr::relocate(exec, .after = atte_t) |>
     dplyr::relocate(exec_t, .after = exec) |>
     dplyr::relocate(iadl, .before = iadl_t)
-  # MoCA-based criteria
+  # MoCA-based algorithms
   moca <- purrr::map_dfr(2:3, function(i) {
     tidyr::crossing(
       group = "moca",
@@ -55,7 +58,7 @@ specify_criteria <- function() {
       )
   }) |>
     mutate(type = paste0("MoCA (", seq_len(n()), ")"))
-  # sMoCA-based criteria
+  # sMoCA-based algorithms
   smoca <- tidyr::crossing(
     group = "smoca",
     glob = "smoca_total",
@@ -66,7 +69,7 @@ specify_criteria <- function() {
       type = paste0("sMoCA (", seq_len(dplyr::n()), ")"),
[…]

### 2025-10-10 R/table_estimands.R

@@ -1,4 +1,4 @@
-#' Generate a summary table of study estimands
+#' Generate a Summary Table of Estimands
 #'
 #' This function creates a pre-specified table that provides a verbal summary
 #' of the study's estimands for inclusion in the Appendix. It is intended as a
@@ -9,6 +9,8 @@
 #' @returns A \code{gt} table object summarising the theoretical estimands,
 #'    empirical estimands, and the corresponding statistical estimates used
 #'    in the study.
+#'
+#' @seealso [gt_apa_table()] is used to format the table.
 #'
 #' @export
 table_estimands <- function() {

### 2025-10-10 R/table_levelII_approximations.R

@@ -1,26 +1,53 @@
-#' Generate summary table of top and bottom Level I algorithms
-#' in predicting Level II-derived probable PDD
+#' Summary Table of Predictors
 #'
-#' This helper function prepares a table showing the five best and
-#' five worst-performing Level I algorithms in predicting a selected
-#' Level II-based diagnosis of probable PDD. It is primarily intended
-#' for use in the manuscript to keep Quarto documents uncluttered.
+#' This helper function prepares a table showing the five best and five
+#' worst-performing Level I algorithms in predicting a selected Level II-based
+#' diagnosis of probable PDD. It is primarily intended for use in the manuscript
+#' to keep Quarto document uncluttered.
 #'
 #' @param values A tibble of concordance statistics generated by
-#'    \code{describe_concordance}.
-#' @param struct A list with two elements:
+#'    \code{describe_concordance}. Can also be provided by the data [concords]
+#'    propagated with the package.
+#' @param struct A list generated by \code{list_algorithms} with two elements:
 #' \describe{
-#'   \item{`[[1]]`}{Character vector of Level I algorithm names derived using FAQ total > 7
-#'   as the IADL criterion.}
-#'   \item{`[[2]]`}{Character vector of Level I algorithm names derived using FAQ item 9 > 1
-#'   as the IADL criterion. Note: order matters.}
+#'   \item{`[[1]]`}{Character vector of Level I algorithm names derived using
+#'   FAQ total > 7 as the IADL criterion.}
+#'   \item{`[[2]]`}{Character vector of Level I algorithm names derived using
+#'   FAQ item 9 > 1 as the IADL criterion. Note: order matters.}
 #' }
-#' @param crit A character string indicating which Level II algorithm is used
-#' as the reference for comparison.
+#' @param crit A character string indicating which concordance statistics is to
+#'    be used for comparisons. Can be one of `"Accuracy_raw"`, `"Balanced Accuracy"`,
+#'    `"Kappa_raw"`, `"Sensitivity"` or `"Specificity"`.
 #'
 #' @returns A \code{gt} table object summarising the concordance statistics
-#' (e.g., accuracy, kappa) of the five best and five worst Level I algorithms
-#' in predicting the specified Level II-based diagnosis of probable PDD.
+#'    (e.g., accuracy, kappa) of the five best and five worst Level I algorithms
[…]

### 2025-10-10 _manuscript/manuscript.qmd
Documentation updated.

@@
 data <- tar_read(raw_data, store = upstore)
-crit <- tar_read(pdd_data, store = upstore)$criteria
+crit <- tar_read(pdd_data, store = upstore)$algorithms
 vars <- tar_read(variables, store = upstore)
@@
 continue <- !any(critnum == 0)
-stopifnot("Some criteria are missing" = continue)
+stopifnot("Some algorithms are missing" = continue)
 ```
@@
 
-table_criteria(vars)
+table_algorithms(vars)
 ```

### 2025-10-11 _manuscript/manuscript.qmd
Extensive answer to R2/3 about missingness added to the manuscript.

@@
 format:
- # apaquarto-pdf:
-   # documentmode: man
-   # keep-tex: true
+ apaquarto-pdf:
+    documentmode: man
+    keep-tex: true
  apaquarto-docx: 
@@
 data <- tar_read(raw_data, store = upstore)
+vars <- tar_read(variables, store = upstore)
 crit <- tar_read(pdd_data, store = upstore)$algorithms
-vars <- tar_read(variables, store = upstore)
+rates <- tar_read(rate_summaries, store = upstore)
+
+pres <- tar_read(pdd_data, store = upstore)$PDD |>
+  mutate(present = if_else(is.na(PDD), 0, 1)) |>
+  select(id, type, present) |>
+  pivot_wider(names_from = type, values_from = present) |>
+  column_to_rownames("id")
 ```
@@
 ```{r}
-#| label: "crit-check"
+#| label: "algo-check"
 
-critnum <- sapply(c("mmse", "moca", "smoca", "lvlII"), function(i) {
+algonum <- sapply(c("mmse", "moca", "smoca", "lvlII"), function(i) {
   subset(crit, group == i) |>
@@
 
-continue <- !any(critnum == 0)
+continue <- !any(algonum == 0)
 stopifnot("Some algorithms are missing" = continue)
@@
 Second, we applied the entire Functional Activities Questionnaire (FAQ) as suggested in the call for change [@kulisevsky2024], employing a cut-off score of 7 points or higher based on Czech normative data [@bezdicek2011].
-These methodologies resulted in a total of `r sum(critnum)` algorithms, which were distributed across different diagnostic criteria: `r critnum["mmse"]` MMSE-based, `r critnum["moca"]` MoCA-based, `r critnum["smoca"]` sMoCA-based, and `r critnum["lvlII"]` based on the Level II battery (see @tbl-crits and Appendix @tbl-algos for the exact specification of each algorithm).
+These methodologies resulted in a total of `r sum(algonum)` algorithms, which were distributed across different diagnostic criteria: `r algonum["mmse"]` MMSE-based, `r algonum["moca"]` MoCA-based, `r algonum["smoca"]` sMoCA-based, and `r algonum["lvlII"]` based on the Level II battery (see @fig-tree, @tbl-crits and Appendix @tbl-algos for the exact specification of each algorithm).
 
@@
 
-To address study objectives, we started by repeatedly assigning each patient the diagnosis of probable PDD based on each PDD algorithm listed in @tbl-crits (see also @tbl-algos) resulting in a `r sum(data$incl)` (patients) $\times$ `r sum(critnum)` (algorithms) matrix where each cell indicates whether a patient (row) meets criteria for probable PDD according to an algorithm (column).
+To address study objectives, we started by repeatedly assigning each patient the diagnosis of probable PDD based on each PDD algorithm listed in @tbl-crits (see also @tbl-algos) resulting in a `r sum(data$incl)` (patients) $\times$ `r sum(algonum)` (algorithms) matrix where each cell indicates whether a patient (row) meets criteria for probable PDD according to an algorithm (column).
 PDD rate estimates were computed as $\frac{N_{PDD}}{N_{total}}$ separately for each algorithm.
@@
 
-Data wrangling and visualizations were done in the *tidyverse* package [@wickham2019] and tables were formatted in the *gt* package [@iannone2024].
+**Missing data were handled by complete cases analysis and pairwise complete cases analysis.
+In other words, each univariate and analysis pairwise comparison used all available data.** Data wrangling and visualizations were done in the *tidyverse* package [@wickham2019] and tables were formatted in the *gt* package [@iannone2024].
 All analyses were conducted within the R (version `r with(version, paste(major, minor, sep="."))`) software environment for statistical computing [@rsoft].
-The software code supporting this article is available at <https://github.com/josefmana/DemCr1t.git>.[^1]
-
-[^1]: **JM: Do not forget to make it public before submitted!**
+The software code supporting this article is available at <https://github.com/josefmana/demcrit.git>.
 
@@
 desc  <- tar_read(sample_description, store = upstore)
-rates <- tar_read(rate_summaries, store = upstore)
 algos <- tar_read(algorithms, store = upstore)
@@
 perc  <- rates$table$perc
+
+# Total numbers:
+n_total <- sum(data$incl)
+id_excl <- names(which(rowSums(pres) == 0))
+n_excl <- length(id_excl)
+n_incl <- n_total - n_excl
+if (n_excl == 1) {
+  excl_txt <- paste0("one patient was excluded due to missing neuropsychological data, resulting in a final sample of ", n_incl, " patients")
+} else if (n_excl > 1) {
+  excl_txt <- paste0(english::english(n_excl), " patients were excluded due to missing neuropsychological data, resulting in a final sample of ", n_incl, " patients")
+} else {
+  excl_txt <- "no patients was excluded."
+}
 
@@
 
-A total of `r sum(data$incl)` patients were included.
+**Total of** `r n_total` patients **were considered for the study, out of which `r excl_txt`**.
 The sample included `r sex` men, with an average of `r age`, `r edu`, `r dur`, `r updrs_off` and `r updrs_on`.
 **Descriptive statistics for neuropsychiatric symptoms indicated mild depressive and anxiety symptoms in our cohort, with the average BDI-II of `r bdi`, average STAI X1 of `r staix1`, and average STAI X2 of `r staix2`.
-However, according to the psychiatric assessment, none of the patients with probable PDD was suffering major depressive disorder, delirium or other neuropsychiatric abnormalities that would disqualify the diagnosis.** Cognitive characteristics of the sample are summarized in @tbl-desc.
+However, according to the psychiatric assessment, none of the patients with probable PDD was suffering from the major depressive disorder, delirium or other neuropsychiatric abnormalities that would disqualify the diagnosis.** Cognitive characteristics of the sample are summarized in @tbl-desc.
 
@@
 
-Algorithm-wise rate of PDD estimates are presented in @tbl-rates.
+```{r}
+#| label: "observations-univariate"
+
+obs <- rates$table|>
+  count(N) |>
+  mutate(perc_missing = glue::glue("{round(100 * (n_incl-N) / n_incl, 1)}%"))
+```
+
+Algorithm-wise rate of PDD estimates **corresponding numbers of patients with available data** are presented in @tbl-rates.
+**In total, there were `r obs[1, "n"]` algorithms based on `r obs[1, "N"]` complete observations (`r obs[1, "perc_missing"]` missing).
+One algorithm used `r obs[2, "N"]`, `r obs[3, "N"]`, and `r obs[4, "N"]` complete observations, respectively (`r obs[2, "perc_missing"]`, `r obs[3, "perc_missing"]` and `r obs[4, "perc_missing"]` missing).
+A further `r obs[5, "n"]` algorithms were based on `r obs[5, "N"]` complete observations (`r obs[5, "perc_missing"]` missing), and `r obs[6, "n"]` algorithms included the full sample of `r obs[6, "N"]` patients.**
+
 On average, estimated PDD rate was `r rates_overall["M"]`% (SD = `r rates_overall["SD"]`, Md = `r rates_overall["Md"]`, range `r rates_overall["minmax"]`).
-Notably, the estimates were substantially lower when FAQ item 9 was used as a criterion of IADL deficit (M = `r rates_FAQ9["M"]`% SD = `r rates_FAQ9["SD"]`, Md = `r rates_FAQ9["Md"]`, range `r rates_FAQ9["minmax"]`) compared to using total FAQ score criterion (M = `r rates_FAQtot["M"]`% SD = `r rates_FAQtot["SD"]`, Md = `r rates_FAQtot["Md"]`, range `r rates_FAQtot["minmax"]`) as demonstrated also in @fig-rates.
+Notably, the estimates were substantially lower when FAQ item 9 was used as a criterion of IADL deficit (M = `r rates_FAQ9["M"]`% SD = `r rates_FAQ9["SD"]`, Md = `r rates_FAQ9["Md"]`, range `r rates_FAQ9["minmax"]`) compared to using total FAQ score criterion (M = `r rates_FAQtot["M"]`% SD = `r rates_FAQtot["SD"]`, Md = `r rates_FAQtot["Md"]`, range `r rates_FAQtot["minmax"]`) as demonstrated in @fig-tree **(see also @fig-rates for per-operationalisation distribution of PDD rate estimates)**.
 Neither age nor sex or their interaction (*p*s $\geq$ `r min_p`) reliably predicted probable PDD classification across algorithms (see @fig-data and @fig-pars).
@@
 ```{r}
-#| label: "fig-rates"
-#| fig-cap: "Summary of the estimates of probable PDD rate."
-#| apa-note: "Vertical lines represent estimates arrived at by using sMoCA (dotted) or Level II (dashed) with FAQ item 9 (orange) or FAQ total score (blue) as criteria for probable PDD."
+#| label: "fig-tree"
+#| fig-cap: "A dendrogram representing algorithms creation."
 
-rates$plot
+make_dendrogram(rates$table)
 ```
@@
 
+```{r}
+#| label: "observations-bivariate"
+
+dims <- glue::glue("{dim(conc$table)[1]} rows x {dim(conc$table)[2]} columns")
+
+obs_pairs <- conc$table |>
+  count(N) |>
+  drop_na() |> # NA for cases where reference == predictor
+  mutate(perc = 100 * n / sum(n))
+
+min_miss <- glue::glue("{round(100 * (n_incl - obs_pairs[1, 'N']) / n_incl, 2)}%")
+
+K <- 200
+perc_above_k <- obs_pairs |>
+  filter(N >= K) |>
+  pull(perc) |>
+  sum() |>
+  do_summary(2)
+```
+
 Results of the analyses of prediction Accuracy, Cohen's $\kappa$, Sensitivity and Specificity are presented in @fig-acc, @fig-kappa, @fig-sens and @fig-spec respectively.
-**Numerical results are available at ...** [^2].
+**The number of complete pairwise cases ranged from `r obs_pairs[1, "N"]` (`r min_miss` missing) to `r obs_pairs[nrow(obs_pairs), "N"]` (full sample).
+Most comparisons (`r perc_above_k`%) were based off of `r K` or more observations.**[^1]
+
+[^1]: **Due to the large number of entries (`r dims` representing pairwise comparisons and metrics of interest respectively), the table with numerical results is not presented here or in the Appendix.
+Instead, we share the table share as data in the accompanying R package available at <https://github.com/josefmana/demcrit.git>.
+To obtain the table in format not dependent on R, follow the tutorial at <https://josefmana.github.io/demcrit/articles/concordance.html>**
+
 Generally, algorithms that employed the same operationalization of IADL deficit showed substantial pairwise concordance, however, algorithms that operationalized IADL deficit differently did not.
@@
 
-[^2]: **JM: Should provide an excel sheet or html table the readers could go through by themselves.**
-
-\[Insert Figure 2 here\]
+\[Insert Figure 3 here\]
 
@@
 
+**Due to the retrospective nature of the study, some patients lacked one or more key measures required for the diagnosis of probable PDD by certain algorithms.
+Missing data were handled using the pairwise complete cases method.
+The main advantages of this approach are its straightforward implementation and preservation of statistical power.
+However, it may introduce bias, particularly when data are not missing completely at random (MCAR) and when causal inference is the goal [@little2019].**
+
+**Modern alternatives, such as multiple imputation techniques, have been shown to produce less biased estimates than case deletion strategies in analyses based on confusion matrices [@karakaya2014], whereas the evidence supporting their superiority in the estimation of Cohen’s $\kappa$ remains limited [@deraadt2019].
+These modern approaches also require careful specification of the imputation model and the underlying causal mechanism of missingness to ensure appropriate covariate selection [@long2011; @bianco2023].
+For the sake of parsimony, we did not employ advanced missing-data techniques but instead explicitly described the observed missingness patterns.
+Consequently, our findings should be regarded as exploratory and primarily serve as a basis for hypothesis generation in future, more controlled studies.**
+
+**Another possible avenue for analysing the present data would be to exploit their hierarchical structure by fitting a multilevel model with algorithms nested within patients (see @fig-tree).
+Such an approach could leverage partial pooling [@gelman2012] and might approximate modern item response theory frameworks [@burkner2021], where PDD is conceptualised as a latent trait and the individual diagnostic algorithms act as items measuring it ().
+If feasible, this psychometric perspective may represent an exciting new direction for conceptualising complex clinical phenomena such as PDD [cf.
+@miller2012, @kiselica2021].**
+
 Furthermore, another limitation is the use of the FAQ questionnaire for IADL assessment.
@@
 anxiety profile) and biomarker correlates.
-Future studies would benefit from focusing on these aspects of PDD.??^[**JM: Why was the previous paragraph added instead of this one?
+Future studies would benefit from focusing on these aspects of PDD.??[^2]
+
+[^2]: **JM: Why was the previous paragraph added instead of this one?
 This looks perfectly functional.
-The same goes largely for the IADL discussion, the original was better (in terms of resources and logical coherence) at most places.**]
+The same goes largely for the IADL discussion, the original was better (in terms of resources and logical coherence) at most places.**
 
@@
 ```{r}
+#| label: "fig-rates"
+#| fig-cap: "Summary of the estimates of probable PDD rate."
+#| apa-note: "Vertical lines represent estimates arrived at by using sMoCA (dotted) or Level II (dashed) with FAQ item 9 (orange) or FAQ total score (blue) as criteria for probable PDD."
+
+rates$plot
+```
+
+```{r}
 #| label: "tbl-rates"

### 2025-10-12 _manuscript/manuscript.qmd
changes to manuscript

--- 2025-10-09 _manuscript/manuscript.qmd
@@
         department: "Department of Neurology and Centre of Clinical Neuroscience"
-abstract: ""
+abstract: "Background: A significant number of patients with Parkinson’s disease (PD) gradually progress to Parkinson’s disease dementia (PDD).
+A recent “Call for Change” (Kulisevsky et al., 2024) proposes updates to the current diagnostic criteria for PDD, including the use of different screening tests and broader functional assessments.
+Objective: We aimed to evaluate the diagnostic concordance between several algorithms for PDD based on Level I (i.e., screening) criteria and to assess their predictive validity for Level II (i.e., neuropsychological battery) diagnosis.
+Methods: We conducted a cross-sectional retrospective analysis of 204 patients diagnosed with PD.
+All patients underwent a comprehensive neuropsychological battery.
+A total of 68 diagnostic algorithms were operationalised based on the combinations of various cognitive assessments, including the Mini Mental State Exam (MMSE), the Montreal Cognitive Assessment (MoCA), and the shortened version of MoCA (sMoCA).
+Functional impairment was based on the Functional Assessment Questionnaire (FAQ), specifically FAQ item 9 or the FAQ total score.
+Concordance between algorithms was assessed using pairwise Cohen’s .
+The predictive validity of Level I algorithms for Level II diagnosis was assessed using Exact Binomial Tests.
+Results: PDD rate estimates varied widely across algorithms from 2.00% up to 16.75%.
+Higher rates were associated with functional impairment based on the definition of FAQ score ( = 0.75, = 0.86).
+Moderate to high concordance was observed among algorithms using the same functional impairment definition.
+MoCA-based algorithms most accurately predicted Level II classification, especially those using Clock Drawing to assess executive function.
+Conclusions: Diagnostic outcomes for PDD are sensitive to the choice of cognitive and functional instruments.
+Our results support some of the proposed changes to PDD diagnostic criteria on Level I, emphasising the consistency of MoCA-based assessments and comprehensive functional impairment evaluation."
 keywords:
   - Parkinson’s disease
-  - Parkinson's disease dementia
-  - level I criteria
+  - dementia
+  - diagnostic criteria
+  - functional assessment
+  - cognitive assessment
 format:
@@
 tar_source(here::here("R"))
-
 data <- tar_read(raw_data, store = upstore)
@@
 According to a recent meta-analysis, approximately one-quarter of PD patients is likely to be diagnosed with PDD [@sousa2022].
-However, reported PDD rate estimates vary widely, ranging from 14% up to 55%, depending on the methodological criteria employed [@sousa2022].
+However, reported PDD rate estimates vary widely, ranging from 14% up to 55%, depending on methodological criteria employed [@sousa2022].
 Moreover, factors such as patients' sex [@cereda2016], age and disease duration appear to modulate the risk of cognitive decline and PDD [@oh2016; @rana2012].
@@
 A milestone in research of PDD was the publication of diagnostic criteria established in 2007 by the International Parkinson and Movement Disorder Society (MDS) [@dubois2007].
-In these criteria, the MDS introduced a two-levelled system for PDD detection.
+In these citeria, the MDS introduced a two-levelled system for PDD detection.
 Level I consists of brief cognitive assessments, while Level II involves comprehensive neuropsychological testing across cognitive domains [@emre2007].
@@
 A recent call for a change pinpoints limitations regarding the original criteria and suggest various updates to enhance their utility [@kulisevsky2024].
-Proposed suggestions include replacement of Mini Mental State Examination (MMSE) by Montreal Cognitive Assessment (MoCA), which is more sensitive to PD specific cognitive impairment; expansion of instrumental activities of daily living (IADL) evaluation; inclusion of language assessment; recognition of anxiety as one **of** the neuropsychiatric symptoms relevant in PDD; and intergration of biomarkers.
+Proposed suggestions include replacement of Mini Mental State Examination (MMSE) by Montreal Cognitive Assessment (MoCA), which is more sensitive to PD specific cognitive impairment; expansion of instrumental activities of daily living (IADL) evaluation; inclusion of language assessment; recognition of anxiety as one the neuropsychiatric symptoms relevant in PDD; and intergration of biomarkers.
 
@@
 
-dates <- sapply(c("min", "max"), function(f) {
+dates <- sapply(
+  c("min", "max"), function(f) {
     date <- do.call(f, list(data$assdate))
-  paste(month(date, label = TRUE, abbr = FALSE), year(date))
-})
+    paste(month(date, label = T, abbr = F), year(date))
+  }
+)
 ```
@@
 
-Cognitive performance was evaluated at both Level I and Level II according to the standard MDS battery for Parkinson’s Disease Mild Cognitive Impairment (PD-MCI) [@litvan2012, @bezdicek2017].
+Cognitive performance was evaluated at both Level I and Level II according to the standard MDS battery for Parkinson's Disease Mild Cognitive Impairment (PD-MCI) [@litvan2012, @bezdicek2017].
 Cognitive performance at Level I was assessed by the Mini-Mental State Examination (MMSE) [@stepankova2014; @folstein1975] and the Montreal Cognitive Assessment (MoCA) [@kopecek2016; @nasreddine2005].
-The comprehensive neuropsychological assessment at Level II evaluated five cognitive domains through specific tests: attention and working memory assessed by Trail Making Test Part A (TMT-A) [@bezdicek2012; @reitan2004], and WAIS Digit Span Backward (WAIS DSB) [@wechsler1997], executive function by Categorical Verbal Fluency (CF) [@nikolai2015], and subtest from the Prague Stroop Test – Colors (PST-C) [@bezdicek2021], language by the WAIS Similarities subtest [@wechsler1997], and the Boston Naming Test (BNT-60) [@zemanova2016; @kaplan1983], memory by the Rey Auditory Verbal Learning Test (RAVLT) [@frydrychova2018; @bezdicek2013; @rey1964] delayed recall, and the Brief Visual Memory Test–Revised (BVMTR) [@havlik2020; @benedict1997] delayed recall, or WAIS Family Pictures subtest [@wechsler1997] delayed recall, visuospatial function assessed by the Judgment of Line Orientation Test (JoLO) [@benton1983], and Clock Drawing Test (CLOX) [@royall1998].
+The comprehensive neuropsychological assessment at Level II evaluated five cognitive domains through specific test: attention and working memory assessed by Trail Making Test Part A (TMT-A) [@bezdicek2012; @reitan2004], and WAIS Digit Span Backward (WAIS DSB) [@wechsler1997], executive function by Categorical Verbal Fluency (CF) [@nikolai2015], and subtest from the Prague Stroop Test – Colors (PST-C) [@bezdicek2021], language by the WAIS Similarities subtest [@wechsler1997], and the Boston Naming Test (BNT-60) [@zemanova2016; @kaplan1983], memory by the Rey Auditory Verbal Learning Test (RAVLT) [@frydrychova2018; @bezdicek2013; @rey1964] delayed recall, and the Brief Visual Memory Test–Revised (BVMTR) [@havlik2020; @benedict1997] delayed recall, or WAIS Family Pictures subtest [@wechsler1997] delayed recall, visuospatial function assessed by the Judgment of Line Orientation Test (JoLO) [@benton1983], and Clock Drawing Test (CLOX) [@royall1998].
 
@@
 The Beck Depression Inventory-II (BDI-II) [@ciharova2020; @beck1996] and State-Trait Anxiety Inventory (STAI) [@spielberg1983; @mullner1980] were used to assess neuropsychiatric status.
-**BDI-II and STAI are a part of the standard methods in our pre-DBS protocol, having well-established psychometric properties in PD populations [@leentjens2000; @dissanayaka2010].
-Other psychiatric scales data, such as the Neuropsychiatric Inventory, were not available for our retrospective dataset.
-However, all patients underwent a comprehensive evaluation by a trained neuropsychiatrist, excluding the presence of major depression, apathy, psychosis, and visual hallucinations, ensuring that confounding neuropsychiatric conditions were not present in our sample.**
+Psychotic symptoms were assessed through structured psychiatric interviews conducted by a trained psychiatrist.
 
@@
 The third approach applied the Czech version of the shortened Montreal Cognitive Assessment (sMoCA) [@bezdicek2020], a time-efficient modification designed to measure global cognitive performance using a reduced testing protocol that omits items providing redundant information.
-**The sMoCA has been validated in Czech PD cohorts [@bezdicek2020] and shown to be sensitive to cognitive deficits while lowering patient burden [@roalf2017a].
-Inclusion of the sMoCA in our study was motivated by its clinical utility in pre-surgical settings, where time restrictions and patients’ fatigue often limit the feasibility of longer assessments.
-Moreover, its inclusion allowed us to compare diagnostic performance across different screening lengths.**
-
 Lastly, the fourth approach followed the Level II protocol for diagnosis of PDD and Mild Cognitive Impairment in PD (PD-MCI) [@dubois2007; @litvan2012].
@@
 
-critnum <- sapply(c("mmse", "moca", "smoca", "lvlII"), function(i) {
+critnum <- sapply(
+  c("mmse", "moca", "smoca", "lvlII"), function(i) {
     subset(crit, group == i) |>
       nrow()
-})
+  }
+)
 
@@
 
-Following the framework proposed @lundberg2021, in this study we explicitly connect our research objectives and their corresponding theoretical (i.e., targets of inference) and empirical (i.e., data-driven) estimands to statistical estimates.
+Following the framework proposed by @lundberg2021, in this study we explicitly connect our research objectives and their corresponding theoretical (i.e., targets of inference) and empirical (i.e., data-driven) estimands to statistical estimates.
 The theoretical estimand refers to a unit-specific quantity defined over a target population and represents the ideal quantity that would address the research question under optimal conditions, such as access to complete population data or perfect experimental control.
 In contrast, the empirical estimand corresponds to the quantity that is actually computable using the available dataset, given real world constraints.
-The full description of the study’s estimands and their relation to our research objectives is presented in the Appendix (see @tbl-estimands).
+The full description of the study's estimands and their relation to our research objectives is presented in the Appendix (see @tbl-estimands).
 
 To address study objectives, we started by repeatedly assigning each patient the diagnosis of probable PDD based on each PDD algorithm listed in @tbl-crits (see also @tbl-algos) resulting in a `r sum(data$incl)` (patients) $\times$ `r sum(critnum)` (algorithms) matrix where each cell indicates whether a patient (row) meets criteria for probable PDD according to an algorithm (column).
-PDD rate estimates were computed as $\frac{N_{PDD}}{N_{total}}$ separately for each algorithm.
+PDD rate estimates were computed as $\frac{N_{PDD}}{N_{total}}$ separately for each algortihm.
 The predictive value of age and sex was then evaluated by fitting a set of logistic regressions, one for each algorithm for probable PDD, whereby the probable PDD was predicted by age, sex and their interaction.
@@
 All analyses were conducted within the R (version `r with(version, paste(major, minor, sep="."))`) software environment for statistical computing [@rsoft].
-The software code supporting this article is available at <https://github.com/josefmana/DemCr1t.git>.[^1]
-
-[^1]: **JM: Do not forget to make it public before submitted!**
+The software code supporting this article is available at <https://github.com/josefmana/DemCr1t.git>.
 
@@
 rates <- tar_read(rate_summaries, store = upstore)
-algos <- tar_read(algorithms, store = upstore)
 pred <- tar_read(demographic_predictors, store = upstore)
 conc <- tar_read(concordance_statistics, store = upstore)
-kappa <- tar_read(kappa_summmaries, store = upstore)
 perc <- rates$table$perc
-
-# Function for writing M (SD):
-msd_txt <- function(x, txt = "", tab = desc$table) {
-  paste0(
-    subset(tab, variable == x)[ , c("M", "SD")] |> paste(collapse = " (SD = "),
-    ")",
-    txt
-  )
-}
 
@@
 sex <- subset(desc$table, variable == "Sex")$N
-age <- msd_txt("Age", " years of age")
-edu <- msd_txt("Education", " years of education")
-dur <- msd_txt("PD duration", " years of disease duration")
-updrs_off <- msd_txt("UPDRS III off state", " Unified Parkinson Disease Rating Scale (UPDRS), part III in medication OFF state")
-updrs_on <- msd_txt("UPDRS III on state", " Unified Parkinson Disease Rating Scale (UPDRS), part III in medication ON state")
-bdi <- msd_txt("BDI")
-staix1 <- msd_txt("STAI X1")
-staix2 <- msd_txt("STAI X2")
+age <- paste0(subset(desc$table, variable == "Age")[ , c("M", "SD")] |> paste(collapse = " (SD = "),") years of age")
+edu <- paste0(subset(desc$table, variable == "Education")[ , c("M", "SD")] |> paste(collapse = " (SD = ")," ) years of education")
+dur <- paste0(subset(desc$table, variable == "PD duration")[ , c("M", "SD")] |> paste(collapse = " (SD = "),") years of disease duration")
+updrs_off <- paste0(subset(desc$table, variable == "UPDRS III off state")[ , c("M", "SD")] |> paste(collapse = " (SD = "),") Unified Parkinson Disease Rating Scale (UPDRS), part III in medication OFF state")
+updrs_on <- paste0(subset(desc$table, variable == "UPDRS III on state")[ , c("M", "SD")] |> paste(collapse = " (SD = "),") UPDRS III in medication ON state")
 
 # In-text rate estimates
-rates_overall <- sapply(c("M", "SD", "Md", "minmax"), \(f) do_summary(perc, 2, f))
-rates_FAQ9 <- sapply(c("M", "SD", "Md", "minmax"), \(f) do_summary(perc[grepl("FAQ 9", rates$table$IADL)], 2, f))
-rates_FAQtot <- sapply(c("M", "SD", "Md", "minmax"), \(f) do_summary(perc[rates$table$IADL == "FAQ > 7"], 2, f))
+rates_overall <- sapply(c("M", "SD", "Md", "minmax"), function(f) do_summary(perc, 2, f))
+rates_FAQ9 <- sapply(c("M", "SD", "Md", "minmax"), function(f) do_summary(perc[grepl("FAQ 9", rates$table$IADL)], 2, f))
+rates_FAQtot <- sapply(c("M", "SD", "Md", "minmax"), function(f) do_summary(perc[rates$table$IADL == "FAQ > 7"], 2, f))
 
 # Minimum p-value of logistic regressions:
-min_p <- subset(pred$values, quantity == "p-value")$estimate |>
+min_p <-
+  subset(pred$values, quantity == "p-value")$estimate |>
   min() |>
   do_summary(3, "p")
+
+# List of algorithms:
+algos <- list(
+  faq_tot = subset(rates$table, IADL == "FAQ > 7")$type,
+  faq_9 = subset(rates$table, IADL == "FAQ 9 > 1")$type
+)
+
+continue <- !(any(algos$faq_tot %in% algos$faq_9) || any(algos$faq_9 %in% algos$faq_tot))
+stopifnot("Some algorithm(s) use both IADL operationalisations!" = continue)
 ```
@@
 The sample included `r sex` men, with an average of `r age`, `r edu`, `r dur`, `r updrs_off` and `r updrs_on`.
-**Descriptive statistics for neuropsychiatric symptoms indicated mild depressive and anxiety symptoms in our cohort, with the average BDI-II of `r bdi`, average STAI X1 of `r staix1`, and average STAI X2 of `r staix2`.
-However, according to the psychiatric assessment, none of the patients with probable PDD was suffering major depressive disorder, delirium or other neuropsychiatric abnormalities that would disqualify the diagnosis.** Cognitive characteristics of the sample are summarized in @tbl-desc.
+Cognitive characteristics of the sample are summarized in @tbl-desc.
 
@@
 
+```{r}
+#| label: "kappas"
+
+# Prepare combinations within IADL operationalisation:
+kappas <- lapply(
+  set_names(names(algos)), function(o) {
+    combn(algos[[o]], 2) |>
+      t() |>
+      as_tibble() |>
+      `colnames<-`(c("predictor", "reference")) |>
+      left_join(conc$table, by = c("predictor", "reference")) |>
+      select(Kappa_raw) |>
+      pull()
+  }
+)
+
+# Add cross-operationalisation cases
+kappas$cross <- subset(conc$table, predictor %in% algos$faq_9 & reference %in% algos$faq_tot)$Kappa_raw
+
+# Extract Kappa summaries:
+kappa_sums <-
+  sapply(
+    names(kappas),
+    function(i)
+      paste0(do_summary(kappas[[i]], 2, "M"),", SD = ", do_summary(kappas[[i]], 2, "SD"))
+  )
+
+# Add Kappa summaries of algos with same everything but different IADL:
+kappa_iadl <-
+  data.frame(
+    predictor = c("Lvl.II (1)", "MMSE (1)", "MMSE (3)", paste0("MoCA (",seq(1,59,2),")"), "sMoCA (1)"),
+    reference = c("Lvl.II (2)", "MMSE (2)", "MMSE (4)", paste0("MoCA (",seq(2,60,2),")"), "sMoCA (2)")
+  ) |>
+    left_join(conc$table) |>
+    select(Kappa_raw) |>
+    pull()
+
+kappa_iadl_sums <- paste0(do_summary(x = kappa_iadl, 2, "M"),", SD = ", do_summary(x = kappa_iadl, 2, "SD"))
+```
+
 Results of the analyses of prediction Accuracy, Cohen's $\kappa$, Sensitivity and Specificity are presented in @fig-acc, @fig-kappa, @fig-sens and @fig-spec respectively.
-**Numerical results are available at ...** [^2].
+Numerical results are available at <https://github.com/josefmana/DemCr1t.git>.
 Generally, algorithms that employed the same operationalization of IADL deficit showed substantial pairwise concordance, however, algorithms that operationalized IADL deficit differently did not.
-Whereas among algorithms with identical IADL deficit operationalization, the agreement judged by Cohen's $\kappa$ was moderately high (operationalization by FAQ total score: $\kappa$ = `r kappa$sums["faq_tot"]`; operationalization by FAQ item 9: $\kappa$ = `r kappa$sums["faq_9"]`), among algorithms that differ in IADL deficit operationalization but are otherwise identical it was low: $\kappa$ = `r kappa$sums["iadl"]`.
-
-[^2]: **JM: Should provide an excel sheet or html table the readers could go through by themselves.**
+Whereas among algorithms with identical IADL deficit operationalization, the agreement judged by Cohen's $\kappa$ was moderately high (operationalization by FAQ total score: $\kappa$ = `r kappa_sums["faq_tot"]`; operationalization by FAQ item 9: $\kappa$ = `r kappa_sums["faq_9"]`), among algorithms that differ in IADL deficit operationalization but are otherwise identical it was low: $\kappa$ = `r kappa_iadl_sums`.
 
@@
 All five Level I algorithms that approximated the Level II classification most accurately were MoCA-based and defined Executive Function deficit by Clock drawing rather than Lexical fluency test.
-On the other hand, two out of the five Level I algorithms with the lowest accuracy were MMSE-based, whereas the remaining three were MoCA-based and defined Executive Function deficit by Lexical fluency test.
+On the other hand, two out of the five Level I algorithms with the lowest accuracy were MMSE-based, whereas the remaining three were MoCA-based and defined Executive function deficit by Lexical fluency test.
 
@@
 
-Finally, if the predictors are sorted by their balanced accuracy (i.e., average of sensitivity and specificity) instead of raw accuracy, the results are similar with the exception that for prediction of Level II with total FAQ score algorithm for probable PDD, the highest balanced accuracy was achieved by the sMoCA algorithm with sensitivity `r smoca_stats["sens"]` and specificity `r smoca_stats["spec"]` (see @tbl-approx-bal).
+Finally, suppose the predictors are sorted by their balanced accuracy (i.e., average of sensitivity and specificity) instead of raw accuracy.
+In that case, the results are similar, with the exception that for the prediction of Level II with total FAQ score algorithm for probable PDD, the highest balanced accuracy was achieved by the sMoCA algorithm with sensitivity `r smoca_stats["sens"]` and specificity `r smoca_stats["spec"]` (see @tbl-approx-bal).
 
@@
 
-This study systematically investigated the application of multiple Level I diagnostic criteria for PDD.
+This study systematically investigated the application of multiple Level I  PDD diagnostic criteria.
 Our results show variability in PDD rate estimates, strongly influenced by the choice of cognitive screening instrument (MMSE, MoCA and sMoCA) and the operationalization of functional impairment.
@@
 This lack of age-dependency may, however, be also stemming from the relatively younger age of our cohort, because previous reports indicate that association between age and PDD is not linear but increases with age and may not reach substantial values before the age of 70 [@rana2012].
-**In our sample, the younger age and preserved cognitive level might limit age-related effects.
-This fact could bias algorithm results toward higher specificity but lower sensitivity [@larvin2023].
-Previous studies report that DBS candidates tend to be, on average, younger and cognitively less impaired than the general PD population [@bronstein2011], which could bias diagnostic algorithm performance toward overestimating specificity.
-Therefore, our findings may not be fully generalizable to older or more cognitively impaired people with PD.**
 
@@
 This demonstrates that even slight methodological differences can yield divergent diagnostic outcomes.
-Such findings are critical for clinicians relying on Level I criteria for eligibility decisions, as the choice of algorithm could lead to contradictory classifications of PDD.
+Such findings are critical for clinicians relying on Level I criteria for eligibility decisions, as the choice of algorithm could lead to contradictory classifications of PDD status.
 
@@
 
-Using Level II diagnosis as the reference, MoCA-based Level I algorithms, particularly those using Clock Drawing to assess executive function, demonstrated the highest predictive accuracy.
-This supports recent proposals to modernize PDD diagnostic frameworks [@kulisevsky2024], favouring MoCA-derived components and broader functional assessment tools.
+Using Level II diagnosis as the gold standard, MoCA-based Level I algorithms,
+particularly those using Clock Drawing to assess executive function, demonstrated the highest predictive accuracy.
+This supports recent proposals to modernise PDD diagnostic frameworks [@kulisevsky2024], favouring MoCA-derived components and more specific, PD-tailored functional assessment tools.
 In contrast, MMSE-based algorithms consistently underperformed, suggesting limited sensitivity in capturing cognitive deficits typical in PDD.
 
-The advantage of MoCA-based algorithm was evident in combination with using the FAQ total score, reaching accuracy rates of 94–99%.
-Furthermore, in the algorithm using sMoCA, the raw accuracy was moderate, however, the balanced accuracy (i.e.
+The diagnostic superiority of the MoCA-based algorithm was evident when using the FAQ total score (accuracy rates of 94–99%).
+Furthermore, in the algorithm using sMoCA, the raw accuracy was moderate; however, the balanced accuracy (i.e.
 combined sensitivity and specificity) was high.
-These findings imply that simplified tools such as sMoCA may be a promising tool for heightening efficiency of the assessment while maintaining diagnostic accuracy.
-**This underscores the clinical utility of sMoCA screening tool.
-Although raw accuracy was moderate, balanced accuracy results ensure avoidance of bias from class imbalance.
-From a practical point of view, it suggests that even brief screening tools can achieve reliable classification, which means an advantage for routine clinical settings and time-restricted assessments.**
+This suggests that simplified tools such as the sMoCA may be a promising tool for increasing the efficiency of the assessment while maintaining diagnostic accuracy.
 
@@
 Nevertheless, the analysis of the algorithm concordance might not be affected by the sample to the same degree allowing for a broader generalization.
-**However, the use of a DBS cohort offered certain methodological advantages.
-For example, all patients underwent standardized and comprehensive neuropsychological testing, which resulted in well-characterized and high-quality dataset, allowing us to systematically evaluate multiple diagnostic algorithms.** Future studies should aim to replicate our results on larger and different cohorts **with more heterogenous sample**.
+Future studies should aim to replicate our results on larger population-based PD cohorts.
 To make this process easier, the code used to generate our results is publicly available and easily applicable to similarly structured data.
-
-Furthermore, another limitation is the use of the FAQ questionnaire for IADL assessment.
-FAQ is subjective and informant-reliant, which may result in biased data.
-**At the same time, the questionnaire may vary across cultures, therefore, the results may not be transferable.
-For example, activities such as financial management or driving, may not be equally practiced across cultures or contexts, which might complicate interpretation of impairment levels.
-The FAQ scores may vary across cultures, and diagnostic thresholds may not be transferable across countries [@odonald2025].
-This may influence both the sensitivity and ecological validity of the functional criteria used.
-Another limitation is that the FAQ scores may correlate with neuropsychological results of certain domains, specifically executive functions or memory [@vanrentergem2020; @moheb2017].
-This overlap could artificially increase concordance between cognitive and functional criteria.
-For future research, we suggest using a PD-specific questionnaire, such as the Penn Parkinson's Daily Activities Questionnaire-15 [@brennan2016] or generally more objective methods, preferably performance-based ones.
+Furthermore, FAQ is subjective and informant-reliant, which may result in biased data.
+For future research, we suggest the use of PD-specific questionnaires, such as the Penn Parkinson&#39;s Daily Activities Questionnaire-15 (Brennan et al., 2016) or more objectiveperformance-based methods [@schmitter-edgecombe2020] or questionnaire adaptation including questions regarding gadgets use and digital literacy [@postema2024].
 Our results underscore the importance of IADL measurement for the PDD diagnosis.
-Therefore, we recommend using more reliable tools with high ecological validity.** For future research, we suggest the use of more objective, modernized tools for IADl assessment, such as performance-based methods [@schmitter-edgecombe2020] or questionnaire adaptation including questions regarding gadgets use and digital literacy [@postema2024].
-Our results emphasize the importance of IADL deficit for the PDD diagnosis, therefore, we suggest exploring more reliable tools with high ecological validity.
-
-**Additionally, the current study did not explore associations of PDD diagnosis with its neuropsychiatric correlates and biomarkers.
-We recommend examining models where psychiatric symptoms and biomarkers are treated as covariates to better capture their impact on PDD diagnosis.**
-
-??Finally, whereas our study systematically investigated how varying definition of global deficit, impaired cognition and IADL deficit affect probable PDD classification, it did not explore associations of PDD diagnosis with its neuropsychiatric (e.g.
+Therefore, we suggest exploring more, reliable tools with high ecological validity.
+Also, the current study did not explore associations of PDD diagnosis with its neuropsychiatric (e.g.
 anxiety profile) and biomarker correlates.
-Future studies would benefit from focusing on these aspects of PDD.??^[**JM: Why was the previous paragraph added instead of this one?
-This looks perfectly functional.
-The same goes largely for the IADL discussion, the original was better (in terms of resources and logical coherence) at most places.**]
+Future studies would benefit from focusing on these aspects of PDD.
 
@@
 
-Our study highlights the variability in PDD classification across Level I diagnostic algorithms, heavily influenced by IADL operationalization and the choice of cognitive screening tools.
-The findings support the call for a change of the current diagnostic criteria [@kulisevsky2024], favouring the use of MoCA-based components and comprehensive IADL assessments.
+In sum,our study systematically investigated how varying definitions of impaired cognition and IADL deficit affect PDD diagnostic accuracy.
+Our study highlights the variability in PDD classification across Level I diagnostic algorithms, significantly influenced by IADL operationalisation and the choice of cognitive screening tools.
+The findings strongly support the call for a change of the current diagnostic criteria [@kulisevsky2024], favouring the use of MoCA-based components and comprehensive IADL assessments.
 
@@
 Specifically, as the rate goes from extremes to 0.5, the variance increases.
-Consequently, if our estimate of the rate was lower than the true population rate, e.g., because our sample includes younger patients compared to the general PD population, our estimate of variance would also be lower than the true variance of PDD rate in the general PD population.
-Nonetheless, the between-algorithm variability may not be affected by this phenomenon as unlike variability of PDD rate, we do not have reason to assume it comes about by summing independent binomial events.
+Consequently, if our estimate of the rate was lower than the true population rate, e.g., because our sample includes younger patients compared to the general PD population, our estimate of variance would also lower than the true variance of PDD rate in the general PD population.
+Nonetheless, the between-algorithm variability may not be affected by this phenomenon as unlike variability of PDD rate, we do not have reason to assume it comes about by summing independent binomial events
 

### 2025-10-12 _manuscript/manuscript.qmd
Incorporatinf further changes,

--- 2025-10-11 _manuscript/manuscript.qmd
@@
 The Beck Depression Inventory-II (BDI-II) [@ciharova2020; @beck1996] and State-Trait Anxiety Inventory (STAI) [@spielberg1983; @mullner1980] were used to assess neuropsychiatric status.
-**BDI-II and STAI are a part of the standard methods in our pre-DBS protocol, having well-established psychometric properties in PD populations [@leentjens2000; @dissanayaka2010].
-Other psychiatric scales data, such as the Neuropsychiatric Inventory, were not available for our retrospective dataset.
-However, all patients underwent a comprehensive evaluation by a trained neuropsychiatrist, excluding the presence of major depression, apathy, psychosis, and visual hallucinations, ensuring that confounding neuropsychiatric conditions were not present in our sample.**
 
@@
 The third approach applied the Czech version of the shortened Montreal Cognitive Assessment (sMoCA) [@bezdicek2020], a time-efficient modification designed to measure global cognitive performance using a reduced testing protocol that omits items providing redundant information.
-**The sMoCA has been validated in Czech PD cohorts [@bezdicek2020] and shown to be sensitive to cognitive deficits while lowering patient burden [@roalf2017a].
-Inclusion of the sMoCA in our study was motivated by its clinical utility in pre-surgical settings, where time restrictions and patients’ fatigue often limit the feasibility of longer assessments.
-Moreover, its inclusion allowed us to compare diagnostic performance across different screening lengths.**
+**The sMoCA has been validated in Czech PD cohort [@bezdicek2020] and shown to be sensitive to cognitive deficits while lowering patient burden [@roalf2017a].
+We included the sMoCA in our study for its clinical utility in pre-surgical settings, where time restrictions and patients’ fatigue often limit the feasibility of longer assessments.
+Moreover, the Czech validation study reported comparable diagnostic accuracy between MoCA (AUC = 0.815) and sMoCA (AUC = 0.796) for distinguishing PD-MCI from PD-NC, supporting the sMoCA as a suitable and efficient alternative.**
 
@@
 These methodologies resulted in a total of `r sum(algonum)` algorithms, which were distributed across different diagnostic criteria: `r algonum["mmse"]` MMSE-based, `r algonum["moca"]` MoCA-based, `r algonum["smoca"]` sMoCA-based, and `r algonum["lvlII"]` based on the Level II battery (see @fig-tree, @tbl-crits and Appendix @tbl-algos for the exact specification of each algorithm).
+
+**Finally, patients were evaluated for psychiatric symptoms, including depression, delirium, apathy, anxiety and psychosis, by a neuropsychiatrist trained in evaluation of movement disorders patients (TU).
+Since severe psychiatric symptoms are exclusion criteria for the diagnosis of probable PDD (conditions 6-8), all patients diagnosed with PDD were double-checked for the presence of psychiatric symptoms in the hospital archives.**
 
@@
 The sample included `r sex` men, with an average of `r age`, `r edu`, `r dur`, `r updrs_off` and `r updrs_on`.
-**Descriptive statistics for neuropsychiatric symptoms indicated mild depressive and anxiety symptoms in our cohort, with the average BDI-II of `r bdi`, average STAI X1 of `r staix1`, and average STAI X2 of `r staix2`.
-However, according to the psychiatric assessment, none of the patients with probable PDD was suffering from the major depressive disorder, delirium or other neuropsychiatric abnormalities that would disqualify the diagnosis.** Cognitive characteristics of the sample are summarized in @tbl-desc.
+**Descriptive statistics for neuropsychiatric symptoms indicated within average level of depressive and anxiety symptoms in our cohort, with the average BDI-II of `r bdi`, average STAI X1 of `r staix1`, and average STAI X2 of `r staix2`.
+However, according to the psychiatric assessment, none of the patients with probable PDD was suffering from the major depressive disorder, delirium or other neuropsychiatric abnormalities that would excluded the diagnosis.** Cognitive characteristics of the sample are summarized in @tbl-desc.
 
@@
 
+# Dimensions of the results table:
 dims <- glue::glue("{dim(conc$table)[1]} rows x {dim(conc$table)[2]} columns")
 
+# Number of observations per algorithm pair:
 obs_pairs <- conc$table |>
@@
 
+# The most missing values present:
 min_miss <- glue::glue("{round(100 * (n_incl - obs_pairs[1, 'N']) / n_incl, 2)}%")
 
+# Percentage of algorithm pairs with N > K-1
 K <- 200
@@
 
+```{r}
+#| label: "age-over-70"
+
+over70 <- data |>
+  mutate(older = 100 * (age >= 70)) |>
+  pull(older) |>
+  mean(na.rm = TRUE) |>
+  round(1) |>
+  paste0("%")
+```
+
 Interestingly, we did not observe any reliable age-related differences in PDD rate within our cohort.
-This lack of age-dependency may, however, be also stemming from the relatively younger age of our cohort, because previous reports indicate that association between age and PDD is not linear but increases with age and may not reach substantial values before the age of 70 [@rana2012].
-**In our sample, the younger age and preserved cognitive level might limit age-related effects.
-This fact could bias algorithm results toward higher specificity but lower sensitivity [@larvin2023].
-Previous studies report that DBS candidates tend to be, on average, younger and cognitively less impaired than the general PD population [@bronstein2011], which could bias diagnostic algorithm performance toward overestimating specificity.
-Therefore, our findings may not be fully generalizable to older or more cognitively impaired people with PD.**
+This lack of age-dependency may, however, be also stemming from the relatively younger age of our cohort, because previous reports indicate that association between age and PDD is not linear but increases with age and may not reach substantial values before **older age.
+In both @rana2012 and @oh2016, nine out of ten patients with dementia were 70 years of age or older.
+In our sample, only `r over70` participants were in this age range.
+Consequently, studies with older cohorts are probably necessary to detect a robust association between age and the risk of probable PDD.**
 
@@
 
+## Constraints on Generality
+
+**Given the single-center nature of our study and the specific cohort, several not yet discussed constraints on generality apply to our findings.
+These can be broadly separated into three topics; age distribution, cognitive phenotype and cultural background.**
+
+**In our sample, the younger age and preserved cognitive level might limit age-related effects.
+This fact could bias algorithm results toward higher specificity but lower sensitivity [@larvin2023].
+Previous studies report that DBS candidates tend to be, on average, younger and cognitively less impaired than the general PD population [@bronstein2011], which could bias diagnostic algorithm performance toward overestimating specificity.
+Therefore, our findings may not be fully generalizable to older or more cognitively impaired people with PD.**
+
+**Previous longitudinal studies of DBS**
+
+**FAQ culturally different.**
+
+**Finally, our results may generalise well to other pre-DBS cohorts.**
+
 ## Limitations and Future Directions
@@
 
-**Another possible avenue for analysing the present data would be to exploit their hierarchical structure by fitting a multilevel model with algorithms nested within patients (see @fig-tree).
-Such an approach could leverage partial pooling [@gelman2012] and might approximate modern item response theory frameworks [@burkner2021], where PDD is conceptualised as a latent trait and the individual diagnostic algorithms act as items measuring it ().
-If feasible, this psychometric perspective may represent an exciting new direction for conceptualising complex clinical phenomena such as PDD [cf.
-@miller2012, @kiselica2021].**
+**Another possible avenue for analysing the present data would be to exploit their hierarchical structure by fitting a multilevel model with algorithms nested within cognitive tests (see @fig-tree).
+Such an approach could leverage partial pooling [@gelman2012] and might approximate modern item response theory frameworks [@burkner2021], where PDD is conceptualised as a latent trait and the individual diagnostic algorithms act as items measuring it.
+If feasible, this psychometric perspective may represent an exciting new direction for conceptualising complex clinical phenomena such as PDD (cf.
+@miller2012, @kiselica2021).**
 

### 2025-10-12 _targets.R

@@ -46,8 +46,12 @@ list(
     command = regress_pdd_on_demographics(raw_data, pdd_data$PDD)
   ),
   tar_target(
-    name = demographic_predictors_adjusted, # Regress probable PDD on demographics adjusting for neuropsychiatry
+    name = demographic_predictors_neuropsychiatry_adjusted, # Regress probable PDD on demographics adjusting for neuropsychiatry per Reviewer 2's demand
     command = regress_pdd_on_demographics(raw_data, pdd_data$PDD, covs = c("bdi", "stai_1"), inter = FALSE)
+  ),
+  tar_target(
+    name = demographic_predictors_cereda_adjusted, # Regress probable PDD on demographics adjusting as per Cereda et al. (2016)
+    command = regress_pdd_on_demographics(raw_data, pdd_data$PDD, covs = c("pd_dur", "edu_years"), inter = FALSE)
   ),
   tar_target(
     name = concordance_statistics, # Describe concordance between different PDD algorithms

### 2025-10-12 _manuscript/manuscript.qmd

@@
 
-Using Level II diagnosis as the reference, MoCA-based Level I algorithms, particularly those using Clock Drawing to assess executive function, demonstrated the highest predictive accuracy.
-This supports recent proposals to modernize PDD diagnostic frameworks [@kulisevsky2024], favouring MoCA-derived components and broader functional assessment tools.
+Using Level II diagnosis as the gold standard, MoCA-based Level I algorithms,
+particularly those using Clock Drawing to assess executive function, demonstrated the highest predictive accuracy.
+This supports recent proposals to modernise PDD diagnostic frameworks [@kulisevsky2024], favouring MoCA-derived components and more specific, PD-tailored functional assessment tools.
 In contrast, MMSE-based algorithms consistently underperformed, suggesting limited sensitivity in capturing cognitive deficits typical in PDD.

### 2025-10-12 _manuscript/manuscript.qmd
JM's changes in response to most reviewers' demands.

@@
 format:
- apaquarto-pdf:
-    documentmode: man
-    keep-tex: true
+ #apaquarto-pdf:
+  #  documentmode: man
+  #  keep-tex: true
  apaquarto-docx: 
@@
 
-**Given the single-center nature of our study and the specific cohort, several not yet discussed constraints on generality apply to our findings.
-These can be broadly separated into three topics; age distribution, cognitive phenotype and cultural background.**
+**This study’s generalizability is limited by the homogeneity of the patient cohort, which does not reflect the diversity of cognitive profiles seen in broader PD populations.
+Specifically, the younger age of the sample, and a possible underrepresentation of high-risk phenotypes for PDD constraints the generality of the presented findings.**
 
-**In our sample, the younger age and preserved cognitive level might limit age-related effects.
-This fact could bias algorithm results toward higher specificity but lower sensitivity [@larvin2023].
-Previous studies report that DBS candidates tend to be, on average, younger and cognitively less impaired than the general PD population [@bronstein2011], which could bias diagnostic algorithm performance toward overestimating specificity.
-Therefore, our findings may not be fully generalizable to older or more cognitively impaired people with PD.**
+**As noted above, the younger age of our sample may partly explain the lower rate of PDD observed compared to previous studies.
+As discussed in the *Theoretical and Empirical Estimands* section of the Appendix, neither estimates of PDD rate nor predictive performance of demographic variables should therefore not be generalized beyond PD patients that are candidates for DBS.
+The extent to which our findings on the concordance between diagnostic algorithms apply to the broader PD population remains to be determined in future studies using different types of cohorts, such as de novo patients or community-based samples.**
 
-**Previous longitudinal studies of DBS**
+**Longitudinal cohort studies suggest that patients considered for DBS represent a subset of the PD population with a distinct cognitive phenotype [@bove2020; @mana2024].
+Specifically, findings of gradual post-surgical cognitive decline predicted by pre-surgical executive deficits indicate that DBS candidates may be preferentially drawn from a fronto-striatal phenotype characterized by slowly progressing executive dysfunction, rather than from a posterior phenotype marked by visuospatial impairment [@kehagia2012].
+Importantly, patients with the posterior phenotype may be at greater risk of developing PDD within as little as five years after disease onset [@williams-gray2009; @summers2024].**
 
-**FAQ culturally different.**
+**In our study, visuospatial function was assessed uniformly across all algorithms within a given screening measure (MoCA cube or MMSE pentagons).
+By contrast, we compared two operationalizations of executive dysfunction, the clock drawing test and lexical fluency.
+The clock drawing test showed stronger predictive value for level II diagnosis than lexical fluency, suggesting that even within our cohort, patients who developed PDD may exhibit features of the posterior phenotype.**
 
-**Finally, our results may generalise well to other pre-DBS cohorts.**
+**However, the use of a DBS cohort also offered several methodological advantages.
+All patients underwent standardized and comprehensive neuropsychological testing, resulting in a well-characterized dataset that enabled a systematic evaluation of multiple diagnostic algorithms.
+Moreover, because dementia is a common exclusion criterion for DBS treatment [@bronstein2011], examining the diagnostic accuracy of algorithms for PDD in a pre-DBS cohort is informative in its own right.**
 
 ## Limitations and Future Directions
-
-This study’s generalizability is limited by its retrospective design and the homogeneity of the patient cohort, which does not reflect broader PD populations with varying cognitive profiles.
-Specifically, the younger age of the sample and possible lack of high-risk phenotypes for PDD compared to the general population of people with PD might have been responsible for comparatively lower PDD rate estimates in our study.
-
-Nevertheless, the analysis of the algorithm concordance might not be affected by the sample to the same degree allowing for a broader generalization.
-**However, the use of a DBS cohort offered certain methodological advantages.
-For example, all patients underwent standardized and comprehensive neuropsychological testing, which resulted in well-characterized and high-quality dataset, allowing us to systematically evaluate multiple diagnostic algorithms.** Future studies should aim to replicate our results on larger and different cohorts **with more heterogenous sample**.
-To make this process easier, the code used to generate our results is publicly available and easily applicable to similarly structured data.
 
@@
 The FAQ scores may vary across cultures, and diagnostic thresholds may not be transferable across countries [@odonald2025].
-This may influence both the sensitivity and ecological validity of the functional criteria used.
+This may influence both the sensitivity and ecological validity of the functional criteria used [@bezdicek2011; @bezdicek2016b].
 Another limitation is that the FAQ scores may correlate with neuropsychological results of certain domains, specifically executive functions or memory [@vanrentergem2020; @moheb2017].
@@
 
-In sum,our study systematically investigated how varying definitions of impaired cognition and IADL deficit affect PDD diagnostic accuracy.
+In sum, our study systematically investigated how varying definitions of impaired cognition and IADL deficit affect PDD diagnostic accuracy.
 Our study highlights the variability in PDD classification across Level I diagnostic algorithms, significantly influenced by IADL operationalisation and the choice of cognitive screening tools.
@@
 Moreover, when using MoCA-based algorithms instead of MMSE-based ones, we can observe better approximations to the Level II battery.
+
+Future studies should aim to replicate our results on larger and different cohorts **with more heterogenous sample**.
+To make this process easier, the code used to generate our results is publicly available and easily applicable to similarly structured data.
 

### 2025-10-12 _manuscript/manuscript.qmd
changes to the manuscript

@@
 
-**Finally, patients were evaluated for psychiatric symptoms, including depression, delirium, apathy, anxiety and psychosis, by a neuropsychiatrist trained in evaluation of movement disorders patients (TU).
-Since severe psychiatric symptoms are exclusion criteria for the diagnosis of probable PDD (conditions 6-8), all patients diagnosed with PDD were double-checked for the presence of psychiatric symptoms in the hospital archives.**
+**Finally, all patients were systematically evaluated for the presence of neuropsychiatric symptoms, including depression, apathy, anxiety, psychosis, and delirium, by a trained neuropsychiatrist (TU) experienced in the assessment of patients with movement disorders.
+Because severe psychiatric symptoms form exclusion criteria for the diagnosis of probable PDD (conditions 6–8; p.
+1), all patients classified as PDD were double-checked in hospital records to confirm the absence of such confounding symptoms.**
 
@@
 
-Using Level II diagnosis as the gold standard, MoCA-based Level I algorithms,
-particularly those using Clock Drawing to assess executive function, demonstrated the highest predictive accuracy.
+Using Level II diagnosis as the gold standard, MoCA-based Level I algorithms, particularly those using Clock Drawing to assess executive function, demonstrated the highest predictive accuracy.
 This supports recent proposals to modernise PDD diagnostic frameworks [@kulisevsky2024], favouring MoCA-derived components and more specific, PD-tailored functional assessment tools.
@@
 
-**Modern alternatives, such as multiple imputation techniques, have been shown to produce less biased estimates than case deletion strategies in analyses based on confusion matrices [@karakaya2014], whereas the evidence supporting their superiority in the estimation of Cohen’s $\kappa$ remains limited [@deraadt2019].
+**Modern alternatives, such as multiple imputation techniques, have been shown to produce less biased estimates than case deletion strategies in analyses based on confusion matrices [@karakaya2014], whereas the evidence supporting their superiority in the estimation of Cohen’s** $\kappa$ remains limited [@deraadt2019].
 These modern approaches also require careful specification of the imputation model and the underlying causal mechanism of missingness to ensure appropriate covariate selection [@long2011; @bianco2023].
 For the sake of parsimony, we did not employ advanced missing-data techniques but instead explicitly described the observed missingness patterns.
-Consequently, our findings should be regarded as exploratory and primarily serve as a basis for hypothesis generation in future, more controlled studies.**
+Consequently, our findings should be regarded as exploratory and primarily serve as a basis for hypothesis generation in future, more controlled studies.
 
@@
 
-Furthermore, another limitation is the use of the FAQ questionnaire for IADL assessment.
-FAQ is subjective and informant-reliant, which may result in biased data.
-**At the same time, the questionnaire may vary across cultures, therefore, the results may not be transferable.
-For example, activities such as financial management or driving, may not be equally practiced across cultures or contexts, which might complicate interpretation of impairment levels.
-The FAQ scores may vary across cultures, and diagnostic thresholds may not be transferable across countries [@odonald2025].
-This may influence both the sensitivity and ecological validity of the functional criteria used [@bezdicek2011; @bezdicek2016b].
-Another limitation is that the FAQ scores may correlate with neuropsychological results of certain domains, specifically executive functions or memory [@vanrentergem2020; @moheb2017].
+**Another limitation concerns the use of the FAQ questionnaire for IADL assessment.
+The FAQ is a subjective and informant-reliant measure, which may contain biased data.
+Moreover, its content can vary across sociocultural contexts, limiting cross-cultural transferability.
+For example, activities such as financial management, cooking or driving, are not uniformly performed across societies.
+Consequently, both the FAQ scores and diagnostic thresholds used to indicate IADL impairment may not be directly transferable between cultural settings [@odonald2025].
+These factors may influence both the sensitivity and ecological validity of the functional criteria used [@bezdicek2011; @bezdicek2016b].**
+
+**Furthermore, the FAQ scores may correlate with neuropsychological results of specific cognitive domains, particularly executive functions or memory [@vanrentergem2020; @moheb2017].
 This overlap could artificially increase concordance between cognitive and functional criteria.
-For future research, we suggest using a PD-specific questionnaire, such as the Penn Parkinson's Daily Activities Questionnaire-15 [@brennan2016] or generally more objective methods, preferably performance-based ones.
+To address thes concerns in future research, we suggest using a PD-specific questionnaire, such as the Penn Parkinson's Daily Activities Questionnaire-15 [@brennan2016] or generally more objective methods, preferably performance-based ones.
 Our results underscore the importance of IADL measurement for the PDD diagnosis.
-Therefore, we recommend using more reliable tools with high ecological validity.** For future research, we suggest the use of more objective, modernized tools for IADl assessment, such as performance-based methods [@schmitter-edgecombe2020] or questionnaire adaptation including questions regarding gadgets use and digital literacy [@postema2024].
+Therefore, we recommend using more reliable tools with high ecological validity.** For future research, we suggest the use of more objective, modernized tools for IADl assessment, such as performance-based methods [@schmitter-edgecombe2020] or questionnaire adaptations including items regarding gadgets use and digital literacy [@postema2024].
 Our results emphasize the importance of IADL deficit for the PDD diagnosis, therefore, we suggest exploring more reliable tools with high ecological validity.
 
-**Additionally, the current study did not explore associations of PDD diagnosis with its neuropsychiatric correlates and biomarkers.
-We recommend examining models where psychiatric symptoms and biomarkers are treated as covariates to better capture their impact on PDD diagnosis.**
-
-??Finally, whereas our study systematically investigated how varying definition of global deficit, impaired cognition and IADL deficit affect probable PDD classification, it did not explore associations of PDD diagnosis with its neuropsychiatric (e.g.
+Finally, whereas our study systematically investigated how varying definition of global deficit, impaired cognition and IADL deficit affect probable PDD classification, it did not explore associations of PDD diagnosis with its neuropsychiatric (e.g.
 anxiety profile) and biomarker correlates.
-Future studies would benefit from focusing on these aspects of PDD.??[^2]
-
-[^2]: **JM: Why was the previous paragraph added instead of this one?
-This looks perfectly functional.
-The same goes largely for the IADL discussion, the original was better (in terms of resources and logical coherence) at most places.**
+Future studies would benefit from focusing on these aspects of PDD.??
 

### 2025-10-12 _manuscript/manuscript.qmd
Grammar check

@@
 This overlap could artificially increase concordance between cognitive and functional criteria.
-To address thes concerns in future research, we suggest using a PD-specific questionnaire, such as the Penn Parkinson's Daily Activities Questionnaire-15 [@brennan2016] or generally more objective methods, preferably performance-based ones.
+To address these concerns in future research, we suggest using a PD-specific questionnaire, such as the Penn Parkinson's Daily Activities Questionnaire-15 [@brennan2016] or generally more objective methods, preferably performance-based ones.
 Our results underscore the importance of IADL measurement for the PDD diagnosis.

### 2025-10-13 _manuscript/manuscript.qmd
NHBC

@@
 All patients underwent a comprehensive neuropsychological battery.
-A total of 68 diagnostic algorithms were operationalised based on the combinations of various cognitive assessments, including the Mini Mental State Exam (MMSE), the Montreal Cognitive Assessment (MoCA), and the shortened version of MoCA (sMoCA).
+A total of 68 diagnostic algorithms were operationalised based on the combinations of various cognitive assessments, including the Mini Mental State Examination (MMSE), the Montreal Cognitive Assessment (MoCA), and the shortened version of MoCA (sMoCA).
 Functional impairment was based on the Functional Assessment Questionnaire (FAQ), specifically FAQ item 9 or the FAQ total score.
@@
 A recent call for a change pinpoints limitations regarding the original criteria and suggest various updates to enhance their utility [@kulisevsky2024].
-Proposed suggestions include replacement of Mini Mental State Examination (MMSE) by Montreal Cognitive Assessment (MoCA), which is more sensitive to PD specific cognitive impairment; expansion of instrumental activities of daily living (IADL) evaluation; inclusion of language assessment; recognition of anxiety as one **of** the neuropsychiatric symptoms relevant in PDD; and intergration of biomarkers.
+Proposed suggestions include replacement of the Mini Mental State Examination (MMSE) by the Montreal Cognitive Assessment (MoCA), which is more sensitive to PD specific cognitive impairment; expansion of instrumental activities of daily living (IADL) evaluation; inclusion of language assessment; recognition of anxiety as one **of** the neuropsychiatric symptoms relevant in PDD; and integration of biomarkers.
 
-In the light of these proposals, the current study aims to evaluate the diagnostic concordance between the original MDS Level I PDD criteria [@emre2007; @dubois2007] and a modified framework based on the recent call for change [@kulisevsky2024].
+In light of these proposals, the current study aims to evaluate the diagnostic concordance between the original MDS Level I PDD criteria [@emre2007; @dubois2007] and a modified framework based on the recent call for change [@kulisevsky2024].
 Furthermore, both Level I diagnostic approaches are compared to PDD diagnosed on Level II.
@@
 Cognitive performance was evaluated at both Level I and Level II according to the standard MDS battery for Parkinson’s Disease Mild Cognitive Impairment (PD-MCI) [@litvan2012, @bezdicek2017].
-Cognitive performance at Level I was assessed by the Mini-Mental State Examination (MMSE) [@stepankova2014; @folstein1975] and the Montreal Cognitive Assessment (MoCA) [@kopecek2016; @nasreddine2005].
+Cognitive performance at Level I was assessed by **the MMSE [@stepankova2014; @folstein1975] and the MoCA** [@kopecek2016; @nasreddine2005].
 The comprehensive neuropsychological assessment at Level II evaluated five cognitive domains through specific tests: attention and working memory assessed by Trail Making Test Part A (TMT-A) [@bezdicek2012; @reitan2004], and WAIS Digit Span Backward (WAIS DSB) [@wechsler1997], executive function by Categorical Verbal Fluency (CF) [@nikolai2015], and subtest from the Prague Stroop Test – Colors (PST-C) [@bezdicek2021], language by the WAIS Similarities subtest [@wechsler1997], and the Boston Naming Test (BNT-60) [@zemanova2016; @kaplan1983], memory by the Rey Auditory Verbal Learning Test (RAVLT) [@frydrychova2018; @bezdicek2013; @rey1964] delayed recall, and the Brief Visual Memory Test–Revised (BVMTR) [@havlik2020; @benedict1997] delayed recall, or WAIS Family Pictures subtest [@wechsler1997] delayed recall, visuospatial function assessed by the Judgment of Line Orientation Test (JoLO) [@benton1983], and Clock Drawing Test (CLOX) [@royall1998].
@@
 The second set of algorithms was based on the recent call for change of dementia diagnostic guidelines [@kulisevsky2024], which advocates for more sensitive cognitive domain assessments in the context of PD.
-This updated approach incorporated specific items from the Montreal Cognitive Assessment (MoCA).
+This updated approach incorporated specific items from the MoCA.
 The third approach applied the Czech version of the shortened Montreal Cognitive Assessment (sMoCA) [@bezdicek2020], a time-efficient modification designed to measure global cognitive performance using a reduced testing protocol that omits items providing redundant information.
@@
 In contrast, the empirical estimand corresponds to the quantity that is actually computable using the available dataset, given real world constraints.
-The full description of the study’s estimands and their relation to our research objectives is presented in the Appendix (see @tbl-estimands).
+A full description of the study’s estimands and their relation to our research objectives is presented in the Appendix (see @tbl-estimands).
 
@@
 
-On average, estimated PDD rate was `r rates_overall["M"]`% (SD = `r rates_overall["SD"]`, Md = `r rates_overall["Md"]`, range `r rates_overall["minmax"]`).
-Notably, the estimates were substantially lower when FAQ item 9 was used as a criterion of IADL deficit (M = `r rates_FAQ9["M"]`% SD = `r rates_FAQ9["SD"]`, Md = `r rates_FAQ9["Md"]`, range `r rates_FAQ9["minmax"]`) compared to using total FAQ score criterion (M = `r rates_FAQtot["M"]`% SD = `r rates_FAQtot["SD"]`, Md = `r rates_FAQtot["Md"]`, range `r rates_FAQtot["minmax"]`) as demonstrated in @fig-tree **(see also @fig-rates for per-operationalisation distribution of PDD rate estimates)**.
-Neither age nor sex or their interaction (*p*s $\geq$ `r min_p`) reliably predicted probable PDD classification across algorithms (see @fig-data and @fig-pars).
+On average, the estimated PDD rate was `r rates_overall["M"]`% (SD = `r rates_overall["SD"]`, Md = `r rates_overall["Md"]`, range `r rates_overall["minmax"]`).
+Notably, the estimates were substantially lower when FAQ item 9 was used as a criterion of IADL deficit (M = `r rates_FAQ9["M"]`% SD = `r rates_FAQ9["SD"]`, Md = `r rates_FAQ9["Md"]`, range `r rates_FAQ9["minmax"]`) compared to using the total FAQ score criterion (M = `r rates_FAQtot["M"]`% SD = `r rates_FAQtot["SD"]`, Md = `r rates_FAQtot["Md"]`, range `r rates_FAQtot["minmax"]`) as demonstrated in @fig-tree **(see also @fig-rates for per-operationalisation distribution of PDD rate estimates)**.
+Neither age, sex nor their interaction (*p*s $\geq$ `r min_p`) reliably predicted probable PDD classification across algorithms (see @fig-data and @fig-pars).
 
@@
 
-For easier interpretability of our results, we next examine cases where Level II algorithms served as a reference and Level I algorithms as a predictor.
+For easier interpretability of our results, we next examined cases where Level II algorithms served as a reference and Level I algorithms as a predictor.
 @tbl-approx shows five Level I algorithms with the highest and five with the lowest accuracy in predicting Level II classification of probable PDD.
@@
 When IADL deficit was defined by total FAQ score, the Level II estimate of PDD rate was `r lvlII_rates["faq_tot"]`.
-All five Level I algorithms that approximated the Level II classification most accurately were MoCA-based and defined Executive Function deficit by Clock drawing rather than Lexical fluency test.
-On the other hand, two out of the five Level I algorithms with the lowest accuracy were MMSE-based, whereas the remaining three were MoCA-based and defined Executive function deficit by Lexical fluency test.
+All five Level I algorithms that approximated the Level II classification most accurately were MoCA-based and defined Executive Function deficit by Clock drawing rather than Verbal fluency test.
+On the other hand, two out of the five Level I algorithms with the lowest accuracy were MMSE-based, whereas the remaining three were MoCA-based and defined Executive function deficit by Verbal fluency test.
 
@@
 by defining executive deficit via clock drawing vs.
-lexical fluency).
+verbal fluency).
 
@@
 **In our study, visuospatial function was assessed uniformly across all algorithms within a given screening measure (MoCA cube or MMSE pentagons).
-By contrast, we compared two operationalizations of executive dysfunction, the clock drawing test and lexical fluency.
-The clock drawing test showed stronger predictive value for level II diagnosis than lexical fluency, suggesting that even within our cohort, patients who developed PDD may exhibit features of the posterior phenotype.**
+By contrast, we compared two operationalizations of executive dysfunction, the clock drawing test and verbal fluency.
+The clock drawing test showed stronger predictive value for level II diagnosis than verbal fluency, suggesting that even within our cohort, patients who developed PDD may exhibit features of the posterior phenotype.**
 
@@
 **Furthermore, the FAQ scores may correlate with neuropsychological results of specific cognitive domains, particularly executive functions or memory [@vanrentergem2020; @moheb2017].
-This overlap could artificially increase concordance between cognitive and functional criteria.
+This overlap could artificially increase concordance between cognitive and functional capacity criteria.
 To address these concerns in future research, we suggest using a PD-specific questionnaire, such as the Penn Parkinson's Daily Activities Questionnaire-15 [@brennan2016] or generally more objective methods, preferably performance-based ones.
@@
 anxiety profile) and biomarker correlates.
-Future studies would benefit from focusing on these aspects of PDD.??
+Future studies would benefit from focusing on these aspects of PDD.
 

### 2025-10-15 _manuscript/manuscript.qmd
Manuscript changes to incorporate all reviewers' comments.

@@
 
-According to a recent meta-analysis, approximately one-quarter of PD patients is likely to be diagnosed with PDD [@sousa2022].
+According to a recent meta-analysis, approximately one-quarter of PD patients are likely to be diagnosed with PDD [@sousa2022].
 However, reported PDD rate estimates vary widely, ranging from 14% up to 55%, depending on the methodological criteria employed [@sousa2022].
@@
 Currently, efforts are focused on refining this PDD diagnostic framework.
-A recent call for a change pinpoints limitations regarding the original criteria and suggest various updates to enhance their utility [@kulisevsky2024].
+A recent call for a change pinpoints limitations regarding the original criteria and suggests various updates to enhance their utility [@kulisevsky2024].
 Proposed suggestions include replacement of the Mini Mental State Examination (MMSE) by the Montreal Cognitive Assessment (MoCA), which is more sensitive to PD specific cognitive impairment; expansion of instrumental activities of daily living (IADL) evaluation; inclusion of language assessment; recognition of anxiety as one **of** the neuropsychiatric symptoms relevant in PDD; and integration of biomarkers.
@@
 Cognitive performance was evaluated at both Level I and Level II according to the standard MDS battery for Parkinson’s Disease Mild Cognitive Impairment (PD-MCI) [@litvan2012, @bezdicek2017].
-Cognitive performance at Level I was assessed by **the MMSE [@stepankova2014; @folstein1975] and the MoCA** [@kopecek2016; @nasreddine2005].
+Cognitive performance at Level I was assessed by the MMSE [@stepankova2014; @folstein1975] and the MoCA [@kopecek2016; @nasreddine2005].
 The comprehensive neuropsychological assessment at Level II evaluated five cognitive domains through specific tests: attention and working memory assessed by Trail Making Test Part A (TMT-A) [@bezdicek2012; @reitan2004], and WAIS Digit Span Backward (WAIS DSB) [@wechsler1997], executive function by Categorical Verbal Fluency (CF) [@nikolai2015], and subtest from the Prague Stroop Test – Colors (PST-C) [@bezdicek2021], language by the WAIS Similarities subtest [@wechsler1997], and the Boston Naming Test (BNT-60) [@zemanova2016; @kaplan1983], memory by the Rey Auditory Verbal Learning Test (RAVLT) [@frydrychova2018; @bezdicek2013; @rey1964] delayed recall, and the Brief Visual Memory Test–Revised (BVMTR) [@havlik2020; @benedict1997] delayed recall, or WAIS Family Pictures subtest [@wechsler1997] delayed recall, visuospatial function assessed by the Judgment of Line Orientation Test (JoLO) [@benton1983], and Clock Drawing Test (CLOX) [@royall1998].
@@
 The third approach applied the Czech version of the shortened Montreal Cognitive Assessment (sMoCA) [@bezdicek2020], a time-efficient modification designed to measure global cognitive performance using a reduced testing protocol that omits items providing redundant information.
-**The sMoCA has been validated in Czech PD cohort [@bezdicek2020] and shown to be sensitive to cognitive deficits while lowering patient burden [@roalf2017a].
+**The sMoCA has been validated in the Czech PD cohort [@bezdicek2020] and shown to be sensitive to cognitive deficits while lowering patient burden [@roalf2017a].
 We included the sMoCA in our study for its clinical utility in pre-surgical settings, where time restrictions and patients’ fatigue often limit the feasibility of longer assessments.
@@
 
-Following the framework proposed @lundberg2021, in this study we explicitly connect our research objectives and their corresponding theoretical (i.e., targets of inference) and empirical (i.e., data-driven) estimands to statistical estimates.
+Following the framework proposed by @lundberg2021, in this study we explicitly connect our research objectives and their corresponding theoretical (i.e., targets of inference) and empirical (i.e., data-driven) estimands to statistical estimates.
 The theoretical estimand refers to a unit-specific quantity defined over a target population and represents the ideal quantity that would address the research question under optimal conditions, such as access to complete population data or perfect experimental control.
@@
 perc  <- rates$table$perc
+
+# Tables of the 'best' and the 'worst' w.r.t. Level II:
+approx_acc <- table_levelII_approximations(conc$table, algos, "Accuracy_raw")
+approx_bal <- table_levelII_approximations(conc$table, algos, "Balanced Accuracy")
 
@@
 
-**Total of** `r n_total` patients **were considered for the study, out of which `r excl_txt`**.
+**A total of** `r n_total` patients **were considered for the study, out of which `r excl_txt`**.
 The sample included `r sex` men, with an average of `r age`, `r edu`, `r dur`, `r updrs_off` and `r updrs_on`.
 **Descriptive statistics for neuropsychiatric symptoms indicated within average level of depressive and anxiety symptoms in our cohort, with the average BDI-II of `r bdi`, average STAI X1 of `r staix1`, and average STAI X2 of `r staix2`.
-However, according to the psychiatric assessment, none of the patients with probable PDD was suffering from the major depressive disorder, delirium or other neuropsychiatric abnormalities that would excluded the diagnosis.** Cognitive characteristics of the sample are summarized in @tbl-desc.
+However, according to the psychiatric assessment, none of the patients with probable PDD was suffering from the major depressive disorder, delirium or other neuropsychiatric abnormalities that would exclude the diagnosis.** Cognitive characteristics of the sample are summarized in @tbl-desc.
 
@@
 #| label: "fig-tree"
-#| fig-cap: "A dendrogram representing algorithms creation."
+#| fig-cap: "A dendrogram representing algorithms for probable Parkinson's Disease Dementia (PDD) construction process."
+#| apa-note: "The dendrogram illustrates the decision process used to construct algorithms for probable Parkinson’s Disease Dementia (PDD).
+The second level depicts the definition of instrumental activities of daily living (IADL) deficit (FAQ total > 7 on the left, FAQ item 9 > 1 on the right).
+The third level indicates the selection of the screening instrument (MMSE, MoCA, sMoCA, or none in the case of Level II).
+Lower branches represent the selection of neuropsychological tests used to define cognitive impairment in executive function, attention, memory, and language, ordered from top to bottom as depicted in the dendrogram.
+Algorithms based on the MMSE are shown in green, those based on the MoCA in blue, on the sMoCA in purple, and Level II algorithms in red.
+The top five screening algorithms according to raw accuracy for predicting diagnosis at Level II (see Table 3) are marked by bold edges.
+Accompanying dots indicate the estimated PDD rate for each algorithm.
+The individual test items comprising each algorithm are listed in Table A1."
 
-make_dendrogram(rates$table)
+make_dendrogram(rates$table, approx_acc$table)
 ```
@@
 **The number of complete pairwise cases ranged from `r obs_pairs[1, "N"]` (`r min_miss` missing) to `r obs_pairs[nrow(obs_pairs), "N"]` (full sample).
-Most comparisons (`r perc_above_k`%) were based off of `r K` or more observations.**[^1]
+Most comparisons (`r perc_above_k`%) were based on `r K` or more observations.**[^1]
 
@@
 
-table_levelII_approximations(conc$table, algos, "Accuracy_raw")
+approx_acc$gtable
 ```
@@
 Our overall PDD rates were consistently lower than previous studies regarding PDD among PD patients, demonstrating wide variability based on various criteria used.
-For instance, a retrospective study reported a PDD rate of 19.7% [@rana2012], while other clinical investigation found even higher rate, reaching up to 30% [@aarsland2005].
+For instance, a retrospective study reported a PDD rate of 19.7% [@rana2012], while other clinical investigations found an even rate higher rate, reaching up to 30% [@aarsland2005].
 A recent meta-analysis synthesizing global data placed the expected PDD rate in PD at 26.30% [@sousa2022].
@@
 
-The advantage of MoCA-based algorithm was evident in combination with using the FAQ total score, reaching accuracy rates of 94–99%.
 Furthermore, in the algorithm using sMoCA, the raw accuracy was moderate, however, the balanced accuracy (i.e.
 combined sensitivity and specificity) was high.
-These findings imply that simplified tools such as sMoCA may be a promising tool for heightening efficiency of the assessment while maintaining diagnostic accuracy.
-**This underscores the clinical utility of sMoCA screening tool.
-Although raw accuracy was moderate, balanced accuracy results ensure avoidance of bias from class imbalance.
-From a practical point of view, it suggests that even brief screening tools can achieve reliable classification, which means an advantage for routine clinical settings and time-restricted assessments.**
+**Consequently, sMoCA appears particularly suitable for approximating Level II PDD diagnosis in populations that differ in PDD prevalence from our cohort, since balanced accuracy, unlike raw accuracy, is independent of prevalence in the sample.
+Moreover, because the sMoCA algorithm demonstrated higher sensitivity while maintaining comparable specificity (see @tbl-approx-bal), it may be especially valuable in contexts where false negatives carry substantial clinical cost.
+In such cases, a neuropsychologist might use sMoCA as an initial screening tool and proceed to a full Level II assessment only for patients who meet criteria for probable PDD in this preliminary stage.**
 
@@
 
-**Modern alternatives, such as multiple imputation techniques, have been shown to produce less biased estimates than case deletion strategies in analyses based on confusion matrices [@karakaya2014], whereas the evidence supporting their superiority in the estimation of Cohen’s** $\kappa$ remains limited [@deraadt2019].
+**Modern alternatives, such as multiple imputation techniques, have been shown to produce less biased estimates than case deletion strategies in analyses based on confusion matrices [@karakaya2014], whereas the evidence supporting their superiority in the estimation of Cohen’s $\kappa$ remains limited [@deraadt2019].
 These modern approaches also require careful specification of the imputation model and the underlying causal mechanism of missingness to ensure appropriate covariate selection [@long2011; @bianco2023].
 For the sake of parsimony, we did not employ advanced missing-data techniques but instead explicitly described the observed missingness patterns.
-Consequently, our findings should be regarded as exploratory and primarily serve as a basis for hypothesis generation in future, more controlled studies.
+Consequently, our findings should be regarded as exploratory and primarily serve as a basis for hypothesis generation in future, more controlled studies.**
 
@@
 
-**Another limitation concerns the use of the FAQ questionnaire for IADL assessment.
-The FAQ is a subjective and informant-reliant measure, which may contain biased data.
+**An additional limitation concerns the use of the FAQ questionnaire for IADL assessment.
+The FAQ is a subjective or informant-reliant measure and thus susceptible to bias.
 Moreover, its content can vary across sociocultural contexts, limiting cross-cultural transferability.
-For example, activities such as financial management, cooking or driving, are not uniformly performed across societies.
+For example, activities such as financial management, cooking or driving, are not universally practised across societies.
 Consequently, both the FAQ scores and diagnostic thresholds used to indicate IADL impairment may not be directly transferable between cultural settings [@odonald2025].
@@
 
-**Furthermore, the FAQ scores may correlate with neuropsychological results of specific cognitive domains, particularly executive functions or memory [@vanrentergem2020; @moheb2017].
-This overlap could artificially increase concordance between cognitive and functional capacity criteria.
-To address these concerns in future research, we suggest using a PD-specific questionnaire, such as the Penn Parkinson's Daily Activities Questionnaire-15 [@brennan2016] or generally more objective methods, preferably performance-based ones.
-Our results underscore the importance of IADL measurement for the PDD diagnosis.
-Therefore, we recommend using more reliable tools with high ecological validity.** For future research, we suggest the use of more objective, modernized tools for IADl assessment, such as performance-based methods [@schmitter-edgecombe2020] or questionnaire adaptations including items regarding gadgets use and digital literacy [@postema2024].
-Our results emphasize the importance of IADL deficit for the PDD diagnosis, therefore, we suggest exploring more reliable tools with high ecological validity.
+**Furthermore, IADL measures may correlate with neuropsychological results of specific cognitive domains, particularly attention/processing speed, and executive function [@reppermund2011; @moheb2017].
+If such correlations were due to shared error variance, they could bias concordance indices measured in this study.
+To examine this possibility, we conducted a post-hoc simulation experiment, available at <https://josefmana.github.io/demcrit/articles/correlation.html>.
+The results indicated that correlations between IADL and neuropsychological measures may either increase or decrease accuracy and balanced accuracy, while consistently inflating Cohen’s $\kappa$ estimates.
+However, given the correlations observed in our dataset, the effect size was small and unlikely to alter the conclusions of our study.**
+
+**To address concerns about measuring IADL outlined above, future research should consider using PD-specific questionnaires or more objective tools.
+Promising options include the Penn Parkinson's Daily Activities Questionnaire-15 [@brennan2016], questionnaire adaptations including items regarding gadgets use and digital literacy [@postema2024] or performance-based assessments [@schmitter-edgecombe2020].
+Our findings underscore the importance of IADL measurement for the PDD diagnosis.
+Therefore, we recommend exploring more reliable tools with high ecological validity.**
 
@@
 anxiety profile) and biomarker correlates.
-Future studies would benefit from focusing on these aspects of PDD.
+**Instead, we only ensured the absence of acute or severe psychiatric symptoms that would preclude a diagnosis of probable PDD, as verified through assessment by a trained neuropsychiatrist.
+However, both the current diagnostic criteria for PDD [@dubois2007] and recent proposals for their revision [@kulisevsky2024] emphasize the use of standardized psychometric instruments for assessing neuropsychiatric symptoms.
+Incorporating such measures could enhance both the efficiency and transparency of the diagnostic process.
+Future research should therefore investigate how integrating structured neuropsychiatric assessments and biomarker data may refine PDD diagnostic accuracy and improve clinical utility.**
 
@@
 
-Future studies should aim to replicate our results on larger and different cohorts **with more heterogenous sample**.
+Future studies should aim to replicate our results on larger and different cohorts **with more heterogeneous sample, and to explore how varying operationalization of other components of probable PDD, namely psychiatric symptoms, affects PDD rates and concordance between diagnostic algorithms**.
 To make this process easier, the code used to generate our results is publicly available and easily applicable to similarly structured data.
@@
 
-In this study, we follow the framework proposed by @lundberg2021 for specifying targets of inference (i.e., the estimands) in qunatitative sciences to increase transparency and connect statistical evidence to relevant theory.
+In this study, we follow the framework proposed by @lundberg2021 for specifying targets of inference (i.e., the estimands) in quantitative sciences to increase transparency and connect statistical evidence to relevant theory.
 @tbl-estimands contains verbal description of the components relating to each of our proclaimed research objectives and map them to the population quantity of interest (the theoretical estimand), data-dependent quantity that could be estimated (the empirical estimand) and quantities that are reported in the study (statistical estimates).
@@
 #| fig-cap: "Summary of the estimates of probable PDD rate."
-#| apa-note: "Vertical lines represent estimates arrived at by using sMoCA (dotted) or Level II (dashed) with FAQ item 9 (orange) or FAQ total score (blue) as criteria for probable PDD."
+#| apa-note: "Vertical lines represent estimates arrived at by using sMoCA (dotted) or Level II (dashed) with FAQ item 9 (orange) or FAQ total score (blue) as criteria for probable PDD.
+The percentages were calculated from all available cases (see Table A3 for numerical summary)."
 
@@
 
-table_levelII_approximations(conc$table, algos, "Balanced Accuracy")
+approx_bal$gtable
 ```

### 2025-10-15 R/compute_descriptives.R

@@ -84,7 +84,7 @@ compute_descriptives <- function(d0, vois) {
     text  <- paste0(dplyr::pull(notes[ , 1]), ": ", dplyr::pull(notes[ , 2])) |> paste(collapse = ", ")
     # Add the text:
     gtab <- gtab |>
-      gt::tab_source_note(html(paste0("<i>Note.</i> ", text)))
+      gt::tab_source_note(html(glue::glue("<i>Note.</i> {text}, All percentages were calculated from the whole sample.")))
   }
   # Return results:
   list(table = demtab, gtable = gtab)

### 2025-10-15 R/describe_concordance.R

@@ -186,7 +186,7 @@ describe_concordance <- function(d0) {
     ggplot2::aes(x = predictor, y = reference, fill = Kappa_raw) +
     ggplot2::geom_tile() +
     ggplot2::scale_fill_gradient2(high = "steelblue", na.value = "white") +
-    ggplot2::labs(x = "Predictor", y = "Reference") +
+    ggplot2::labs(x = "Predictor", y = "Reference", fill = "Cohen's \u03ba") +
     ggplot2::theme(
       axis.text.x = ggplot2::element_text(colour = dplyr::if_else(ord$xcol == "black", "black", "red3"), angle = 66, hjust = 1),
       axis.text.y = ggplot2::element_text(colour = dplyr::if_else(ord$ycol == "black", "black", "red3"))

### 2025-10-15 R/diagnose_pdd_case.R

@@ -37,7 +37,7 @@ diagnose_pdd_case <- function(x, c) {
     crit2 = TRUE, # 2. Parkinson’s disease developed before dementia
     crit3 = dplyr::if_else(x[[c$glob]] < c$glob_t, TRUE, FALSE), # 3. Global deficit
     crit4 = dplyr::if_else(x[[c$iadl]] > c$iadl_t, TRUE, FALSE), # 4. Dementia has Impact on ADLs
-    crit5 = case_when( # 5. Impaired cognition (for Yes, at least of 2 tests abnormal)
+    crit5 = dplyr::case_when( # 5. Impaired cognition (for Yes, at least of 2 tests abnormal)
       c$group == "mmse" ~ dplyr::if_else(rowSums(dplyr::across(tidyselect::starts_with("impaired_orig"))) > 1, TRUE, FALSE),
       c$group == "moca" ~ dplyr::if_else(rowSums(dplyr::across(tidyselect::starts_with("impaired"))) > 1, TRUE, FALSE),
       c$group %in% c("smoca","lvlII") ~ crit3
@@ -47,5 +47,5 @@ diagnose_pdd_case <- function(x, c) {
     crit8 = TRUE, # 8. Absence of other abnormalities that obscure diagnosis
     PDD = dplyr::if_else(rowSums(dplyr::across(tidyselect::starts_with("crit"))) == 8, TRUE, FALSE)
   ) |>
-    select(id, PDD, tidyselect::starts_with("impaired"), tidyselect::starts_with("crit"))
+    dplyr::select(id, PDD, tidyselect::starts_with("impaired"), tidyselect::starts_with("crit"))
 }

### 2025-10-15 R/make_dendrogram.r

@@ -8,15 +8,37 @@
 #'
 #' @param tab A table containing all criteria used as generated by
 #'    \code{summarise_rates}.
-#'
+#' @param top A tibble containing top criteria for predicting PDD
+#'    diagnosis by Level II a generated by \code{table_levelII_approximations}.
 #'
 #' @returns A ggplot object with a dendrogram created
 #'    by {ggraph}.
 #'
-#'@seealso [summarise_rates()] which is fed into this function.
+#'@seealso
+#' * [summarise_rates()] prepares `tab`.
+#' * [table_levelII_approximations()] prepares `top`.
+#'
+#' @examples
+#' \dontrun{
+#' # Raw data:
+#' p <- data_paths("data-raw")
+#' data <- prepare_data(p)
+#' pdd  <- diagnose_pdd_sample(data)
+#'
+#' # PDD rates for point size:
+#' vars <- here::here("data-raw", "VariablesOfInterest.csv")
+#' rates <- summarise_rates(pdd, vars)
+#'
+#' # Concordance with Level II for edge width:
+#' algos <- list_algorithms(rates$table)
+#' data(concords)
+#' best <- table_levelII_approximations(concords, algos, "Accuracy_raw")
+#'
+#' tree <- make_dendrogram(rates$table, best$table)
+#' }
 #'
 #' @export
-make_dendrogram <- function(tab) {
+make_dendrogram <- function(tab, top) {
   # Prepare data for plot edges:
   data <- tab |>
     dplyr::mutate(
@@ -53,6 +75,19 @@ make_dendrogram <- function(tab) {
       tidyselect::starts_with("lvl"), IADL, Global, Attention, Executive,
       Construction, Memory, Language, type, N, Rate, perc
     )
+  # Extract 'the best' algorithms:
+  bold <- top |>
+    dplyr::filter(stringr::str_detect(part, "Top")) |>
+    dplyr::select(tidyselect::starts_with("predictor")) |>
+    tidyr::pivot_longer(tidyselect::everything()) |>
+    dplyr::pull(value)
+  # Find links of 'the best' algorithms:
+  links <- data |>
+    dplyr::filter(type %in% bold) |>
+    dplyr::select(tidyselect::starts_with("lvl")) |>
+    tidyr::pivot_longer(tidyselect::everything()) |>
+    dplyr::distinct() |>
[…]

### 2025-10-15 R/specify_algorithms.R

@@ -51,14 +51,14 @@ specify_algorithms <- function() {
       lang = c("moca_anim", "moca_abs"),
       iadl = c("faq", "faq_9")
     ) |>
-      mutate(
+      dplyr::mutate(
         exec_t = dplyr::case_when(exec == "vf_k" ~ 11, exec == "moca_cloc" ~ i),
         iadl_t = dplyr::case_when(iadl == "faq" ~ 7, iadl == "faq_9" ~ 1),
         lang_t = dplyr::case_when(lang == "moca_abs" ~ 2 , lang == "moca_anim" ~ 3)
       )
   }) |>
     dplyr::distinct() |>
-    mutate(type = paste0("MoCA (", seq_len(n()), ")"))
+    dplyr::mutate(type = paste0("MoCA (", seq_len(dplyr::n()), ")"))
   # sMoCA-based algorithms
   smoca <- tidyr::crossing(
     group = "smoca",
@@ -83,5 +83,5 @@ specify_algorithms <- function() {
     )
   # Return all-in-one:
   purrr::map_dfr(list(mmse, moca, smoca, lvlII), \(x) x) |>
-    mutate(dplyr::across(tidyselect::ends_with("_t"), \(x) dplyr::if_else(is.na(x), 0, x)))
+    dplyr::mutate(dplyr::across(tidyselect::ends_with("_t"), \(x) dplyr::if_else(is.na(x), 0, x)))
 }

### 2025-10-15 R/summarise_rates.R

@@ -94,7 +94,8 @@ summarise_rates <- function(d0, vars, descending = TRUE) {
     dplyr::arrange(dplyr::desc(perc)) |>
     dplyr::select(type, N, Rate) |>
     gt_apa_table() |>
-    gt::cols_label(type ~ "Algorithm")
+    gt::cols_label(type ~ "Algorithm") |>
+    gt::tab_source_note(gt::html("<i>Note.</i> Percentages were calculated from all available cases. The items comprising each listed algorithm can be found in Table A1."))
   # Make a table showing algorithms' specification:
   gtab_algos <- tab |>
     dplyr::select(-perc, -N, -Rate) |>
@@ -116,7 +117,7 @@ summarise_rates <- function(d0, vars, descending = TRUE) {
     text <- paste0(dplyr::pull(notes[ , 1]), ": ", dplyr::pull(notes[ , 2])) |> paste(collapse = ", ")
     # Add the text:
     gtab_algos <- gtab_algos |>
-      gt::tab_source_note(gt::html(paste0("<i>Note.</i> ", text)))
+      gt::tab_source_note(gt::html(glue::glue("<i>Note.</i> {text}")))
   }
   # Visualisation code:
   smoca_9 <- subset(prevs, type == subset(d0$algorithms, group == "smoca" & iadl == "faq_9")$type)$perc

### 2025-10-15 R/table_levelII_approximations.R

@@ -19,9 +19,14 @@
 #'    be used for comparisons. Can be one of `"Accuracy_raw"`, `"Balanced Accuracy"`,
 #'    `"Kappa_raw"`, `"Sensitivity"` or `"Specificity"`.
 #'
-#' @returns A \code{gt} table object summarising the concordance statistics
-#'    (e.g., accuracy, kappa) of the five best and five worst Level I algorithms
-#'    in predicting the specified Level II-based diagnosis of probable PDD.
+#' @returns A list containing:
+#'   \describe{
+#'     \item{\code{table}}{A tibble with the five best and five worst Level I algorithms
+#'     in predicting the specified Level II-based diagnosis of probable PDD.}
+#'     \item{\code{gtable}}{A \code{gt} table object summarising the concordance statistics
+#'     (e.g., accuracy, kappa) of the five best and five worst Level I algorithms in predicting
+#'     the specified Level II-based diagnosis of probable PDD.}
+#'   }
 #'
 #' @seealso
 #' * [concords] contains data for `values`.
@@ -51,7 +56,7 @@
 #'
 #' @export
 table_levelII_approximations <- function(values, struct, crit) {
-  purrr::map_dfr(1:2, function(ii) {
+  tab <- purrr::map_dfr(1:2, function(ii) {
     values |>
       dplyr::filter(reference == paste0("Lvl.II (", ii, ")")) |>
       dplyr::filter(predictor %in% struct[[ii]]) |>
@@ -74,7 +79,9 @@ table_levelII_approximations <- function(values, struct, crit) {
       dplyr::across(tidyselect::where(is.numeric), \(x) do_summary(x, 2))
     ) |>
     dplyr::filter(!is.na(part)) |>
-    dplyr::select(part, tidyselect::ends_with("(1)"), tidyselect::ends_with("(2)")) |>
+    dplyr::select(part, tidyselect::ends_with("(1)"), tidyselect::ends_with("(2)"))
+  # Prepare gt table as well:
+  gtab <- tab |>
     gt_apa_table(grp = "part") |>
     gt::tab_spanner(columns = tidyselect::ends_with("(1)"), label = "Level II (1)", gather = FALSE) |>
     gt::tab_spanner(columns = tidyselect::ends_with("(2)"), label = "Level II (2)", gather = FALSE) |>
@@ -88,18 +95,20 @@ table_levelII_approximations <- function(values, struct, crit) {
     ) |>
     gt::tab_footnote(
       locations = gt::cells_column_spanners("Level II (1)"),
[…]

### 2025-10-18 R/data_concords.r
Finishing the two primary vignettes.

@@ -14,8 +14,9 @@
 #' * [psych::cohen.kappa()] computes and documents Cohen’s kappa and related
 #'   metrics.
 #' * [caret::confusionMatrix()] computes and documents the remaining metrics.
+#' * \link{rates} contains definitions of each algorithm.
 #'
-#' @format ## `concords`
+#' @format
 #' A tibble with 7,744 rows and 21 columns:
 #' \describe{
 #'   \item{predictor}{Character; name of the algorithm used as a predictor.}

### 2025-10-18 R/data_rates.r

@@ -1,0 +1,26 @@
+#' Algorithm Specifications and PDD Rates
+#'
+#' A comprehensive table containing all algorithms evaluated in the article. It
+#' containing algorithm names as well as their constituting components including
+#' operationalisations of IADL deficit, global cognitive deficit and cognitive
+#' impairment according to monitored domains.
+#'
+#' @seealso \link{concords} for data containing pairwise concordance statistics.
+#'
+#' @format
+#' A tibble with 68 rows and 11 columns:
+#' \describe{
+#'   \item{type}{Character; name of the algorithm.}
+#'   \item{perc}{Numeric; raw PDD rate.}
+#'   \item{Global}{Character; operationalisation of global cognitive deficit.}
+#'   \item{Executive}{Character; operationalisation of executive deficit.}
+#'   \item{Construction}{Character; operationalisation of visuoconstruction deficit.}
+#'   \item{Memory}{Character; operationalisation of memory deficit.}
+#'   \item{Language}{Character; operationalisation of language deficit.}
+#'   \item{IADL}{Character; operationalisation of IADL deficit.}
+#'   \item{N}{Integer; number of observations for a given algorithm.}
+#'   \item{Rate}{Character; number and percentage of patients with probable PDD.}
+#' }
+#'
+#' @source Generated by the targets pipeline in `_targets.R`.
+"rates"

### 2025-10-18 R/init.r

@@ -1,6 +1,0 @@
-.onLoad <- function(libname, pkgname) {
-  # Load the tidyverse package
-  if (!requireNamespace("tidyverse", quietly = TRUE)) {
-    stop("The tidyverse package must be installed to use this package.")
-  }
-}

### 2025-10-18 R/prepare_defaults.r

@@ -1,0 +1,54 @@
+#' Prepare Defaults for Data Simulation
+#'
+#' Prepares a list with default values for data simulation. These defaults are
+#' derived from the data used as a basis of the demcrit article. This function
+#' is a helper with no inputs and is typically called within \code{simulate_PDD_data}.
+#'
+#' @returns A list containing:
+#'   \describe{
+#'     \item{\code{Mu}}{Named numeric vector containing means for FAQ, MMSE, MoCA,
+#'     and sMoCA.}
+#'     \item{\code{Sigma}}{Named matrix containing variance-covariance of the variables
+#'     (FAQ, MMSE, MoCA, and sMoCA).}
+#'     \item{\code{cens}}{A matrix with named rows denoting scales (FAQ, MMSE, MoCA,
+#'     and sMoCA) and two columns - the first containing lower bound, the second
+#'     containing upper bound of each scale.}
+#'     \item{\code{crits}}{A data.frame with four columns denoting IADL measure to use
+#'     (`IADL`), threshold for the IADL measure (`IADL_thres`), cognitive measure to
+#'     use (`cognition`), and threshold for the cognitive measure (`cognition_thres`).
+#'     Score above `IADL_thres` and below `cognition_thres` are regarder to indicate
+#'     PDD symptoms. Rownames contain algorithm names.}
+#'   }
+#'
+#' @seealso [simulate_pdd_data()] uses outputs of this function unless specified
+#'    by the user.
+#'
+#' @export
+prepare_defaults <- function() {
+  # Data means:
+  mu = c(FAQ = 4.05, MMSE = 26.69, MoCA = 24.07, sMoCA = 11.26)
+  # Data variance-covariance matrix:
+  sd <- c(FAQ = 4.89, MMSE = 2.22, MoCA = 3.48, sMoCA = 2.74)
+  corrs <- matrix(
+    c(1, -0.21, -0.27, -0.26, -0.21, 1, 0.63, 0.54, -0.27, 0.63, 1, 0.95, -0.26, 0.54, 0.95, 1),
+    nrow = length(sd), # If the defaults ever changed.
+    dimnames = lapply(seq_len(2), \(i) c("FAQ", "MMSE", "MoCA", "sMoCA"))
+  )
+  sigma <- cor2cov(corrs, sd)
[…]

### 2025-10-18 R/simulate_pdd_data.r

@@ -1,0 +1,108 @@
+#' Make Fake Data
+#'
+#' Prepares a minimal example of data in the right format for analysis via the demcrit
+#' package functions. It is compulsory to specify FAQ summary data (mean, variance and
+#' covariance, see below) and at least one additional global cognitive measure summary
+#' data. The function then simulates `N` synthetic patients' PDD status via the
+#' following steps:
+#'
+#' 1. generate latent FAQ and cognitive scores via [MASS::mvrnorm()],
+#' 2. generate observed FAQ item 9 data via the [rbinom()] following tau-equivalence
+#'    assumption applied to the FAQ questionnaire,
+#' 3. censor the scores according to `cens` and round to nearest integer to obtain
+#'    observed data
+#' 4. apply criteria for PDD from `crits` to get per algorithm/patient pair probable
+#'    PDD.
+#'
+#' @param N A numeric indicating number of simulated patients. Defaults to 203.
+#' @param defaults A logical indicator whether default values derived from the
+#'    observed data should be used (`TRUE`, default) or the user will provide their
+#'    values instead (`FALSE`).
+#' @param Mu Named numeric vector of means. Needs to contain one component denoting
+#'    FAQ. If `NULL` (default), it is set to values from observed data..
+#' @param Sigma Named matrix of variances and covariances. If `NULL` (default), it
+#'    is set at empirical values from the observed data.
+#' @param cens A matrix or data frame with two columns, the first indicating lower
+#'    bound, the second the upper bound of a scale in the rowname. Defaults to
+#'    `NULL` which does not censor data (and prints latent scores instead).
+#' @param crits description
+#'
+#' @returns A data.frame
+#'
+#' @seealso [prepare_defaults()] for correct format of simulation parameters specification.
+#'
+#' @examples
+#' \dontrun{
+#' # Simulate using defaults:
+#' sim1 <- simulate_pdd_data()
[…]

### 2025-10-19 _manuscript/manuscript.qmd
Changes in manuscript, mainly formating

@@
 All patients underwent a comprehensive neuropsychological battery.
-A total of 68 diagnostic algorithms were operationalised based on the combinations of various cognitive assessments, including the Mini Mental State Examination (MMSE), the Montreal Cognitive Assessment (MoCA), and the shortened version of MoCA (sMoCA).
+A total of 68 diagnostic algorithms were operationalized based on the combinations of various cognitive assessments, including the Mini-Mental State Examination (MMSE), the Montreal Cognitive Assessment (MoCA), and the shortened version of MoCA (sMoCA).
 Functional impairment was based on the Functional Assessment Questionnaire (FAQ), specifically FAQ item 9 or the FAQ total score.
-Concordance between algorithms was assessed using pairwise Cohen’s .
+Concordance between algorithms was assessed using pairwise Cohen’s κ.
 The predictive validity of Level I algorithms for Level II diagnosis was assessed using Exact Binomial Tests.
@@
 Conclusions: Diagnostic outcomes for PDD are sensitive to the choice of cognitive and functional instruments.
-Our results support some of the proposed changes to PDD diagnostic criteria on Level I, emphasising the consistency of MoCA-based assessments and comprehensive functional impairment evaluation."
+Our results support some of the proposed changes to PDD diagnostic criteria on Level I,  emphasizing the consistency of MoCA-based assessments and comprehensive functional impairment evaluation."
 keywords:
@@
   disclosures:
-    financial-support: "Supported by The project National Institute for Neurological Research (Programme EXCELES, ID Project No.
+    financial-support: "Supported by the project National Institute for Neurological Research (Programme EXCELES, ID Project No.
 LX22NPO5107) - Funded by the European Union – Next Generation EU; Charles University: Cooperatio Program in Neuroscience; General University Hospital in Prague project MH CZ-DRO-VFN64165."
@@
 According to a recent meta-analysis, approximately one-quarter of PD patients are likely to be diagnosed with PDD [@sousa2022].
-However, reported PDD rate estimates vary widely, ranging from 14% up to 55%, depending on the methodological criteria employed [@sousa2022].
+However, reported PDD rate estimates vary widely, ranging from 14% up to 55%, depending on the methodological criteria employed.
 Moreover, factors such as patients' sex [@cereda2016], age and disease duration appear to modulate the risk of cognitive decline and PDD [@oh2016; @rana2012].
@@
 A recent call for a change pinpoints limitations regarding the original criteria and suggests various updates to enhance their utility [@kulisevsky2024].
-Proposed suggestions include replacement of the Mini Mental State Examination (MMSE) by the Montreal Cognitive Assessment (MoCA), which is more sensitive to PD specific cognitive impairment; expansion of instrumental activities of daily living (IADL) evaluation; inclusion of language assessment; recognition of anxiety as one **of** the neuropsychiatric symptoms relevant in PDD; and integration of biomarkers.
+Proposed suggestions include replacement of the Mini-Mental State Examination (MMSE) by the Montreal Cognitive Assessment (MoCA), which is more sensitive to PD-specific cognitive impairment; expansion of instrumental activities of daily living (IADL) evaluation; inclusion of language assessment; recognition of anxiety as one **of** the neuropsychiatric symptoms relevant in PDD; and integration of biomarkers.
 
@@
 Cognitive performance at Level I was assessed by the MMSE [@stepankova2014; @folstein1975] and the MoCA [@kopecek2016; @nasreddine2005].
-The comprehensive neuropsychological assessment at Level II evaluated five cognitive domains through specific tests: attention and working memory assessed by Trail Making Test Part A (TMT-A) [@bezdicek2012; @reitan2004], and WAIS Digit Span Backward (WAIS DSB) [@wechsler1997], executive function by Categorical Verbal Fluency (CF) [@nikolai2015], and subtest from the Prague Stroop Test – Colors (PST-C) [@bezdicek2021], language by the WAIS Similarities subtest [@wechsler1997], and the Boston Naming Test (BNT-60) [@zemanova2016; @kaplan1983], memory by the Rey Auditory Verbal Learning Test (RAVLT) [@frydrychova2018; @bezdicek2013; @rey1964] delayed recall, and the Brief Visual Memory Test–Revised (BVMTR) [@havlik2020; @benedict1997] delayed recall, or WAIS Family Pictures subtest [@wechsler1997] delayed recall, visuospatial function assessed by the Judgment of Line Orientation Test (JoLO) [@benton1983], and Clock Drawing Test (CLOX) [@royall1998].
+The comprehensive neuropsychological assessment at Level II evaluated five cognitive domains through specific tests: attention and working memory assessed by Trail Making Test Part A (TMT-A) [@bezdicek2012; @reitan2004], and WAIS Digit Span Backward (WAIS DSB) [@wechsler1997], executive function by Categorical Verbal Fluency - Animals (CF-A) [@nikolai2015], and subtest from the Prague Stroop Test – Colors (PST-C) [@bezdicek2021], language by the WAIS Similarities subtest [@wechsler1997], and the Boston Naming Test (BNT-60) [@zemanova2016; @kaplan1983], memory by the Rey Auditory Verbal Learning Test (RAVLT-DR) [@frydrychova2018; @bezdicek2013; @rey1964] delayed recall, and the Brief Visual Memory Test–Revised (BVMTR-DR) [@havlik2020; @benedict1997] delayed recall, or WAIS Family Pictures subtest [@wechsler1997] delayed recall, visuospatial function assessed by the Judgment of Line Orientation Test (JoL) [@benton1983], and Clock Drawing Test (CLOX-I) [@royall1998].
 
@@
 In this study, we applied three distinct sets of diagnostic algorithms for probable PDD at Level I.
-The first set was based on the original framework [@dubois2007], which utilized the Mini-Mental State Examination (MMSE) as a global cognitive screening tool, supplemented by assessments of attention, executive function, visuospatial abilities, and memory.
+The first set was based on the original framework [@dubois2007], which utilized the MMSE as a global cognitive screening tool, supplemented by assessments of attention, executive function, visuospatial abilities, and memory.
 The second set of algorithms was based on the recent call for change of dementia diagnostic guidelines [@kulisevsky2024], which advocates for more sensitive cognitive domain assessments in the context of PD.
@@
 The theoretical estimand refers to a unit-specific quantity defined over a target population and represents the ideal quantity that would address the research question under optimal conditions, such as access to complete population data or perfect experimental control.
-In contrast, the empirical estimand corresponds to the quantity that is actually computable using the available dataset, given real world constraints.
+In contrast, the empirical estimand corresponds to the quantity that is actually computable using the available dataset, given real-world constraints.
 A full description of the study’s estimands and their relation to our research objectives is presented in the Appendix (see @tbl-estimands).
@@
 For each pair of algorithms, the analysis was repeated twice such that each variable of the pair served once as the reference and once as the predictor.
-Following measures were used to evaluate pairwise concordance between different algorithms for probable PDD: 1) Cohen's $\kappa$ with its 95% confidence interval (CI) computed via the `cohen.kappa()` function from the R package *psych* [@revelle2024]; 2) Accuracy (i.e., the proportion of correct predictions, both true positives and true negatives, among the total number of cases) with its 95% CI; 3) Sensitivity/Recall (i.e., the proportion of true positives); and 4) Specificity (i.e., the proportion of true negatives).
+The following measures were used to evaluate pairwise concordance between different algorithms for probable PDD: 1) Cohen's $\kappa$ with its 95% confidence interval (CI) computed via the `cohen.kappa()` function from the R package *psych* [@revelle2024]; 2) Accuracy (i.e., the proportion of correct predictions, both true positives and true negatives, among the total number of cases) with its 95% CI; 3) Sensitivity/Recall (i.e., the proportion of true positives); and 4) Specificity (i.e., the proportion of true negatives).
 
@@
 NIR is the accuracy that could be obtained by always predicting the majority class and in our case it is equivalent to the complement of the PDD rate estimate according to the reference algorithm.
-Accuracy of prediction was compared to the NIR via a one-sided Exact Binomial Test as implemented by the `binom.test()` R stats function.
+The accuracy of prediction was compared to the NIR via a one-sided Exact Binomial Test as implemented by the `binom.test()` R stats function.
 Reference/predictor pairs associated with p \< .05 were considered to show significantly better accuracy than NIR.
@@
 **Missing data were handled by complete cases analysis and pairwise complete cases analysis.
-In other words, each univariate and analysis pairwise comparison used all available data.** Data wrangling and visualizations were done in the *tidyverse* package [@wickham2019] and tables were formatted in the *gt* package [@iannone2024].
+In other words, each univariate analysis and pairwise comparison used all available data.** Data wrangling and visualizations were done in the *tidyverse* package [@wickham2019] and tables were formatted in the *gt* package [@iannone2024].
 All analyses were conducted within the R (version `r with(version, paste(major, minor, sep="."))`) software environment for statistical computing [@rsoft].
@@
 The sample included `r sex` men, with an average of `r age`, `r edu`, `r dur`, `r updrs_off` and `r updrs_on`.
-**Descriptive statistics for neuropsychiatric symptoms indicated within average level of depressive and anxiety symptoms in our cohort, with the average BDI-II of `r bdi`, average STAI X1 of `r staix1`, and average STAI X2 of `r staix2`.
+**Descriptive statistics for neuropsychiatric symptoms indicated within average levels of depressive and anxiety symptoms in our cohort, with the average BDI-II of `r bdi`, average STAI X1 of `r staix1`, and average STAI X2 of `r staix2`.
 However, according to the psychiatric assessment, none of the patients with probable PDD was suffering from the major depressive disorder, delirium or other neuropsychiatric abnormalities that would exclude the diagnosis.** Cognitive characteristics of the sample are summarized in @tbl-desc.
@@
 On average, the estimated PDD rate was `r rates_overall["M"]`% (SD = `r rates_overall["SD"]`, Md = `r rates_overall["Md"]`, range `r rates_overall["minmax"]`).
-Notably, the estimates were substantially lower when FAQ item 9 was used as a criterion of IADL deficit (M = `r rates_FAQ9["M"]`% SD = `r rates_FAQ9["SD"]`, Md = `r rates_FAQ9["Md"]`, range `r rates_FAQ9["minmax"]`) compared to using the total FAQ score criterion (M = `r rates_FAQtot["M"]`% SD = `r rates_FAQtot["SD"]`, Md = `r rates_FAQtot["Md"]`, range `r rates_FAQtot["minmax"]`) as demonstrated in @fig-tree **(see also @fig-rates for per-operationalisation distribution of PDD rate estimates)**.
+Notably, the estimates were substantially lower when FAQ item 9 was used as a criterion of IADL deficit (M = `r rates_FAQ9["M"]`% SD = `r rates_FAQ9["SD"]`, Md = `r rates_FAQ9["Md"]`, range `r rates_FAQ9["minmax"]`) compared to using the total FAQ score criterion (M = `r rates_FAQtot["M"]`% SD = `r rates_FAQtot["SD"]`, Md = `r rates_FAQtot["Md"]`, range `r rates_FAQtot["minmax"]`) as demonstrated in @fig-tree **(see also @fig-rates for per-operationalization distribution of PDD rate estimates)**.
 Neither age, sex nor their interaction (*p*s $\geq$ `r min_p`) reliably predicted probable PDD classification across algorithms (see @fig-data and @fig-pars).
@@
 
-When IADL deficit was defined by total FAQ score, the Level II estimate of PDD rate was `r lvlII_rates["faq_tot"]`.
-All five Level I algorithms that approximated the Level II classification most accurately were MoCA-based and defined Executive Function deficit by Clock drawing rather than Verbal fluency test.
-On the other hand, two out of the five Level I algorithms with the lowest accuracy were MMSE-based, whereas the remaining three were MoCA-based and defined Executive function deficit by Verbal fluency test.
+When the IADL deficit was defined by total FAQ score, the Level II estimate of PDD rate was `r lvlII_rates["faq_tot"]`.
+All five Level I algorithms that approximated the Level II classification most accurately were MoCA-based and defined Executive Function deficit by Clock drawing rather than the Verbal fluency test.
+On the other hand, two out of the five Level I algorithms with the lowest accuracy were MMSE-based, whereas the remaining three were MoCA-based and defined Executive function deficit by the Verbal fluency test.
 
-When IADL deficit was defined by FAQ item 9 score, the Level II estimate of PDD rate was `r lvlII_rates["faq_9"]`.
+When the IADL deficit was defined by the FAQ item 9 score, the Level II estimate of PDD rate was `r lvlII_rates["faq_9"]`.
 Overall, the difference between the most accurate and the least accurate Level I algorithms was lower than in the case of IADL deficit being defined by FAQ total score (see @tbl-approx).
-The five most accurate algorithms were all MoCA-based, defined Executive Function deficit by Clock drawing (with threshold \< 2) and in majority of cases defined Language deficit by Animal naming.
+The five most accurate algorithms were all MoCA-based, defined Executive Function deficit by Clock drawing (with threshold \< 2) and in the majority of cases defined Language deficit by Animal naming.
 Two out of the five Level I algorithms with the lowest accuracy were MMSE-based, whereas the remaining three were MoCA-based and defined Executive Function deficit by Clock drawing (with threshold \< 3) and Language deficit by Abstraction.
@@
 
-Our overall PDD rates were consistently lower than previous studies regarding PDD among PD patients, demonstrating wide variability based on various criteria used.
-For instance, a retrospective study reported a PDD rate of 19.7% [@rana2012], while other clinical investigations found an even rate higher rate, reaching up to 30% [@aarsland2005].
+Our overall PDD rates were consistently lower than previous studies regarding **dementia** among PD patients, demonstrating wide variability based on various criteria used.
+For instance, a retrospective study reported a PDD rate of 19.7% [@rana2012], while other clinical investigations found even higher rate, reaching up to 30% [@aarsland2005].
 A recent meta-analysis synthesizing global data placed the expected PDD rate in PD at 26.30% [@sousa2022].
@@
 Interestingly, we did not observe any reliable age-related differences in PDD rate within our cohort.
-This lack of age-dependency may, however, be also stemming from the relatively younger age of our cohort, because previous reports indicate that association between age and PDD is not linear but increases with age and may not reach substantial values before **older age.
+This lack of age-dependency may, however, also stem from the relatively younger age of our cohort, because previous reports indicate that the association between age and PDD is not linear but increases with age and may not reach substantial values before **older age.
 In both @rana2012 and @oh2016, nine out of ten patients with dementia were 70 years of age or older.
@@
 One possible explanation of this difference follows from the observation that algorithms using the full-scale definition yielded higher PDD rate estimates.
-Because there was a higher probability of being diagnosed with IADL deficit based on the full FAQ scale, there was also a bigger room for disagreement in the Cognitive Impairment status when different indexes were used (e.g.
-by defining executive deficit via clock drawing vs.
-verbal fluency).
+Because there was a higher probability of being diagnosed with IADL deficit based on the full FAQ scale, there was also a bigger room for disagreement in the cognitive impairment status when different indexes were used (e.g.
+by defining executive deficit via Clock Drawing vs.
+Verbal Fluency).
 
@@
 Using Level II diagnosis as the gold standard, MoCA-based Level I algorithms, particularly those using Clock Drawing to assess executive function, demonstrated the highest predictive accuracy.
-This supports recent proposals to modernise PDD diagnostic frameworks [@kulisevsky2024], favouring MoCA-derived components and more specific, PD-tailored functional assessment tools.
+This supports recent proposals to modernize PDD diagnostic frameworks [@kulisevsky2024], favoring MoCA-derived components and more ADL-specific, PD-tailored functional assessment tools.
 In contrast, MMSE-based algorithms consistently underperformed, suggesting limited sensitivity in capturing cognitive deficits typical in PDD.
@@
 **This study’s generalizability is limited by the homogeneity of the patient cohort, which does not reflect the diversity of cognitive profiles seen in broader PD populations.
-Specifically, the younger age of the sample, and a possible underrepresentation of high-risk phenotypes for PDD constraints the generality of the presented findings.**
+Specifically, the younger age of the sample, and a possible underrepresentation of high-risk phenotypes for PDD constrain the generality of the presented findings.**
 
 **As noted above, the younger age of our sample may partly explain the lower rate of PDD observed compared to previous studies.
-As discussed in the *Theoretical and Empirical Estimands* section of the Appendix, neither estimates of PDD rate nor predictive performance of demographic variables should therefore not be generalized beyond PD patients that are candidates for DBS.
+As discussed in the *Theoretical and Empirical Estimands* section of the Appendix, neither estimates of PDD rate nor predictive performance of demographic variables therefore should not be generalized beyond PD patients who are DBS candidates.
 The extent to which our findings on the concordance between diagnostic algorithms apply to the broader PD population remains to be determined in future studies using different types of cohorts, such as de novo patients or community-based samples.**
@@
 **Another possible avenue for analysing the present data would be to exploit their hierarchical structure by fitting a multilevel model with algorithms nested within cognitive tests (see @fig-tree).
-Such an approach could leverage partial pooling [@gelman2012] and might approximate modern item response theory frameworks [@burkner2021], where PDD is conceptualised as a latent trait and the individual diagnostic algorithms act as items measuring it.
-If feasible, this psychometric perspective may represent an exciting new direction for conceptualising complex clinical phenomena such as PDD (cf.
+Such an approach could leverage partial pooling [@gelman2012] and might approximate modern item response theory frameworks [@burkner2021], where PDD is conceptualized as a latent trait and the individual diagnostic algorithms act as items measuring it.
+If feasible, this psychometric perspective may represent an exciting new direction for conceptualizing complex clinical phenomena such as PDD (cf.
 @miller2012, @kiselica2021).**
@@
 The FAQ is a subjective or informant-reliant measure and thus susceptible to bias.
-Moreover, its content can vary across sociocultural contexts, limiting cross-cultural transferability.
-For example, activities such as financial management, cooking or driving, are not universally practised across societies.
+Moreover, its content can vary across sociocultural contexts, which limits its cross-cultural transferability.
+For example, activities such as financial management, cooking or driving, are not universally practiced across societies.
 Consequently, both the FAQ scores and diagnostic thresholds used to indicate IADL impairment may not be directly transferable between cultural settings [@odonald2025].
@@
 **To address concerns about measuring IADL outlined above, future research should consider using PD-specific questionnaires or more objective tools.
-Promising options include the Penn Parkinson's Daily Activities Questionnaire-15 [@brennan2016], questionnaire adaptations including items regarding gadgets use and digital literacy [@postema2024] or performance-based assessments [@schmitter-edgecombe2020].
+Promising options include the Penn Parkinson's Daily Activities Questionnaire-15 [@brennan2016], questionnaire adaptations including items regarding gadget use and digital literacy [@postema2024] or performance-based assessments [@schmitter-edgecombe2020].
 Our findings underscore the importance of IADL measurement for the PDD diagnosis.
@@
 
-Finally, whereas our study systematically investigated how varying definition of global deficit, impaired cognition and IADL deficit affect probable PDD classification, it did not explore associations of PDD diagnosis with its neuropsychiatric (e.g.
+Finally, whereas our study systematically investigated how varying definitions of global deficit, impaired cognition and IADL deficit affect probable PDD classification, it did not explore associations of PDD diagnosis with its neuropsychiatric (e.g.
 anxiety profile) and biomarker correlates.
@@
 In sum, our study systematically investigated how varying definitions of impaired cognition and IADL deficit affect PDD diagnostic accuracy.
-Our study highlights the variability in PDD classification across Level I diagnostic algorithms, significantly influenced by IADL operationalisation and the choice of cognitive screening tools.
-The findings strongly support the call for a change of the current diagnostic criteria [@kulisevsky2024], favouring the use of MoCA-based components and comprehensive IADL assessments.
+Our study highlights the variability in PDD classification across Level I diagnostic algorithms, significantly influenced by IADL operationalization and the choice of cognitive screening tools.
+The findings strongly support the call for a change of the current diagnostic criteria [@kulisevsky2024], favoring the use of MoCA-based components and comprehensive IADL assessments.
 
@@
 
-$Attention = \{z(TMT\ A) < -1.5\ \cup z(WAIS\ DSB) < -1.5\}$
+$Attention = \{z(\textit{TMT-A}) < -1.5\ \cup z(WAIS\ DSB) < -1.5\}$
 
-$Executive = \{z(CF\ A) < -1.5\ \cup z(PST\ C) < -1.5\}$
+$Executive = \{z(CF\ A) < -1.5\ \cup z(\textit{PST-C}) < -1.5\}$
 
-$Construction = \{z(JoLO) < -1.5\ \cup z(CLOXI) < -1.5\}$
+$Construction = \{z(JoL) < -1.5\ \cup z(\textit{CLOX-I}) < -1.5\}$
 
-$Memory = \{z(RAVLT\ DR) < -1.5\ \cup z(BVMTR\ DR) < -1.5\ \cup z(WMS\text-III\ Family\ Pictures) < -1.5\}$
+$Memory = \{z(\textit{RAVLT-DR}) < -1.5\ \cup z(\textit{BVMTR- DR}) < -1.5\ \cup z(WMS\text-III\ Family\ Pictures) < -1.5\}$
 
-$Language = \{z(WAIS\ Similarities) < -1.5 \ \cup z(BNT\ 60) < -1.5\}$
+$Language = \{z(WAIS\ Similarities) < -1.5 \ \cup z(\textit{BNT-60}) < -1.5\}$
 

### 2025-10-19 _manuscript/manuscript.qmd
Finishing vignettes just before submission. v1.0.0 finished!

@@
 updrs_on <- msd_txt("UPDRS III on state", " Unified Parkinson Disease Rating Scale (UPDRS), part III in medication ON state")
-bdi <- msd_txt("BDI")
+bdi <- msd_txt("BDI-II")
 staix1 <- msd_txt("STAI X1")

### 2025-10-19 R/compute_descriptives.R

@@ -84,7 +84,7 @@ compute_descriptives <- function(d0, vois) {
     text  <- paste0(dplyr::pull(notes[ , 1]), ": ", dplyr::pull(notes[ , 2])) |> paste(collapse = ", ")
     # Add the text:
     gtab <- gtab |>
-      gt::tab_source_note(html(glue::glue("<i>Note.</i> {text}, All percentages were calculated from the whole sample.")))
+      gt::tab_source_note(html(glue::glue("<i>Note.</i> {text}, all percentages were calculated from the whole sample.")))
   }
   # Return results:
   list(table = demtab, gtable = gtab)

### 2025-10-19 R/gt_apa_table.R

@@ -47,7 +47,7 @@ gt_apa_table <- function(x, grp = NULL, nms = NULL, tit = "") {
         rows = tidyselect::everything()
       )
     ) |>
-    tab_header( # title setup
+    gt::tab_header( # title setup
       title = gt::html(tit)
     ) |>
     gt::opt_align_table_header(align = "left")

### 2025-10-19 R/prepare_defaults.r

@@ -34,10 +34,10 @@ prepare_defaults <- function() {
     nrow = length(sd), # If the defaults ever changed.
     dimnames = lapply(seq_len(2), \(i) c("FAQ", "MMSE", "MoCA", "sMoCA"))
   )
-  sigma <- cor2cov(corrs, sd)
+  sigma <- MBESS::cor2cov(corrs, sd)
   # Censoring values:
   cens <- matrix(
-    c(rep(0, 4), rep(30, 3), 18),
+    c(rep(0, 4), rep(30, 3), 16),
     nrow = 4,
     dimnames = list(c("FAQ", "MMSE", "MoCA", "sMoCA"))
   )
@@ -47,7 +47,7 @@ prepare_defaults <- function() {
     IADL = rep(c("FAQ", "FAQ9"), 3),
     IADL_thres = rep(c(7, 1), 3),
     cognition = c(rep("MMSE", 2), rep("MoCA", 2), rep("sMoCA", 2)),
-    cognition_thres = c(rep(26, 2), rep(26, 2), rep(12, 2))
+    cognition_thres = c(rep(26, 2), rep(26, 2), rep(13, 2))
   )
   # Return:
   list(Mu = mu, Sigma = sigma, cens = cens, crits = crits)

### 2025-10-19 R/simulate_pdd_data.r

@@ -3,8 +3,10 @@
 #' Prepares a minimal example of data in the right format for analysis via the demcrit
 #' package functions. It is compulsory to specify FAQ summary data (mean, variance and
 #' covariance, see below) and at least one additional global cognitive measure summary
-#' data. The function then simulates `N` synthetic patients' PDD status via the
-#' following steps:
+#' data.
+#'
+#' @details
+#' The function simulates `N` synthetic patients' PDD status via the following steps:
 #'
 #' 1. generate latent FAQ and cognitive scores via [MASS::mvrnorm()],
 #' 2. generate observed FAQ item 9 data via the [rbinom()] following tau-equivalence
@@ -19,15 +21,29 @@
 #'    observed data should be used (`TRUE`, default) or the user will provide their
 #'    values instead (`FALSE`).
 #' @param Mu Named numeric vector of means. Needs to contain one component denoting
-#'    FAQ. If `NULL` (default), it is set to values from observed data..
-#' @param Sigma Named matrix of variances and covariances. If `NULL` (default), it
+#'    FAQ. If `defaults = TRUE`, it is set to values from observed data.
+#' @param Sigma Named matrix of variances and covariances. If `defaults = TRUE`, it
 #'    is set at empirical values from the observed data.
 #' @param cens A matrix or data frame with two columns, the first indicating lower
-#'    bound, the second the upper bound of a scale in the rowname. Defaults to
-#'    `NULL` which does not censor data (and prints latent scores instead).
-#' @param crits description
+#'    bound, the second the upper bound of a scale in the rowname. If `defaults == TRUE`,
+#'    or `param = NULL` which does not censor data (and prints latent scores instead).
+#' @param crits A data.frame with named rows and four columns:
+#'    - rownames indicate algorithm labels,
+#'    - `IADL` indicate variable from the data used for IADL deficit calculation,
+#'    - `IADL_thres` indicate threshold above which there is IADL deficit,
+#'    - `cognition` indicate variable from the data used for cognitive deficit calculation,
+#'    - `cognition_thres` indicate threshold below which there is cognitive deficit.
 #'
-#' @returns A data.frame
[…]

### 2025-10-19 R/summarise_rates.R

@@ -17,6 +17,8 @@
 #'    such a table.
 #' @param descending A logical indicating whether algorithms with the highest
 #'    estimated PDD rates should be listed first (default is `TRUE`).
+#' @param plot A logical indicating whether to plot PDD rates as densities.
+#'    The code is dataset-specific (default is `TRUE`).
 #'
 #' @returns A list with:
 #' \describe{
@@ -44,7 +46,7 @@
 #' }
 #'
 #' @export
-summarise_rates <- function(d0, vars, descending = TRUE) {
+summarise_rates <- function(d0, vars, descending = TRUE, plot = TRUE) {
   # Get variables mapping:
   if (is.character(vars)) {
     v <- readr::read_delim(vars, delim = ";", col_types = readr::cols())
@@ -95,7 +97,7 @@ summarise_rates <- function(d0, vars, descending = TRUE) {
     dplyr::select(type, N, Rate) |>
     gt_apa_table() |>
     gt::cols_label(type ~ "Algorithm") |>
-    gt::tab_source_note(gt::html("<i>Note.</i> Percentages were calculated from all available cases. The items comprising each listed algorithm can be found in Table A1."))
+    gt::tab_source_note(gt::html("<i>Note.</i> Percentages were calculated from all available cases, the items comprising each listed algorithm can be found in Table A1."))
   # Make a table showing algorithms' specification:
   gtab_algos <- tab |>
     dplyr::select(-perc, -N, -Rate) |>
@@ -113,51 +115,55 @@ summarise_rates <- function(d0, vars, descending = TRUE) {
     )
   # If there are notes, add them to the table:
   if (ncol(v) > 4 && any(!is.na(v$note))) {
-    notes <- v[complete.cases(v[ , 5]), c(2,5)]
-    text <- paste0(dplyr::pull(notes[ , 1]), ": ", dplyr::pull(notes[ , 2])) |> paste(collapse = ", ")
+    notes <- v[complete.cases(v[, 5]), c(2, 5)]
+    text <- paste0(dplyr::pull(notes[, 1]), ": ", dplyr::pull(notes[, 2])) |> paste(collapse = ", ")
     # Add the text:
     gtab_algos <- gtab_algos |>
       gt::tab_source_note(gt::html(glue::glue("<i>Note.</i> {text}")))
   }
[…]
